### 2019-05-29 README.Rmd
initial commit

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# hotspots

<!-- badges: start -->
<!-- badges: end -->

The goal of hotspots is to ...

## Installation

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
#library(hotspots)
## basic example code
```

## Code of Conduct

Please note that the 'hotspots' project is released with a
[Contributor Code of Conduct](.github/CODE_OF_CONDUCT.md).
By contributing to this project, you agree to abide by its terms.

### 2019-05-29 analysis/paper/paper.Rmd
added analysis subdir

---
title: "Title Goes Here"
author:
  - Author One
  - Author Two
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::html_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
bibliography: references.bib
csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
abstract: |
  Text of abstract
keywords: |
  keyword 1; keyword 2; keyword 3
highlights: |
  These are the highlights. 
---

<!-- This is the format for text comments that will be ignored during renderings. Do not put R code in these comments because it will not be ignored. -->

```{r, setup, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  comment = "#>",
  fig.path = "../figures/"
)

library(hotspots) # Or use devtools::load_all('.', quiet = T) if your code is in script files, rather than as functions in the `/R` diretory
```

# Introduction

Here is a citation [@Marwick2017]

# Background

```{r demo, eval = FALSE}
1 + 1
```

# Methods

# Results

```{r get-data, eval = FALSE}
# Note the path that we need to use to access our data files when rendering this document
my_data <- read.csv(here::here('analysis', 'data', 'raw_data', 'my_csv_file.csv'))
```

# Discussion

# Conclusion

# Acknowledgements

<!-- The following line inserts a page break when the output is MS Word. For page breaks in PDF, use \newpage on its own line.  -->
##### pagebreak

# References 
<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->
<div id="refs"></div>

##### pagebreak

### Colophon

This report was generated on `r Sys.time()` using the following computational environment and dependencies:

```{r colophon, cache = FALSE}
# which R packages and versions?
devtools::session_info()
```

The current Git commit details are:

```{r}
# what commit is this file at? 
git2r::repository(here::here())
```

### 2019-06-06 analysis/paper/paper.Rmd
added call to install R package here
added abstract

@@
 abstract: |
-  Text of abstract
+  High-dimensional traits in multiparental populations of model organisms offer 
+  unprecedented opportunities to understand relationships between genotypic 
+  variation and phenotypic variation. The ability to infer founder haplotypes in 
+  multiparental populations combined with the measurement of molecular phenotypes 
+  in hundreds of mammalian subjects has produced insights into diverse biological 
+  processes.
+
+  Quantitative trait locus (QTL) mapping of one trait at a time provides insights 
+  into genetic underpinnings of complex traits. Furthermore, multivariate QTL
+  mapping enables identification of pleiotropic QTL. Mediation analysis is widely 
+  used to identify causal relationships among pairs of biomolecular traits. We 
+  examine and compare two strategies, mediation analysis and pleiotropy testing, 
+  for understanding relationships among traits. We apply our approach to five 
+  expression trait hotspots that Keller, et al. (2018) identified in pancreatic 
+  islets from nearly 400 Diversity Outbred mice.
+
+  At each hotspot, we identify local transcripts and nonlocal transcripts. We then 
+  perform bivariate pleiotropy tests for all pairs involving one local and one 
+  nonlocal transcript. Our mediation analyses consider each local transcript as a 
+  candidate mediator for each nonlocal transcript. We summarize our findings with 
+  the pleiotropy test statistics and LOD difference statistics. We conclude with 
+  recommended practices for dissection of expression trait hotspots.
 keywords: |
@@
 # Results
+
+```{r}
+install.packages("here")
+```
 

### 2019-06-06 README.Rmd
added travis badge
updated README
added Makefile

@@
 <!-- badges: start -->
+[![Travis-CI Build Status](https://travis-ci.org/fboehm/hotspots.svg?branch=master)](https://travis-ci.org/fboehm/hotspots)
+
 <!-- badges: end -->
@@
 ## Installation
+
+Install from Dockerhub with this command:
+
+```{bash, docker_pull, eval = FALSE}
+docker pull fjboehm/hotspots
+```
+
+Alternatively, if you wish to install the package directly from Github, use `devtools`.
+
+```{r, devtools_stuff, eval = FALSE}
+install.packages("devtools")
+devtools::install_github("fboehm/hotspots")
+```
 
@@
 
+## Build the manuscript `paper.Rmd`
+
+To render the manuscript `paper.Rmd`, into html, tex, and pdf, use a shell to enter the following text from the 'hotspots' directory:
+
+```{bash, docker_build, eval = FALSE}
+docker build .
+```
+
+If you're not using Docker, and have cloned the repository from Github, you can use the Makefile to render `paper.Rmd`. Just type `make` in the shell:
+
+```{bash, make, eval = FALSE}
+make
+```
+
 ## Code of Conduct

### 2019-06-06 analysis/paper/paper.Rmd
started intro
updated README
added Makefile

@@
       reference_docx: "../templates/template.docx" # Insert path for the DOCX file
+    pdf_document:
+      keep_tex: yes
 bibliography: references.bib
@@
 
+Studies of complex traits in mammalian multiparental populations, such as the Diversity Outbred mice, have generated insights into many biological processes, including pancreatic islet gene expression patterns, <!-- look at articles that cite DO papers  -->
+
 # Background

### 2019-06-11 analysis/paper/find-missing-jobs-chtc.Rmd
wrote Rmd file to identify failed chtc jobs

---
title: "Initial analysis of pleiotropy test results"
author: "Frederick Boehm"
date: "6/10/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
path_to_results <- "~/Box Sync/attie/keller-hotspots/results"
```

First, how many results files do I have?

Recall that I submitted 10332 (Chr 11), 1768 (Chr 13), 7644 (Chr 2), 6734 (Chr 5), and 8979 (Chr 7) jobs.

Now, let's count the files in our subdirectories.

```{bash}
cd ~/Box\ Sync/attie/keller-hotspots/results/
cd pvl-run-9901; ls | wc
cd ../pvl-run-9911; ls | wc
cd ../pvl-run-9921; ls | wc
cd ../pvl-run-9931; ls | wc
cd ../pvl-run-9941; ls | wc
```

```{r}
library(tidyverse)
```

```{r}
library(qtl2chtc)
```

```{r}
find_missing_jobs(DIR = "~/Box Sync/attie/keller-hotspots/results/pvl-run-9901", njobs = 6734, outfile = "~/Box Sync/attie/keller-hotspots/submit_files/run9901-bad")
find_missing_jobs(DIR = "~/Box Sync/attie/keller-hotspots/results/pvl-run-9911", njobs = 8979, outfile = "~/Box Sync/attie/keller-hotspots/submit_files/run9911-bad")
find_missing_jobs(DIR = "~/Box Sync/attie/keller-hotspots/results/pvl-run-9921", njobs = 10332, outfile = "~/Box Sync/attie/keller-hotspots/submit_files/run9921-bad")
```

```{r}

```

### 2019-06-11 analysis/paper/paper.Rmd
added repos argument
copied bib file from diss-latex.

@@
       keep_tex: yes
-bibliography: references.bib
+bibliography: research.bib
 csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
@@
 ```{r}
-install.packages("here")
+install.packages("here", repos = "https://cloud.r-project.org")
 ```

### 2019-06-17 analysis/paper/Chr11-prelim.Rmd
added Rmd files for initial analyses - mediation v pleiotropy test statistic

---
title: "Chr11 hotspot preliminary analyses"
author: "Frederick Boehm"
date: "6/16/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
hot_chr <- 11
run_num <- 9921
hot_mid <- 71
```

```{r, load_attie}
load("~/Box Sync/attie/keller2018-chr2-hotspot-chtc/data-to-ignore/Attie_DO378_eQTL_viewer_v1.Rdata")
ls()
library(tidyverse)
```

We need to get the names of the nonlocal traits that map to Chromosome 2 hotspot.

I previously documented the issue with the Attie data set's hotspot annotations. Namely, when that file was created, they permitted the "hotspot" field to be only a single value per transcript. Thus, if a transcript mapped to multiple hotspots, it would still have only a single value in this "hotspot" field. Because of this, I need to first redo the mediation analyses that Keller did.

We want the ensembl ids for the traits that map to Chr `r hot_chr` hotspot. First, define the boundaries of the hotspot. The supplemental table in Keller et al. 2018 lists only the midpoint of the hotspot. The half-width is 2 Mb.

```{r}
hot_lower <- hot_mid - 2
hot_upper <- hot_mid + 2
```

Now, filter to get those traits that map to the hotspot. Note that we're likely to get local traits, too, in this step.

```{r}
dataset.islet.rnaseq$lod.peaks %>%
  filter(chrom == hot_chr, pos <= hot_upper, pos >= hot_lower) %>%
  left_join(y = dataset.islet.rnaseq$annots, by = c("annot.id" = "gene_id")) %>%
  filter(!duplicated(annot.id)) %>%
  filter(lod >= 7.18)
```

The above yields expression traits that map to the hotspot. The list includes both local and nonlocal traits. We'll create two distinct objects, one for locals and one for nonlocal traits.

```{r}
local_annot <- dataset.islet.rnaseq$lod.peaks %>%
  filter(chrom == hot_chr, pos <= hot_upper, pos >= hot_lower) %>%
  left_join(y = dataset.islet.rnaseq$annots, by = c("annot.id" = "gene_id")) %>%
  filter(!duplicated(annot.id)) %>%
  filter(lod >= 7.18) %>%
  mutate(local_tx = chr == hot_chr) %>%
  filter(local_tx)
```

```{r}
dim(local_annot)
```

```{r}
nonlocal_annot <- dataset.islet.rnaseq$lod.peaks %>%
  filter(chrom == hot_chr, pos <= hot_upper, pos >= hot_lower) %>%
  left_join(y = dataset.islet.rnaseq$annots, by = c("annot.id" = "gene_id")) %>%
  filter(!duplicated(annot.id)) %>%
  filter(lod >= 7.18) %>% #genomewide threshold
  mutate(local_tx = chr == hot_chr) %>%
  filter(!local_tx) %>% 
  filter(!is.na(hotspot))
```

```{r}
dim(nonlocal_annot)
```

We'll use only those nonlocal traits that Keller et al. annotated as mapping to a hotspot.

This gives us the nonlocal traits for the Chr `r hot_chr` hotspot, which is what Keller et al. reported.

We also want to look at the local traits for Chr `r hot_chr` hotspot. We want to identify a subset of those for two-dimensional scans. Let's first limit to those near the hotspot, ie, within 5Mb of either end.

```{r, make_local_annot2}
local_annot2 <- local_annot %>%
  filter(middle <= hot_upper + 5, middle >= hot_lower - 5) %>%
  arrange(desc(lod))
```

Also, I'm limiting consideration of local genes to those with middle within 5Mb of the hotspot.

Now, I make the two expression trait matrices that I need for use with Condor.

```{r}
hot_local <- dataset.islet.rnaseq$expr[, colnames(dataset.islet.rnaseq$expr) %in% local_annot2$annot.id]
hot_nonlocal <- dataset.islet.rnaseq$expr[, colnames(dataset.islet.rnaseq$expr) %in% nonlocal_annot$annot.id]
```

Prepare the covariates matrix for input to `fit1_med`.

```{r define_addcovar}
addcovar <- dataset.islet.rnaseq$covar[rownames(dataset.islet.rnaseq$covar) %in% rownames(hot_local), 1:4] 
```

Calls to `fit1_med`

```{r install_intermediate}
devtools::install_github("fboehm/qtl2mediate")
```

We then define the driver object. Note that we use the marker that corresponds to each nonlocal trait's peak position.

```{r define_dri}
# need the index of the peak for every nonlocal trait
indices <- nonlocal_annot %>%
  mutate(marker_index = purrr::map_int(.x = marker.id, .f = function(x){which(names(map[[as.character(hot_chr)]]) == x)})) %>%
  select(marker_index) %>%
  unlist() %>%
  as.integer()
```

```{r}
nonlocal_annot$marker.id[1] -> aa
which(names(map[[as.character(hot_chr)]]) == aa)
```

```{r fit1_med}
nlocal <- ncol(hot_local)
nnonlocal <- ncol(hot_nonlocal)
out <- list()

for (j in 1:nlocal){
  foo <- list()
  for (i in 1:nnonlocal){
    foo[[i]] <- qtl2mediate:::fit1_med_qtl2(driver = genoprobs[[as.character(hot_chr)]][, , indices[i]],
                                             target = hot_nonlocal[ , i],
                                             mediator = hot_local[ , j],
                                             addcovar = addcovar
                                             )
  }
  out[[j]] <- foo
}
```

We need to be careful when creating the `local_gene_id` and `nonlocal_gene_id` objects.

```{r tabulate, message=FALSE, warning = FALSE}
tib <- out %>%
  purrr::map(bind_rows) %>%
  bind_rows() %>%
  as_tibble() %>%
  rename(lod_without_med = lod_no_med, lod_with_med = lod_med) %>%
  mutate(lod_diff = lod_without_med - lod_with_med, lod_diff_proportion = lod_diff / lod_without_med) %>%
  mutate(local_gene_id = rep(colnames(hot_local), each = nnonlocal), nonlocal_gene_id = rep(colnames(hot_nonlocal), times = nlocal)) %>%
  left_join(local_annot2, by = c("local_gene_id" = "annot.id")) %>%
  left_join(nonlocal_annot, by = c("nonlocal_gene_id" = "annot.id")) %>%
  select(symbol.x, symbol.y, lod_without_med, lod_with_med, lod_diff, lod_diff_proportion, local_gene_id, nonlocal_gene_id) %>%
  rename(local_symbol = symbol.x, nonlocal_symbol = symbol.y)
```

We also added gene symbols to the tibble.

## Reading pleiotropy test results

```{r read_pleio_results}
results_dir <- paste0("~/Box Sync/attie/keller-hotspots/results/pvl-run-", run_num)
dir(results_dir) -> fns
mylist <- list()
for (i in 1:length(fns)){
  mylist[[i]] <- read.table(file.path(results_dir, fns[i])) %>%
  as_tibble() %>%
  mutate(filename = fns[i]) %>%
  mutate(local_gene_id = stringr::str_split(fns[i], "_")[[1]][[3]], nonlocal_gene_id = stringr::str_split(stringr::str_split(fns[i], "_")[[1]][[4]], ".txt")[[1]][[1]]) %>%
  filter(lod == max(lod), trace == "profile1") %>%
  select(local_gene_id, nonlocal_gene_id, lod, filename)
}
```

## Scatter plots of LOD difference & pleiotropy test statistics

```{r plot_all}
p <- mylist %>%
  bind_rows() %>%
  rename(pleiotropy_lod = lod) %>%
  left_join(tib, by = c("local_gene_id", "nonlocal_gene_id")) %>%
  select(local_symbol, nonlocal_symbol, pleiotropy_lod, lod_diff, lod_diff_proportion, lod_without_med, lod_with_med, filename, nonlocal_gene_id, local_gene_id) %>%
  ggplot() + geom_point(mapping = aes(x = pleiotropy_lod, y = lod_diff, colour = local_symbol, nonlocal = nonlocal_symbol), size = 0.1)
plotly::ggplotly(p)

```

```{r, plotly_by_local, fig.height=ceiling(nlocal / 2)}
p <- mylist %>%
  bind_rows() %>%
  rename(pleiotropy_lod = lod) %>%
  left_join(tib, by = c("local_gene_id", "nonlocal_gene_id")) %>%
  select(local_symbol, nonlocal_symbol, pleiotropy_lod, lod_diff, lod_diff_proportion, lod_without_med, lod_with_med, filename, nonlocal_gene_id, local_gene_id) %>%
  ggplot() + geom_point(mapping = aes(x = pleiotropy_lod, y = lod_diff, colour = local_symbol, nonlocal = nonlocal_symbol), size = 0.1) + facet_wrap(. ~ local_symbol, nrow = ceiling(nlocal / 2), ncol = 2) + theme(legend.position = "none")
plotly::ggplotly(p)
```

```{r, plotly_by_nonlocal, fig.height=ceiling(nnonlocal / 4)}
p <- mylist %>%
  bind_rows() %>%
  rename(pleiotropy_lod = lod) %>%
  left_join(tib, by = c("local_gene_id", "nonlocal_gene_id")) %>%
  select(local_symbol, nonlocal_symbol, pleiotropy_lod, lod_diff, lod_diff_proportion, lod_without_med, lod_with_med, filename, nonlocal_gene_id, local_gene_id) %>%
  ggplot() + geom_point(mapping = aes(x = pleiotropy_lod, y = lod_diff, colour = local_symbol, nonlocal = nonlocal_symbol), size = 0.1) + facet_wrap(. ~ nonlocal_symbol, ncol = 4, nrow = ceiling(nnonlocal / 4)) + theme(legend.position = "none")
plotly::ggplotly(p)
```

### 2019-06-17 analysis/paper/Chr13-prelim.Rmd
added Rmd files for initial analyses - mediation v pleiotropy test statistic

--- 2019-06-17 analysis/paper/Chr11-prelim.Rmd
@@
 ---
-title: "Chr11 hotspot preliminary analyses"
+title: "Chr13 hotspot preliminary analyses"
 author: "Frederick Boehm"
@@
 knitr::opts_chunk$set(echo = TRUE)
-hot_chr <- 11
-run_num <- 9921
-hot_mid <- 71
+hot_chr <- 13
+run_num <- 9931
+hot_mid <- 112.5
 ```

### 2019-06-17 analysis/paper/Chr2-prelim.Rmd
added Rmd files for initial analyses - mediation v pleiotropy test statistic

--- 2019-06-17 analysis/paper/Chr11-prelim.Rmd
@@
 ---
-title: "Chr11 hotspot preliminary analyses"
+title: "Chr2 hotspot preliminary analyses"
 author: "Frederick Boehm"
@@
 knitr::opts_chunk$set(echo = TRUE)
-hot_chr <- 11
-run_num <- 9921
-hot_mid <- 71
+hot_chr <- 2
+run_num <- 9941
+hot_mid <- 165.5
 ```

### 2019-06-17 analysis/paper/Chr5-prelim.Rmd
added Rmd files for initial analyses - mediation v pleiotropy test statistic

---
title: "Chr5 hotspot preliminary analyses"
author: "Frederick Boehm"
date: "6/11/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, load_attie}
load("~/Box Sync/attie/keller2018-chr2-hotspot-chtc/data-to-ignore/Attie_DO378_eQTL_viewer_v1.Rdata")
ls()
library(tidyverse)
```

```{r checks}
dataset.islet.rnaseq$expr %>%
  dim
colnames(dataset.islet.rnaseq$expr) %>%
  head
```

We need to get the names of the nonlocal traits that map to Chromosome 5 hotspot.

```{r, make_Chr5_tib}
dataset.islet.rnaseq$annots %>%
  colnames
#dataset.islet.rnaseq$annots$hotspot
dataset.islet.rnaseq$annots %>%
  filter(hotspot == "chr5") -> hotspot_annot_tib
dim(hotspot_annot_tib)
```

Keller et al., in their supplementary files, reported 182 transcripts, yet I've isolated only 180 here.

I previously documented the issue with the Attie data set's hotspot annotations. Namely, when that file was created, they permitted the "hotspot" field to be only a single value per transcript. Thus, if a transcript mapped to multiple hotspots, it would still have only a single value in this "hotspot" field. Because of this, I need to first redo the mediation analyses that Keller did.

Specifically, I'll isolate all ~600 or so transcripts that map to at least one hotspot per the hotspot annotation. I'll then perform mediation analyses with each (for the appropriate local transcript).

```{r}
dataset.islet.rnaseq$annots %>%
  filter(!is.na(hotspot)) -> hotspot_all_tx
dim(hotspot_all_tx)
```

I have 671 transcripts. Let's examine all 671 at each of the four hotspots that I have yet to study, ie, those on Chromosomes 5, 7, 11, and 13. Actually, before I do that, I want to get the genome-wide scan results for these 671 transcripts to see, for each, which hotspots they map to.

```{r get_671_tx}
tx671 <- dataset.islet.rnaseq$expr[ , colnames(dataset.islet.rnaseq$expr) %in% hotspot_all_tx$gene_id]
dim(tx671)  
```

```{r load_qtl2}
library(qtl2)
```

I forgot that the LOD peaks for all traits are saved in `dataset.islet.rnaseq$lod.peaks`. It has nearly 40,000 peaks recorded. I really just need to look at it. That is, for now, I don't really need the output of `scan1()`.

We want the ensembl ids for the traits that map to Chr 5 hotspot. First, define the boundaries of the hotspot. The supplemental table in Keller et al. 2018 lists only the midpoint of the hotspot. The half-width is 2 Mb.

```{r}
c5hot_lower <- 144
c5hot_upper <- 148
```

Now, filter to get those traits that map to the hotspot. Note that we're likely to get local traits, too, in this step.

```{r}
dataset.islet.rnaseq$lod.peaks %>%
  filter(chrom == 5, pos <= c5hot_upper, pos >= c5hot_lower) %>%
  left_join(y = dataset.islet.rnaseq$annots, by = c("annot.id" = "gene_id")) %>%
  filter(!duplicated(annot.id)) %>%
  filter(lod >= 7.18)
```

The above yields 226 expression traits that map to the hotspot on Chr 5. The list includes both local and nonlocal traits. We'll create two distinct objects, one for locals and one for nonlocal traits.

```{r}
local_annot <- dataset.islet.rnaseq$lod.peaks %>%
  filter(chrom == 5, pos <= c5hot_upper, pos >= c5hot_lower) %>%
  left_join(y = dataset.islet.rnaseq$annots, by = c("annot.id" = "gene_id")) %>%
  filter(!duplicated(annot.id)) %>%
  filter(lod >= 7.18) %>%
  mutate(local_tx = chr ==5) %>%
  filter(local_tx)
```

```{r}
dim(local_annot)
```

```{r}
nonlocal_annot <- dataset.islet.rnaseq$lod.peaks %>%
  filter(chrom == 5, pos <= c5hot_upper, pos >= c5hot_lower) %>%
  left_join(y = dataset.islet.rnaseq$annots, by = c("annot.id" = "gene_id")) %>%
  filter(!duplicated(annot.id)) %>%
  filter(lod >= 7.18) %>% #genomewide threshold
  mutate(local_tx = chr == 5) %>%
  filter(!local_tx) %>% 
  filter(!is.na(hotspot))
```

```{r}
dim(nonlocal_annot)
```

We'll use only those nonlocal traits that Keller et al. annotated as mapping to a hotspot.

This gives us 182 traits for Chr 5 hotspot, which is what Keller et al. reported.

We also want to look at the local traits for Chr5 hotspot. We have 42 local traits. We want to identify a subset of those for two-dimensional scans. Let's first limit to those near the hotspot, ie, within 5Mb of either end.

```{r}
local_annot2 <- local_annot %>%
  filter(middle <= c5hot_upper + 5, middle >= c5hot_lower - 5) %>%
  arrange(desc(lod))
```

Recall that *Pdx1* is the gene that Keller et al. identified as the key mediator for Chr 5 hotspot. Its LOD is only 15, which seems to indicate that I should be looking at a broad set of local traits, not just those, eg., with LOD > 40!

Also, I'm limiting consideration of local genes to those with middle within 5Mb of the hotspot.

Now, I make the two expression trait matrices that I need for use with Condor.

```{r}
Chr5hot_local <- dataset.islet.rnaseq$expr[, colnames(dataset.islet.rnaseq$expr) %in% local_annot2$annot.id]
Chr5hot_nonlocal <- dataset.islet.rnaseq$expr[, colnames(dataset.islet.rnaseq$expr) %in% nonlocal_annot$annot.id]
```

Prepare the covariates matrix for input to `fit1_med`.

```{r define_addcovar}
addcovar <- dataset.islet.rnaseq$covar[rownames(dataset.islet.rnaseq$covar) %in% rownames(Chr5hot_local), 1:4] 
```

Calls to `fit1_med`

```{r install_intermediate}
devtools::install_github("fboehm/qtl2mediate")
```

We then define the driver object. Note that we use the marker that corresponds to each nonlocal trait's peak position.

```{r define_dri}
# need the index of the peak for every nonlocal trait
indices <- nonlocal_annot %>%
  mutate(marker_index = purrr::map_int(.x = marker.id, .f = function(x){which(names(map$`5`) == x)})) %>%
  select(marker_index) %>%
  unlist() %>%
  as.integer()
```

```{r}
nonlocal_annot$marker.id[1] -> aa
which(names(map$`5`) == aa)
```

```{r fit1_med}
nlocal <- ncol(Chr5hot_local)
nnonlocal <- ncol(Chr5hot_nonlocal)
out <- list()
out2 <- list()
for (j in 1:nlocal){
  foo <- list()
  foo2 <- list()
  for (i in 1:nnonlocal){
    foo[[i]]<- qtl2mediate::fit1_med(driver = genoprobs$`5`[rownames(genoprobs$`5`) %in% rownames(Chr5hot_local) , , indices[i]], 
                                      # genoprobs at peak position for nonlocal trait
                                      target = Chr5hot_nonlocal[ , i], 
                                      mediator = Chr5hot_local[ , j], 
                                      addcovar = addcovar
         )
    foo2[[i]] <- qtl2mediate:::fit1_med_qtl2(driver = genoprobs$`5`[, , indices[i]],
                                             target = Chr5hot_nonlocal[ , i],
                                             mediator = Chr5hot_local[ , j],
                                             addcovar = addcovar
                                             )
  }
  out[[j]] <- foo
  out2[[j]] <- foo2
}
```

We need to be careful when creating the `local_gene_id` and `nonlocal_gene_id` objects.

```{r tabulate, message=FALSE, warning = FALSE}
tib <- out %>%
  purrr::map(bind_rows) %>%
  bind_rows() %>%
  as_tibble() %>%
  mutate(local_gene_id = rep(colnames(Chr5hot_local), each = nnonlocal), nonlocal_gene_id = rep(colnames(Chr5hot_nonlocal), times = nlocal)) %>%
  mutate(lod_without_med = (ll_d - ll_1) / log(10), lod_with_med = (ll_dm - ll_1m) / log(10), lod_diff = lod_without_med - lod_with_med, lod_diff_proportion = lod_diff / lod_without_med) %>%
  left_join(local_annot2, by = c("local_gene_id" = "annot.id")) %>%
  left_join(nonlocal_annot, by = c("nonlocal_gene_id" = "annot.id")) %>%
  select(symbol.x, symbol.y, lod_without_med, lod_with_med, lod_diff, lod_diff_proportion, local_gene_id, nonlocal_gene_id) %>%
  rename(local_symbol = symbol.x, nonlocal_symbol = symbol.y)
```

We also added gene symbols to the tibble.

## Reading pleiotropy test results

```{r read_pleio_results}
results_dir <- "~/Box Sync/attie/keller-hotspots/results/pvl-run-9901"
dir(results_dir) -> fns
mylist <- list()
for (i in 1:length(fns)){
  mylist[[i]] <- read.table(file.path(results_dir, fns[i])) %>%
  as_tibble() %>%
  mutate(filename = fns[i]) %>%
  mutate(local_gene_id = stringr::str_split(fns[i], "_")[[1]][[3]], nonlocal_gene_id = stringr::str_split(stringr::str_split(fns[i], "_")[[1]][[4]], ".txt")[[1]][[1]]) %>%
  filter(lod == max(lod), trace == "profile1") %>%
  select(local_gene_id, nonlocal_gene_id, lod, filename)
}
mylist %>%
  bind_rows() %>%
  rename(pleiotropy_lod = lod) %>%
  left_join(tib, by = c("local_gene_id", "nonlocal_gene_id")) %>%
  select(local_symbol, nonlocal_symbol, pleiotropy_lod, lod_diff, lod_diff_proportion, lod_without_med, lod_with_med, filename, nonlocal_gene_id, local_gene_id) %>%
  group_by(nonlocal_symbol) %>%
  filter(lod_diff == max(lod_diff)) %>%
  arrange(local_symbol, desc(lod_diff))
```

We want to identify, for each nonlocal transcript, which local transcript gives the greatest `lod_diff` value.

```{r}
(counts1 <- mylist %>%
  bind_rows() %>%
  rename(pleiotropy_lod = lod) %>%
  left_join(tib, by = c("local_gene_id", "nonlocal_gene_id")) %>%
  select(local_symbol, nonlocal_symbol, pleiotropy_lod, lod_diff, lod_diff_proportion, lod_without_med, lod_with_med, filename, nonlocal_gene_id, local_gene_id) %>%
  group_by(nonlocal_symbol) %>%
  filter(lod_diff == max(lod_diff)) %>%
  arrange(local_symbol, desc(lod_diff)) %>%
  plyr::ddply(.variables = "local_symbol", .fun = count) %>%
  arrange(desc(n)) %>%
  rename(n_no_pleio_thresh = n)
)
```

```{r}
pleiotropy_threshold <- 3
mylist %>%
  bind_rows() %>%
  rename(pleiotropy_lod = lod) %>%
  left_join(tib, by = c("local_gene_id", "nonlocal_gene_id")) %>%
  select(local_symbol, nonlocal_symbol, pleiotropy_lod, lod_diff, lod_diff_proportion, lod_without_med, lod_with_med, filename, nonlocal_gene_id, local_gene_id) %>%
  group_by(nonlocal_symbol) %>%
  filter(pleiotropy_lod < pleiotropy_threshold) %>%
  filter(lod_diff == max(lod_diff)) %>%
  arrange(local_symbol, desc(lod_diff)) %>%
  plyr::ddply(.variables = "local_symbol", .fun = count) %>%
  arrange(desc(n)) %>%
  rename(n_with_pleio_thresh_3 = n) %>%
  left_join(counts1)
```

```{r}
mylist %>%
  bind_rows() %>%
  rename(pleiotropy_lod = lod) %>%
  left_join(tib, by = c("local_gene_id", "nonlocal_gene_id")) %>%
  select(local_symbol, nonlocal_symbol, pleiotropy_lod, lod_diff, lod_diff_proportion, lod_without_med, lod_with_med, filename, nonlocal_gene_id, local_gene_id) %>%
  filter(local_symbol == "Rpl21") %>%
  arrange(desc(lod_diff)) %>%
  filter(pleiotropy_lod < pleiotropy_threshold)
```

```{r}
mylist %>%
  bind_rows() %>%
  rename(pleiotropy_lod = lod) %>%
  left_join(tib, by = c("local_gene_id", "nonlocal_gene_id")) %>%
  select(local_symbol, nonlocal_symbol, pleiotropy_lod, lod_diff, lod_diff_proportion, lod_without_med, lod_with_med, filename, nonlocal_gene_id, local_gene_id) %>%
  filter(local_symbol == "Pdx1") %>%
  arrange(desc(lod_without_med)) %>%
  filter(pleiotropy_lod < pleiotropy_threshold)

```

## Scatter plots of LOD difference & pleiotropy test statistics

```{r}
p <- mylist %>%
  bind_rows() %>%
  rename(pleiotropy_lod = lod) %>%
  left_join(tib, by = c("local_gene_id", "nonlocal_gene_id")) %>%
  select(local_symbol, nonlocal_symbol, pleiotropy_lod, lod_diff, lod_diff_proportion, lod_without_med, lod_with_med, filename, nonlocal_gene_id, local_gene_id) %>%
  ggplot() + geom_point(mapping = aes(x = pleiotropy_lod, y = lod_diff, colour = local_symbol, nonlocal = nonlocal_symbol), size = 0.1)
plotly::ggplotly(p)

```

```{r, plotly_by_local, fig.height=20}
p <- mylist %>%
  bind_rows() %>%
  rename(pleiotropy_lod = lod) %>%
  left_join(tib, by = c("local_gene_id", "nonlocal_gene_id")) %>%
  select(local_symbol, nonlocal_symbol, pleiotropy_lod, lod_diff, lod_diff_proportion, lod_without_med, lod_with_med, filename, nonlocal_gene_id, local_gene_id) %>%
  ggplot() + geom_point(mapping = aes(x = pleiotropy_lod, y = lod_diff, colour = local_symbol, nonlocal = nonlocal_symbol), size = 0.1) + facet_wrap(. ~ local_symbol, nrow = 20, ncol = 2) + theme(legend.position = "none")
plotly::ggplotly(p)
```

```{r, plotly_by_nonlocal, fig.height=46}
p <- mylist %>%
  bind_rows() %>%
  rename(pleiotropy_lod = lod) %>%
  left_join(tib, by = c("local_gene_id", "nonlocal_gene_id")) %>%
  select(local_symbol, nonlocal_symbol, pleiotropy_lod, lod_diff, lod_diff_proportion, lod_without_med, lod_with_med, filename, nonlocal_gene_id, local_gene_id) %>%
  ggplot() + geom_point(mapping = aes(x = pleiotropy_lod, y = lod_diff, colour = local_symbol, nonlocal = nonlocal_symbol), size = 0.1) + facet_wrap(. ~ nonlocal_symbol, nrow = 46, ncol = 4) + theme(legend.position = "none")
plotly::ggplotly(p)
```

## Are the lods from fit1_med correct?

Yes.

I see that a large number of nonlocal transcripts have `lod_without_med` values below 7.18. While the 7.18 threshold was determined with the mixed model, I still find the large number of transcripts unusual.

First, a given nonlocal transcript should have a single value for `lod_without_med`, regardless of which local transcript it's paired with. Let's verify this.

```{r, check1}
mylist %>%
  bind_rows() %>%
  rename(pleiotropy_lod = lod) %>%
  left_join(tib, by = c("local_gene_id", "nonlocal_gene_id")) %>%
  select(local_symbol, nonlocal_symbol, pleiotropy_lod, lod_diff, lod_diff_proportion, lod_without_med, lod_with_med, filename, nonlocal_gene_id, local_gene_id) %>% 
  arrange(nonlocal_symbol)
```

```{r, check_table}
mylist %>%
  bind_rows() %>%
  select(lod) %>%
  mutate(indic = lod > 5) %>%
  select(indic) %>%
  table()
```

### 2019-06-17 analysis/paper/Chr7-prelim.Rmd
added Rmd files for initial analyses - mediation v pleiotropy test statistic

--- 2019-06-17 analysis/paper/Chr11-prelim.Rmd
@@
 ---
-title: "Chr11 hotspot preliminary analyses"
+title: "Chr7 hotspot preliminary analyses"
 author: "Frederick Boehm"
@@
 knitr::opts_chunk$set(echo = TRUE)
-hot_chr <- 11
-run_num <- 9921
-hot_mid <- 71
+hot_chr <- 7
+run_num <- 9911
+hot_mid <- 46
 ```

### 2019-06-17 analysis/paper/find-missing-jobs-chtc.Rmd
added Rmd files for initial analyses - mediation v pleiotropy test statistic

@@
 
-```{r}
+```{r, eval = TRUE}
 find_missing_jobs(DIR = "~/Box Sync/attie/keller-hotspots/results/pvl-run-9901", njobs = 6734, outfile = "~/Box Sync/attie/keller-hotspots/submit_files/run9901-bad")

### 2019-06-17 analysis/paper/paper.Rmd
removed Marwick citation
added Rmd files for initial analyses - mediation v pleiotropy test statistic

@@
 
-Here is a citation [@Marwick2017]
+Studies of complex traits in mammalian multiparental populations, such as the Diversity Outbred mice, have generated insights into many biological processes, including pancreatic islet gene expression patterns, <!-- look at articles that cite DO papers  -->
 
-Studies of complex traits in mammalian multiparental populations, such as the Diversity Outbred mice, have generated insights into many biological processes, including pancreatic islet gene expression patterns, <!-- look at articles that cite DO papers  -->
+@chick2016defining
 
@@
 
-```{r demo, eval = FALSE}
-1 + 1
-```
+# Methods
 
-# Methods
+@keller2018genetic studied gene expression in pancreatic islets of 378 Diversity Outbred mice.
+They used RNA sequencing technology to quantify 21,771 distinct transcripts.
+They publicly shared these data at this url.
+
+We performed pleiotropy tests and mediation analyses at each of five hotspots that @keller2018genetic identified. They are located on Chromosomes 2, 5, 7, 11, and 13 .
 
@@
 
-```{r}
-install.packages("here", repos = "https://cloud.r-project.org")
-```
+## Chr 2 hotspot
+
+## Chr 5 hotspot
+
+## Chr 7 hotspot
+
+## Chr 11 hotspot
+
+## Chr 13 hotspot
 

### 2019-06-17 analysis/paper/paper.Rnw
added Rmd files for initial analyses - mediation v pleiotropy test statistic

\documentclass[oneside]{book}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage[utf8]{inputenc}
%\usepackage{lineno}
%\linenumbers

\usepackage{bbm}
\usepackage{setspace}
\usepackage[left = 1in, right = 1in, top = 1.5in, bottom = 1in]{geometry}
\usepackage{multirow}
\usepackage[framemethod=TikZ]{mdframed}
\usepackage{float}
\usepackage{newfloat}
\usepackage{multicol}
%\usepackage[lite]{mtpro2} %contains \widetilde
\usepackage[titletoc]{appendix}
\usepackage{amsmath}
\usepackage{hyperref}
\hypersetup{unicode=true,
%            pdftitle={Chapter 1},
%            pdfauthor={Frederick Boehm},
%            pdfborder={0 0 0},
            colorlinks=true,
            breaklinks=true}

\usepackage[
    backend=biber,
    style=chicago-authordate,
%    citestyle=authoryear,
    natbib=true, %citet and citep
    url=true,
    doi=true,
    eprint=true,
    maxbibnames=10,
    backref=true
]{biblatex}
\DeclareLanguageMapping{english}{english-apa}
\addbibresource{research.bib}
\usepackage{graphicx}
\usepackage{color, colortbl}
\definecolor{LightCyan}{rgb}{0.88,1,1}
\usepackage{subcaption}
\usepackage{array}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{longtable}
\usepackage[outdir=./figs/]{epstopdf}
\usepackage{etoolbox} % for messing with page style & title pages
\usepackage{fancyhdr}
\setlength{\headheight}{15.2pt}
\pagestyle{fancyplain}
\fancyhf{}
\lhead{}
\chead{}
%\rhead{}
\rhead{ \fancyplain{}{\thepage} } % put page number in upper right corner
\renewcommand{\headrulewidth}{0pt} % remove top line
\renewcommand{\footrulewidth}{0pt} % remove bottom line
\renewcommand{\headrule}{}
\title{The title}
\author{An Author}
\patchcmd{\titlepage}{empty}{fancy}{}{}
\patchcmd{\chapter}{plain}{fancy}{}{}
\DeclareCaptionType{equ}[][]
%https://stackoverflow.com/questions/149479/adding-a-caption-to-an-equation-in-latex
%\captionsetup[equ]{labelformat=empty}

\DeclareFloatingEnvironment[fileext=frm,placement={!ht},name=Frame]{myfloat}
\captionsetup[myfloat]{labelfont=bf}
\newenvironment{frameenv}[1]
    {\begin{myfloat}[tb]
    \begin{mdframed}[roundcorner=10pt,backgroundcolor=blue!10]
    \caption{#1}
    %\begin{multicols*}{2}
    }
    {%\end{multicols*}
    \end{mdframed}\end{myfloat}
    }

\title{}
\author{Frederick J. Boehm, Brian S. Yandell, and Karl W. Broman}
\date{\today}

%% from vignette tex file
\urlstyle{same}  % don't use monospace font for urls
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}

%% end from vignette tex file

\begin{document}

<<include=FALSE>>=
library(knitr)
opts_chunk$set(
concordance=TRUE, echo = FALSE, fig.lp = "fig:", warning = FALSE, message = FALSE, results = "hide"
)
@

\doublespacing
\begin{titlepage}\thispagestyle{empty}
\tableofcontents
%\listoffigures
%\listoftables
%\pagestyle{headings} % page numbers appear at bottom of page, centered!
%\markright{\hfill}
\chapter{Introduction}

Genetics studies in model organisms like mice can identify
genomic regions that affect complex traits, such as systolic blood pressure and body weight \citep{sax1923association,soller1976power,lander1989mapping,broman2009guide,jansen2007quantitative}. These genomic regions are called ``quantitative trait loci'' or ``QTL''.
A genome-wide QTL ``scan'' reveals
associations between genotypes and phenotypes by considering each position,
one at a time, as a candidate QTL for the trait of interest.
A region with strong evidence of association with a complex trait, then, defines a QTL (for that trait).
Because nearby markers have correlated
genotypes, a QTL in a two-parent cross often spans multiple megabases in length
and may contain more than a hundred genes.
Identification of the causal gene (for a given complex trait) from among those genes near the QTL is
challenging and may require costly and time-consuming experiments.
The growing need for greater QTL mapping resolution fueled development over the
last two decades of model organism multiparental populations for high-resolution QTL mapping \citep{de2017back,churchill2004collaborative,svenson2012high,huang2012multiparent,shivakumar2018soybean,huang2011analysis,kover2009multiparent,tisne2017identification,stanley2017genetic}.
With experimentalists now measuring tens of thousands of biomolecular traits
in multiparental populations, the systems genetics community needs multivariate
statistical tools to fully examine the large volumes of data \citep{keller2018genetic,chick2016defining}.
Identifying loci that affect multiple traits can aid in identifying biomolecular interactions and in clarifying complex trait genetic architecture. A QTL that affects more than one trait is called a ``pleiotropic'' QTL.
A test of pleiotropy vs. separate QTL is one multivariate statistical tool that will inform
complex trait genetics by enabling researchers to identify the number of
unique QTL in a genomic region of interest.
In this thesis, I develop a test of pleiotropy vs. separate QTL in multiparental
populations. I study its statistical properties and demonstrate its utility
by analyzing experimental data.

<<load-broman>>=
library(broman)
@

Due to the breeding scheme, each Diversity Outbred mouse is a highly
heterozygous and essentially unique mosaic of
founders' DNA. Figure~\ref{fig:do} presents color-coded chromosomes for a collection of
subjects from a multiparental population like the Diversity Outbred mice. DNA from each founder
line is color-coded. Each pair of vertical bars corresponds to one subject. Each row corresponds to a single generation, with generation labels on the left. Generation number increases from top to bottom.
With each new generation, the Diversity Outbred mouse genomes are shuffled
in smaller and smaller chunks.

\section{QTL mapping in Diversity Outbred mice}\label{sec:do-qtl}

QTL mapping in Diversity Outbred mice, like that in mice from two-parent
crosses, is a multi-step procedure: 1. data acquisition, 2. inference of
missing genotypes, and 3. modeling phenotypes as a function of genotypes.
Data acquisition involves measurement of phenotypes and, at specified
genetic markers, termed ``single nucleotide polymorphism'' or SNP markers,
measurement of two-allele genotypes. Often, the SNP marker genotypes are
obtained by use of a microarray, such as the GigaMUGA SNP microarray
\citep{morgan2015mouse}.

The next step, missing genotypes inference, is needed because of the
``missing data problem'' \citep{broman2009guide}. It takes as input the two-allele
genotypes at the measured SNP markers. Hidden Markov model methods, developed by
\citet{broman2012haplotype} and \citet{broman2012genotype} and implemented in the \texttt{qtl2} R
package \citep{broman2019rqtl2}, output 36-state genotype probabilities for all
nuclear autosomal markers and pseudomarkers. The outputted genotypes have 36 states
because each position has eight homozygote states (one for each founder line) and $\binom{8}{2}$ heterozygote states. Pseudomarkers are
arbitrary positions at which the researcher wants 36-state genotype probabilities.
Finally, I collapse the 36-state genotype probabilities to eight founder allele
dosages at each marker. This last step is optional, but often helpful, because the
simplified models require specification of fewer parameters. We then treat the
founder allele dosages as known quantities in subsequent steps.

% model form - connecting genotype to phenotype

After inferring founder allele dosages, I address the
second major statistical challenge of
QTL mapping: the ``model selection'' problem \citep{broman2009guide}.
A genetically ``additive'' linear model, in which
I assume a linear relationship between
a trait's values and each founder allele's dosage, is a popular default model.
\citep{gatti2014quantitative,broman2019rqtl2}.

% model form - random effects incorporation & relatedness
% First, relatedness
Due to the complexity of the breeding design, Diversity Outbred mice
have
complicated pairwise relationships. A given study cohort may include pairs of first
cousins, parent-offspring pairs, grandparent-grandchild pairs, second cousin pairs, and others.
Because relatedness can confound genotype-phenotype associations
\citep{yang2014advantages}, researchers have developed methods to account
for relatedness in their statistical models. One popular approach involves
use of a polygenic random effect in the statistical model \citep{kang2008efficient}.
The inclusion of a random effect in a model with fixed (\emph{i.e.}, nonrandom)
effects results in what statisticians call a ``linear mixed effects model''.

One then fits a linear mixed effects model and calculates the log$_{10}$ likelihood at each position. As in two-parent crosses (Section~\ref{sec:qtl-two-parent}), one summarizes evidence for a QTL by plotting LOD scores across the genome. Because multiple hypothesis tests are performed, one typically wants to control family-wise error rate. One strategy to do this involves a permutation test, much like that in two-parent crosses \citep{churchill1994empirical}.

%%% CHAPTER 1 End
%%%%%%%%%%%%%%%%%
%%%%% 3A start
\section{Expression trait hotspot dissection}\label{sec:hotspot-dissection}
\subsection{Introduction}

A central goal of systems genetics studies is to identify causal relationships between biomolecules.
Recent work by \citet{chick2016defining}, which builds on research from \citet{baron1986moderator}, has popularized linear regression-based mediation analysis for causal inference in systems genetics.
Because of the great successes of mediation analysis in systems genetics
\citep{chick2016defining,keller2018genetic}, we need to clarify a role for our pleiotropy test.
We argue below that our pleiotropy test complements mediation analysis.

Second, when regression-based mediation analysis fails to identify a mediator, our
pleiotropy test still provides information on the number of QTL, which may aid biological
understanding and inform subsequent studies. Below, we first review the prerequisite
molecular biology, including the ``central dogma'', before discussing in some detail
regression-based mediation analysis in the systems genetics context.

\citet{crick1958protein} articulated a pathway for transmission of biological information
that is now known as the central dogma of molecular biology (Figure~\ref{fig:dogma})
\citep{crick1970central}.
In it, he argued that information encoded in DNA sequence is transmitted via transcription
to
RNA molecules, which, in turn, transfer the information to proteins via translation.
The process of transcription uses DNA as a template for creating a RNA molecule that
conveys the information encoded in DNA.
In this sense, every gene leads to a unique RNA molecule.
Translation is the molecular biology process by which an RNA molecule's information is
transferred to a protein.
As in transcription, the nucleic acid (DNA in transcription and RNA in translation) serves
as a template for synthesis of the product (RNA in transcription and protein in translation).
Thus, RNA molecules from distinct genes lead to different proteins.

This sequence of information transfer, from DNA to RNA to protein, provides a natural
setting by which to examine mediation analysis. If a DNA variant affects protein
concentrations only through its gene's RNA transcripts, then conditioning on RNA transcript
levels would greatly reduce the strength of association between DNA variant and protein
concentration. Before we continue our discussion, we define key terms and discuss an
example below.

We continue by stating what it means for one trait to mediate a relationship between a
DNA variant and another trait. To clarify our discussion, we refer to an example from
\citet{chick2016defining} (Figure~\ref{fig:Dhtkd1}). \citet{chick2016defining}, in
studying livers of 192 Diversity Outbred mice, found evidence that \emph{Dhtkd1}
transcript levels associated with a Chromosome 2 marker near the \emph{Dhtkd1} gene.
They also found that
the same marker affected DHTKD1 protein concentrations. (Note that DHTKD1 protein is the
product of translation of \emph{Dhtkd1} transcripts.)
As anticipated, mediation analysis, in which the DHTKD1 protein concentrations are
regressed on founder allele dosages (at the Chromosome 2 marker) demonstrated that
\emph{Dhtkd1} transcript levels act as a mediator between DHTKD1 protein concentrations
and founder allele dosages. In fact, the extent of the reduction in association
strength indicates that the primary pathway by which the genetic marker affects DHTKD1
protein concentrations is through \emph{Dhtkd1} transcript levels.

\begin{figure}
  \centering
  \includegraphics[width = 0.7\textwidth]{../QTLfigs/Figs/central_dogma.pdf}
  \caption{Biological information is encoded in DNA. This information is passed, via transcription, to sequence-specific RNA molecules. The process of translation transmits the information to sequence-specific proteins.}\label{fig:dogma}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width = 0.7\textwidth]{../QTLfigs/Figs/central_dogma-Dhtkd1.pdf}
  \caption{A DNA variant in the \emph{Dhtkd1} gene affects \emph{Dhtkd1} transcript abundances which, in turn, affect DHTKD1 protein concentrations.}\label{fig:Dhtkd1}
\end{figure}

In our studies below, we follow \citet{keller2018genetic} by generalizing this setting to
the case where a DNA variant affects a local transcript level, which then affects a
nonlocal transcript level.
We term a transcript ``local'' to a marker when its gene is near that marker.
We use a threshold of no more than 2 Mb to restrict the
number of local transcripts for a given marker.
A nonlocal transcript, then, is either one that arises from a gene on another chromosome or
from a distant gene on the same chromosome as the marker.

It is highly plausible that concentration variations in one transcript may
affect abundances of a second transcript.
For example, the first transcript may encode a transcription factor protein.
In this case, the transcription factor protein may influence expression
patterns of the second transcript (and perhaps other transcripts, too).

To determine whether a local transcript level mediates the relationship between
a nonlocal transcript level and a DNA variant, we perform a series of regression analyses,
which we detail below (Frame~\ref{frame3}).
In brief, we regress the nonlocal transcript levels on founder allele dosages at the DNA
variant, with and without conditioning on the candidate mediator (the local transcript levels).
If the LOD score measuring association between nonlocal trait and QTL genotype diminishes
sufficiently upon conditioning on a candidate,
then we declare the candidate a mediator.

The rationale behind this strategy follows.
If the DNA variant affects the nonlocal transcript levels solely by way of local
transcript levels, then conditioning on the local transcript levels would nullify the
relationship between DNA variant and the nonlocal transcript levels.
At the other extreme, if the DNA variant affects nonlocal transcript levels solely
through mechanisms that don't involve the local transcript levels, then conditioning
on local transcript levels would not affect the association between the DNA variant and
nonlocal transcript levels.

%\subsection{Regression-based mediation methods}

\begin{frameenv}{Four regressions for a single mediation analysis}\label{frame3}

%\begin{equ}[!ht]
\begin{equation}
Y = b1 + WC + E
\label{model1}
\end{equation}
%\caption{Linear model with intercept and covariates only.}
%\end{equ}

%\begin{equ}[!ht]
\begin{equation}
Y = XB + WC + E
\label{model2}
\end{equation}
%\caption{Linear model with founder allele dosages and covariates.}
%\end{equ}

%\begin{equ}[!ht]
\begin{equation}
Y = b1 + WC + M\beta + E
\label{model3}
\end{equation}
%\caption{Linear model with intercept, covariates, and candidate mediator.}
%\end{equ}

%\begin{equ}[!ht]
\begin{equation}
Y = XB + WC + M\beta + E
\label{model4}
\end{equation}
%\caption{Linear model with founder allele dosages, covariates, and candidate mediator.}
%\end{equ}
\end{frameenv}

We now consider the procedures needed for a mediation analysis in systems genetics.
In the four linear regression models (Frame~\ref{frame3}), $X$ is an $n$ by $8$ matrix of
founder allele dosages at a single marker, $B$ is an $8$ by $1$ matrix of founder allele
effects, $E$ is an $n$ by $1$ matrix of random errors, $b$
is a number, $1$ is
an $n$ by $1$ matrix with all entries set to $1$, $Y$ is
an $n$ by $1$ matrix of
phenotype values (for a single trait), and $M$ is an $n$
by $1$ matrix of values for a putative mediator.
$C$ is a $c$ by 1 matrix of covariate effects, and $W$ is
a $n$ by $c$ matrix of covariates.
We denote the coefficient of the mediator by $\beta$.

We assume that the vector $E$ is (multivariate) normally
distributed with zero vector as mean and covariance matrix
$\Sigma = \sigma^2I_n$, where $I_n$ is the $n$ by $n$
identity matrix.

In the above models with normally distributed random
errors, the log-likelihoods are easily calculated. For
example, in Equation \ref{model1}, the vector $Y$ follows
a multivariate normal distribution with mean $(b1 + WC)$
and covariance $\Sigma = \sigma^2I$. Thus, we can write
the likelihood for Model \ref{model1} as:

\begin{equation}
    L(b, C, \sigma^2| Y, W) = (2\pi)^{- \frac{n}{2}}\exp{ \left(- \frac{1}{2}(Y - b1 - WC)^T\Sigma^{-1}(Y - b1 - WC)\right)}
\end{equation}

We thus have the following equation (~\ref{eq:ll} for the
log-likelihood for Model~\ref{model1}:

\begin{equation}
    \log L(b, C, \sigma^2 | Y, W) = - \frac{n}{2}\log (2\pi) - \frac{1}{2} (Y - b1 - WC)^T\Sigma^{-1}(Y - b1 - WC)\label{eq:ll}
\end{equation}

\citet{chick2016defining} calculated the log$_{10}$
likelihoods for all four models before determining two LOD
scores (Equations~\ref{eq:lod1} and~\ref{eq:lod2}).

\begin{equation}
LOD_1 = log_{10}(\text{Model~\ref{model2} likelihood}) - log_{10}(\text{Model~\ref{model1} likelihood})\label{eq:lod1}
\end{equation}

\begin{equation}
LOD_2 = log_{10}(\text{Model~\ref{model4} likelihood}) - log_{10}(\text{Model~\ref{model3} likelihood})\label{eq:lod2}
\end{equation}

And, finally, \citet{chick2016defining} calculated the LOD
difference statistic (Equation~\ref{eq:lod-diff}).

\begin{equation}
\text{LOD difference} = LOD_1 - LOD_2\label{eq:lod-diff}
\end{equation}

LOD difference values need not be positive.
For example, in the setting where the putative mediator has no effect on the phenotype-QTL association, $LOD$ difference may be negative.

In our analyses below, we consider the LOD difference
proportion (Equation~\ref{eq:lod-diff-prop}).

\begin{equation}
\text{LOD difference proportion} = \frac{(LOD_1 - LOD_2)}{LOD_1}
\label{eq:lod-diff-prop}
\end{equation}

In other words, we consider what proportion of the association strength, on the LOD scale, is diminished by conditioning on a putative mediator.
Scaling the LOD difference statistic by LOD$_1$ accommodates the diversity of LODs in our data, which will be useful for graphical comparisons.
For example, a LOD difference statistic of 5 may be relevant when
a trait has a LOD$_1$ of 10, but unimportant if the trait's LOD$_1$ is 100. Now that we've defined our summary statistics for a mediation analysis in systems genetics, we turn attention to two statistical challenges in this area of research, 1. confounding and 2. significance thresholds.

%\subsection{Four assumptions for causal inference}

\begin{frameenv}{Four assumptions for causal inference}\label{frame1}
  \begin{enumerate}
\item No unmeasured confounding of the DNA variant-nonlocal transcript levels relationship
\item No unmeasured confounding of the local transcript levels-nonlocal transcript levels relationship
\item No unmeasured confounding of the DNA variant-local transcript levels relationship
\item No local transcript levels-nonlocal transcript levels confounder that is affected by the DNA variant
\end{enumerate}

\end{frameenv}

Our first statistical challenge is due to confounding. To claim that the LOD
difference statistic reflects a causal relationship, four assumptions about
confounding are needed (Frame~\ref{frame1}) \citep{vanderweele2015explanation},
yet
it is often difficult or impossible to recognize unmeasured confounders. Unmeasured
confounders, as I use the term here, are unmeasured random variables that confound a causal
relationship between two other random variables.
For example, Figure~\ref{fig:confounding} shows a directed graph. Arrows indicate causal
directions. For example, ``QTL'' causes ``M''. ``M'' designates the mediator (local transcript). ``Y'' denotes the nonlocal transcript
level in this setting, while ``C1'', ``C2'', and ``C3'' are covariates that confound,
respectively, the DNA variant - nonlocal transcript level relationship, the local transcript
level - nonlocal transcript level relationship, and the DNA variant - local transcript level
relationship. Failure to measure C1, C2, or C3 leads to violations of Assumption 1,
Assumption 2, or Assumption 3, respectively (Frame~\ref{frame1}).

\begin{figure}
\includegraphics{figs/confounding.pdf}
\caption{Example causal diagram showing relationships between random variables in an expression QTL study.}\label{fig:confounding}
\end{figure}

In studies of Diversity Outbred mice, relatedness is a possible confounder,
yet the linear models (Equations~\ref{model1},~\ref{model2},~\ref{model3},~\ref{model4}
in Frame~\ref{frame3}) fail to account for the complex relatedness patterns
among Diversity Outbred mice.
In fact, the only covariates in our models are for wave membership and sex.
The 400 mice were shipped from the Jackson Laboratory in four waves of 100 mice.
Because each wave was processed at a different time, I include wave membership
as a covariate with the goal of accounting for wave-related effects.
Other sources of confounding, such as batch effects in the phenotyping, may be
unmeasured.

In efforts to quantify the potential impact of unmeasured confounding, scientists have
developed a suite of sensitivity analysis tools for use in regression-based mediation analysis.
While a discussion of sensitivity analysis is beyond the scope of this thesis,
it may be useful in future systems genetics studies to assess robustness of mediation analysis
results in the presence of unmeasured confounders.
\citet{vanderweele2015explanation} discusses sensitivity analysis in the context of epidemiological studies.

%\subsection{Declaring mediators}

Besides accounting for confounding, a second statistical challenge in mediation analysis is assessing significance of the LOD difference statistic.
\citet{chick2016defining} used individual transcript levels as sham mediators and tabulated their LOD difference statistics.
They then compared the observed LOD difference statistics for putative mediators to the empirical distribution of LOD difference statistics obtained from the collection of sham mediators.
\citet{keller2018genetic}, on the other hand, in their study of pancreatic islet
biology, declared mediators to be those local transcripts that diminished the LOD score
of nonlocal transcripts by at least 1.5.
While significance threshold determination remains an active area of research,
we proceed below by examining all 147 nonlocal transcript levels that
\citet{keller2018genetic} identified as mapping to the Chromosome 2 hotspot.

\subsection{Methods}

We examined the potential that the two methods, 1. pleiotropy vs. separate QTL testing and
2. mediation analysis, play complementary roles in efforts to dissect gene expression trait hotspots.
We use the term ``hotspot'' to refer to a contiguous genomic region,
typically no more than 5 Mb in length, that affects many expression traits.
After describing our data below, we detail our statistical analyses involving 13 local
gene expression traits and 147 nonlocal gene expression traits,
all of which map to a 4-Mb hotspot on Chromosome 2.

%\subsection{Data description}

We analyzed data from 378 Diversity Outbred mice \citep{keller2018genetic}.
\citet{keller2018genetic} genotyped tail biopsies with the GigaMUGA microarray \citep{morgan2015mouse}.
They also used RNA sequencing to measure genome-wide pancreatic islet gene expression
for each mouse at the time of sacrifice \citep{keller2018genetic}.

We examined the Chromosome 2 pancreatic islet expression trait hotspot that \citet{keller2018genetic} identified.
\citet{keller2018genetic} found that 147 nonlocal traits map to the 4-Mb region
centered at 165.5 Mb on Chromosome 2.
The 147 nonlocal traits all exceeded the genome-wide LOD significance threshold,
7.18 \citep{keller2018genetic}.
With regression-based mediation analyses, they identified transcript levels of local gene \emph{Hnf4a} as a mediator of 88 of these 147 nonlocal traits.

We designed a study to examine the possible roles for mediation analysis and pleiotropy testing.
Because \citet{keller2018genetic} reported that some nonlocal traits that map to
the Chromosome 2 hotspot did not demonstrate evidence of mediation by \emph{Hnf4a} expression
levels, we elected to study a collection of local gene expression traits,
rather than \emph{Hnf4a} alone.
This strategy enabled us to ask whether one of twelve other local traits mediates those
nonlocal hotspot traits that are not mediated by \emph{Hnf4a}.
Our set of local gene expression traits includes \emph{Hnf4a} and 12
other local genes (Table \ref{tab:annot}).
Our 13 local genes are the only genes that met three criteria (Frame~\ref{frame2}).
The 147 nonlocal traits that we studied all had LOD peak heights above 7.18 and QTL positions within 2 Mb of the center of the hotspot (at 165.5 Mb).

\begin{frameenv}{Local gene inclusion criteria}\label{frame2}
\begin{enumerate}
    \item QTL peak with LOD $>$ 40
    \item QTL peak position within 2 Mb of hotspot center (165.5 Mb)
    \item Gene midpoint within 2 Mb of hotspot center (165.5 Mb)
\end{enumerate}
\end{frameenv}

%\subsection{Statistical analyses}

We now describe our statistical analyses.
After univariate QTL mapping to identify expression traits that
map to the Chromosome 2 hotspot,
we performed both bivariate QTL scans and mediation analyses of all 13 * 147 = 1911 pairs
involving one local expression trait and one nonlocal expression trait.

%\subsubsection{Bivariate QTL scans}

Our bivariate QTL analyses involved the same 13 local expression traits and 147 nonlocal expression traits.
We described above (Frame~\ref{frame2}) the criteria for choosing these expression traits.

We performed a series of two-dimensional QTL scans in which we paired each local gene's
transcript levels with each nonlocal gene's transcript levels,
for a total of 13 x 147 = 1,911 two-dimensional scans.
Each scan examined the same set of 180 markers, which spanned the interval from 163.1 Mb to 167.8 Mb and included univariate peaks for all 13 + 147 = 160 expression traits.
We performed these analyses with the R package \texttt{qtl2pleio} \citep{qtl2pleio}.

For each bivariate QTL scan, we fitted a collection of bivariate models for
all 180 * 180 = 32,400 ordered pairs of markers.
For each ordered pair of markers, we fitted a bivariate linear mixed
effects model using the methods of Chapter 2.

%\subsubsection{Mediation analyses}

We performed mediation analyses for
all 1,911 local-nonlocal trait pairs in which
we probed the extent to which each nonlocal trait's association
strength diminished upon conditioning on transcript levels of a putative mediator.

Each of the 13 local expression traits, considered one at a time, served as putative mediators.
We thus fitted the four linear regression models that we describe
above (Equations \ref{model1}, \ref{model2}, \ref{model3}, \ref{model4}).

One question that needs clarification is the choice of genetic marker
for each mediation analysis.
We elected to use the founder allele dosages at the marker (or pseudomarker) that demonstrated the
univariate LOD peak for the nonlocal trait.
Alternative analyses, in which one uses the founder allele dosages at which the
local trait has its univariate peak, are also possible.

%\subsubsection{Local gene analyses}

% latex table generated in R 3.5.1 by xtable 1.8-3 package
% Mon Dec 17 10:04:32 2018
\begin{table}[ht]
\centering
\begin{tabular}{lrrrr}
  \hline
Gene & Start & End & QTL peak position & LOD\\
  \hline
Pkig & 163.66 & 163.73 & 163.52 & 51.68 \\
  Serinc3 & 163.62 & 163.65 & 163.58 & 126.93 \\
  Hnf4a & 163.51 & 163.57 & 164.02 & 48.98 \\
  Stk4 & 164.07 & 164.16 & 164.03 & 60.39 \\
  Pabpc1l & 164.03 & 164.05 & 164.03 & 52.50 \\
  Slpi & 164.35 & 164.39 & 164.61 & 40.50 \\
  Neurl2 & 164.83 & 164.83 & 164.64 & 64.58 \\
  Cdh22 & 165.11 & 165.23 & 165.05 & 53.84 \\
  2810408M09Rik & 165.49 & 165.49 & 165.57 & 67.34 \\
  Eya2 & 165.60 & 165.77 & 165.72 & 98.89 \\
  Prex1 & 166.57 & 166.71 & 166.75 & 46.91 \\
  Ptgis & 167.19 & 167.24 & 167.27 & 56.25 \\
  Gm14291 & 167.20 & 167.20 & 167.27 & 73.72 \\
   \hline
\end{tabular}
\caption{Local gene annotations for analysis of Chromosome 2 expression trait hotspot.
All positions are in units of Mb on Chromosome 2.
LOD peak position and LOD peak height refer to those obtained from univariate analyses.
``Start'' and ``end'' refer to the local gene's DNA start and end positions, as annotated by Ensembl version 75.}
\label{tab:annot}
\end{table}

To visualize the summary statistics for the two methods, I plotted,
for each local gene, a scatterplot of LOD difference proportion values
against pleiotropy test statistics.

%\subsubsection{Nonlocal gene analyses}

We also examined the 1,911 pairs by considering one nonlocal gene at a time.
For each of the 147 nonlocal genes, I plotted LOD difference proportion
against pleiotropy test statistic values.
I present below (Figure \ref{fig:4nonlocal}) examples that illustrate some of the
observed patterns between pleiotropy test statistics and LOD difference proportion values.

\subsection{Results}

%\subsection{Scatter plot for all 1911 pairs}
Figure \ref{fig:lod-diff-prop-v-lrt-all} is a scatterplot of the results from all pleiotropy tests and mediation analyses.
Each point contains one local expression trait and one nonlocal expression trait.
I see that points with high values of LOD difference proportion tend to have small
values of pleiotropy test statistic, and those points with high values of the
pleiotropy test statistic tend to have small values of LOD difference proportion.

Some points demonstrate large values of pleiotropy test statistic (\emph{eg.}, $>$ 10),
yet still have sizeable LOD difference proportion statistics (Figure~\ref{fig:lod-diff-prop-v-lrt-all}).
One possible explanation for such points is that the univariate LOD (LOD$_1$) value is small,
so that even a modest LOD difference gives rise to a sizeable LOD difference proportion value.
A second possible explanation is that the phenotype variances for both phenotypes in the pair
are large enough to skew the null distribution of the test statistic towards larger values.
To distinguish between these two, I would
examine the LOD difference statistics and obtain bootstrap $p$-values.

In Figure~\ref{fig:lod-diff-prop-v-lrt-all}, I colored red the points that involve \emph{Hnf4a};
points with other local genes are blue.
The most striking feature of the coloring is that many red points have small
values of the pleiotropy test statistics and very high values of LOD difference
proportion. This is what I expect when \emph{Hnf4a} mediates the relationship
between a nonlocal trait and a QTL; namely, \emph{Hnf4a} and a nonlocal trait
share a pleiotropic QTL and conditioning on \emph{Hnf4a} diminishes the LOD
score for the association between nonlocal trait and QTL.

\begin{figure}
    \centering
    \includegraphics[width = 0.7\textwidth]{../keller2018-chr2-hotspot-chtc/Rmd/lod-diff-prop-v-lrt.eps}
    \caption[LOD difference proportion vs. pleiotropy test statistic]{LOD difference proportion against pleiotropy vs. separate QTL test statistics for 1911 pairs of traits.
    Each point represents a single pair. Pairs that involve local expression trait \emph{Hnf4a} are colored red, while all others are colored blue.
    Note the high concentration of red points in the upper left quadrant of the figure.
    These points, with low values of the pleiotropy vs. separate QTL test statistic and
    high values of LOD difference proportion, are consistent with \emph{Hnf4a} transcript
    levels mediating the effect of \emph{Hnf4a} genetic variants affecting nonlocal
    transcript abundances.}
    \label{fig:lod-diff-prop-v-lrt-all}
\end{figure}

%\subsection{Local gene analyses}

To more thoroughly examine the pleiotropy test and mediation analysis
relationships across the 13 local genes,
I created 13 plots of LOD difference proportion against pleiotropy test
statistic (Figures~\ref{fig:hnf4a} and~\ref{fig:nothnf4a-12}).
They reveal common patterns.
First, I see no points in the upper right quadrant of each plot (Figure~\ref{fig:nothnf4a-12}).
This tells us that those nonlocal genes with high values of pleiotropy test statistic
(when paired with the specified local gene) have low values of LOD difference proportion.
Similarly, those nonlocal genes with high values of LOD difference proportion tend
to have small values of the pleiotropy test statistic.
Finally, some trait pairs demonstrate low values of both
the LOD difference proportion and pleiotropy test statistic.
This observation suggests that, for a given local expression trait, the nonlocal trait
is not mediated by the local expression trait yet it shares a pleiotropic locus.
Low power to resolve multiple nearby QTL may give rise to such findings.

In comparing the \emph{Hnf4a} plot (Figure \ref{fig:hnf4a}) with the other
12 plots (Figure \ref{fig:nothnf4a-12}),
I see that none of the 12 plots in Figure \ref{fig:nothnf4a-12} closely resembles Figure \ref{fig:hnf4a}.
\emph{Serinc3}, \emph{Stk4}, \emph{Neurl2}, and \emph{Cdh22} are closest in appearance to the plot of \emph{Hnf4a}.
However, each of \emph{Serinc3}, \emph{Stk4}, \emph{Neurl2}, and \emph{Cdh22}
has very few points with LOD difference proportion above 0.5,
while \emph{Hnf4a} has many points with LOD difference proportion above 0.5.

\begin{figure}
    \centering
    \includegraphics[width = \textwidth]{../keller2018-chr2-hotspot-chtc/Rmd/Hnf4a-lod-diff-prop-v-lrt.eps}
    \caption[LOD difference proportion vs. pleiotropy test statistic.]{Scatter plot of LOD difference proportion against pleiotropy vs.
    separate QTL test statistic for 147 pairs of traits.
    Each pair includes \emph{Hnf4a} and one of the nonlocal gene expression traits that map to the Chromosome 2 hotspot.}
    \label{fig:hnf4a}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width = 0.8\textwidth]{../keller2018-chr2-hotspot-chtc/Rmd/12local-facet_grid-no-strip.eps}
    \caption[LOD difference proportion vs. pleiotropy test statistic from the per-local gene perspective.]{Scatter plots of LOD difference proportion against pleiotropy vs. separate QTL
    test statistic with 147 pairs of traits per panel.
    Each panel includes a local gene expression trait
    (per the label in the upper right quadrant) and one of the 147 nonlocal gene
    expression traits that map to the Chromosome 2 hotspot.}
    \label{fig:nothnf4a-12}
\end{figure}

%\subsection{Nonlocal gene analyses}

Now that I've examined our results from the perspective of every one of 13 local gene expression traits,
I turn attention to the nonlocal gene expression trait perspective.
I present four nonlocal gene expression trait plots in Figure~\ref{fig:4nonlocal}. I chose
these four from the 147 to demonstrate two of the observed patterns. In a single scatterplot,
I've plotted 13 points, one for each local gene expression trait, paired with a single
nonlocal gene expression trait. Local expression trait \emph{Hnf4a} is colored in red,
while the other local genes are colored in blue. The top row of Figure~\ref{fig:4nonlocal}
features scatterplots for nonlocal expression traits \emph{Abcb4} and \emph{Smim5}. I see
that local expression trait \emph{Hnf4a} has, by far, the greatest value of LOD difference
proportion. In the second row of Figure~\ref{fig:4nonlocal}, I see a different pattern in
the plots for nonlocal expression traits \emph{Tmprss4} and \emph{Gm3095}. Both demonstrate,
for \emph{Hnf4a}, small values for the pleiotropy test statistic and small values of LOD
difference proportion. This tells me that \emph{Hnf4a} shares a pleiotropic locus with both
\emph{Tmprss4} and \emph{Gm3095}, yet none of the 13 local gene expression traits mediates
the relationship between these two nonlocal traits and the QTL.

\begin{figure}
    \centering
    \caption{Scatter plots for four nonlocal expression traits. Each plot features 13
    points, one for each local gene expression trait. The vertical axis denotes LOD
    difference proportion values, while the horizontal axis corresponds to pleiotropy test
    statistics. Red points represent the pairing with local gene expression trait
    \emph{Hnf4a}. Blue points represent the other 12 local gene expression traits.}
    \begin{subfigure}[t]{.45\textwidth}
        \includegraphics[width = \textwidth]{../keller2018-chr2-hotspot-chtc/Rmd/nonlocal-scatter_97.eps}
        \caption[\emph{Hnf4a} transcript levels mediate \emph{Abcb4} transcript levels.]{\emph{Hnf4a} transcript levels mediate \emph{Abcb4} transcript levels. The high LOD difference proportion and the very small pleiotropy test statistic together provide evidence that \emph{Hnf4a} mediates \emph{Abcb4} transcript levels.}
    \end{subfigure}\hspace{0.05\textwidth}
    \begin{subfigure}[t]{.45\textwidth}
        \includegraphics[width = \textwidth]{../keller2018-chr2-hotspot-chtc/Rmd/nonlocal-scatter_105.eps}
        \caption[\emph{Hnf4a} transcript levels mediate \emph{Smim5} transcript levels.]{\emph{Hnf4a} transcript levels mediate \emph{Smim5} transcript levels. The high LOD difference proportion and the very small pleiotropy test statistic together provide evidence that \emph{Hnf4a} mediates \emph{Smim5} transcript levels.}
    \end{subfigure}
    \begin{subfigure}[t]{.45\textwidth}
        \includegraphics[width = \textwidth]{../keller2018-chr2-hotspot-chtc/Rmd/nonlocal-scatter_64.eps}
        \caption[\emph{Tmprss4} transcript levels are not mediated by any of the 13 local gene expression traits.]{\emph{Tmprss4} transcript levels are not mediated by any of the 13 local gene expression traits. However, several local gene expression traits arise from separate QTL, as evidenced by their large (greater than 5) values of the pleiotropy test statistic.}
    \end{subfigure}\hspace{0.05\textwidth}
    \begin{subfigure}[t]{.45\textwidth}
        \includegraphics[width = \textwidth]{../keller2018-chr2-hotspot-chtc/Rmd/nonlocal-scatter_141.eps}
        \caption[\emph{Gm3095} transcript levels are not mediated by any of the 13 local genes.]{\emph{Gm3095} transcript levels are not mediated by any of the 13 local genes. The tests of pleiotropy demonstrate evidence of separate QTL for some local expression traits when paired with \emph{Gm3095}.}
    \end{subfigure}
    \label{fig:4nonlocal}
\end{figure}

\subsection{Discussion}

Our pairwise analyses with both mediation analyses and pleiotropy tests provide additional
evidence for the importance of \emph{Hnf4a} in the biology of the Chromosome 2 hotspot
in pancreatic islets.
Our analyses, and, specifically, the test of pleiotropy vs. separate QTL, may be more useful
when studying nonlocal traits that map to a hotspot yet don't show strong evidence
of mediation by local expression traits.
In such a setting, the pleiotropy test can, at least, provide some information about
the genetic architecture at the hotspot.
Specifically, our pleiotropy test may inform inferences about the number of
underlying QTL in a given expression trait hotspot.
Additionally, our test may limit the number of expression traits that are
potential intermediates between a QTL and a specified nonlocal expression trait.
This relies on the assumption that a causally intermediate (local) expression trait and a
target (nonlocal) expression trait presumably share a QTL. On the other hand, mediation analyses, when they provide evidence for mediation of a nonlocal
trait by a local expression trait, are possibly more valuable than the test of pleiotropy vs.
separate QTL, since the mediation analyses identify precisely the intermediate expression trait.

I recommend using both tests of pleiotropy vs separate QTL and mediation analyses when
dissecting an expression trait QTL hotspot.
Future researchers may use the test of pleiotropy vs. separate QTL as a
preliminary screen to limit the set of candidate mediators to those that have
test results that are consistent with pleiotropy. Subsequently, fewer mediation
analyses are needed, which would reduce the magnitude of the multiple testing
adjustment. This, in turn, would lead to detection of more true positive signals
in mediation analysis.

Future research may investigate the use of polygenic random effects in the statistical models
for mediation analysis. Additional methodological questions include approaches for declaring
significant a mediation LOD difference proportion and consideration of other possible measures
and scales of extent of mediation. Additionally, future researchers may wish to consider
biological models that contain two mediators.

The social and health sciences have witnessed much methods research in mediation analysis.
The field of statistical genetics has not fully adopted these strategies yet, but, given the
nature of current and future data, many opportunities exist for translation of approaches from
causal inference in epidemiology to systems genetics.
For example, \citet{vanderweele2015explanation} contains detailed discussions of many methods
issues that arose in mediation analyses in
epidemiology studies.

%%%%% 3A end
\printbibliography

<<r install_here>>=
install.packages("here", repos = "https://cloud.r-project.org")
@

\subsection{Colophon}

This report was generated on `r Sys.time()` using the following computational environment and dependencies:

<<r colophon, cache = FALSE>>=
# which R packages and versions?
devtools::session_info()
@

The current Git commit details are:

<<r git2r, cache = FALSE>>=
# what commit is this file at?
git2r::repository(here::here())
@

\end{document}

### 2019-06-17 paper.tex
added Rmd files for initial analyses - mediation v pleiotropy test statistic

--- 2019-06-17 analysis/paper/paper.Rnw
@@
-\documentclass[oneside]{book}
+\documentclass[oneside]{book}\usepackage[]{graphicx}\usepackage[]{color}
+% maxwidth is the original width if it is less than linewidth
+% otherwise use linewidth (to make sure the graphics do not exceed the margin)
+\makeatletter
+\def\maxwidth{ %
+  \ifdim\Gin@nat@width>\linewidth
+    \linewidth
+  \else
+    \Gin@nat@width
+  \fi
+}
+\makeatother
+
+\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
+\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
+\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
+\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
+\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
+\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
+\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
+\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
+\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
+\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
+\let\hlipl\hlkwb
+
+\usepackage{framed}
+\makeatletter
+\newenvironment{kframe}{%
+ \def\at@end@of@kframe{}%
+ \ifinner\ifhmode%
+  \def\at@end@of@kframe{\end{minipage}}%
+  \begin{minipage}{\columnwidth}%
+ \fi\fi%
+ \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
+ \colorbox{shadecolor}{##1}\hskip-\fboxsep
+     % There is no \\@totalrightmargin, so:
+     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
+ \MakeFramed {\advance\hsize-\width
+   \@totalleftmargin\z@ \linewidth\hsize
+   \@setminipage}}%
+ {\par\unskip\endMakeFramed%
+ \at@end@of@kframe}
+\makeatother
+
+\definecolor{shadecolor}{rgb}{.97, .97, .97}
+\definecolor{messagecolor}{rgb}{0, 0, 0}
+\definecolor{warningcolor}{rgb}{1, 0, 1}
+\definecolor{errorcolor}{rgb}{1, 0, 0}
+\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX
+
+\usepackage{alltt}
 \usepackage[colorinlistoftodos]{todonotes}
@@
 %% end from vignette tex file
-
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
 \begin{document}
-
-<<include=FALSE>>=
-library(knitr)
-opts_chunk$set(
-concordance=TRUE, echo = FALSE, fig.lp = "fig:", warning = FALSE, message = FALSE, results = "hide"
-)
-@
 
@@
 by analyzing experimental data.
-
-<<load-broman>>=
-library(broman)
-@
 
@@
 
-<<r install_here>>=
-install.packages("here", repos = "https://cloud.r-project.org")
-@
-
 \subsection{Colophon}
@@
 
-<<r colophon, cache = FALSE>>=
-# which R packages and versions?
-devtools::session_info()
-@
-
 The current Git commit details are:
 
-<<r git2r, cache = FALSE>>=
-# what commit is this file at?
-git2r::repository(here::here())
-@
-
 \end{document}

### 2019-06-18 analysis/paper/paper.Rnw
commented out line for todonotes latex package
fixed Makefile & paper.Rnw

@@
 \documentclass[oneside]{book}
-\usepackage[colorinlistoftodos]{todonotes}
+%\usepackage[colorinlistoftodos]{todonotes}
 \usepackage[utf8]{inputenc}
@@
 \doublespacing
-\begin{titlepage}\thispagestyle{empty}
 \tableofcontents
@@
 protein concentrations is through \emph{Dhtkd1} transcript levels.
-
-\begin{figure}
-  \centering
-  \includegraphics[width = 0.7\textwidth]{../QTLfigs/Figs/central_dogma.pdf}
-  \caption{Biological information is encoded in DNA.
-This information is passed, via transcription, to sequence-specific RNA molecules.
-The process of translation transmits the information to sequence-specific proteins.}\label{fig:dogma}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \includegraphics[width = 0.7\textwidth]{../QTLfigs/Figs/central_dogma-Dhtkd1.pdf}
-  \caption{A DNA variant in the \emph{Dhtkd1} gene affects \emph{Dhtkd1} transcript abundances which, in turn, affect DHTKD1 protein concentrations.}\label{fig:Dhtkd1}
-\end{figure}
 
@@
 Assumption 2, or Assumption 3, respectively (Frame~\ref{frame1}).
-
-\begin{figure}
-\includegraphics{figs/confounding.pdf}
-\caption{Example causal diagram showing relationships between random variables in an expression QTL study.}\label{fig:confounding}
-\end{figure}
 
@@
 
-\begin{figure}
-    \centering
-    \includegraphics[width = 0.7\textwidth]{../keller2018-chr2-hotspot-chtc/Rmd/lod-diff-prop-v-lrt.eps}
-    \caption[LOD difference proportion vs. pleiotropy test statistic]{LOD difference proportion against pleiotropy vs. separate QTL test statistics for 1911 pairs of traits.
-    Each point represents a single pair. Pairs that involve local expression trait \emph{Hnf4a} are colored red, while all others are colored blue.
-    Note the high concentration of red points in the upper left quadrant of the figure.
-    These points, with low values of the pleiotropy vs. separate QTL test statistic and
-    high values of LOD difference proportion, are consistent with \emph{Hnf4a} transcript
-    levels mediating the effect of \emph{Hnf4a} genetic variants affecting nonlocal
-    transcript abundances.}
-    \label{fig:lod-diff-prop-v-lrt-all}
-\end{figure}
-
 %\subsection{Local gene analyses}
@@
 
-\begin{figure}
-    \centering
-    \includegraphics[width = \textwidth]{../keller2018-chr2-hotspot-chtc/Rmd/Hnf4a-lod-diff-prop-v-lrt.eps}
-    \caption[LOD difference proportion vs. pleiotropy test statistic.]{Scatter plot of LOD difference proportion against pleiotropy vs.
-    separate QTL test statistic for 147 pairs of traits.
-    Each pair includes \emph{Hnf4a} and one of the nonlocal gene expression traits that map to the Chromosome 2 hotspot.}
-    \label{fig:hnf4a}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[width = 0.8\textwidth]{../keller2018-chr2-hotspot-chtc/Rmd/12local-facet_grid-no-strip.eps}
-    \caption[LOD difference proportion vs. pleiotropy test statistic from the per-local gene perspective.]{Scatter plots of LOD difference proportion against pleiotropy vs. separate QTL
-    test statistic with 147 pairs of traits per panel.
-    Each panel includes a local gene expression trait
-    (per the label in the upper right quadrant) and one of the 147 nonlocal gene
-    expression traits that map to the Chromosome 2 hotspot.}
-    \label{fig:nothnf4a-12}
-\end{figure}
-
 %\subsection{Nonlocal gene analyses}
@@
 the relationship between these two nonlocal traits and the QTL.
-
-\begin{figure}
-    \centering
-    \caption{Scatter plots for four nonlocal expression traits. Each plot features 13
-    points, one for each local gene expression trait. The vertical axis denotes LOD
-    difference proportion values, while the horizontal axis corresponds to pleiotropy test
-    statistics. Red points represent the pairing with local gene expression trait
-    \emph{Hnf4a}. Blue points represent the other 12 local gene expression traits.}
-    \begin{subfigure}[t]{.45\textwidth}
-        \includegraphics[width = \textwidth]{../keller2018-chr2-hotspot-chtc/Rmd/nonlocal-scatter_97.eps}
-        \caption[\emph{Hnf4a} transcript levels mediate \emph{Abcb4} transcript levels.]{\emph{Hnf4a} transcript levels mediate \emph{Abcb4} transcript levels.
-The high LOD difference proportion and the very small pleiotropy test statistic together provide evidence that \emph{Hnf4a} mediates \emph{Abcb4} transcript levels.}
-    \end{subfigure}\hspace{0.05\textwidth}
-    \begin{subfigure}[t]{.45\textwidth}
-        \includegraphics[width = \textwidth]{../keller2018-chr2-hotspot-chtc/Rmd/nonlocal-scatter_105.eps}
-        \caption[\emph{Hnf4a} transcript levels mediate \emph{Smim5} transcript levels.]{\emph{Hnf4a} transcript levels mediate \emph{Smim5} transcript levels.
-The high LOD difference proportion and the very small pleiotropy test statistic together provide evidence that \emph{Hnf4a} mediates \emph{Smim5} transcript levels.}
-    \end{subfigure}
-    \begin{subfigure}[t]{.45\textwidth}
-        \includegraphics[width = \textwidth]{../keller2018-chr2-hotspot-chtc/Rmd/nonlocal-scatter_64.eps}
-        \caption[\emph{Tmprss4} transcript levels are not mediated by any of the 13 local gene expression traits.]{\emph{Tmprss4} transcript levels are not mediated by any of the 13 local gene expression traits.
-However, several local gene expression traits arise from separate QTL, as evidenced by their large (greater than 5) values of the pleiotropy test statistic.}
-    \end{subfigure}\hspace{0.05\textwidth}
-    \begin{subfigure}[t]{.45\textwidth}
-        \includegraphics[width = \textwidth]{../keller2018-chr2-hotspot-chtc/Rmd/nonlocal-scatter_141.eps}
-        \caption[\emph{Gm3095} transcript levels are not mediated by any of the 13 local genes.]{\emph{Gm3095} transcript levels are not mediated by any of the 13 local genes.
-The tests of pleiotropy demonstrate evidence of separate QTL for some local expression traits when paired with \emph{Gm3095}.}
-    \end{subfigure}
-    \label{fig:4nonlocal}
-\end{figure}
 

### 2019-06-18 analysis/paper/paper.Rnw
started using tinytex & switched to results equals default value for knitr
removed much unneeded latex text from Rnw file

@@
 \documentclass[oneside]{book}
-%\usepackage[colorinlistoftodos]{todonotes}
+\usepackage[colorinlistoftodos]{todonotes}
 \usepackage[utf8]{inputenc}
-%\usepackage{lineno}
-%\linenumbers
+\usepackage{lineno}
+\linenumbers
 
@@
 \usepackage{multirow}
-\usepackage[framemethod=TikZ]{mdframed}
 \usepackage{float}
@@
             breaklinks=true}
+\usepackage[yyyymmdd,hhmmss]{datetime}
 
@@
 \chead{}
-%\rhead{}
 \rhead{ \fancyplain{}{\thepage} } % put page number in upper right corner
+\rfoot{Compiled on \today\ at \currenttime}
 \renewcommand{\headrulewidth}{0pt} % remove top line
@@
 \renewcommand{\headrule}{}
-\title{The title}
-\author{An Author}
-\patchcmd{\titlepage}{empty}{fancy}{}{}
-\patchcmd{\chapter}{plain}{fancy}{}{}
-\DeclareCaptionType{equ}[][]
-%https://stackoverflow.com/questions/149479/adding-a-caption-to-an-equation-in-latex
-%\captionsetup[equ]{labelformat=empty}
 
-\DeclareFloatingEnvironment[fileext=frm,placement={!ht},name=Frame]{myfloat}
-\captionsetup[myfloat]{labelfont=bf}
-\newenvironment{frameenv}[1]
-    {\begin{myfloat}[tb]
-    \begin{mdframed}[roundcorner=10pt,backgroundcolor=blue!10]
-    \caption{#1}
-    %\begin{multicols*}{2}
-    }
-    {%\end{multicols*}
-    \end{mdframed}\end{myfloat}
-    }
-
-\title{}
+\title{Expression QTL hotspot dissection}
 \author{Frederick J. Boehm, Brian S. Yandell, and Karl W. Broman}
-\date{\today}
-
-%% from vignette tex file
-\urlstyle{same}  % don't use monospace font for urls
-\usepackage{color}
-\usepackage{fancyvrb}
-\newcommand{\VerbBar}{|}
-\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
-\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
-% Add ',fontsize=\small' for more characters per line
-\usepackage{framed}
-\definecolor{shadecolor}{RGB}{248,248,248}
-\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
-\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
-\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
-\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\BuiltInTok}[1]{#1}
-\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
-\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
-\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
-\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
-\newcommand{\ExtensionTok}[1]{#1}
-\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\ImportTok}[1]{#1}
-\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
-\newcommand{\NormalTok}[1]{#1}
-\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
-\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
-\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
-\newcommand{\RegionMarkerTok}[1]{#1}
-\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\usepackage{graphicx,grffile}
-\makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
-
-%% end from vignette tex file
+\date{}
 
@@
 opts_chunk$set(
-concordance=TRUE, echo = FALSE, fig.lp = "fig:", warning = FALSE, message = FALSE, results = "hide"
-)
+concordance=TRUE,
+echo = TRUE,
+fig.lp = "../figures:",
+warning = FALSE,
+message = FALSE)
 @

### 2019-06-18 analysis/paper/paper.Rnw
added lists of figs & tables; removed references to nonexistent figures

@@
 \tableofcontents
-%\listoffigures
-%\listoftables
+\listoffigures
+\listoftables
 %\pagestyle{headings} % page numbers appear at bottom of page, centered!
@@
 heterozygous and essentially unique mosaic of
-founders' DNA. Figure~\ref{fig:do} presents color-coded chromosomes for a collection of
+founders' DNA. Figure presents color-coded chromosomes for a collection of
 subjects from a multiparental population like the Diversity Outbred mice. DNA from each founder
@@
 \citet{crick1958protein} articulated a pathway for transmission of biological information
-that is now known as the central dogma of molecular biology (Figure~\ref{fig:dogma})
+that is now known as the central dogma of molecular biology (Figure)
 \citep{crick1970central}.
@@
 DNA variant and another trait. To clarify our discussion, we refer to an example from
-\citet{chick2016defining} (Figure~\ref{fig:Dhtkd1}). \citet{chick2016defining}, in
+\citet{chick2016defining} (Figure). \citet{chick2016defining}, in
 studying livers of 192 Diversity Outbred mice, found evidence that \emph{Dhtkd1}
@@
 a nonlocal transcript level and a DNA variant, we perform a series of regression analyses,
-which we detail below (Frame~\ref{frame3}).
+which we detail below.
 In brief, we regress the nonlocal transcript levels on founder allele dosages at the DNA
@@
 %\subsection{Regression-based mediation methods}
-
-\begin{frameenv}{Four regressions for a single mediation analysis}\label{frame3}
 
@@
 %\end{equ}
-\end{frameenv}
 
 We now consider the procedures needed for a mediation analysis in systems genetics.
-In the four linear regression models (Frame~\ref{frame3}), $X$ is an $n$ by $8$ matrix of
+In the four linear regression models, $X$ is an $n$ by $8$ matrix of
 founder allele dosages at a single marker, $B$ is an $8$ by $1$ matrix of founder allele
@@
 
-\begin{frameenv}{Four assumptions for causal inference}\label{frame1}
   \begin{enumerate}
@@
 
-\end{frameenv}
-
 Our first statistical challenge is due to confounding. To claim that the LOD
 difference statistic reflects a causal relationship, four assumptions about
-confounding are needed (Frame~\ref{frame1}) \citep{vanderweele2015explanation},
+confounding are needed \citep{vanderweele2015explanation},
 yet
@@
 relationship between two other random variables.
-For example, Figure~\ref{fig:confounding} shows a directed graph. Arrows indicate causal
+For example, Figure shows a directed graph. Arrows indicate causal
 directions. For example, ``QTL'' causes ``M''. ``M'' designates the mediator (local transcript). ``Y'' denotes the nonlocal transcript
@@
 relationship. Failure to measure C1, C2, or C3 leads to violations of Assumption 1,
-Assumption 2, or Assumption 3, respectively (Frame~\ref{frame1}).
+Assumption 2, or Assumption 3, respectively.
 
 In studies of Diversity Outbred mice, relatedness is a possible confounder,
-yet the linear models (Equations~\ref{model1},~\ref{model2},~\ref{model3},~\ref{model4}
-in Frame~\ref{frame3}) fail to account for the complex relatedness patterns
+yet the linear models (Equations~\ref{model1},~\ref{model2},~\ref{model3},~\ref{model4}) fail to account for the complex relatedness patterns
 among Diversity Outbred mice.
@@
 other local genes (Table \ref{tab:annot}).
-Our 13 local genes are the only genes that met three criteria (Frame~\ref{frame2}).
+Our 13 local genes are the only genes that met three criteria.
 The 147 nonlocal traits that we studied all had LOD peak heights above 7.18 and QTL positions within 2 Mb of the center of the hotspot (at 165.5 Mb).
 
-\begin{frameenv}{Local gene inclusion criteria}\label{frame2}
 \begin{enumerate}
@@
 \end{enumerate}
-\end{frameenv}
 
@@
 Our bivariate QTL analyses involved the same 13 local expression traits and 147 nonlocal expression traits.
-We described above (Frame~\ref{frame2}) the criteria for choosing these expression traits.
+We described above the criteria for choosing these expression traits.
 
@@
 against pleiotropy test statistic values.
-I present below (Figure \ref{fig:4nonlocal}) examples that illustrate some of the
+I present below (Figure ) examples that illustrate some of the
 observed patterns between pleiotropy test statistics and LOD difference proportion values.
@@
 %\subsection{Scatter plot for all 1911 pairs}
-Figure \ref{fig:lod-diff-prop-v-lrt-all} is a scatterplot of the results from all pleiotropy tests and mediation analyses.
+Figure is a scatterplot of the results from all pleiotropy tests and mediation analyses.
 Each point contains one local expression trait and one nonlocal expression trait.
@@
 Some points demonstrate large values of pleiotropy test statistic (\emph{eg.}, $>$ 10),
-yet still have sizeable LOD difference proportion statistics (Figure~\ref{fig:lod-diff-prop-v-lrt-all}).
+yet still have sizeable LOD difference proportion statistics (Figure).
 One possible explanation for such points is that the univariate LOD (LOD$_1$) value is small,
@@
 
-In Figure~\ref{fig:lod-diff-prop-v-lrt-all}, I colored red the points that involve \emph{Hnf4a};
+In Figure, I colored red the points that involve \emph{Hnf4a};
 points with other local genes are blue.
@@
 I created 13 plots of LOD difference proportion against pleiotropy test
-statistic (Figures~\ref{fig:hnf4a} and~\ref{fig:nothnf4a-12}).
+statistic (Figures).
 They reveal common patterns.
-First, I see no points in the upper right quadrant of each plot (Figure~\ref{fig:nothnf4a-12}).
+First, I see no points in the upper right quadrant of each plot (Figure).
 This tells us that those nonlocal genes with high values of pleiotropy test statistic
@@
 
-In comparing the \emph{Hnf4a} plot (Figure \ref{fig:hnf4a}) with the other
-12 plots (Figure \ref{fig:nothnf4a-12}),
-I see that none of the 12 plots in Figure \ref{fig:nothnf4a-12} closely resembles Figure \ref{fig:hnf4a}.
+In comparing the \emph{Hnf4a} plot (Figure ) with the other
+12 plots (Figure ),
+I see that none of the 12 plots in Figure  closely resembles Figure .
 \emph{Serinc3}, \emph{Stk4}, \emph{Neurl2}, and \emph{Cdh22} are closest in appearance to the plot of \emph{Hnf4a}.
@@
 I turn attention to the nonlocal gene expression trait perspective.
-I present four nonlocal gene expression trait plots in Figure~\ref{fig:4nonlocal}. I chose
+I present four nonlocal gene expression trait plots in Figure. I chose
 these four from the 147 to demonstrate two of the observed patterns. In a single scatterplot,
@@
 nonlocal gene expression trait. Local expression trait \emph{Hnf4a} is colored in red,
-while the other local genes are colored in blue. The top row of Figure~\ref{fig:4nonlocal}
+while the other local genes are colored in blue. The top row of Figure
 features scatterplots for nonlocal expression traits \emph{Abcb4} and \emph{Smim5}. I see
 that local expression trait \emph{Hnf4a} has, by far, the greatest value of LOD difference
-proportion. In the second row of Figure~\ref{fig:4nonlocal}, I see a different pattern in
+proportion. In the second row of Figure, I see a different pattern in
 the plots for nonlocal expression traits \emph{Tmprss4} and \emph{Gm3095}. Both demonstrate,
@@
 
-This report was generated on `r Sys.time()` using the following computational environment and dependencies:
+This report was generated on \Sexpr{Sys.time()} using the following computational environment and dependencies:
 

### 2019-06-19 analysis/paper/paper.Rnw
added \maketitle to Rnw
started writing methods section

@@
 \author{Frederick J. Boehm, Brian S. Yandell, and Karl W. Broman}
-\date{}
+\date{\today}
 
 \begin{document}
+
+\maketitle
 
@@
 echo = TRUE,
-fig.lp = "../figures:",
+fig.lp = "fig:",
 warning = FALSE,
@@
 unique QTL in a genomic region of interest.
-In this thesis, I develop a test of pleiotropy vs. separate QTL in multiparental
-populations. I study its statistical properties and demonstrate its utility
-by analyzing experimental data.
 
@@
 
-Due to the breeding scheme, each Diversity Outbred mouse is a highly
-heterozygous and essentially unique mosaic of
-founders' DNA. Figure presents color-coded chromosomes for a collection of
-subjects from a multiparental population like the Diversity Outbred mice. DNA from each founder
-line is color-coded.
-Each pair of vertical bars corresponds to one subject.
-Each row corresponds to a single generation, with generation labels on the left.
-Generation number increases from top to bottom.
-With each new generation, the Diversity Outbred mouse genomes are shuffled
-in smaller and smaller chunks.
+\section{Methods}
 
-\section{QTL mapping in Diversity Outbred mice}\label{sec:do-qtl}
+\citet{keller2018genetic} identified five expression trait hotspots in their
+study of pancreatic islet gene expression in 378 Diversity Outbred mice. They
+defined a hotspot to be a 4-Mb region that affects at least 100 expression
+traits.
 
-QTL mapping in Diversity Outbred mice, like that in mice from two-parent
-crosses, is a multi-step procedure: 1. data acquisition, 2. inference of
-missing genotypes, and 3. modeling phenotypes as a function of genotypes.
-Data acquisition involves measurement of phenotypes and, at specified
-genetic markers, termed ``single nucleotide polymorphism'' or SNP markers,
-measurement of two-allele genotypes. Often, the SNP marker genotypes are
-obtained by use of a microarray, such as the GigaMUGA SNP microarray
-\citep{morgan2015mouse}.
+For each hotspot, we performed a series of mediation analyses and pleiotropy
+tests. For each hotspot, we identified a set of local expression traits and a
+set of nonlocal expression traits. All traits demonstrated LOD peaks that met
+the genome-wide significance threshold. The nonlocal expression traits arose from genes
+on other chromosomes, while the local traits reflected expression levels for
+genes on the same chromosome as the hotspot and within 7 Mb of the hotspot
+center.
 
-The next step, missing genotypes inference, is needed because of the
-``missing data problem'' \citep{broman2009guide}. It takes as input the two-allele
-genotypes at the measured SNP markers. Hidden Markov model methods, developed by
-\citet{broman2012haplotype} and \citet{broman2012genotype} and implemented in the \texttt{qtl2} R
-package \citep{broman2019rqtl2}, output 36-state genotype probabilities for all
-nuclear autosomal markers and pseudomarkers. The outputted genotypes have 36 states
-because each position has eight homozygote states (one for each founder line) and $\binom{8}{2}$ heterozygote states. Pseudomarkers are
-arbitrary positions at which the researcher wants 36-state genotype probabilities.
-Finally, I collapse the 36-state genotype probabilities to eight founder allele
-dosages at each marker. This last step is optional, but often helpful, because the
-simplified models require specification of fewer parameters. We then treat the
-founder allele dosages as known quantities in subsequent steps.
+\subsection{Pleiotropy testing}
 
-% model form - connecting genotype to phenotype
-
-After inferring founder allele dosages, I address the
-second major statistical challenge of
-QTL mapping: the ``model selection'' problem \citep{broman2009guide}.
-A genetically ``additive'' linear model, in which
-I assume a linear relationship between
-a trait's values and each founder allele's dosage, is a popular default model.
-\citep{gatti2014quantitative,broman2019rqtl2}.
-
-% model form - random effects incorporation & relatedness
-% First, relatedness
-Due to the complexity of the breeding design, Diversity Outbred mice
-have
-complicated pairwise relationships. A given study cohort may include pairs of first
-cousins, parent-offspring pairs, grandparent-grandchild pairs, second cousin pairs, and others.
-Because relatedness can confound genotype-phenotype associations
-\citep{yang2014advantages}, researchers have developed methods to account
-for relatedness in their statistical models. One popular approach involves
-use of a polygenic random effect in the statistical model \citep{kang2008efficient}.
-The inclusion of a random effect in a model with fixed (\emph{i.e.}, nonrandom)
-effects results in what statisticians call a ``linear mixed effects model''.
-
-One then fits a linear mixed effects model and calculates the log$_{10}$ likelihood at each position.
-As in two-parent crosses (Section~\ref{sec:qtl-two-parent}), one summarizes evidence for a QTL by plotting LOD scores across the genome.
-Because multiple hypothesis tests are performed, one typically wants to control family-wise error rate.
-One strategy to do this involves a permutation test, much like that in two-parent crosses \citep{churchill1994empirical}.
-
-%%% CHAPTER 1 End
-%%%%%%%%%%%%%%%%%
-%%%%% 3A start
-\section{Expression trait hotspot dissection}\label{sec:hotspot-dissection}
-\subsection{Introduction}
-
-A central goal of systems genetics studies is to identify causal relationships between biomolecules.
-Recent work by \citet{chick2016defining}, which builds on research from \citet{baron1986moderator}, has popularized linear regression-based mediation analysis for causal inference in systems genetics.
-Because of the great successes of mediation analysis in systems genetics
-\citep{chick2016defining,keller2018genetic}, we need to clarify a role for our pleiotropy test.
-We argue below that our pleiotropy test complements mediation analysis.
-
-Second, when regression-based mediation analysis fails to identify a mediator, our
-pleiotropy test still provides information on the number of QTL, which may aid biological
-understanding and inform subsequent studies. Below, we first review the prerequisite
-molecular biology, including the ``central dogma'', before discussing in some detail
-regression-based mediation analysis in the systems genetics context.
-
-\citet{crick1958protein} articulated a pathway for transmission of biological information
-that is now known as the central dogma of molecular biology (Figure)
-\citep{crick1970central}.
-In it, he argued that information encoded in DNA sequence is transmitted via transcription
-to
-RNA molecules, which, in turn, transfer the information to proteins via translation.
-The process of transcription uses DNA as a template for creating a RNA molecule that
-conveys the information encoded in DNA.
-In this sense, every gene leads to a unique RNA molecule.
-Translation is the molecular biology process by which an RNA molecule's information is
-transferred to a protein.
-As in transcription, the nucleic acid (DNA in transcription and RNA in translation) serves
-as a template for synthesis of the product (RNA in transcription and protein in translation).
-Thus, RNA molecules from distinct genes lead to different proteins.
-
-This sequence of information transfer, from DNA to RNA to protein, provides a natural
-setting by which to examine mediation analysis. If a DNA variant affects protein
-concentrations only through its gene's RNA transcripts, then conditioning on RNA transcript
-levels would greatly reduce the strength of association between DNA variant and protein
-concentration. Before we continue our discussion, we define key terms and discuss an
-example below.
-
-We continue by stating what it means for one trait to mediate a relationship between a
-DNA variant and another trait. To clarify our discussion, we refer to an example from
-\citet{chick2016defining} (Figure). \citet{chick2016defining}, in
-studying livers of 192 Diversity Outbred mice, found evidence that \emph{Dhtkd1}
-transcript levels associated with a Chromosome 2 marker near the \emph{Dhtkd1} gene.
-They also found that
-the same marker affected DHTKD1 protein concentrations. (Note that DHTKD1 protein is the
-product of translation of \emph{Dhtkd1} transcripts.)
-As anticipated, mediation analysis, in which the DHTKD1 protein concentrations are
-regressed on founder allele dosages (at the Chromosome 2 marker) demonstrated that
-\emph{Dhtkd1} transcript levels act as a mediator between DHTKD1 protein concentrations
-and founder allele dosages. In fact, the extent of the reduction in association
-strength indicates that the primary pathway by which the genetic marker affects DHTKD1
-protein concentrations is through \emph{Dhtkd1} transcript levels.
-
-In our studies below, we follow \citet{keller2018genetic} by generalizing this setting to
-the case where a DNA variant affects a local transcript level, which then affects a
-nonlocal transcript level.
-We term a transcript ``local'' to a marker when its gene is near that marker.
-We use a threshold of no more than 2 Mb to restrict the
-number of local transcripts for a given marker.
-A nonlocal transcript, then, is either one that arises from a gene on another chromosome or
-from a distant gene on the same chromosome as the marker.
-
-It is highly plausible that concentration variations in one transcript may
-affect abundances of a second transcript.
-For example, the first transcript may encode a transcription factor protein.
-In this case, the transcription factor protein may influence expression
-patterns of the second transcript (and perhaps other transcripts, too).
-
-To determine whether a local transcript level mediates the relationship between
-a nonlocal transcript level and a DNA variant, we perform a series of regression analyses,
-which we detail below.
-In brief, we regress the nonlocal transcript levels on founder allele dosages at the DNA
-variant, with and without conditioning on the candidate mediator (the local transcript levels).
-If the LOD score measuring association between nonlocal trait and QTL genotype diminishes
-sufficiently upon conditioning on a candidate,
-then we declare the candidate a mediator.
-
-The rationale behind this strategy follows.
-If the DNA variant affects the nonlocal transcript levels solely by way of local
-transcript levels, then conditioning on the local transcript levels would nullify the
-relationship between DNA variant and the nonlocal transcript levels.
-At the other extreme, if the DNA variant affects nonlocal transcript levels solely
-through mechanisms that don't involve the local transcript levels, then conditioning
-on local transcript levels would not affect the association between the DNA variant and
-nonlocal transcript levels.
-
-%\subsection{Regression-based mediation methods}
-
-%\begin{equ}[!ht]
-\begin{equation}
-Y = b1 + WC + E
-\label{model1}
-\end{equation}
-%\caption{Linear model with intercept and covariates only.}
-%\end{equ}
-
-%\begin{equ}[!ht]
-\begin{equation}
-Y = XB + WC + E
-\label{model2}
-\end{equation}
-%\caption{Linear model with founder allele dosages and covariates.}
-%\end{equ}
-
-%\begin{equ}[!ht]
-\begin{equation}
-Y = b1 + WC + M\beta + E
-\label{model3}
-\end{equation}
-%\caption{Linear model with intercept, covariates, and candidate mediator.}
-%\end{equ}
-
-%\begin{equ}[!ht]
-\begin{equation}
-Y = XB + WC + M\beta + E
-\label{model4}
-\end{equation}
-%\caption{Linear model with founder allele dosages, covariates, and candidate mediator.}
-%\end{equ}
-
-We now consider the procedures needed for a mediation analysis in systems genetics.
-In the four linear regression models, $X$ is an $n$ by $8$ matrix of
-founder allele dosages at a single marker, $B$ is an $8$ by $1$ matrix of founder allele
-effects, $E$ is an $n$ by $1$ matrix of random errors, $b$
-is a number, $1$ is
-an $n$ by $1$ matrix with all entries set to $1$, $Y$ is
-an $n$ by $1$ matrix of
-phenotype values (for a single trait), and $M$ is an $n$
-by $1$ matrix of values for a putative mediator.
-$C$ is a $c$ by 1 matrix of covariate effects, and $W$ is
-a $n$ by $c$ matrix of covariates.
-We denote the coefficient of the mediator by $\beta$.
-
-We assume that the vector $E$ is (multivariate) normally
-distributed with zero vector as mean and covariance matrix
-$\Sigma = \sigma^2I_n$, where $I_n$ is the $n$ by $n$
-identity matrix.
-
-In the above models with normally distributed random
-errors, the log-likelihoods are easily calculated. For
-example, in Equation \ref{model1}, the vector $Y$ follows
-a multivariate normal distribution with mean $(b1 + WC)$
-and covariance $\Sigma = \sigma^2I$. Thus, we can write
-the likelihood for Model \ref{model1} as:
+For each hotspot, we paired each local trait with each nonlocal trait and performed a
+two-dimensional, two-QTL scan over a genomic region that included the entirety of the
+hotspot. We fitted bivariate linear mixed effects models at each ordered pair of markers \citep{boehm2019testing}.
 
 \begin{equation}
-    L(b, C, \sigma^2| Y, W) = (2\pi)^{- \frac{n}{2}}\exp{ \left(- \frac{1}{2}(Y - b1 - WC)^T\Sigma^{-1}(Y - b1 - WC)\right)}
+vec(Y) = X vec(B) + vec(G) + vec(E)
+\label{eq:mvlmm}
 \end{equation}
 
-We thus have the following equation (~\ref{eq:ll} for the
-log-likelihood for Model~\ref{model1}:
+From the two-dimensional scans, we calculated pleiotropy test statistics for each pair of traits \citep{jiang1995multiple,boehm2019testing}.
 
 \begin{equation}
-    \log L(b, C, \sigma^2 | Y, W) = - \frac{n}{2}\log (2\pi) - \frac{1}{2} (Y - b1 - WC)^T\Sigma^{-1}(Y - b1 - WC)\label{eq:ll}
+\Lambda = \max_{\lambda_1, \lambda_2}ll(\lambda_1, \lambda_2) - \max_{\lambda} ll(\lambda, \lambda)
+\label{eq:lrt}
 \end{equation}
 
-\citet{chick2016defining} calculated the log$_{10}$
-likelihoods for all four models before determining two LOD
-scores (Equations~\ref{eq:lod1} and~\ref{eq:lod2}).
+\section{Results}
 
-\begin{equation}
-LOD_1 = log_{10}(\text{Model~\ref{model2} likelihood}) - log_{10}(\text{Model~\ref{model1} likelihood})\label{eq:lod1}
-\end{equation}
+\section{Discussion}
 
-\begin{equation}
-LOD_2 = log_{10}(\text{Model~\ref{model4} likelihood}) - log_{10}(\text{Model~\ref{model3} likelihood})\label{eq:lod2}
-\end{equation}
-
-And, finally, \citet{chick2016defining} calculated the LOD
-difference statistic (Equation~\ref{eq:lod-diff}).
-
-\begin{equation}
-\text{LOD difference} = LOD_1 - LOD_2\label{eq:lod-diff}
-\end{equation}
-
-LOD difference values need not be positive.
-For example, in the setting where the putative mediator has no effect on the phenotype-QTL association, $LOD$ difference may be negative.
-
-In our analyses below, we consider the LOD difference
-proportion (Equation~\ref{eq:lod-diff-prop}).
-
-\begin{equation}
-\text{LOD difference proportion} = \frac{(LOD_1 - LOD_2)}{LOD_1}
-\label{eq:lod-diff-prop}
-\end{equation}
-
-In other words, we consider what proportion of the association strength, on the LOD scale, is diminished by conditioning on a putative mediator.
-Scaling the LOD difference statistic by LOD$_1$ accommodates the diversity of LODs in our data, which will be useful for graphical comparisons.
-For example, a LOD difference statistic of 5 may be relevant when
-a trait has a LOD$_1$ of 10, but unimportant if the trait's LOD$_1$ is 100.
-Now that we've defined our summary statistics for a mediation analysis in systems genetics, we turn attention to two statistical challenges in this area of research, 1.
-confounding and 2.
-significance thresholds.
-
-%\subsection{Four assumptions for causal inference}
-
-  \begin{enumerate}
-\item No unmeasured confounding of the DNA variant-nonlocal transcript levels relationship
-\item No unmeasured confounding of the local transcript levels-nonlocal transcript levels relationship
-\item No unmeasured confounding of the DNA variant-local transcript levels relationship
-\item No local transcript levels-nonlocal transcript levels confounder that is affected by the DNA variant
-\end{enumerate}
-
-Our first statistical challenge is due to confounding. To claim that the LOD
-difference statistic reflects a causal relationship, four assumptions about
-confounding are needed \citep{vanderweele2015explanation},
-yet
-it is often difficult or impossible to recognize unmeasured confounders. Unmeasured
-confounders, as I use the term here, are unmeasured random variables that confound a causal
-relationship between two other random variables.
-For example, Figure shows a directed graph. Arrows indicate causal
-directions. For example, ``QTL'' causes ``M''. ``M'' designates the mediator (local transcript). ``Y'' denotes the nonlocal transcript
-level in this setting, while ``C1'', ``C2'', and ``C3'' are covariates that confound,
-respectively, the DNA variant - nonlocal transcript level relationship, the local transcript
-level - nonlocal transcript level relationship, and the DNA variant - local transcript level
-relationship. Failure to measure C1, C2, or C3 leads to violations of Assumption 1,
-Assumption 2, or Assumption 3, respectively.
-
-In studies of Diversity Outbred mice, relatedness is a possible confounder,
-yet the linear models (Equations~\ref{model1},~\ref{model2},~\ref{model3},~\ref{model4}) fail to account for the complex relatedness patterns
-among Diversity Outbred mice.
-In fact, the only covariates in our models are for wave membership and sex.
-The 400 mice were shipped from the Jackson Laboratory in four waves of 100 mice.
-Because each wave was processed at a different time, I include wave membership
-as a covariate with the goal of accounting for wave-related effects.
-Other sources of confounding, such as batch effects in the phenotyping, may be
-unmeasured.
-
-In efforts to quantify the potential impact of unmeasured confounding, scientists have
-developed a suite of sensitivity analysis tools for use in regression-based mediation analysis.
-While a discussion of sensitivity analysis is beyond the scope of this thesis,
-it may be useful in future systems genetics studies to assess robustness of mediation analysis
-results in the presence of unmeasured confounders.
-\citet{vanderweele2015explanation} discusses sensitivity analysis in the context of epidemiological studies.
-
-%\subsection{Declaring mediators}
-
-Besides accounting for confounding, a second statistical challenge in mediation analysis is assessing significance of the LOD difference statistic.
-\citet{chick2016defining} used individual transcript levels as sham mediators and tabulated their LOD difference statistics.
-They then compared the observed LOD difference statistics for putative mediators to the empirical distribution of LOD difference statistics obtained from the collection of sham mediators.
-\citet{keller2018genetic}, on the other hand, in their study of pancreatic islet
-biology, declared mediators to be those local transcripts that diminished the LOD score
-of nonlocal transcripts by at least 1.5.
-While significance threshold determination remains an active area of research,
-we proceed below by examining all 147 nonlocal transcript levels that
-\citet{keller2018genetic} identified as mapping to the Chromosome 2 hotspot.
-
-\subsection{Methods}
-
-We examined the potential that the two methods, 1. pleiotropy vs. separate QTL testing and
-2. mediation analysis, play complementary roles in efforts to dissect gene expression trait hotspots.
-We use the term ``hotspot'' to refer to a contiguous genomic region,
-typically no more than 5 Mb in length, that affects many expression traits.
-After describing our data below, we detail our statistical analyses involving 13 local
-gene expression traits and 147 nonlocal gene expression traits,
-all of which map to a 4-Mb hotspot on Chromosome 2.
-
-%\subsection{Data description}
-
-We analyzed data from 378 Diversity Outbred mice \citep{keller2018genetic}.
-\citet{keller2018genetic} genotyped tail biopsies with the GigaMUGA microarray \citep{morgan2015mouse}.
-They also used RNA sequencing to measure genome-wide pancreatic islet gene expression
-for each mouse at the time of sacrifice \citep{keller2018genetic}.
-
-We examined the Chromosome 2 pancreatic islet expression trait hotspot that \citet{keller2018genetic} identified.
-\citet{keller2018genetic} found that 147 nonlocal traits map to the 4-Mb region
-centered at 165.5 Mb on Chromosome 2.
-The 147 nonlocal traits all exceeded the genome-wide LOD significance threshold,
-7.18 \citep{keller2018genetic}.
-With regression-based mediation analyses, they identified transcript levels of local gene \emph{Hnf4a} as a mediator of 88 of these 147 nonlocal traits.
-
-We designed a study to examine the possible roles for mediation analysis and pleiotropy testing.
-Because \citet{keller2018genetic} reported that some nonlocal traits that map to
-the Chromosome 2 hotspot did not demonstrate evidence of mediation by \emph{Hnf4a} expression
-levels, we elected to study a collection of local gene expression traits,
-rather than \emph{Hnf4a} alone.
-This strategy enabled us to ask whether one of twelve other local traits mediates those
-nonlocal hotspot traits that are not mediated by \emph{Hnf4a}.
-Our set of local gene expression traits includes \emph{Hnf4a} and 12
-other local genes (Table \ref{tab:annot}).
-Our 13 local genes are the only genes that met three criteria.
-The 147 nonlocal traits that we studied all had LOD peak heights above 7.18 and QTL positions within 2 Mb of the center of the hotspot (at 165.5 Mb).
-
-\begin{enumerate}
-    \item QTL peak with LOD $>$ 40
-    \item QTL peak position within 2 Mb of hotspot center (165.5 Mb)
-    \item Gene midpoint within 2 Mb of hotspot center (165.5 Mb)
-\end{enumerate}
-
-%\subsection{Statistical analyses}
-
-We now describe our statistical analyses.
-After univariate QTL mapping to identify expression traits that
-map to the Chromosome 2 hotspot,
-we performed both bivariate QTL scans and mediation analyses of all 13 * 147 = 1911 pairs
-involving one local expression trait and one nonlocal expression trait.
-
-%\subsubsection{Bivariate QTL scans}
-
-Our bivariate QTL analyses involved the same 13 local expression traits and 147 nonlocal expression traits.
-We described above the criteria for choosing these expression traits.
-
-We performed a series of two-dimensional QTL scans in which we paired each local gene's
-transcript levels with each nonlocal gene's transcript levels,
-for a total of 13 x 147 = 1,911 two-dimensional scans.
-Each scan examined the same set of 180 markers, which spanned the interval from 163.1 Mb to 167.8 Mb and included univariate peaks for all 13 + 147 = 160 expression traits.
-We performed these analyses with the R package \texttt{qtl2pleio} \citep{qtl2pleio}.
-
-For each bivariate QTL scan, we fitted a collection of bivariate models for
-all 180 * 180 = 32,400 ordered pairs of markers.
-For each ordered pair of markers, we fitted a bivariate linear mixed
-effects model using the methods of Chapter 2.
-
-%\subsubsection{Mediation analyses}
-
-We performed mediation analyses for
-all 1,911 local-nonlocal trait pairs in which
-we probed the extent to which each nonlocal trait's association
-strength diminished upon conditioning on transcript levels of a putative mediator.
-
-Each of the 13 local expression traits, considered one at a time, served as putative mediators.
-We thus fitted the four linear regression models that we describe
-above (Equations \ref{model1}, \ref{model2}, \ref{model3}, \ref{model4}).
-
-One question that needs clarification is the choice of genetic marker
-for each mediation analysis.
-We elected to use the founder allele dosages at the marker (or pseudomarker) that demonstrated the
-univariate LOD peak for the nonlocal trait.
-Alternative analyses, in which one uses the founder allele dosages at which the
-local trait has its univariate peak, are also possible.
-
-%\subsubsection{Local gene analyses}
-
-% latex table generated in R 3.5.1 by xtable 1.8-3 package
-% Mon Dec 17 10:04:32 2018
-\begin{table}[ht]
-\centering
-\begin{tabular}{lrrrr}
-  \hline
-Gene & Start & End & QTL peak position & LOD\\
-  \hline
-Pkig & 163.66 & 163.73 & 163.52 & 51.68 \\
-  Serinc3 & 163.62 & 163.65 & 163.58 & 126.93 \\
-  Hnf4a & 163.51 & 163.57 & 164.02 & 48.98 \\
-  Stk4 & 164.07 & 164.16 & 164.03 & 60.39 \\
-  Pabpc1l & 164.03 & 164.05 & 164.03 & 52.50 \\
-  Slpi & 164.35 & 164.39 & 164.61 & 40.50 \\
-  Neurl2 & 164.83 & 164.83 & 164.64 & 64.58 \\
-  Cdh22 & 165.11 & 165.23 & 165.05 & 53.84 \\
-  2810408M09Rik & 165.49 & 165.49 & 165.57 & 67.34 \\
-  Eya2 & 165.60 & 165.77 & 165.72 & 98.89 \\
-  Prex1 & 166.57 & 166.71 & 166.75 & 46.91 \\
-  Ptgis & 167.19 & 167.24 & 167.27 & 56.25 \\
-  Gm14291 & 167.20 & 167.20 & 167.27 & 73.72 \\
-   \hline
-\end{tabular}
-\caption{Local gene annotations for analysis of Chromosome 2 expression trait hotspot.
-All positions are in units of Mb on Chromosome 2.
-LOD peak position and LOD peak height refer to those obtained from univariate analyses.
-``Start'' and ``end'' refer to the local gene's DNA start and end positions, as annotated by Ensembl version 75.}
-\label{tab:annot}
-\end{table}
-
-To visualize the summary statistics for the two methods, I plotted,
-for each local gene, a scatterplot of LOD difference proportion values
-against pleiotropy test statistics.
-
-%\subsubsection{Nonlocal gene analyses}
-
-We also examined the 1,911 pairs by considering one nonlocal gene at a time.
-For each of the 147 nonlocal genes, I plotted LOD difference proportion
-against pleiotropy test statistic values.
-I present below (Figure ) examples that illustrate some of the
-observed patterns between pleiotropy test statistics and LOD difference proportion values.
-
-\subsection{Results}
-
-%\subsection{Scatter plot for all 1911 pairs}
-Figure is a scatterplot of the results from all pleiotropy tests and mediation analyses.
-Each point contains one local expression trait and one nonlocal expression trait.
-I see that points with high values of LOD difference proportion tend to have small
-values of pleiotropy test statistic, and those points with high values of the
-pleiotropy test statistic tend to have small values of LOD difference proportion.
-
-Some points demonstrate large values of pleiotropy test statistic (\emph{eg.}, $>$ 10),
-yet still have sizeable LOD difference proportion statistics (Figure).
-One possible explanation for such points is that the univariate LOD (LOD$_1$) value is small,
-so that even a modest LOD difference gives rise to a sizeable LOD difference proportion value.
-A second possible explanation is that the phenotype variances for both phenotypes in the pair
-are large enough to skew the null distribution of the test statistic towards larger values.
-To distinguish between these two, I would
-examine the LOD difference statistics and obtain bootstrap $p$-values.
-
-In Figure, I colored red the points that involve \emph{Hnf4a};
-points with other local genes are blue.
-The most striking feature of the coloring is that many red points have small
-values of the pleiotropy test statistics and very high values of LOD difference
-proportion. This is what I expect when \emph{Hnf4a} mediates the relationship
-between a nonlocal trait and a QTL; namely, \emph{Hnf4a} and a nonlocal trait
-share a pleiotropic QTL and conditioning on \emph{Hnf4a} diminishes the LOD
-score for the association between nonlocal trait and QTL.
-
-%\subsection{Local gene analyses}
-
-To more thoroughly examine the pleiotropy test and mediation analysis
-relationships across the 13 local genes,
-I created 13 plots of LOD difference proportion against pleiotropy test
-statistic (Figures).
-They reveal common patterns.
-First, I see no points in the upper right quadrant of each plot (Figure).
-This tells us that those nonlocal genes with high values of pleiotropy test statistic
-(when paired with the specified local gene) have low values of LOD difference proportion.
-Similarly, those nonlocal genes with high values of LOD difference proportion tend
-to have small values of the pleiotropy test statistic.
-Finally, some trait pairs demonstrate low values of both
-the LOD difference proportion and pleiotropy test statistic.
-This observation suggests that, for a given local expression trait, the nonlocal trait
-is not mediated by the local expression trait yet it shares a pleiotropic locus.
-Low power to resolve multiple nearby QTL may give rise to such findings.
-
-In comparing the \emph{Hnf4a} plot (Figure ) with the other
-12 plots (Figure ),
-I see that none of the 12 plots in Figure  closely resembles Figure .
-\emph{Serinc3}, \emph{Stk4}, \emph{Neurl2}, and \emph{Cdh22} are closest in appearance to the plot of \emph{Hnf4a}.
-However, each of \emph{Serinc3}, \emph{Stk4}, \emph{Neurl2}, and \emph{Cdh22}
-has very few points with LOD difference proportion above 0.5,
-while \emph{Hnf4a} has many points with LOD difference proportion above 0.5.
-
-%\subsection{Nonlocal gene analyses}
-
-Now that I've examined our results from the perspective of every one of 13 local gene expression traits,
-I turn attention to the nonlocal gene expression trait perspective.
-I present four nonlocal gene expression trait plots in Figure. I chose
-these four from the 147 to demonstrate two of the observed patterns. In a single scatterplot,
-I've plotted 13 points, one for each local gene expression trait, paired with a single
-nonlocal gene expression trait. Local expression trait \emph{Hnf4a} is colored in red,
-while the other local genes are colored in blue. The top row of Figure
-features scatterplots for nonlocal expression traits \emph{Abcb4} and \emph{Smim5}. I see
-that local expression trait \emph{Hnf4a} has, by far, the greatest value of LOD difference
-proportion. In the second row of Figure, I see a different pattern in
-the plots for nonlocal expression traits \emph{Tmprss4} and \emph{Gm3095}. Both demonstrate,
-for \emph{Hnf4a}, small values for the pleiotropy test statistic and small values of LOD
-difference proportion. This tells me that \emph{Hnf4a} shares a pleiotropic locus with both
-\emph{Tmprss4} and \emph{Gm3095}, yet none of the 13 local gene expression traits mediates
-the relationship between these two nonlocal traits and the QTL.
-
-\subsection{Discussion}
-
-Our pairwise analyses with both mediation analyses and pleiotropy tests provide additional
-evidence for the importance of \emph{Hnf4a} in the biology of the Chromosome 2 hotspot
-in pancreatic islets.
-Our analyses, and, specifically, the test of pleiotropy vs. separate QTL, may be more useful
-when studying nonlocal traits that map to a hotspot yet don't show strong evidence
-of mediation by local expression traits.
-In such a setting, the pleiotropy test can, at least, provide some information about
-the genetic architecture at the hotspot.
-Specifically, our pleiotropy test may inform inferences about the number of
-underlying QTL in a given expression trait hotspot.
-Additionally, our test may limit the number of expression traits that are
-potential intermediates between a QTL and a specified nonlocal expression trait.
-This relies on the assumption that a causally intermediate (local) expression trait and a
-target (nonlocal) expression trait presumably share a QTL. On the other hand, mediation analyses, when they provide evidence for mediation of a nonlocal
-trait by a local expression trait, are possibly more valuable than the test of pleiotropy vs.
-separate QTL, since the mediation analyses identify precisely the intermediate expression trait.
-
-I recommend using both tests of pleiotropy vs separate QTL and mediation analyses when
-dissecting an expression trait QTL hotspot.
-Future researchers may use the test of pleiotropy vs. separate QTL as a
-preliminary screen to limit the set of candidate mediators to those that have
-test results that are consistent with pleiotropy. Subsequently, fewer mediation
-analyses are needed, which would reduce the magnitude of the multiple testing
-adjustment. This, in turn, would lead to detection of more true positive signals
-in mediation analysis.
-
-Future research may investigate the use of polygenic random effects in the statistical models
-for mediation analysis. Additional methodological questions include approaches for declaring
-significant a mediation LOD difference proportion and consideration of other possible measures
-and scales of extent of mediation. Additionally, future researchers may wish to consider
-biological models that contain two mediators.
-
-The social and health sciences have witnessed much methods research in mediation analysis.
-The field of statistical genetics has not fully adopted these strategies yet, but, given the
-nature of current and future data, many opportunities exist for translation of approaches from
-causal inference in epidemiology to systems genetics.
-For example, \citet{vanderweele2015explanation} contains detailed discussions of many methods
-issues that arose in mediation analyses in
-epidemiology studies.
-
-%%%%% 3A end
 \printbibliography

### 2019-06-19 analysis/paper/paper.Rnw
documentclass{book} -> {article}
removed unneeded packages

@@
-\documentclass[oneside]{book}
+\documentclass[]{article}
 \usepackage[colorinlistoftodos]{todonotes}
@@
 \usepackage{graphicx}
-\usepackage{color, colortbl}
-\definecolor{LightCyan}{rgb}{0.88,1,1}
-\usepackage{subcaption}
-\usepackage{array}
+\usepackage{color, colortbl, xcolor}
 \usepackage{amssymb}
@@
 \usepackage{longtable}
-\usepackage[outdir=./figs/]{epstopdf}
+%\usepackage[outdir=./figs/]{epstopdf}
 \usepackage{etoolbox} % for messing with page style & title pages
-\usepackage{fancyhdr}
-\setlength{\headheight}{15.2pt}
-\pagestyle{fancyplain}
-\fancyhf{}
-\lhead{}
-\chead{}
-\rhead{ \fancyplain{}{\thepage} } % put page number in upper right corner
-\rfoot{Compiled on \today\ at \currenttime}
-\renewcommand{\headrulewidth}{0pt} % remove top line
-\renewcommand{\footrulewidth}{0pt} % remove bottom line
-\renewcommand{\headrule}{}
+%\usepackage{fancyhdr}
+%\setlength{\headheight}{15.2pt}
+%\pagestyle{fancyplain}
+%\fancyhf{}
+%\lhead{}
+%\chead{}
+%\rhead{ \fancyplain{}{\thepage} } % put page number in upper right corner
+%\rfoot{Compiled on \today\ at \currenttime}
+%\renewcommand{\headrulewidth}{0pt} % remove top line
+%\renewcommand{\footrulewidth}{0pt} % remove bottom line
+%\renewcommand{\headrule}{}%
 
-\title{Expression QTL hotspot dissection}
+\title{Expression QTL hotspot dissection with mediation analysis and pleiotropy testing}
 \author{Frederick J. Boehm, Brian S. Yandell, and Karl W. Broman}
@@
 
-<<include=FALSE>>=
+<<setup, include=FALSE>>=
 library(knitr)
@@
 \doublespacing
+\listoftodos
 \tableofcontents
@@
 %\markright{\hfill}
-\chapter{Introduction}
+\section{Introduction}
 
@@
 
+\subsection{Mediation analysis}
+
+For each hotspot and each local trait - nonlocal trait pair, we also performed a
+mediation analysis. Each mediation analysis involved fitting four linear models (Equations~\ref{model1}, \ref{model2}, \ref{model3}, \ref{model4}).
+
+\begin{equation}
+Y = \alpha 1 + WC + E
+\label{model1}
+\end{equation}
+
+\begin{equation}
+Y = XB + WC + E
+\label{model2}
+\end{equation}
+
+\begin{equation}
+Y = \alpha 1 + WC + M\beta + E
+\label{model3}
+\end{equation}
+
+\begin{equation}
+Y = XB + WC + M\beta + E
+\label{model4}
+\end{equation}
+
+In all mediation analyses, the nonlocal transcripts are the targets, while the local transcripts serve as candidate mediators of the QTL-target relationship.
+
 \section{Results}
@@
 
-<<r install_here>>=
+<<install_here>>=
 install.packages("here", repos = "https://cloud.r-project.org")
@@
 
-<<r colophon, cache = FALSE>>=
+<<colophon, cache = FALSE>>=
 # which R packages and versions?
@@
 
-<<r git2r, cache = FALSE>>=
+<<git2r, cache = FALSE>>=
 # what commit is this file at?

### 2019-06-19 analysis/paper/paper.Rnw
removed more unnecessary packages
fixed spacing issue with session_info()

@@
 \linenumbers
-
+\usepackage{setspace}
 \usepackage{bbm}
-\usepackage{setspace}
 \usepackage[left = 1in, right = 1in, top = 1.5in, bottom = 1in]{geometry}
-\usepackage{multirow}
-\usepackage{float}
-\usepackage{newfloat}
-\usepackage{multicol}
-%\usepackage[lite]{mtpro2} %contains \widetilde
-\usepackage[titletoc]{appendix}
+%\usepackage[titletoc]{appendix}
 \usepackage{amsmath}
@@
 \usepackage{graphicx}
-\usepackage{color, colortbl, xcolor}
+\usepackage{color, colortbl}
 \usepackage{amssymb}
@@
 \usepackage{amsfonts}
-\usepackage{longtable}
+%\usepackage{longtable}
 %\usepackage[outdir=./figs/]{epstopdf}
-\usepackage{etoolbox} % for messing with page style & title pages
-%\usepackage{fancyhdr}
-%\setlength{\headheight}{15.2pt}
-%\pagestyle{fancyplain}
-%\fancyhf{}
-%\lhead{}
-%\chead{}
-%\rhead{ \fancyplain{}{\thepage} } % put page number in upper right corner
-%\rfoot{Compiled on \today\ at \currenttime}
-%\renewcommand{\headrulewidth}{0pt} % remove top line
-%\renewcommand{\footrulewidth}{0pt} % remove bottom line
-%\renewcommand{\headrule}{}%
+%\usepackage{etoolbox} % for messing with page style & title pages
+\usepackage{fancyhdr}
+\setlength{\headheight}{15.2pt}
+\pagestyle{fancyplain}
+\fancyhf{}
+\lhead{}
+\chead{}
+\rhead{ \fancyplain{}{\thepage} } % put page number in upper right corner
+\rfoot{Compiled on \today\ at \currenttime}
+\renewcommand{\headrulewidth}{0pt} % remove top line
+\renewcommand{\footrulewidth}{0pt} % remove bottom line
+\renewcommand{\headrule}{}%
 
@@
 \date{\today}
+
+\doublespacing
 
@@
 message = FALSE)
+
+options(tinytex.verbose = TRUE)
 @
 
-\doublespacing
 \listoftodos
@@
 
+\begin{singlespacing}
+
 <<colophon, cache = FALSE>>=
@@
 
+\end{singlespacing}
+
 \end{document}

### 2019-06-24 R/id_pleio_sets.R
added R functions to package (and added tests & man pages)

@@ -1,0 +1,42 @@
+#' Cluster pleiotropy test statistics
+#'
+#' @param tib a tibble with three columns; the first two are gene ids and the last is pleiotropy test statistics
+#' @param binarize logical of length one indicating whether to binarize the matrix of pleiotropy test statistics based on pleio_threshold value
+#' @param pleio_threshold a positive number to distinguish pleiotropy from separate QTL
+
+cluster_pleio_stats <- function(tib, binarize = FALSE, pleio_threshold = NULL){
+  # convert tibble to matrix
+  mat <- tibble_to_matrix(tib)
+  # binarize matrix, optionally
+  if (binarize) mat <- mat < pleio_threshold
+  # sort matrix
+  hclust(dist(mat))$order -> ord
+  bmat[ord, ord]
+}
+
+
+#' Convert tibble to matrix
+#'
+#' @param tib
+
+tibble_to_matrix <- function(tib){
+  rn <- unique(tib[[1]])
+  cn <- unique(tib[[2]])
+  out <- matrix(NA, nrow = length(rn), ncol = length(cn))
+  rownames(out) <- rn
+  colnames(out) <- cn
+  for (i in 1:length(rn)){
+    for (j in 1:length(cn)){
+      out[i, j] <- tib[[3]][(tib[[1]] == rownames(out)[i] & tib[[2]] == colnames(out)[j])]
+    }
+  }
+  return(out)
+}
+
+#' Sort a binary matrix into something close to a block diagonal
+#'
+#' @param bmat a binary matrix
+
+sort_binary_matrix <- function(bmat){
+  bmat[ , order(colSums(bmat), decreasing = TRUE)]
+}

### 2019-06-24 tests/testthat/test-tibble_to_matrix.R

@@ -1,0 +1,9 @@
+library(hotspots)
+library(testthat)
+
+foo <- tibble::tibble(v1 = letters[c(1, 1, 2, 2)], v2 = letters[c(3, 4, 3, 4)], lrt = runif(4))
+
+test_that("tibble converts to matrix with correct dimensions", {
+  expect_equal(nrow(tibble_to_matrix(foo)), 2)
+  expect_equal(ncol(tibble_to_matrix(foo)), 2)
+})

### 2019-06-24 analysis/paper/Chr2-prelim.Rmd
updated Chr2 prelim analysis code

@@
 ```
+
+## Filter to remove local traits that are outside hotspot
+
+In initial analyses for each hotspot, I included too many local traits.
+Namely, I considered all local traits that are within 7 Mb of the hotspot midpoint.
+I wish to limit this to traits that are within 3 Mb of the hotspot midpoint.
+
+```{r define_foo}
+foo <- mylist %>%
+  bind_rows() %>%
+  rename(pleiotropy_lod = lod) %>%
+  left_join(tib, by = c("local_gene_id", "nonlocal_gene_id")) %>%
+  select(local_symbol, nonlocal_symbol, pleiotropy_lod, lod_diff, lod_diff_proportion, lod_without_med, lod_with_med, filename, nonlocal_gene_id, local_gene_id) 
+```
+
+```{r add_local_peak_positions}
+foo %>%
+  left_join(local_annot2, by = c("local_gene_id" = "annot.id")) %>%
+  rename(local_peak_pos = pos, local_peak_lod = lod) %>%
+  select(- marker.id, - chrom, - strand, - biotype, - local_tx) %>%
+  filter(local_peak_pos < hot_upper, local_peak_pos > hot_lower)
+
+```
+
+```{r}
+foo %>%
+  left_join(local_annot2, by = c("local_gene_id" = "annot.id")) %>%
+  rename(local_peak_pos = pos, local_peak_lod = lod) %>%
+  select(- marker.id, - chrom, - strand, - biotype, - local_tx) %>%
+  filter(local_peak_pos < hot_upper, local_peak_pos > hot_lower) %>%
+  select(middle) %>% unlist() %>% range()
+
+```
+
+## Convert tibble to matrix for clustering of pleiotropy test statistics
+
+```{r, hclust_pleio_stats}
+bar <- foo %>%
+  select(nonlocal_gene_id, local_gene_id, pleiotropy_lod) %>%
+hotspots::tibble_to_matrix() %>%
+  dist() %>%
+  hclust() 
+```
+
+```{r load_colors}
+library(RColorBrewer) 
+hmcol <- colorRampPalette(brewer.pal(9, "GnBu"))(100)
+```
+
+```{r heat_pleio_stats}
+foo %>%
+  select(nonlocal_gene_id, local_gene_id, pleiotropy_lod, lod_diff) %>%
+hotspots::tibble_to_matrix() %>%
+  t() %>%
+  gplots::heatmap.2(cexRow = 0.1, cexCol = 0.1, col = hmcol, trace = "none")
+```
+
+```{r heat_lod_diff}
+foo %>%
+  select(nonlocal_gene_id, local_gene_id, lod_diff) %>%
+hotspots::tibble_to_matrix() %>%
+  t() %>%
+  gplots::heatmap.2(cexRow = 0.1, cexCol = 0.1, col = hmcol, trace = "none")
+```
+
+```{r heat_lod_diff0}
+foo %>%
+  select(nonlocal_gene_id, local_gene_id, lod_diff) %>%
+  mutate(lod_diff = lod_diff * (lod_diff > 0)) %>%
+hotspots::tibble_to_matrix() %>%
+  t() %>%
+  gplots::heatmap.2(cexRow = 0.1, cexCol = 0.1, col = hmcol, trace = "none")
+```
+
+```{r, define_hld, eval = FALSE}
+hld <- foo %>%
+  select(nonlocal_gene_id, local_gene_id, lod_diff)
+plotly::plot_ly(as.data.frame(hld), z = as.data.frame(hld)$lod_diff, type = "heatmap")
+```
+
+## Compare dendograms
+
+Should we create a lod difference dendogram based on the mediations with a single local trait? Or with the full matrix of mediation analyses?
+
+```{r load_dendextend}
+suppressPackageStartupMessages(library(dendextend))
+```
+
+```{r define_hc_objects}
+hc_lod_diff <- foo %>%
+  select(nonlocal_gene_id, local_gene_id, lod_diff) %>%
+  hotspots::tibble_to_matrix() %>%
+  dist() %>%
+  hclust() 
+hc_pleio <- foo %>%
+  select(nonlocal_gene_id, local_gene_id, pleiotropy_lod) %>%
+  hotspots::tibble_to_matrix() %>%
+  dist() %>%
+  hclust() 
+```
+
+```{r load_dendextend}
+library(dendextend)
+```
+
+```{r dend_dist}
+dend_diff(as.dendrogram(hc_pleio), as.dendrogram(hc_lod_diff))
+```
+
+## 2 x 2 tables
+
+We binarize the continuum of mediation LOD difference values and that of pleiotropy test statistics.
+
+```{r 2x2_table}
+mediation_threshold <- 1.5
+mediation_threshold_proportion <- 0.5
+pleiotropy_threshold <- 4
+table(med = foo$lod_diff > mediation_threshold, pleio = foo$pleiotropy_lod <= pleiotropy_threshold)
+```
+
+```{r 2x2_table_proportion}
+table(med = foo$lod_diff_proportion > mediation_threshold_proportion, pleio = foo$pleiotropy_lod <= pleiotropy_threshold)
+```
+
+```{r 2x2_table_proportion}
+table(med = foo$lod_diff_proportion > 0.4, pleio = foo$pleiotropy_lod <= pleiotropy_threshold)
+```
+
+```{r counts_lod_diff}
+foo %>%
+  filter(lod_diff > mediation_threshold, pleiotropy_lod <= pleiotropy_threshold) %>%
+  group_by(local_symbol) %>%
+  count() %>%
+  arrange(desc(n))
+
+```
+
+```{r counts_lod_diff_prop_first}
+foo %>%
+  filter(lod_diff_proportion > mediation_threshold_proportion, pleiotropy_lod <= pleiotropy_threshold) %>%
+  group_by(local_symbol) %>%
+  count() %>%
+  arrange(desc(n))
+
+```
+
+```{r counts_both}
+foo %>%
+  filter(lod_diff > mediation_threshold, lod_diff_proportion > mediation_threshold_proportion, pleiotropy_lod <= pleiotropy_threshold) %>%
+  group_by(local_symbol) %>%
+  count() %>%
+  arrange(desc(n))
+```
+
+```{r counts_either}
+foo %>%
+  filter(lod_diff_proportion > mediation_threshold_proportion | lod_diff > mediation_threshold, pleiotropy_lod <= pleiotropy_threshold) %>%
+  group_by(local_symbol) %>%
+  count() %>%
+  arrange(desc(n))
+
+```
+
+```{r counts_lod_diff_prop}
+foo %>%
+  filter(lod_diff_proportion > 0.2, pleiotropy_lod <= pleiotropy_threshold) %>%
+  group_by(local_symbol) %>%
+  count() %>%
+  arrange(desc(n))
+
+```
+
+```{r counts_more}
+foo %>%
+  filter(lod_diff_proportion > 0.2, pleiotropy_lod <= pleiotropy_threshold) %>%
+  group_by(nonlocal_symbol) %>%
+  count() %>%
+  arrange(desc(n))
+foo %>% 
+  filter(lod_diff_proportion > 0.2, pleiotropy_lod <= pleiotropy_threshold) %>%
+  select(local_symbol, nonlocal_symbol, pleiotropy_lod, lod_diff_proportion, lod_diff)
+  arrange(desc(nonlocal_symbol)) 
+```

### 2019-06-24 analysis/paper/paper.Rnw
minor fixes to paper text

@@
 
-For each hotspot, we performed a series of mediation analyses and pleiotropy
-tests. For each hotspot, we identified a set of local expression traits and a
+For each hotspot, we identified a set of local expression traits and a
 set of nonlocal expression traits. All traits demonstrated LOD peaks that met
-the genome-wide significance threshold. The nonlocal expression traits arose from genes
+the genome-wide significance threshold, 7.18 \citep{keller2018genetic}.
+The nonlocal expression traits arose from genes
 on other chromosomes, while the local traits reflected expression levels for
-genes on the same chromosome as the hotspot and within 7 Mb of the hotspot
-center.
+genes on the same chromosome as the hotspot and within 4 Mb of the hotspot
+center. We ignored transcripts for genes on the hotspot chromosome but not within 4 Mb of
+the hotspot center.
 
@@
 two-dimensional, two-QTL scan over a genomic region that included the entirety of the
-hotspot. We fitted bivariate linear mixed effects models at each ordered pair of markers \citep{boehm2019testing}.
+hotspot. We fitted bivariate linear mixed effects models at each ordered pair of markers in a two-dimensional grid \citep{boehm2019testing}.
 
@@
 
-From the two-dimensional scans, we calculated pleiotropy test statistics for each pair of traits \citep{jiang1995multiple,boehm2019testing}.
+From the two-dimensional scans, we calculated pleiotropy likelihood ratio test statistics for each
+pair of traits \citep{jiang1995multiple,boehm2019testing, boehm2019thesis}. Because of the required
+computing time for bootstrap analyses, we didn't obtain p-values for the pleiotropy tests. Instead,
+we worked directly with the test statistic values.
+
+\todo[inline]{copy equations from G3 paper}
 
 \begin{equation}
-\Lambda = \max_{\lambda_1, \lambda_2}ll(\lambda_1, \lambda_2) - \max_{\lambda} ll(\lambda, \lambda)
+LOD = \log_{10} \left(\frac{\max_{\lambda_1, \lambda_2}L(B, \Sigma, \lambda_1, \lambda_2)}{\max_{\lambda}L(B, \Sigma, \lambda, \lambda)} \right)
 \label{eq:lrt}
+\end{equation}
+
+\begin{equation}
+ll(\lambda_1, \lambda_2)
 \end{equation}
@@
 For each hotspot and each local trait - nonlocal trait pair, we also performed a
-mediation analysis. Each mediation analysis involved fitting four linear models (Equations~\ref{model1}, \ref{model2}, \ref{model3}, \ref{model4}).
+mediation analysis to assess the extent which the local trait mediates the QTL effect on the nonlocal trait.
+Each mediation analysis involved fitting four linear models (Equations~\ref{model1}, \ref{model2}, \ref{model3}, \ref{model4}) \citep{chick2016defining}.
 
@@
 
-In all mediation analyses, the nonlocal transcripts are the targets, while the local transcripts serve as candidate mediators of the QTL-target relationship.
+In all mediation analyses, the nonlocal transcripts are the targets, while the local
+transcripts serve as candidate mediators of the QTL-target relationship.
+In Equations~\ref{model1} through \ref{model4}, $X$ is a matrix of founder allele dosages,
+$B$ is a matrix
+of founder allele effects, $W$ is a matrix of covariates, $C$ is a matrix of covariate
+effects, $M$ is a vector of values for a single candidate mediator, $\beta$ is the effect
+of the mediator on the complex trait, 1 is a vector with all entries set to 1, $\alpha$
+is the mean trait value, and $E$ is a vector of normally distributed random errors. \todo[inline]{better so say 'column matrix' or 'vector'? I hate to worry about this.}
 

### 2019-06-24 analysis/paper/paper.Rnw
started writing a discussion

@@
 
+We generated scatter plots of pleiotropy test statistics vs.
+mediation LOD difference proportion values.
+Each point represents a pairing of a local trait with a nonlocal trait.
+In all five scatter plots, we observe few points in the upper right quadrant of the figure.
+That is, few pairs have high pleiotropy test statistics and high mediation LOD difference values.
+
 \section{Discussion}
+
+Pleiotropy tests complement mediation analyses in the following manner. When considering a
+collection of potential mediators for a nonlocal trait, mediation analyses clearly identify the
+mediator when there is, in fact, a mediator among the candidates. However, when there is no
+mediator, mediation analyses provide little information about the genetics of the complex trait.
+Pleiotropy testing is particularly useful in this setting, since a pleiotropy test always gives
+an answer about the number of underlying QTL.
+A drawback of pleiotropy testing is that the scientific question that it addresses - do two traits share a single QTL - informs, but doesn't fully resolve, the genetics of the nonlocal complex trait under study.
 

### 2019-06-27 R/id_pleio_sets.R
added @export to functions

@@ -3,6 +3,7 @@
 #' @param tib a tibble with three columns; the first two are gene ids and the last is pleiotropy test statistics
 #' @param binarize logical of length one indicating whether to binarize the matrix of pleiotropy test statistics based on pleio_threshold value
 #' @param pleio_threshold a positive number to distinguish pleiotropy from separate QTL
+#' @export
 
 cluster_pleio_stats <- function(tib, binarize = FALSE, pleio_threshold = NULL){
   # convert tibble to matrix
@@ -18,6 +19,7 @@ cluster_pleio_stats <- function(tib, binarize = FALSE, pleio_threshold = NULL){
 #' Convert tibble to matrix
 #'
 #' @param tib
+#' @export
 
 tibble_to_matrix <- function(tib){
   rn <- unique(tib[[1]])
@@ -33,10 +35,11 @@ tibble_to_matrix <- function(tib){
   return(out)
 }
 
-#' Sort a binary matrix into something close to a block diagonal
+#' Determine adjacency matrix from a n local by m nonlocal traits binary matrix
 #'
-#' @param bmat a binary matrix
+#' @param mat a binary matrix with rownames and colnames
+#' @export
 
-sort_binary_matrix <- function(bmat){
-  bmat[ , order(colSums(bmat), decreasing = TRUE)]
+make_adjacency_matrix <- function(mat){
+  mat %*% t(mat)
 }

### 2019-06-27 analysis/paper/Chr11-prelim.Rmd
added 2 by 2 tables

@@
 ```
+
+## 2 x 2 tables
+
+```{r define_foo}
+foo <- mylist %>%
+  bind_rows() %>%
+  rename(pleiotropy_lod = lod) %>%
+  left_join(tib, by = c("local_gene_id", "nonlocal_gene_id")) %>%
+  select(local_symbol, nonlocal_symbol, pleiotropy_lod, lod_diff, lod_diff_proportion, lod_without_med, lod_with_med, filename, nonlocal_gene_id, local_gene_id) 
+```
+
+We binarize the continuum of mediation LOD difference values and that of pleiotropy test statistics.
+
+```{r 2x2_table}
+mediation_threshold <- 1.5
+mediation_threshold_proportion <- 0.5
+pleiotropy_threshold <- 4
+table(med = foo$lod_diff > mediation_threshold, pleio = foo$pleiotropy_lod <= pleiotropy_threshold)
+```
+
+```{r make_mylist}
+med_vals <- 2 * (1:5)
+pleio_vals <- 2 * (2:5)
+mylist <- list()
+for(i in 1:length(med_vals)){
+  my_other_list <- list()
+  for (j in 1:length(pleio_vals)){
+    uniq_cts_nonlocal <- foo %>% 
+      mutate(med = lod_diff > med_vals[i], pleio = pleiotropy_lod <= pleio_vals[j]) %>%
+      group_by(med, pleio) %>%
+      select(nonlocal_gene_id, med, pleio) %>%
+      unique() %>%
+      count() %>%
+      rename(n_unique_nonlocal_gene_id = n)
+    uniq_cts <- foo %>% 
+      mutate(med = lod_diff > med_vals[i], pleio = pleiotropy_lod <= pleio_vals[j]) %>%
+      group_by(med, pleio) %>%
+      select(local_gene_id, med, pleio) %>%
+      unique() %>%
+      count() %>%
+      rename(n_unique_local_gene_id = n) %>%
+      left_join(uniq_cts_nonlocal, by = c("pleio", "med"))
+
+    my_other_list[[j]] <- foo %>%
+      mutate(med = lod_diff > med_vals[i], pleio = pleiotropy_lod <= pleio_vals[j]) %>%
+      group_by(med, pleio) %>%
+      count() %>%
+      left_join(uniq_cts, by = c("med", "pleio"))
+  }
+  mylist[[i]] <- my_other_list
+} 
+mylist
+```
+
+```{r}
+mylist %>%
+  map_df(bind_rows)
+```
+
+```{r}
+mylist %>%
+  map_df(bind_rows) %>%
+  filter(as.logical(med), !as.logical(pleio))
+```
+
+```{r}
+mylist %>%
+  map_df(bind_rows) %>%
+  filter(as.logical(med), as.logical(pleio))
+```
+
+```{r 2x2_table_proportion}
+table(med = foo$lod_diff_proportion > mediation_threshold_proportion, pleio = foo$pleiotropy_lod <= pleiotropy_threshold)
+```
+
+```{r counts_lod_diff}
+foo %>%
+  filter(lod_diff > mediation_threshold, pleiotropy_lod <= pleiotropy_threshold) %>%
+  group_by(local_symbol) %>%
+  count() %>%
+  arrange(desc(n))
+
+```
+
+```{r counts_lod_diff_prop_first}
+foo %>%
+  filter(lod_diff_proportion > mediation_threshold_proportion, pleiotropy_lod <= pleiotropy_threshold) %>%
+  group_by(local_symbol) %>%
+  count() %>%
+  arrange(desc(n))
+
+```
+
+```{r counts_both}
+foo %>%
+  filter(lod_diff > mediation_threshold, lod_diff_proportion > mediation_threshold_proportion, pleiotropy_lod <= pleiotropy_threshold) %>%
+  group_by(local_symbol) %>%
+  count() %>%
+  arrange(desc(n))
+```
+
+```{r counts_either}
+foo %>%
+  filter(lod_diff_proportion > mediation_threshold_proportion | lod_diff > mediation_threshold, pleiotropy_lod <= pleiotropy_threshold) %>%
+  group_by(local_symbol) %>%
+  count() %>%
+  arrange(desc(n))
+
+```
+
+```{r counts_lod_diff_prop}
+foo %>%
+  filter(lod_diff_proportion > 0.2, pleiotropy_lod <= pleiotropy_threshold) %>%
+  group_by(local_symbol) %>%
+  count() %>%
+  arrange(desc(n))
+
+```

### 2019-06-27 analysis/paper/Chr2-prelim.Rmd
added 2 by 2 tables

@@
 library(tidyverse)
+options(tibble.print_max = 50, tibble.print_min = 20)
 ```
@@
 
-```{r load_dendextend}
-suppressPackageStartupMessages(library(dendextend))
-```
-
 ```{r define_hc_objects}
@@
 
-```{r 2x2_table_proportion}
-table(med = foo$lod_diff_proportion > mediation_threshold_proportion, pleio = foo$pleiotropy_lod <= pleiotropy_threshold)
+```{r make_mylist}
+med_vals <- 2 * (1:5)
+pleio_vals <- 2 * (2:5)
+mylist <- list()
+for(i in 1:length(med_vals)){
+  my_other_list <- list()
+  for (j in 1:length(pleio_vals)){
+    uniq_cts_nonlocal <- foo %>% 
+      mutate(med = lod_diff > med_vals[i], pleio = pleiotropy_lod <= pleio_vals[j]) %>%
+      group_by(med, pleio) %>%
+      select(nonlocal_gene_id, med, pleio) %>%
+      unique() %>%
+      count() %>%
+      rename(n_unique_nonlocal_gene_id = n)
+    uniq_cts <- foo %>% 
+      mutate(med = lod_diff > med_vals[i], pleio = pleiotropy_lod <= pleio_vals[j]) %>%
+      group_by(med, pleio) %>%
+      select(local_gene_id, med, pleio) %>%
+      unique() %>%
+      count() %>%
+      rename(n_unique_local_gene_id = n) %>%
+      left_join(uniq_cts_nonlocal, by = c("pleio", "med"))
+
+    my_other_list[[j]] <- foo %>%
+      mutate(med = lod_diff > med_vals[i], pleio = pleiotropy_lod <= pleio_vals[j]) %>%
+      group_by(med, pleio) %>%
+      count() %>%
+      left_join(uniq_cts, by = c("med", "pleio"))
+  }
+  mylist[[i]] <- my_other_list
+} 
+mylist
+```
+
+```{r}
+mylist %>%
+  map_df(bind_rows)
+```
+
+```{r}
+mylist %>%
+  map_df(bind_rows) %>%
+  filter(as.logical(med), !as.logical(pleio))
+```
+
+```{r}
+mylist %>%
+  map_df(bind_rows) %>%
+  filter(as.logical(med), as.logical(pleio))
 ```
@@
 ```{r 2x2_table_proportion}
-table(med = foo$lod_diff_proportion > 0.4, pleio = foo$pleiotropy_lod <= pleiotropy_threshold)
+table(med = foo$lod_diff_proportion > mediation_threshold_proportion, pleio = foo$pleiotropy_lod <= pleiotropy_threshold)
 ```
@@
 
-```{r counts_more}
-foo %>%
-  filter(lod_diff_proportion > 0.2, pleiotropy_lod <= pleiotropy_threshold) %>%
-  group_by(nonlocal_symbol) %>%
-  count() %>%
-  arrange(desc(n))
-foo %>% 
-  filter(lod_diff_proportion > 0.2, pleiotropy_lod <= pleiotropy_threshold) %>%
-  select(local_symbol, nonlocal_symbol, pleiotropy_lod, lod_diff_proportion, lod_diff)
-  arrange(desc(nonlocal_symbol)) 
+## More figures
+
+Let's plot a matrix of pleiotropy test statistics for nonlocal - local pairs, ordered by local gene's chromosomal position (ie, coordinate of the middle of the gene per ENSEMBL).
+We could, for example, color cells by pleiotropy test statistic value.
+Maybe truncate for those over 10.
+
+## Still more figures: graph theory package
+
+Can we represent, say, the collection of nonlocal traits in an undirected graph theory graph?
+We might say that nodes i and j are connected by an edge if they are both pleiotropic in pairwise tests with a single local trait.
+
+```{r load_hotspots}
+library(hotspots)
 ```
+
+```{r}
+mymat <- foo %>%
+  tibble_to_matrix() %>%
+  t() %>%
+  (function(mat) mat <= 4) %>%
+  make_adjacency_matrix()
+myg <- mymat %>%
+  igraph::graph_from_adjacency_matrix(mode = "undirected", weighted = TRUE, diag = FALSE) 
+comm2 <- igraph::fastgreedy.community(myg)
+```
+
+### Find communities
+
+```{r}
+library(igraph)
+```
+
+```{r}
+comm <- foo %>%
+  tibble_to_matrix() %>%
+  t() %>%
+  make_adjacency_matrix() %>%
+  igraph::graph_from_adjacency_matrix(mode = "undirected", weighted = TRUE, diag = FALSE) %>%
+  fastgreedy.community()
+
+```

### 2019-06-27 analysis/paper/paper.Rnw
minor spacing fixes

@@
 
-We generated scatter plots of pleiotropy test statistics vs.
-mediation LOD difference proportion values.
-Each point represents a pairing of a local trait with a nonlocal trait.
-In all five scatter plots, we observe few points in the upper right quadrant of the figure.
-That is, few pairs have high pleiotropy test statistics and high mediation LOD difference values.
+We generated scatter plots of pleiotropy test statistics vs. mediation LOD difference proportion values. Each
+point represents a pairing of a local trait with a nonlocal trait. In all five scatter plots, we observe few
+points in the upper right quadrant of the figure. That is, few pairs have high pleiotropy test statistics and
+high mediation LOD difference values.
 
@@
 an answer about the number of underlying QTL.
-A drawback of pleiotropy testing is that the scientific question that it addresses - do two traits share a single QTL - informs, but doesn't fully resolve, the genetics of the nonlocal complex trait under study.
+A drawback of pleiotropy testing is that the scientific question that it addresses - do two traits share a single QTL - informs, but doesn't fully resolve, the genetics of the nonlocal complex traits under study.
 

### 2019-06-27 analysis/paper/Chr11-prelim.Rmd
added threshold values to tibbles

@@
       count() %>%
-      left_join(uniq_cts, by = c("med", "pleio"))
+      left_join(uniq_cts, by = c("med", "pleio")) %>%
+      mutate(med_thr = med_vals[i], pleio_thr = pleio_vals[j])
   }
@@
 ```
-
 ```{r}

### 2019-06-28 analysis/paper/paper.Rnw
updated paper methods section

@@
 
-For each hotspot, we paired each local trait with each nonlocal trait and performed a
+For each hotspot, we tested all pairs of traits, regardless of whether the traits were local or nonlocal. We performed a
 two-dimensional, two-QTL scan over a genomic region that included the entirety of the
@@
 
+\begin{equation}
+G \sim MN_{n x 2}(0, K, V_g)
+\end{equation}
+
+\begin{equation}
+E \sim MN_{n x 2}(0, I_n, V_e)
+\end{equation}
+
+and $G$ and $E$ are independent.
+
+\begin{equation}
+\hat B = (X^T\hat \Sigma^{-1}X)^{-1}X^T\hat\Sigma^{-1}Y
+\label{eq:Bhat}
+\end{equation}
+
+\begin{equation}
+\hat \Sigma = \hat V_g \otimes \hat K + \hat V_e \otimes I_n
+\label{eq:Sigmahat}
+\end{equation}
+
+\begin{equation}
+l_{10}(\hat B, \hat \Sigma) = - \frac{1}{2}\log_{10}\left((Y - X\hat B)^T\hat \Sigma^{-1}(Y - X\hat B)\right) - n\log_{10}(2\pi) - \log_{10}|\hat \Sigma|
+\label{eq:loglik}
+\end{equation}
+
+For $n$ mice, $\Sigma$ is a $2n$ by $2n$ covariance matrix, while $V_g$ and $V_e$ are $2$ by $2$ covariance matrices.
+Estimates $\hat V_g$ and $\hat V_e$ are obtained via an expectation-maximization algorithm \citep{zhou2014efficient,gemma2}
+
+\subsubsection{Calculating likelihood ratio test statistics}
 From the two-dimensional scans, we calculated pleiotropy likelihood ratio test statistics for each

### 2019-07-01 R/id_pleio_sets.R
fixed tibble_to_matrix and added tests for it

@@ -18,18 +18,35 @@ cluster_pleio_stats <- function(tib, binarize = FALSE, pleio_threshold = NULL){
 
 #' Convert tibble to matrix
 #'
-#' @param tib
+#' @param tib a tibble with pleiotropy test results
+#' @param symmetric logical to indicate if the resulting matrix should be symmetric
 #' @export
 
-tibble_to_matrix <- function(tib){
-  rn <- unique(tib[[1]])
-  cn <- unique(tib[[2]])
-  out <- matrix(NA, nrow = length(rn), ncol = length(cn))
-  rownames(out) <- rn
-  colnames(out) <- cn
-  for (i in 1:length(rn)){
-    for (j in 1:length(cn)){
-      out[i, j] <- tib[[3]][(tib[[1]] == rownames(out)[i] & tib[[2]] == colnames(out)[j])]
+tibble_to_matrix <- function(tib, symmetric = FALSE){
+  if (!symmetric){
+    rn <- unique(tib[[1]])
+    cn <- unique(tib[[2]])
+    out <- matrix(NA, nrow = length(rn), ncol = length(cn))
+    rownames(out) <- rn
+    colnames(out) <- cn
+    for (i in 1:length(rn)){
+      for (j in 1:length(cn)){
+        out[i, j] <- tib[[3]][(tib[[1]] == rownames(out)[i] & tib[[2]] == colnames(out)[j])]
+      }
+    }
+  }
+  if (symmetric) {
+    rn <- unique(union(tib[[1]], tib[[2]]))
+    out <- matrix(NA, nrow = length(rn), ncol = length(rn))
+    rownames(out) <- rn
+    colnames(out) <- rn
+    for (i in 2:length(rn)){
+      for (j in 1:(i - 1)){
+        ind1 <- tib[[1]] == rownames(out)[i] & tib[[2]] == colnames(out)[j]
+        ind2 <- tib[[2]] == rownames(out)[i] & tib[[1]] == colnames(out)[j]
+        if (sum(ind1) | sum(ind2)) out[i, j] <- tib[[3]][ind1 | ind2]
+        out[j, i] <- out[i, j]
+      }
     }
   }
   return(out)

### 2019-07-01 tests/testthat/test-tibble_to_matrix.R

@@ -3,7 +3,22 @@ library(testthat)
 
 foo <- tibble::tibble(v1 = letters[c(1, 1, 2, 2)], v2 = letters[c(3, 4, 3, 4)], lrt = runif(4))
 
-test_that("tibble converts to matrix with correct dimensions", {
-  expect_equal(nrow(tibble_to_matrix(foo)), 2)
-  expect_equal(ncol(tibble_to_matrix(foo)), 2)
+test_that("tibble converts to rectangular matrix with correct dimensions", {
+  expect_equal(nrow(tibble_to_matrix(foo, symmetric = FALSE)), 2)
+  expect_equal(ncol(tibble_to_matrix(foo, symmetric = FALSE)), 2)
+})
+
+bar <- foo
+bar$lrt[3] <- NA
+
+
+test_that("tibble converts to symmetric matrix with correct dimensions", {
+  expect_equal(nrow(tibble_to_matrix(bar, symmetric = TRUE)), 4)
+  expect_equal(ncol(tibble_to_matrix(foo, symmetric = TRUE)), 4)
+})
+
+test_that("tibble converts to symmetric matrix with correct dimensions", {
+  expect_equal(sum(is.na(tibble_to_matrix(bar, symmetric = TRUE))), 10)
+  expect_equal(sum(is.na(tibble_to_matrix(foo, symmetric = TRUE))), 8)
+  expect_true(isSymmetric(tibble_to_matrix(foo, symmetric = TRUE)))
 })

### 2019-07-01 analysis/paper/Chr2-prelim-fewer-figs.Rmd
started working with "...-fewer-figs.Rmd" file to speed up interactive work.

--- 2019-06-27 analysis/paper/Chr2-prelim.Rmd
@@
 
-## Reading pleiotropy test results
+## Reading pleiotropy test results (for local-nonlocal pairs)
 
@@
 
-```{r, plotly_by_local, fig.height=ceiling(nlocal / 2)}
+```{r, plotly_by_local, fig.height=ceiling(nlocal / 2), eval = FALSE}
 p <- mylist %>%
@@
 
-```{r, plotly_by_nonlocal, fig.height=ceiling(nnonlocal / 4)}
+```{r, plotly_by_nonlocal, fig.height=ceiling(nnonlocal / 4), eval = FALSE}
 p <- mylist %>%
@@
 
+## Reading results for additional pleiotropy tests
+
+Above we considered only those pleiotropy tests that paired a nonlocal trait with
+a local trait. We also have results for the pairwise pleiotropy tests that involve pairs of 
+local traits and pairs of nonlocal traits.
+
+```{r read_pleio_results_local}
+results_dir <- paste0("~/Box Sync/attie/keller-hotspots/results/pvl-run-", run_num * 10 + 1)
+dir(results_dir) -> fns
+mylist_local <- list()
+for (i in 1:length(fns)){
+  mylist_local[[i]] <- read.table(file.path(results_dir, fns[i])) %>%
+  as_tibble() %>%
+  mutate(filename = fns[i]) %>%
+  mutate(gene1_id = stringr::str_split(fns[i], "_")[[1]][[3]], gene2_id = stringr::str_split(stringr::str_split(fns[i], "_")[[1]][[4]], ".txt")[[1]][[1]]) %>%
+  filter(lod == max(lod), trace == "profile1") %>%
+  select(gene1_id, gene2_id, lod, filename)
+}
+```
+
+```{r read_pleio_results_nonlocal}
+results_dir <- paste0("~/Box Sync/attie/keller-hotspots/results/pvl-run-", run_num * 10)
+dir(results_dir) -> fns
+mylist_nonlocal <- list()
+for (i in 1:length(fns)){
+  mylist_nonlocal[[i]] <- read.table(file.path(results_dir, fns[i])) %>%
+  as_tibble() %>%
+  mutate(filename = fns[i]) %>%
+  mutate(gene1_id = stringr::str_split(fns[i], "_")[[1]][[3]], 
+         gene2_id = stringr::str_split(stringr::str_split(fns[i], "_")[[1]][[4]], ".txt")[[1]][[1]]) %>%
+  filter(lod == max(lod), trace == "profile1") %>%
+  select(gene1_id, gene2_id, lod, filename)
+}
+```
+
+### Convert pleiotropy results lists to tibbles
+
+```{r}
+local_annot3 <- local_annot2 %>%
+  select(annot.id, symbol)
+local_pleio <- mylist_local %>%
+  bind_rows() %>%
+  rename(pleiotropy_lod = lod) %>%
+  left_join(local_annot3, by = c("gene1_id" = "annot.id")) %>%
+  rename(gene1_symbol = symbol) %>%
+  left_join(local_annot3, by = c("gene2_id" = "annot.id")) %>%
+  rename(gene2_symbol = symbol)
+
+nonlocal_annot3 <- nonlocal_annot %>%
+  select(annot.id, symbol)
+nonlocal_pleio <- mylist_nonlocal %>%
+  bind_rows() %>%
+  rename(pleiotropy_lod = lod) %>%
+  left_join(nonlocal_annot3, by = c("gene1_id" = "annot.id")) %>%
+  rename(gene1_symbol = symbol) %>%
+  left_join(nonlocal_annot3, by = c("gene2_id" = "annot.id")) %>%
+  rename(gene2_symbol = symbol)
+
+```
+
+I now bind all three tibbles that contain pleiotropy results into a single tibble.
+
+```{r}
+pleio_results_mat <- foo %>%
+  select(local_gene_id, nonlocal_gene_id, pleiotropy_lod, local_symbol, nonlocal_symbol) %>%
+  rename(gene1_id = local_gene_id, 
+         gene2_id = nonlocal_gene_id, 
+         gene1_symbol = local_symbol,
+         gene2_symbol = nonlocal_symbol
+         ) %>%
+  bind_rows(local_pleio, nonlocal_pleio) %>%
+  hotspots::tibble_to_matrix(symmetric = TRUE)
+```
+
+```{r}
+pleio_results_mat %>%
+  is.na() %>%
+  sum()
+```
+
+```{r}
+pleio_results_mat %>%
+  dist() %>%
+  hclust()
+```
+
+```{r}
+foobar <- pleio_results_mat %>%
+  igraph::graph_from_adjacency_matrix(mode = "undirected", weighted = TRUE, diag = FALSE) %>%
+  fastgreedy.community()
+foobar$membership %>% table()
+t1 <- tibble(gene = foobar$names, 
+       community_membership = foobar$membership, 
+       local = foobar$names %in% local_annot3$annot.id
+       )
+  table(t1$community_membership, t1$local)
+```
+
+## Clustering
+
+```{r, hclust1}
+pleio_results_mat %>%
+  dist() %>%
+  hclust() %>% 
+  plot(cex = 0.1)
+```
+
+Check out https://genomicsclass.github.io/book/pages/clustering_and_heatmaps.html for helpful code on clustering gene expressions.
+
+```{r rafalib_plot}
+pleio_results_mat %>%
+  dist() %>%
+  hclust() %>% 
+  rafalib::myplclust(
+                     lab.col = 1 + (rownames(pleio_results_mat) %in% local_annot3$annot.id), cex = 0.1
+                     )
+```
+
+```{r}
+hclusters <- pleio_results_mat %>%
+  dist() %>%
+  hclust() %>% 
+  cutree(h = 30)
+table(true=rownames(pleio_results_mat) %in% local_annot3$annot.id, cluster=hclusters)
+```
+
+## Clustering of the rectangular nlocal by nnonlocal matrix
+
+First, we make the matrix from the tibble `foo`.
+
+```{r}
+foomat <- foo %>%
+  tibble_to_matrix()
+```
+
+```{r}
+foomat_hc <- foomat %>%
+  dist() %>%
+  hclust
+```
+
+```{r}
+plot(foomat_hc, cex = 0.7)
+rafalib::myplclust(foomat_hc, cex = 0.7)
+```
+
+```{r}
+groups <- foomat_hc %>%
+  cutree(k = 6) %>%
+  as_tibble(rownames = "gene_symbol") %>%
+  rename(group_number = value) %>%
+  arrange(desc(group_number))
+
+```
+
+## Heatmaps
+
+```{r, load_colors}
+library(RColorBrewer) 
+library(genefilter)
+library(gplots)
+```
+
+```{r define_color}
+hmcol <- colorRampPalette(brewer.pal(9, "GnBu"))(100)
+heatmap.2(foomat, 
+          #labCol=tissue,
+          trace="none", 
+          #ColSideColors=cols, 
+          col=hmcol,
+          cexCol = 0.2, cexRow = 0.4
+      )
+
+```
+
+## Mediation statistic clustering
+
+```{r, loddiff_clust}
+foo %>%
+  select(1, 2, 4) %>% # choose lod_diff as the stat of interest
+  tibble_to_matrix() %>%
+  heatmap.2( 
+          #labCol=tissue,
+          trace="none", 
+          #ColSideColors=cols, 
+          col=hmcol,
+          cexCol = 0.2, 
+          cexRow = 0.4
+      )
+
+```
+
+```{r, loddiffprop_clust}
+foo %>%
+  select(1, 2, 5) %>% # choose lod_diff_proportion as stat of interest
+  tibble_to_matrix() %>%
+  heatmap.2( 
+          #labCol=tissue,
+          trace="none", 
+          #ColSideColors=cols, 
+          col=hmcol,
+          cexCol = 0.2, cexRow = 0.4
+      )
+
+```
+
+```{r load_d3heat}
+library(d3heatmap)
+```
+
+### Binarizing mediation matrix before clustering
+
+```{r threshold_clust}
+med_thresh <- 2
+foomat <- foo %>%
+  select(1, 2, 4) %>% # choose lod_diff_proportion as stat of interest
+  tibble_to_matrix()
+  apply(FUN = function(x) x <= med_thresh, X = foomat, MARGIN = 2) %>%
+    d3heatmap()
+```
+
+```{r}
+foo %>%
+  select(1, 2, 4) %>% # choose lod_diff_proportion as stat of interest
+  tibble_to_matrix() %>%
+  d3heatmap(cexRow = 0.2, cexCol = 0.2)
+```
+
+## IDEAS
+
+Maybe we think about clustering of subsets. So, subset on, say, pleiotropy values then cluster the resulting subsets separately??
+
+Also, think about writing a procedure for identifying "pleiotropy groups".
+
+Also, we might first remove those nonlocal traits that are clearly mediated by, eg, Hnf4a. Can we then use pleiotropy tests to inform what remains?
+
+## kmeans
+
+```{r kmeans}
+set.seed(01-28-1978)
+# replace NA with zeros
+pleio_resd0 <- apply(FUN = function(x){x[is.na(x)] <- 0; return(x)}, X = pleio_results_mat, MARGIN = 2)
+#km <- kmeans(t(e[1:2,]), centers=7)
+#names(km)
+
+```
+
 ## More figures

### 2019-07-08 analysis/paper/Chr11.Rmd
updated Rmd files for each hotspot

--- 2019-07-01 analysis/paper/Chr2-prelim-fewer-figs.Rmd
@@
 ---
-title: "Chr2 hotspot preliminary analyses"
+title: "Chr11 hotspot analyses"
 author: "Frederick Boehm"
-date: "6/16/2019"
+date: "July 8, 2019"
 output: html_document
@@
 
-```{r setup, include=FALSE}
+```{r setup}
 knitr::opts_chunk$set(echo = TRUE)
-hot_chr <- 2
-run_num <- 9941
-hot_mid <- 165.5
+hot_chr <- 11
+run_num <- 9921
+hot_mid <- 71
+keller_mediator <- "Sat2"
+library(hotspots)
 ```
@@
 out <- list()
-
+medfile <- paste0("_cache/Chr", hot_chr, "_mediations.rds")
+if (!file.exists(medfile)){
   for (j in 1:nlocal){
@@
 }
+saveRDS(out, medfile)
+} else {out <- readRDS(medfile)}
 ```
@@
 ```{r read_pleio_results}
+pleio_file <- paste0("_cache/Chr", hot_chr, "_pleio.rds")
+if (file.exists(pleio_file)) {mylist <- readRDS(pleio_file)} else {
   results_dir <- paste0("~/Box Sync/attie/keller-hotspots/results/pvl-run-", run_num)
@@
   }
+  saveRDS(mylist, pleio_file)
+}
 ```
@@
   select(local_symbol, nonlocal_symbol, pleiotropy_lod, lod_diff, lod_diff_proportion, lod_without_med, lod_with_med, filename, nonlocal_gene_id, local_gene_id) %>%
-  ggplot() + geom_point(mapping = aes(x = pleiotropy_lod, y = lod_diff, colour = local_symbol, nonlocal = nonlocal_symbol), size = 0.1)
+  ggplot() + geom_point(mapping = aes(x = pleiotropy_lod, y = lod_diff, colour = local_symbol, nonlocal = nonlocal_symbol), size = 0.05)
 plotly::ggplotly(p)
@@
 ~ local_symbol, nrow = ceiling(nlocal / 2), ncol = 2) + theme(legend.position = "none")
-plotly::ggplotly(p)
+#plotly::ggplotly(p)
 ```
@@
 ~ nonlocal_symbol, ncol = 4, nrow = ceiling(nnonlocal / 4)) + theme(legend.position = "none")
-plotly::ggplotly(p)
+#plotly::ggplotly(p)
 ```
@@
 
-```{r heat_pleio_stats}
+```{r heat_pleio_stats, fig.height=4, fig.width = 6}
 foo %>%
@@
 dir(results_dir) -> fns
+pleio_file_local <- paste0("_cache/Chr", hot_chr, "_pleio_local.rds")
+if (file.exists(pleio_file_local)) {mylist_local <- readRDS(pleio_file_local)} else {
   mylist_local <- list()
@@
   }
+  saveRDS(mylist_local, pleio_file_local)
+}
+
 ```
@@
 dir(results_dir) -> fns
+pleio_file_nonlocal <- paste0("_cache/Chr", hot_chr, "_pleio_nonlocal.rds")
+if (file.exists(pleio_file_nonlocal)) {mylist_nonlocal <- readRDS(pleio_file_nonlocal)} else {
   mylist_nonlocal <- list()
@@
     select(gene1_id, gene2_id, lod, filename)
+  }
+  saveRDS(mylist_nonlocal, pleio_file_nonlocal)
 }
@@
 ```{r}
+library(igraph)
+```
+
+```{r}
 foobar <- pleio_results_mat %>%
@@
   hclust() %>% 
-  rafalib::myplclust(
-                     lab.col = 1 + (rownames(pleio_results_mat) %in% local_annot3$annot.id), cex = 0.1
+  rafalib::myplclust(cex = 0.1
                      )
 ```
+
+```{r rafalib_plot2}
+bar <- foo %>%
+  filter(local_symbol == keller_mediator) # gene symbol from the keller supplementary table
+
+pleio_hc <- foo %>%
+  tibble_to_matrix() %>%
+  t() %>%
+  dist() %>%
+  hclust()
+
+rafalib::myplclust(pleio_hc,
+                     lab.col = 1+ round(bar$pleiotropy_lod[match(x = bar$nonlocal_symbol, table = pleio_hc$labels)]),
+                     cex = 0.3
+                     )
+```
+
+```{r}
+rafalib::myplclust(pleio_hc,
+                   lab.col = 1 + (bar$lod_diff[match(x = bar$nonlocal_symbol,
+                                                     table = pleio_hc$labels)] >
+                                    mediation_threshold),
+                     cex = 0.3
+                     )
+
+```
+
+```{r}
+library(RColorBrewer) 
+library(genefilter)
+library(gplots) ##Available from CRAN
+cols <- palette(RColorBrewer::brewer.pal(8, "Dark2"))
+#head(cbind(colnames(e),cols))
+```
+
+```{r heatmap.2}
+pleio_mat <- foo %>%
+  tibble_to_matrix() %>%
+  t()
+ld <- bar$lod_diff
+idx1 <- match(x = bar$nonlocal_symbol, table = rownames(pleio_mat))
+idx <- 1 + floor((ld * (ld > 0))[idx1])
+cols[idx] -> indexed_colors
+ic2 <- 1 + (ld[idx1] > mediation_threshold)
+#indexed_colors[is.na(indexed_colors)] <- cols[8]
+gplots::heatmap.2(pleio_mat, 
+                  trace="none", 
+                  #RowSideColors = indexed_colors, 
+                  RowSideColors = c("red", "green")[ic2],
+                  col=hmcol,
+                  cexRow = 0.5,
+                  cexCol = 0.5
+                  )
+```
+
+```{r}
+d3heatmap::d3heatmap(t(pleio_mat) <= 4, cexRow = 0.2, cexCol = 0.3)
+```
+
+```{r}
+gplots::heatmap.2(t(pleio_mat), 
+                  trace="none", 
+                  #RowSideColors = indexed_colors, 
+                  ColSideColors = c("red", "green")[ic2],
+                  col=hmcol,
+                  cexRow = 0.5,
+                  cexCol = 0.5
+                  )
+
+```
+
+## Cutting the dendrogram
 
@@
 
-## Clustering of the rectangular nlocal by nnonlocal matrix
+## Clustering of the rectangular nlocal by nnonlocal matrix of pleiotropy results
 
@@
 ## Heatmaps
-
-```{r, load_colors}
-library(RColorBrewer) 
-library(genefilter)
-library(gplots)
-```
 
@@
 foomat <- foo %>%
-  select(1, 2, 4) %>% # choose lod_diff_proportion as stat of interest
+  select(1, 2, 4) %>% # choose lod_diff as stat of interest
   tibble_to_matrix()
@@
 foo %>%
-  select(1, 2, 4) %>% # choose lod_diff_proportion as stat of interest
+  select(1, 2, 4) %>% # choose lod_diff as stat of interest
   tibble_to_matrix() %>%
@@
 
-## kmeans
-
-```{r kmeans}
-set.seed(01-28-1978)
-# replace NA with zeros
-pleio_resd0 <- apply(FUN = function(x){x[is.na(x)] <- 0; return(x)}, X = pleio_results_mat, MARGIN = 2)
-#km <- kmeans(t(e[1:2,]), centers=7)
-#names(km)
-
-```
+Consider, also, truncating negative values at zero for the mediation analyses.
 
@@
 ```
+
+## Biclustering with `biclust` R package
+
+```{r}
+library(biclust)
+```
+
+```{r}
+bccc_out<- biclust(pleio_mat, method = BCCC())
+c1 <- bicluster(pleio_mat, bccc_out, number = 1)
+d3heatmap(c1$Bicluster1, "none", Rowv = FALSE, cexRow = 0.3, cexCol = 0.3)
+```
+
+```{r}
+d3heatmap(c1$Bicluster1, cexRow = 0.3, cexCol = 0.3)
+
+```
+
+```{r}
+bccc_out2 <- biclust(pleio_mat > pleiotropy_threshold, method = BCCC())
+bccc_out2
+```
+
+```{r}
+biclust(pleio_mat > pleiotropy_threshold, method = BCBimax())
+```
+
+## Clustering only those nonlocal traits that aren't mediated by `r keller_mediator`
+
+Another approach is to cluster only those traits that aren't 
+mediated by `r keller_mediator`. For now, let's use the lod difference threshold 
+1.5 to define mediation.
+
+```{r}
+# subset to get nonlocals that aren't mediated by Hnf4a
+# first, get symbols for nonlocals that aren't mediated by Hnf4a
+pmat <- bar %>%
+  filter(lod_diff <= mediation_threshold) %>%
+  dplyr::select(nonlocal_symbol) %>%
+  left_join(foo) %>%
+  tibble_to_matrix()
+dim(pmat)
+```
+
+```{r d3heat}
+d3heatmap(pmat, cexRow = 0.3, cexCol = 0.3)
+```
+```{r heat_binary}
+hc1 <- hclust(dist(pmat))
+plot(hc1, cex = 0.3)
+hc2 <- hclust(dist(t(pmat)))
+plot(hc2, cex = 0.3)
+pmat %>%
+  dist() %>%
+  hclust()
+
+```
+
+```{r d3heat_logical}
+d3heatmap(pmat > pleiotropy_threshold, cexRow = 0.3, cexCol = 0.3)
+```
+
+```{r}
+dim(pmat)
+length(hc1$order)
+d3heatmap(pmat[hc1$order, hc2$order] > pleiotropy_threshold, 
+          Rowv = FALSE, 
+          Colv = FALSE, 
+          cexRow = 0.3, 
+          cexCol = 0.3)
+d3heatmap(pmat[hc1$order, hc2$order], 
+          Rowv = FALSE, 
+          Colv = FALSE, 
+          cexRow = 0.3, 
+          cexCol = 0.3)
+
+```
+
+## Local-local pleiotropy test results
+
+```{r local-local-pleio}
+local_pleio %>%
+  dplyr::select(5, 6, 3) %>%
+  tibble_to_matrix(symmetric = TRUE) %>%
+  d3heatmap()
+```
+
+## Nonlocal-nonlocal pleiotropy test results
+
+```{r nonlocal-nonlocal-pleio}
+nonlocal_pleio %>%
+  dplyr::select(5, 6, 3) %>%
+  tibble_to_matrix(symmetric = TRUE) %>%
+  d3heatmap(cexRow = 0.3, cexCol = 0.3)
+```

### 2019-07-08 analysis/paper/Chr13.Rmd
updated Rmd files for each hotspot

--- 2019-07-08 analysis/paper/Chr11.Rmd
@@
 ---
-title: "Chr11 hotspot analyses"
+title: "Chr13 hotspot analyses"
 author: "Frederick Boehm"
-date: "July 8, 2019"
+date: "7/08/2019"
 output: html_document
@@
 knitr::opts_chunk$set(echo = TRUE)
-hot_chr <- 11
-run_num <- 9921
-hot_mid <- 71
-keller_mediator <- "Sat2"
+hot_chr <- 13
+run_num <- 9931
+hot_mid <- 112.5
+keller_mediator <- "Il6st"
 library(hotspots)

### 2019-07-08 analysis/paper/Chr2.Rmd
updated Rmd files for each hotspot

--- 2019-07-08 analysis/paper/Chr11.Rmd
@@
 ---
-title: "Chr11 hotspot analyses"
+title: "Chr2 hotspot analyses"
 author: "Frederick Boehm"
-date: "July 8, 2019"
+date: "6/16/2019"
 output: html_document
@@
 
-```{r setup}
+```{r setup, include=FALSE}
 knitr::opts_chunk$set(echo = TRUE)
-hot_chr <- 11
-run_num <- 9921
-hot_mid <- 71
-keller_mediator <- "Sat2"
+hot_chr <- 2
+run_num <- 9941
+hot_mid <- 165.5
 library(hotspots)
+library(igraph)
+library(d3heatmap)
 ```
@@
 out <- list()
-medfile <- paste0("_cache/Chr", hot_chr, "_mediations.rds")
+medfile <- "_cache/Chr2_mediations.rds"
 if (!file.exists(medfile)){
@@
 ```{r read_pleio_results}
-pleio_file <- paste0("_cache/Chr", hot_chr, "_pleio.rds")
+pleio_file <- "_cache/Chr2_pleio.rds"
 if (file.exists(pleio_file)) {mylist <- readRDS(pleio_file)} else {
@@
 
-```{r heat_pleio_stats, fig.height=4, fig.width = 6}
+```{r heat_pleio_stats}
 foo %>%
@@
 dir(results_dir) -> fns
-pleio_file_local <- paste0("_cache/Chr", hot_chr, "_pleio_local.rds")
+pleio_file_local <- "_cache/Chr2_pleio_local.rds"
 if (file.exists(pleio_file_local)) {mylist_local <- readRDS(pleio_file_local)} else {
@@
 dir(results_dir) -> fns
-pleio_file_nonlocal <- paste0("_cache/Chr", hot_chr, "_pleio_nonlocal.rds")
+pleio_file_nonlocal <- "_cache/Chr2_pleio_nonlocal.rds"
 if (file.exists(pleio_file_nonlocal)) {mylist_nonlocal <- readRDS(pleio_file_nonlocal)} else {
@@
 ```{r}
-library(igraph)
-```
-
-```{r}
 foobar <- pleio_results_mat %>%
@@
 bar <- foo %>%
-  filter(local_symbol == keller_mediator) # gene symbol from the keller supplementary table
+  filter(local_symbol == "Hnf4a")
 
@@
 ```{r}
-d3heatmap::d3heatmap(t(pleio_mat) <= 4, cexRow = 0.2, cexCol = 0.3)
+d3heatmap(t(pleio_mat) <= 4, cexRow = 0.2, cexCol = 0.3)
 ```
@@
 
-## Clustering only those nonlocal traits that aren't mediated by `r keller_mediator`
+## Clustering only those nonlocal traits that aren't mediated by Hnf4a
 
 Another approach is to cluster only those traits that aren't 
-mediated by `r keller_mediator`. For now, let's use the lod difference threshold 
+mediated by Hnf4a. For now, let's use the lod difference threshold 
 1.5 to define mediation.

### 2019-07-08 analysis/paper/Chr5-prelim.Rmd
updated Rmd files for each hotspot

@@
 ```{r, load_attie}
+hot_chr <- 5
 load("~/Box Sync/attie/keller2018-chr2-hotspot-chtc/data-to-ignore/Attie_DO378_eQTL_viewer_v1.Rdata")
@@
 ```{r}
-Chr5hot_local <- dataset.islet.rnaseq$expr[, colnames(dataset.islet.rnaseq$expr) %in% local_annot2$annot.id]
-Chr5hot_nonlocal <- dataset.islet.rnaseq$expr[, colnames(dataset.islet.rnaseq$expr) %in% nonlocal_annot$annot.id]
+hot_local <- dataset.islet.rnaseq$expr[, colnames(dataset.islet.rnaseq$expr) %in% local_annot2$annot.id]
+hot_nonlocal <- dataset.islet.rnaseq$expr[, colnames(dataset.islet.rnaseq$expr) %in% nonlocal_annot$annot.id]
 ```
@@
 ```{r fit1_med}
-nlocal <- ncol(Chr5hot_local)
-nnonlocal <- ncol(Chr5hot_nonlocal)
+nlocal <- ncol(hot_local)
+nnonlocal <- ncol(hot_nonlocal)
+medfile <- "_cache/Chr5_mediations.rds"
+if (!file.exists(medfile)){
   out <- list()
-out2 <- list()
   for (j in 1:nlocal){
   foo <- list()
-  foo2 <- list()
   for (i in 1:nnonlocal){
-    foo[[i]]<- qtl2mediate::fit1_med(driver = genoprobs$`5`[rownames(genoprobs$`5`) %in% rownames(Chr5hot_local) , , indices[i]], 
-                                      # genoprobs at peak position for nonlocal trait
-                                      target = Chr5hot_nonlocal[ , i], 
-                                      mediator = Chr5hot_local[ , j], 
-                                      addcovar = addcovar
-         )
-    foo2[[i]] <- qtl2mediate:::fit1_med_qtl2(driver = genoprobs$`5`[, , indices[i]],
-                                             target = Chr5hot_nonlocal[ , i],
-                                             mediator = Chr5hot_local[ , j],
+    foo[[i]] <- qtl2mediate:::fit1_med_qtl2(driver = genoprobs[[as.character(hot_chr)]][, , indices[i]],
+                                             target = hot_nonlocal[ , i],
+                                             mediator = hot_local[ , j],
                                              addcovar = addcovar
@@
   out[[j]] <- foo
-  out2[[j]] <- foo2
 }
+saveRDS(out, medfile)
+} else {out <- readRDS(medfile)}
 ```
@@
 
-## Reading pleiotropy test results
+## Reading pleiotropy test results (for local-nonlocal pairs)
 
 ```{r read_pleio_results}
-results_dir <- "~/Box Sync/attie/keller-hotspots/results/pvl-run-9901"
+run_num <- 9901
+pleio_file <- "_cache/Chr5_pleio.rds"
+
+if (file.exists(pleio_file)) {mylist <- readRDS(pleio_file)} else {
+  results_dir <- paste0("~/Box Sync/attie/keller-hotspots/results/pvl-run-", run_num)
   dir(results_dir) -> fns
@@
   }
-mylist %>%
-  bind_rows() %>%
-  rename(pleiotropy_lod = lod) %>%
-  left_join(tib, by = c("local_gene_id", "nonlocal_gene_id")) %>%
-  select(local_symbol, nonlocal_symbol, pleiotropy_lod, lod_diff, lod_diff_proportion, lod_without_med, lod_with_med, filename, nonlocal_gene_id, local_gene_id) %>%
-  group_by(nonlocal_symbol) %>%
-  filter(lod_diff == max(lod_diff)) %>%
-  arrange(local_symbol, desc(lod_diff))
+  saveRDS(mylist, pleio_file)
+}
+```
+
+```{r read_pleio_results_local}
+results_dir <- paste0("~/Box Sync/attie/keller-hotspots/results/pvl-run-", run_num * 10 + 1)
+dir(results_dir) -> fns
+pleio_file_local <- "_cache/Chr5_pleio_local.rds"
+if (file.exists(pleio_file_local)) {mylist_local <- readRDS(pleio_file_local)} else {
+  mylist_local <- list()
+  for (i in 1:length(fns)){
+    mylist_local[[i]] <- read.table(file.path(results_dir, fns[i])) %>%
+    as_tibble() %>%
+    mutate(filename = fns[i]) %>%
+    mutate(gene1_id = stringr::str_split(fns[i], "_")[[1]][[3]], gene2_id = stringr::str_split(stringr::str_split(fns[i], "_")[[1]][[4]], ".txt")[[1]][[1]]) %>%
+    filter(lod == max(lod), trace == "profile1") %>%
+    select(gene1_id, gene2_id, lod, filename)
+  }
+  saveRDS(mylist_local, pleio_file_local)
+}
+
+```
+
+```{r read_pleio_results_nonlocal}
+results_dir <- paste0("~/Box Sync/attie/keller-hotspots/results/pvl-run-", run_num * 10)
+dir(results_dir) -> fns
+pleio_file_nonlocal <- "_cache/Chr5_pleio_nonlocal.rds"
+if (file.exists(pleio_file_nonlocal)) {mylist_nonlocal <- readRDS(pleio_file_nonlocal)} else {
+  mylist_nonlocal <- list()
+  for (i in 1:length(fns)){
+    mylist_nonlocal[[i]] <- read.table(file.path(results_dir, fns[i])) %>%
+    as_tibble() %>%
+    mutate(filename = fns[i]) %>%
+    mutate(gene1_id = stringr::str_split(fns[i], "_")[[1]][[3]], 
+           gene2_id = stringr::str_split(stringr::str_split(fns[i], "_")[[1]][[4]], ".txt")[[1]][[1]]) %>%
+    filter(lod == max(lod), trace == "profile1") %>%
+    select(gene1_id, gene2_id, lod, filename)
+  }
+  saveRDS(mylist_nonlocal, pleio_file_nonlocal)
+}
 ```
@@
 plotly::ggplotly(p)
-```
-
-## Are the lods from fit1_med correct?
-
-Yes.
-
-I see that a large number of nonlocal transcripts have `lod_without_med` values below 7.18.
-While the 7.18 threshold was determined with the mixed model, I still find the large number of transcripts unusual.
-
-First, a given nonlocal transcript should have a single value for `lod_without_med`, regardless of which local transcript it's paired with. Let's verify this.
-
-```{r, check1}
-mylist %>%
-  bind_rows() %>%
-  rename(pleiotropy_lod = lod) %>%
-  left_join(tib, by = c("local_gene_id", "nonlocal_gene_id")) %>%
-  select(local_symbol, nonlocal_symbol, pleiotropy_lod, lod_diff, lod_diff_proportion, lod_without_med, lod_with_med, filename, nonlocal_gene_id, local_gene_id) %>% 
-  arrange(nonlocal_symbol)
-```
-
-```{r, check_table}
-mylist %>%
-  bind_rows() %>%
-  select(lod) %>%
-  mutate(indic = lod > 5) %>%
-  select(indic) %>%
-  table()
 ```
@@
 ```
+
+## Heatmaps of pleiotropy results
+
+### Convert pleiotropy results lists to tibbles
+
+```{r}
+local_annot3 <- local_annot2 %>%
+  select(annot.id, symbol)
+local_pleio <- mylist_local %>%
+  bind_rows() %>%
+  rename(pleiotropy_lod = lod) %>%
+  left_join(local_annot3, by = c("gene1_id" = "annot.id")) %>%
+  rename(gene1_symbol = symbol) %>%
+  left_join(local_annot3, by = c("gene2_id" = "annot.id")) %>%
+  rename(gene2_symbol = symbol)
+
+nonlocal_annot3 <- nonlocal_annot %>%
+  select(annot.id, symbol)
+nonlocal_pleio <- mylist_nonlocal %>%
+  bind_rows() %>%
+  rename(pleiotropy_lod = lod) %>%
+  left_join(nonlocal_annot3, by = c("gene1_id" = "annot.id")) %>%
+  rename(gene1_symbol = symbol) %>%
+  left_join(nonlocal_annot3, by = c("gene2_id" = "annot.id")) %>%
+  rename(gene2_symbol = symbol)
+
+```
+
+I now bind all three tibbles that contain pleiotropy results into a single tibble.
+
+```{r}
+pleio_results_mat <- foo %>%
+  dplyr::select(local_gene_id, nonlocal_gene_id, pleiotropy_lod, local_symbol, nonlocal_symbol) %>%
+  rename(gene1_id = local_gene_id, 
+         gene2_id = nonlocal_gene_id, 
+         gene1_symbol = local_symbol,
+         gene2_symbol = nonlocal_symbol
+         ) %>%
+  bind_rows(local_pleio, nonlocal_pleio) %>%
+  hotspots::tibble_to_matrix(symmetric = TRUE)
+```
+
+```{r}
+pleio_results_mat %>%
+  is.na() %>%
+  sum()
+```
+
+```{r}
+pleio_results_mat %>%
+  dist() %>%
+  hclust()
+```
+
+```{r}
+foobar <- pleio_results_mat %>%
+  igraph::graph_from_adjacency_matrix(mode = "undirected", weighted = TRUE, diag = FALSE) %>%
+  fastgreedy.community()
+foobar$membership %>% table()
+t1 <- tibble(gene = foobar$names, 
+       community_membership = foobar$membership, 
+       local = foobar$names %in% local_annot3$annot.id
+       )
+  table(t1$community_membership, t1$local)
+```
+
+## Clustering
+
+```{r, hclust1}
+pleio_results_mat %>%
+  dist() %>%
+  hclust() %>% 
+  plot(cex = 0.1)
+```
+
+Check out https://genomicsclass.github.io/book/pages/clustering_and_heatmaps.html for helpful code on clustering gene expressions.
+
+```{r rafalib_plot}
+pleio_results_mat %>%
+  dist() %>%
+  hclust() %>% 
+  rafalib::myplclust(cex = 0.1
+                     )
+```
+
+```{r rafalib_plot2}
+bar <- foo %>%
+  filter(local_symbol == "Hnf4a")
+
+pleio_hc <- foo %>%
+  tibble_to_matrix() %>%
+  t() %>%
+  dist() %>%
+  hclust()
+
+rafalib::myplclust(pleio_hc,
+                     lab.col = 1+ round(bar$pleiotropy_lod[match(x = bar$nonlocal_symbol, table = pleio_hc$labels)]),
+                     cex = 0.3
+                     )
+```
+
+```{r}
+rafalib::myplclust(pleio_hc,
+                   lab.col = 1 + (bar$lod_diff[match(x = bar$nonlocal_symbol,
+                                                     table = pleio_hc$labels)] >
+                                    mediation_threshold),
+                     cex = 0.3
+                     )
+
+```
+
+```{r}
+library(RColorBrewer) 
+library(genefilter)
+library(gplots) ##Available from CRAN
+cols <- palette(RColorBrewer::brewer.pal(8, "Dark2"))
+#head(cbind(colnames(e),cols))
+```
+
+```{r heatmap.2}
+pleio_mat <- foo %>%
+  tibble_to_matrix() %>%
+  t()
+ld <- bar$lod_diff
+idx1 <- match(x = bar$nonlocal_symbol, table = rownames(pleio_mat))
+idx <- 1 + floor((ld * (ld > 0))[idx1])
+cols[idx] -> indexed_colors
+ic2 <- 1 + (ld[idx1] > mediation_threshold)
+#indexed_colors[is.na(indexed_colors)] <- cols[8]
+gplots::heatmap.2(pleio_mat, 
+                  trace="none", 
+                  #RowSideColors = indexed_colors, 
+                  RowSideColors = c("red", "green")[ic2],
+                  col=hmcol,
+                  cexRow = 0.5,
+                  cexCol = 0.5
+                  )
+```
+
+```{r}
+d3heatmap(t(pleio_mat) <= 4, cexRow = 0.2, cexCol = 0.3)
+```
+
+```{r}
+gplots::heatmap.2(t(pleio_mat), 
+                  trace="none", 
+                  #RowSideColors = indexed_colors, 
+                  ColSideColors = c("red", "green")[ic2],
+                  col=hmcol,
+                  cexRow = 0.5,
+                  cexCol = 0.5
+                  )
+
+```

### 2019-07-08 analysis/paper/Chr5.Rmd
updated Rmd files for each hotspot

--- 2019-07-08 analysis/paper/Chr11.Rmd
@@
 ---
-title: "Chr11 hotspot analyses"
+title: "Chr5 hotspot analyses"
 author: "Frederick Boehm"
-date: "July 8, 2019"
+date: "7/08/2019"
 output: html_document
@@
 knitr::opts_chunk$set(echo = TRUE)
-hot_chr <- 11
-run_num <- 9921
-hot_mid <- 71
-keller_mediator <- "Sat2"
+hot_chr <- 5
+run_num <- 9901
+hot_mid <- 146
+keller_mediator <- "Pdx1"
 library(hotspots)
+library(igraph)
 ```

### 2019-07-08 analysis/paper/Chr7.Rmd
updated Rmd files for each hotspot

--- 2019-07-08 analysis/paper/Chr11.Rmd
@@
 ---
-title: "Chr11 hotspot analyses"
+title: "Chr7 hotspot analyses"
 author: "Frederick Boehm"
-date: "July 8, 2019"
+date: "7/08/2019"
 output: html_document
@@
 knitr::opts_chunk$set(echo = TRUE)
-hot_chr <- 11
-run_num <- 9921
-hot_mid <- 71
-keller_mediator <- "Sat2"
+hot_chr <- 7
+run_num <- 9911
+hot_mid <- 46
+keller_mediator <- "Fam83e"
 library(hotspots)
@@
 
+Another possibility is to look at the local traits' pleiotropy test results only.
+Assign each local trait to a "pleiotropy group".
+Then look at each pleiotropy group, and see how similar their LOD difference values (binarized or not) are.
+That is, does every member of a "pleiotropy group" mediate the same set of traits?
+
 ## More figures

### 2019-07-09 analysis/paper/Chr11-hclusterings.Rmd
updated Rmd files

--- 2019-07-08 analysis/paper/Chr7.Rmd
@@
 ---
-title: "Chr7 hotspot analyses"
+title: "Chr11 hotspot analyses"
 author: "Frederick Boehm"
@@
 knitr::opts_chunk$set(echo = TRUE)
-hot_chr <- 7
-run_num <- 9911
-hot_mid <- 46
-keller_mediator <- "Fam83e"
+hot_chr <- 11
+run_num <- 9921
+hot_mid <- 71
+keller_mediator <- "Sat2"
 library(hotspots)
+fp <- paste0("../figures/Chr", hot_chr, "-")
+knitr::opts_chunk$set(fig.path = fp)
+
 ```
@@
 Now, filter to get those traits that map to the hotspot. Note that we're likely to get local traits, too, in this step.
-
-```{r}
-dataset.islet.rnaseq$lod.peaks %>%
-  filter(chrom == hot_chr, pos <= hot_upper, pos >= hot_lower) %>%
-  left_join(y = dataset.islet.rnaseq$annots, by = c("annot.id" = "gene_id")) %>%
-  filter(!duplicated(annot.id)) %>%
-  filter(lod >= 7.18)
-```
 
@@
   mutate(marker_index = purrr::map_int(.x = marker.id, .f = function(x){which(names(map[[as.character(hot_chr)]]) == x)})) %>%
-  select(marker_index) %>%
+  dplyr::select(marker_index) %>%
   unlist() %>%
@@
   left_join(nonlocal_annot, by = c("nonlocal_gene_id" = "annot.id")) %>%
-  select(symbol.x, symbol.y, lod_without_med, lod_with_med, lod_diff, lod_diff_proportion, local_gene_id, nonlocal_gene_id) %>%
+  dplyr::select(symbol.x, symbol.y, lod_without_med, lod_with_med, lod_diff, lod_diff_proportion, local_gene_id, nonlocal_gene_id) %>%
   rename(local_symbol = symbol.x, nonlocal_symbol = symbol.y)
@@
     filter(lod == max(lod), trace == "profile1") %>%
-    select(local_gene_id, nonlocal_gene_id, lod, filename)
+    dplyr::select(local_gene_id, nonlocal_gene_id, lod, filename)
   }
@@
   left_join(tib, by = c("local_gene_id", "nonlocal_gene_id")) %>%
-  select(local_symbol, nonlocal_symbol, pleiotropy_lod, lod_diff, lod_diff_proportion, lod_without_med, lod_with_med, filename, nonlocal_gene_id, local_gene_id) %>%
+  dplyr::select(local_symbol, nonlocal_symbol, pleiotropy_lod, lod_diff, lod_diff_proportion, lod_without_med, lod_with_med, filename, nonlocal_gene_id, local_gene_id) %>%
   ggplot() + geom_point(mapping = aes(x = pleiotropy_lod, y = lod_diff, colour = local_symbol, nonlocal = nonlocal_symbol), size = 0.05)
@@
   left_join(tib, by = c("local_gene_id", "nonlocal_gene_id")) %>%
-  select(local_symbol, nonlocal_symbol, pleiotropy_lod, lod_diff, lod_diff_proportion, lod_without_med, lod_with_med, filename, nonlocal_gene_id, local_gene_id) %>%
+  dplyr::select(local_symbol, nonlocal_symbol, pleiotropy_lod, lod_diff, lod_diff_proportion, lod_without_med, lod_with_med, filename, nonlocal_gene_id, local_gene_id) %>%
   ggplot() + geom_point(mapping = aes(x = pleiotropy_lod, y = lod_diff, colour = local_symbol, nonlocal = nonlocal_symbol), size = 0.1) + facet_wrap(.
@@
   left_join(tib, by = c("local_gene_id", "nonlocal_gene_id")) %>%
-  select(local_symbol, nonlocal_symbol, pleiotropy_lod, lod_diff, lod_diff_proportion, lod_without_med, lod_with_med, filename, nonlocal_gene_id, local_gene_id) 
+  dplyr::select(local_symbol, nonlocal_symbol, pleiotropy_lod, lod_diff, lod_diff_proportion, lod_without_med, lod_with_med, filename, nonlocal_gene_id, local_gene_id) 
 ```
@@
   rename(local_peak_pos = pos, local_peak_lod = lod) %>%
-  select(- marker.id, - chrom, - strand, - biotype, - local_tx) %>%
+  dplyr::select(- marker.id, - chrom, - strand, - biotype, - local_tx) %>%
   filter(local_peak_pos < hot_upper, local_peak_pos > hot_lower)
@@
   rename(local_peak_pos = pos, local_peak_lod = lod) %>%
-  select(- marker.id, - chrom, - strand, - biotype, - local_tx) %>%
+  dplyr::select(- marker.id, - chrom, - strand, - biotype, - local_tx) %>%
   filter(local_peak_pos < hot_upper, local_peak_pos > hot_lower) %>%
-  select(middle) %>% unlist() %>% range()
+  dplyr::select(middle) %>% unlist() %>% range()
 
@@
 bar <- foo %>%
-  select(nonlocal_gene_id, local_gene_id, pleiotropy_lod) %>%
+  dplyr::select(nonlocal_gene_id, local_gene_id, pleiotropy_lod) %>%
 hotspots::tibble_to_matrix() %>%
@@
 
-```{r heat_pleio_stats, fig.height=4, fig.width = 6}
-foo %>%
-  select(nonlocal_gene_id, local_gene_id, pleiotropy_lod, lod_diff) %>%
-hotspots::tibble_to_matrix() %>%
-  t() %>%
-  gplots::heatmap.2(cexRow = 0.1, cexCol = 0.1, col = hmcol, trace = "none")
-```
-
-```{r heat_lod_diff}
-foo %>%
-  select(nonlocal_gene_id, local_gene_id, lod_diff) %>%
-hotspots::tibble_to_matrix() %>%
-  t() %>%
-  gplots::heatmap.2(cexRow = 0.1, cexCol = 0.1, col = hmcol, trace = "none")
-```
-
-```{r heat_lod_diff0}
-foo %>%
-  select(nonlocal_gene_id, local_gene_id, lod_diff) %>%
-  mutate(lod_diff = lod_diff * (lod_diff > 0)) %>%
-hotspots::tibble_to_matrix() %>%
-  t() %>%
-  gplots::heatmap.2(cexRow = 0.1, cexCol = 0.1, col = hmcol, trace = "none")
-```
-
 ```{r, define_hld, eval = FALSE}
 hld <- foo %>%
-  select(nonlocal_gene_id, local_gene_id, lod_diff)
+  dplyr::select(nonlocal_gene_id, local_gene_id, lod_diff)
 plotly::plot_ly(as.data.frame(hld), z = as.data.frame(hld)$lod_diff, type = "heatmap")
@@
 hc_lod_diff <- foo %>%
-  select(nonlocal_gene_id, local_gene_id, lod_diff) %>%
+  dplyr::select(nonlocal_gene_id, local_gene_id, lod_diff) %>%
   hotspots::tibble_to_matrix() %>%
@@
 hc_pleio <- foo %>%
-  select(nonlocal_gene_id, local_gene_id, pleiotropy_lod) %>%
+  dplyr::select(nonlocal_gene_id, local_gene_id, pleiotropy_lod) %>%
   hotspots::tibble_to_matrix() %>%
@@
       group_by(med, pleio) %>%
-      select(nonlocal_gene_id, med, pleio) %>%
+      dplyr::select(nonlocal_gene_id, med, pleio) %>%
       unique() %>%
@@
       group_by(med, pleio) %>%
-      select(local_gene_id, med, pleio) %>%
+      dplyr::select(local_gene_id, med, pleio) %>%
       unique() %>%
@@
     filter(lod == max(lod), trace == "profile1") %>%
-    select(gene1_id, gene2_id, lod, filename)
+    dplyr::select(gene1_id, gene2_id, lod, filename)
   }
@@
 local_annot3 <- local_annot2 %>%
-  select(annot.id, symbol)
+  dplyr::select(annot.id, symbol)
 local_pleio <- mylist_local %>%
@@
 nonlocal_annot3 <- nonlocal_annot %>%
-  select(annot.id, symbol)
+  dplyr::select(annot.id, symbol)
 nonlocal_pleio <- mylist_nonlocal %>%
@@
 pleio_results_mat <- foo %>%
-  select(local_gene_id, nonlocal_gene_id, pleiotropy_lod, local_symbol, nonlocal_symbol) %>%
+  dplyr::select(local_gene_id, nonlocal_gene_id, pleiotropy_lod, local_symbol, nonlocal_symbol) %>%
   rename(gene1_id = local_gene_id, 
@@
 #indexed_colors[is.na(indexed_colors)] <- cols[8]
-gplots::heatmap.2(pleio_mat, 
-                  trace="none", 
-                  #RowSideColors = indexed_colors, 
-                  RowSideColors = c("red", "green")[ic2],
-                  col=hmcol,
-                  cexRow = 0.5,
-                  cexCol = 0.5
-                  )
-```
-
-```{r}
-d3heatmap::d3heatmap(t(pleio_mat) <= 4, cexRow = 0.2, cexCol = 0.3)
-```
-
-```{r}
-gplots::heatmap.2(t(pleio_mat), 
-                  trace="none", 
-                  #RowSideColors = indexed_colors, 
-                  ColSideColors = c("red", "green")[ic2],
-                  col=hmcol,
-                  cexRow = 0.5,
-                  cexCol = 0.5
-                  )
-
 ```
@@
 hmcol <- colorRampPalette(brewer.pal(9, "GnBu"))(100)
-heatmap.2(foomat, 
-          #labCol=tissue,
-          trace="none", 
-          #ColSideColors=cols, 
-          col=hmcol,
-          cexCol = 0.2, cexRow = 0.4
-      )
-
 ```
@@
 foo %>%
-  select(1, 2, 4) %>% # choose lod_diff as the stat of interest
-  tibble_to_matrix() %>%
-  heatmap.2( 
-          #labCol=tissue,
-          trace="none", 
-          #ColSideColors=cols, 
-          col=hmcol,
-          cexCol = 0.2, 
-          cexRow = 0.4
-      )
-
+  dplyr::select(1, 2, 4) %>% # choose lod_diff as the stat of interest
+  tibble_to_matrix() 
 ```
@@
 foo %>%
-  select(1, 2, 5) %>% # choose lod_diff_proportion as stat of interest
-  tibble_to_matrix() %>%
-  heatmap.2( 
-          #labCol=tissue,
-          trace="none", 
-          #ColSideColors=cols, 
-          col=hmcol,
-          cexCol = 0.2, cexRow = 0.4
-      )
-
+  dplyr::select(1, 2, 5) %>% # choose lod_diff_proportion as stat of interest
+  tibble_to_matrix() 
 ```
@@
 foomat <- foo %>%
-  select(1, 2, 4) %>% # choose lod_diff as stat of interest
+  dplyr::select(1, 2, 4) %>% # choose lod_diff as stat of interest
   tibble_to_matrix()
-  apply(FUN = function(x) x <= med_thresh, X = foomat, MARGIN = 2) %>%
-    d3heatmap()
+  apply(FUN = function(x) x <= med_thresh, X = foomat, MARGIN = 2) 
 ```
@@
 foo %>%
-  select(1, 2, 4) %>% # choose lod_diff as stat of interest
-  tibble_to_matrix() %>%
-  d3heatmap(cexRow = 0.2, cexCol = 0.2)
+  dplyr::select(1, 2, 4) %>% # choose lod_diff as stat of interest
+  tibble_to_matrix()
 ```
@@
 c1 <- bicluster(pleio_mat, bccc_out, number = 1)
-d3heatmap(c1$Bicluster1, "none", Rowv = FALSE, cexRow = 0.3, cexCol = 0.3)
-```
-
-```{r}
-d3heatmap(c1$Bicluster1, cexRow = 0.3, cexCol = 0.3)
-
 ```
@@
 
-```{r}
-# subset to get nonlocals that aren't mediated by Hnf4a
-# first, get symbols for nonlocals that aren't mediated by Hnf4a
-pmat <- bar %>%
-  filter(lod_diff <= mediation_threshold) %>%
-  dplyr::select(nonlocal_symbol) %>%
-  left_join(foo) %>%
-  tibble_to_matrix()
-dim(pmat)
-```
-
-```{r d3heat}
-d3heatmap(pmat, cexRow = 0.3, cexCol = 0.3)
-```
-```{r heat_binary}
-hc1 <- hclust(dist(pmat))
-plot(hc1, cex = 0.3)
-hc2 <- hclust(dist(t(pmat)))
-plot(hc2, cex = 0.3)
-pmat %>%
-  dist() %>%
-  hclust()
-
-```
-
-```{r d3heat_logical}
-d3heatmap(pmat > pleiotropy_threshold, cexRow = 0.3, cexCol = 0.3)
-```
-
-```{r}
-dim(pmat)
-length(hc1$order)
-d3heatmap(pmat[hc1$order, hc2$order] > pleiotropy_threshold, 
-          Rowv = FALSE, 
-          Colv = FALSE, 
-          cexRow = 0.3, 
-          cexCol = 0.3)
-d3heatmap(pmat[hc1$order, hc2$order], 
-          Rowv = FALSE, 
-          Colv = FALSE, 
-          cexRow = 0.3, 
-          cexCol = 0.3)
-
-```
-
 ## Local-local pleiotropy test results
@@
   tibble_to_matrix(symmetric = TRUE) %>%
-  d3heatmap()
+  d3heatmap(cexRow = 0.3, cexCol = 0.3, symm = TRUE)
+
+```
+
+```{r local-local-hclust}
+l_l_hclust <- local_pleio %>%
+  dplyr::select(5, 6, 3) %>%
+  tibble_to_matrix(symmetric = TRUE) %>%
+  (function(x){x[is.null(x)] <- 0; return(x)}) %>%
+  dist() %>%
+  hclust()
+l_l_hclust %>%  
+  cutree(h = 40) %>% table()
+
+l_l_hclust %>%
+  rafalib::myplclust(cex = 0.4, lab.col = 1 + as.numeric(l_l_hclust$labels == keller_mediator))
 ```
@@
   tibble_to_matrix(symmetric = TRUE) %>%
+  d3heatmap(cexRow = 0.3, cexCol = 0.3, symm = TRUE)
+```
+
+```{r nl-nl-hclust}
+nl_nl_hclust <- nonlocal_pleio %>%
+  dplyr::select(5, 6, 3) %>%
+  tibble_to_matrix(symmetric = TRUE) %>%
+  dist() %>% 
+  hclust()
+plot(nl_nl_hclust, cex = 0.2)
+nl_nl_hclust %>%
+  cutree(h = 10) %>%
+  table()
+```
+
+```{r rafalib-plot1}
+# make color labels for dendro plot
+precols <- tibble(nonlocal_symbol = nl_nl_hclust$labels) %>%
+  left_join(bar) %>% # bar has local_symbol set to keller_mediator
+  dplyr::arrange(match(x = nonlocal_symbol, table = nl_nl_hclust$labels))
+# check ordering of nonlocal_symbol against nl_nl_hclust$labels
+# IT'S THE SAME, but only after including the call to dplyr::arrange()
+
+cols <- 1 + (precols$lod_diff > 1.5)
+# dendro plot
+nl_nl_hclust %>%
+  rafalib::myplclust(cex = 0.3, lab.col = cols)
+```
+
+## Pleiotropy test results for *all* trait pairs
+
+```{r all-pleio}
+foo %>%
+  dplyr::select(9, 10, 3, 8, 1, 2) %>%
+  rename(gene1_id = nonlocal_gene_id,
+         gene2_id = local_gene_id,
+         gene1_symbol = nonlocal_symbol, 
+         gene2_symbol = local_symbol
+         ) %>%
+  bind_rows(local_pleio) %>%
+  bind_rows(nonlocal_pleio) %>%
+  dplyr::select(5, 6, 3) %>%
+  tibble_to_matrix(symmetric = TRUE) %>%
+  d3heatmap(cexRow = 0.2, cexCol = 0.2)
+```
+
+```{r, all-all-hclust}
+all_all_hclust <- foo %>%
+  dplyr::select(9, 10, 3, 8, 1, 2) %>%
+  rename(gene1_id = nonlocal_gene_id,
+         gene2_id = local_gene_id,
+         gene1_symbol = nonlocal_symbol, 
+         gene2_symbol = local_symbol
+         ) %>%
+  bind_rows(local_pleio) %>%
+  bind_rows(nonlocal_pleio) %>%
+  dplyr::select(5, 6, 3) %>%
+  tibble_to_matrix(symmetric = TRUE) %>%
+  dist() %>%
+  hclust()
+```
+
+```{r rafalib-plot2}
+cols <- 1 + (all_all_hclust$labels %in% bar$nonlocal_symbol)
+cols[all_all_hclust$labels == keller_mediator] <- 3
+# dendro plot
+all_all_hclust %>%
+  rafalib::myplclust(cex = 0.2, 
+                     lab.col = cols, 
+                     main = paste0("Chr", 
+                                   hot_chr, 
+                                   " dendrogram"
+                                   ), 
+                     sub = "pleiotropy test statistics"
+                       )
+
+```
+
+## Mediation (LOD diff) results (truncated at zero) for local-nonlocal pairs
+
+```{r med-trunc-heat}
+foo %>%
+  dplyr::select(1, 2, 4) %>%
+  mutate(lod_diff_trunc = lod_diff * (lod_diff > 0)) %>%
+  dplyr::select(-3) %>%
+  tibble_to_matrix(symmetric = FALSE) %>%
   d3heatmap(cexRow = 0.3, cexCol = 0.3)
 ```
+
+## Mediation (LOD diff proportion) results (truncated at zero) for local-nonlocal pairs
+
+```{r med-prop-trunc-heat}
+foo %>%
+  dplyr::select(1, 2, 5) %>%
+  mutate(lod_diff_prop_trunc = lod_diff_proportion * (lod_diff_proportion > 0)) %>%
+  dplyr::select(-3) %>%
+  tibble_to_matrix(symmetric = FALSE) %>%
+  d3heatmap(cexRow = 0.3, cexCol = 0.3)
+```

### 2019-07-09 analysis/paper/Chr11.Rmd
updated Rmd files

@@
 foo %>%
-  select(nonlocal_gene_id, local_gene_id, pleiotropy_lod, lod_diff) %>%
+  dplyr::select(nonlocal_gene_id, local_gene_id, pleiotropy_lod, lod_diff) %>%
 hotspots::tibble_to_matrix() %>%
@@
 foo %>%
-  select(nonlocal_gene_id, local_gene_id, lod_diff) %>%
+  dplyr::select(nonlocal_gene_id, local_gene_id, lod_diff) %>%
 hotspots::tibble_to_matrix() %>%
@@
 foo %>%
-  select(nonlocal_gene_id, local_gene_id, lod_diff) %>%
+  dplyr::select(nonlocal_gene_id, local_gene_id, lod_diff) %>%
   mutate(lod_diff = lod_diff * (lod_diff > 0)) %>%
@@
 hld <- foo %>%
-  select(nonlocal_gene_id, local_gene_id, lod_diff)
+  dplyr::select(nonlocal_gene_id, local_gene_id, lod_diff)
 plotly::plot_ly(as.data.frame(hld), z = as.data.frame(hld)$lod_diff, type = "heatmap")
@@
 foo %>%
-  select(1, 2, 4) %>% # choose lod_diff as the stat of interest
+  dplyr::select(1, 2, 4) %>% # choose lod_diff as the stat of interest
   tibble_to_matrix() %>%
@@
 foo %>%
-  select(1, 2, 5) %>% # choose lod_diff_proportion as stat of interest
+  dplyr::select(1, 2, 5) %>% # choose lod_diff_proportion as stat of interest
   tibble_to_matrix() %>%
@@
 foomat <- foo %>%
-  select(1, 2, 4) %>% # choose lod_diff as stat of interest
+  dplyr::select(1, 2, 4) %>% # choose lod_diff as stat of interest
   tibble_to_matrix()
@@
 foo %>%
-  select(1, 2, 4) %>% # choose lod_diff as stat of interest
+  dplyr::select(1, 2, 4) %>% # choose lod_diff as stat of interest
   tibble_to_matrix() %>%
@@
 ```
+
+## Pleiotropy test results for *all* trait pairs
+
+```{r all-pleio}
+foo %>%
+  dplyr::select(9, 10, 3, 8, 1, 2) %>%
+  rename(gene1_id = nonlocal_gene_id,
+         gene2_id = local_gene_id,
+         gene1_symbol = nonlocal_symbol, 
+         gene2_symbol = local_symbol
+         ) %>%
+  bind_rows(local_pleio) %>%
+  bind_rows(nonlocal_pleio) %>%
+  dplyr::select(5, 6, 3) %>%
+  tibble_to_matrix(symmetric = TRUE) %>%
+  d3heatmap(cexRow = 0.2, cexCol = 0.2)
+```
+
+## Mediation (LOD diff) results (truncated at zero) for local-nonlocal pairs
+
+```{r med-trunc-heat}
+foo %>%
+  dplyr::select(1, 2, 4) %>%
+  mutate(lod_diff_trunc = lod_diff * (lod_diff > 0)) %>%
+  dplyr::select(-3) %>%
+  tibble_to_matrix(symmetric = FALSE) %>%
+  d3heatmap(cexRow = 0.3, cexCol = 0.3)
+```
+
+## Mediation (LOD diff proportion) results (truncated at zero) for local-nonlocal pairs
+
+```{r med-prop-trunc-heat}
+foo %>%
+  dplyr::select(1, 2, 5) %>%
+  mutate(lod_diff_prop_trunc = lod_diff_proportion * (lod_diff_proportion > 0)) %>%
+  dplyr::select(-3) %>%
+  tibble_to_matrix(symmetric = FALSE) %>%
+  d3heatmap(cexRow = 0.3, cexCol = 0.3)
+```

### 2019-07-09 analysis/paper/Chr13-hclusterings.Rmd
updated Rmd files

--- 2019-07-09 analysis/paper/Chr11-hclusterings.Rmd
@@
 ---
-title: "Chr11 hotspot analyses"
+title: "Chr13 hotspot analyses"
 author: "Frederick Boehm"
@@
 # dendro plot
+# dendro plot
 all_all_hclust %>%

### 2019-07-09 analysis/paper/Chr2-hclusterings.Rmd
updated Rmd files

--- 2019-07-09 analysis/paper/Chr11-hclusterings.Rmd
@@
 ---
-title: "Chr11 hotspot analyses"
+title: "Chr2 hotspot analyses"
 author: "Frederick Boehm"
@@
 knitr::opts_chunk$set(echo = TRUE)
-hot_chr <- 11
-run_num <- 9921
-hot_mid <- 71
-keller_mediator <- "Sat2"
+hot_chr <- 2
+run_num <- 9941
+hot_mid <- 165.5
+keller_mediator <- "Hnf4a"
 library(hotspots)
@@
 
-```{r}
-foo %>%
-  dplyr::select(1, 2, 4) %>% # choose lod_diff as stat of interest
-  tibble_to_matrix()
-```
-
 ## IDEAS

### 2019-07-09 analysis/paper/Chr2.Rmd
updated Rmd files

@@
   left_join(tib, by = c("local_gene_id", "nonlocal_gene_id")) %>%
-  select(local_symbol, nonlocal_symbol, pleiotropy_lod, lod_diff, lod_diff_proportion, lod_without_med, lod_with_med, filename, nonlocal_gene_id, local_gene_id) 
+  dplyr::select(local_symbol, nonlocal_symbol, pleiotropy_lod, lod_diff, lod_diff_proportion, lod_without_med, lod_with_med, filename, nonlocal_gene_id, local_gene_id)
 ```
@@
 pleio_results_mat <- foo %>%
-  select(local_gene_id, nonlocal_gene_id, pleiotropy_lod, local_symbol, nonlocal_symbol) %>%
+  dplyr::select(local_gene_id, nonlocal_gene_id, pleiotropy_lod, local_symbol, nonlocal_symbol) %>%
   rename(gene1_id = local_gene_id,

### 2019-07-09 analysis/paper/Chr5-hclusterings.Rmd
updated Rmd files

--- 2019-07-09 analysis/paper/Chr11-hclusterings.Rmd
@@
 ---
-title: "Chr11 hotspot analyses"
+title: "Chr5 hotspot analyses"
 author: "Frederick Boehm"
@@
 knitr::opts_chunk$set(echo = TRUE)
-hot_chr <- 11
-run_num <- 9921
-hot_mid <- 71
-keller_mediator <- "Sat2"
+hot_chr <- 5
+run_num <- 9901
+hot_mid <- 146
+keller_mediator <- "Pdx1"
 library(hotspots)

### 2019-07-09 analysis/paper/Chr7-hclusterings.Rmd
updated Rmd files

--- 2019-07-09 analysis/paper/Chr5-hclusterings.Rmd
@@
 ---
-title: "Chr5 hotspot analyses"
+title: "Chr7 hotspot analyses"
 author: "Frederick Boehm"
@@
 knitr::opts_chunk$set(echo = TRUE)
-hot_chr <- 5
-run_num <- 9901
-hot_mid <- 146
-keller_mediator <- "Pdx1"
+hot_chr <- 7
+run_num <- 9911
+hot_mid <- 46
+keller_mediator <- "Fam83e"
 library(hotspots)

### 2019-07-09 analysis/paper/Chr7.Rmd
updated Rmd files

@@
   tibble_to_matrix(symmetric = TRUE) %>%
+  d3heatmap(cexRow = 0.3, cexCol = 0.3, symm = TRUE)
+```
+
+```{r nl-nl-hclust}
+nl_nl_hclust <- nonlocal_pleio %>%
+  dplyr::select(5, 6, 3) %>%
+  tibble_to_matrix(symmetric = TRUE) %>%
+  dist() %>% 
+  hclust()
+plot(nl_nl_hclust, cex = 0.2)
+nl_nl_hclust %>%
+  cutree(h = 10) %>%
+  table()
+```
+
+```{r rafalib-plot1}
+# make color labels for dendro plot
+precols <- tibble(nonlocal_symbol = nl_nl_hclust$labels) %>%
+  left_join(bar) %>% # bar has local_symbol set to keller_mediator
+  dplyr::arrange(match(x = nonlocal_symbol, table = nl_nl_hclust$labels))
+# check ordering of nonlocal_symbol against nl_nl_hclust$labels
+# IT'S THE SAME, but only after including the call to dplyr::arrange()
+
+cols <- 1 + (precols$lod_diff > 1.5)
+# dendro plot
+nl_nl_hclust %>%
+  rafalib::myplclust(cex = 0.3, lab.col = cols)
+```
+
+## Pleiotropy test results for *all* trait pairs
+
+```{r all-pleio}
+foo %>%
+  dplyr::select(9, 10, 3, 8, 1, 2) %>%
+  rename(gene1_id = nonlocal_gene_id,
+         gene2_id = local_gene_id,
+         gene1_symbol = nonlocal_symbol, 
+         gene2_symbol = local_symbol
+         ) %>%
+  bind_rows(local_pleio) %>%
+  bind_rows(nonlocal_pleio) %>%
+  dplyr::select(5, 6, 3) %>%
+  tibble_to_matrix(symmetric = TRUE) %>%
+  d3heatmap(cexRow = 0.2, cexCol = 0.2)
+```
+
+```{r, all-all-hclust}
+all_all_hclust <- foo %>%
+  dplyr::select(9, 10, 3, 8, 1, 2) %>%
+  rename(gene1_id = nonlocal_gene_id,
+         gene2_id = local_gene_id,
+         gene1_symbol = nonlocal_symbol, 
+         gene2_symbol = local_symbol
+         ) %>%
+  bind_rows(local_pleio) %>%
+  bind_rows(nonlocal_pleio) %>%
+  dplyr::select(5, 6, 3) %>%
+  tibble_to_matrix(symmetric = TRUE) %>%
+  dist() %>%
+  hclust()
+```
+
+```{r rafalib-plot2}
+cols <- 1 + (all_all_hclust$labels %in% bar$nonlocal_symbol)
+cols[all_all_hclust$labels == keller_mediator] <- 3
+# dendro plot
+all_all_hclust %>%
+  rafalib::myplclust(cex = 0.2, lab.col = cols)
+
+```
+
+## Mediation (LOD diff) results (truncated at zero) for local-nonlocal pairs
+
+```{r med-trunc-heat}
+foo %>%
+  dplyr::select(1, 2, 4) %>%
+  mutate(lod_diff_trunc = lod_diff * (lod_diff > 0)) %>%
+  dplyr::select(-3) %>%
+  tibble_to_matrix(symmetric = FALSE) %>%
   d3heatmap(cexRow = 0.3, cexCol = 0.3)
 ```
+
+## Mediation (LOD diff proportion) results (truncated at zero) for local-nonlocal pairs
+
+```{r med-prop-trunc-heat}
+foo %>%
+  dplyr::select(1, 2, 5) %>%
+  mutate(lod_diff_prop_trunc = lod_diff_proportion * (lod_diff_proportion > 0)) %>%
+  dplyr::select(-3) %>%
+  tibble_to_matrix(symmetric = FALSE) %>%
+  d3heatmap(cexRow = 0.3, cexCol = 0.3)
+```

### 2019-07-15 analysis/paper/Chr2-hclusterings.Rmd
updated Chr2 analyses.

@@
 ```{r all-pleio}
-foo %>%
+foohc <- foo %>%
   dplyr::select(9, 10, 3, 8, 1, 2) %>%
@@
   tibble_to_matrix(symmetric = TRUE) %>%
-  d3heatmap(cexRow = 0.2, cexCol = 0.2)
+  (function(x) {
+    hc <- dist(x) %>% hclust()
+    return(x[hc$order, hc$order])
+    }
+   ) %>%
+  (function(x) {
+    cols <- c("green", "orange")[1 + (colnames(x) %in% bar$nonlocal_symbol)]
+    heatmap.2(x, symm = TRUE, dendrogram = "both", na.rm = TRUE, ColSideColors = cols, cexRow = 0.2, cexCol = 0.2)
+              }
+   )
 ```

### 2019-07-15 analysis/paper/heatmaps-dendro-check.Rmd
added file to do checks on dendrograms & heatmaps

---
title: "Checking heatmaps and dendrograms"
author: "Frederick Boehm"
date: "7/14/2019"
output: html_document
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
library(hotspots)
fp <- paste0("../figures/Chr", hot_chr, "-")
knitr::opts_chunk$set(fig.path = fp)

```

The questions that I want to address today deal with heatmaps and dendrograms. Basically, I want to ensure that I'm correctly sorting & ordering a data matrix before passing it to a heatmap function.

I need to check a few things. First, what is the ordering of rows of a matrix when I create a dendrogram? What is the ordering of rows when I create a heatmap (without attempting any manual reordering)?

Can I order the data matrix according to the dendrogram so that, when calling the heatmap function *while preventing the automated reordering of rows*, do I get the same ordering as in the dendrogram?

Other questions that I want to consider include the *method* for dendrogram creation. I also want to think about coloring schemes in heatmaps.

## Reading Chr 11 intermediate result files

Let's examine the Chr 11 files for pleiotropy results.

```{r}
pleio <- readRDS("_cache/Chr11_pleio.rds")
```

### 2019-07-16 analysis/paper/Chr11-heat.Rmd
added Rmd files for heatmaps

---
title: "Heatmaps of pleiotropy test statistics"
author: "Frederick Boehm"
date: "7/14/2019"
output: html_document
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
library(hotspots)
```

The questions that I want to address today deal with heatmaps and dendrograms. Basically, I want to ensure that I'm correctly sorting & ordering a data matrix before passing it to a heatmap function.

I need to check a few things. First, what is the ordering of rows of a matrix when I create a dendrogram? What is the ordering of rows when I create a heatmap (without attempting any manual reordering)?

Can I order the data matrix according to the dendrogram so that, when calling the heatmap function *while preventing the automated reordering of rows*, do I get the same ordering as in the dendrogram?

Other questions that I want to consider include the *method* for dendrogram creation. I also want to think about coloring schemes in heatmaps.

# Chr 11

```{r setup_continued}
hot_chr <- 11
hot_mid <- 71
```

```{r more_setup}
hot_lower <- hot_mid - 2
hot_upper <- hot_mid + 2
fp <- paste0("../figures/Chr", hot_chr, "-")
knitr::opts_chunk$set(fig.path = fp)
```

## Reading Chr 11 intermediate result files

Let's examine the Chr 11 files for pleiotropy results.

```{r readRDS}
pleio_mixed_list <- readRDS("_cache/Chr11_pleio.rds")
pleio_local_list <- readRDS("_cache/Chr11_pleio_local.rds")
pleio_nonlocal_list <- readRDS("_cache/Chr11_pleio_nonlocal.rds")
med_list <- readRDS("_cache/Chr11_mediations.rds")
```

```{r, load_attie}
load("~/Box Sync/attie/keller2018-chr2-hotspot-chtc/data-to-ignore/Attie_DO378_eQTL_viewer_v1.Rdata")
library(tidyverse)
options(tibble.print_max = 50, tibble.print_min = 20)
```

```{r make_local_annot}
local_annot <- dataset.islet.rnaseq$lod.peaks %>%
  filter(chrom == hot_chr, pos <= hot_upper, pos >= hot_lower) %>%
  left_join(y = dataset.islet.rnaseq$annots, by = c("annot.id" = "gene_id")) %>%
  filter(!duplicated(annot.id)) %>%
  filter(lod >= 7.18) %>%
  mutate(local_tx = chr == hot_chr) %>%
  filter(local_tx)
```

```{r make_nonlocal_annot}
nonlocal_annot <- dataset.islet.rnaseq$lod.peaks %>%
  filter(chrom == hot_chr, pos <= hot_upper, pos >= hot_lower) %>%
  left_join(y = dataset.islet.rnaseq$annots, by = c("annot.id" = "gene_id")) %>%
  filter(!duplicated(annot.id)) %>%
  filter(lod >= 7.18) %>% #genomewide threshold
  mutate(local_tx = chr == hot_chr) %>%
  filter(!local_tx) %>% 
  filter(!is.na(hotspot))
```

```{r, make_local_annot2}
local_annot2 <- local_annot %>%
  filter(middle <= hot_upper + 5, middle >= hot_lower - 5) %>%
  arrange(desc(lod))
```

```{r subset_expr_matrix}
hot_local <- dataset.islet.rnaseq$expr[, colnames(dataset.islet.rnaseq$expr) %in% local_annot2$annot.id]
hot_nonlocal <- dataset.islet.rnaseq$expr[, colnames(dataset.islet.rnaseq$expr) %in% nonlocal_annot$annot.id]
nnonlocal <- ncol(hot_nonlocal)
nlocal <- ncol(hot_local)
```

```{r tabulate, message=FALSE, warning = FALSE}
med_tib <- med_list %>%
  purrr::map(bind_rows) %>%
  bind_rows() %>%
  as_tibble() %>%
  rename(lod_without_med = lod_no_med, lod_with_med = lod_med) %>%
  mutate(lod_diff = lod_without_med - lod_with_med, lod_diff_proportion = lod_diff / lod_without_med) %>%
  mutate(local_gene_id = rep(colnames(hot_local), each = nnonlocal), nonlocal_gene_id = rep(colnames(hot_nonlocal), times = nlocal)) %>%
  left_join(local_annot2, by = c("local_gene_id" = "annot.id")) %>%
  left_join(nonlocal_annot, by = c("nonlocal_gene_id" = "annot.id")) %>%
  dplyr::select(symbol.x, symbol.y, lod_without_med, lod_with_med, lod_diff, lod_diff_proportion, local_gene_id, nonlocal_gene_id) %>%
  rename(local_symbol = symbol.x, nonlocal_symbol = symbol.y)
```

```{r make_pleio_tib}
hot_annot <- local_annot2 %>%
  bind_rows(nonlocal_annot) %>%
  select(annot.id, symbol)
(pleio_tib <- pleio_mixed_list %>%
  purrr::map(.f = function(x) rename(x, gene1_id = local_gene_id, gene2_id = nonlocal_gene_id)) %>%
  c(pleio_local_list, pleio_nonlocal_list) %>%
  bind_rows() %>%
  rename(pleiotropy_lod = lod) %>%
  left_join(hot_annot, by = c("gene1_id" = "annot.id")) %>%
  rename(gene1_symbol = symbol) %>%
  left_join(hot_annot, by = c("gene2_id" = "annot.id")) %>%
  rename(gene2_symbol = symbol) %>%
  select(gene1_symbol, gene2_symbol, pleiotropy_lod)
)  
```

```{r hier_clust}
pm <- pleio_tib %>%
  tibble_to_matrix(symmetric = TRUE)
hc <- pm %>% 
  dist() %>%
  hclust()
hc %>%
  plot(cex = 0.2)
```

```{r}
hc$labels == rownames(pm)[hc$order]
```

```{r}
tibble(hc$labels, rownames(pm)[hc$order], hc$labels[hc$order], rownames(pm))
```

The tibble above shows the needed result. `hc$labels` is, by default definition, equal to `rownames(pm)`, while `rownames(pm)[hc$order]` gives the sequence of symbols that align with the leaves of the dendrogram.

```{r load_heatmaply}
library(heatmaply)
```

The vignette for `heatmaply` is quite helpful and includes discussion of its methods for ordering rows.

```{r heat-pleio}
heatmaply(pm, symm = TRUE)
```

```{r heat1}
heatmaply(percentize(pm), cexRow = 0.2, cexCol = 0.2, seriate = "OLO", symm = TRUE)
```

```{r heat2}
heatmaply(percentize(pm), cexRow = 0.2, cexCol = 0.2, seriate = "mean", symm = TRUE)
```

```{r heat3}
heatmaply(normalize(pm), 
          cexRow = 0.2, 
          cexCol = 0.2, 
          seriate = "OLO", 
          symm = TRUE, dendrogram = "both"
          )
```

## Heatmaps for matrix containing only nonlocal traits

```{r heat4}
pm_nonlocal <- pleio_tib %>%
  filter(gene1_symbol %in% nonlocal_annot$symbol, gene2_symbol %in% nonlocal_annot$symbol) %>%
  tibble_to_matrix(symmetric = TRUE)
pm_nonlocal %>%
  heatmaply(symm = TRUE, cexRow = 0.3, cexCol = 0.3)

```

```{r heat5}
(pm_nonlocal > 3) %>%
  apply(FUN = as.integer, MARGIN = 2) %>%
  heatmaply(symm = TRUE, cexRow = 0.3, cexCol = 0.3)
```

In the matrix of nonlocal - nonlocal pairs of pleiotropy test statistics, the fact that the maximum is only `r round(max(pm_nonlocal, na.rm = TRUE))` tells us that the heatmap of binary values is not going to be informative, since all or nearly all entries are less than any reasonable threshold.

Regarding possible transformations of the rows of the matrix, I don't really like transformations here. Transformations aim to preserve, in some sense, the distribution of values - test statistics, here - and I don't think that we should do that. I don't really care if a given row comes from a normal distribution, for example. I'm inclined to prefer heatmaps of the raw values.

### 2019-07-16 analysis/paper/heatmaps-pleiotropy-statistics-LOD.Rmd
added Rmd files for heatmaps

---
title: "Heatmaps of pleiotropy test statistics - all traits above LOD 10"
author: "Frederick Boehm"
date: "7/15/2019"
output: 
  html_document:
    toc: true
    code_folding: hide
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
library(hotspots)
pleio_threshold <- 4
ult <- 10 # univariate lod threshold
```

Below, we produce heatmaps for every hotspot. We consider, for every hotspot, all local and nonlocal traits with univariate LOD at least `r ult`. Note that the choice of `r ult` is arbitrary and limits the sets of traits to those that are rather strongly associated.

In future analyses, we might subset only the local traits... ie, use only the strongest local traits and all of the nonlocals

# Chr 2

```{r setup_continued-2}
hot_chr <- 2
hot_mid <- 165.5
```

```{r more_setup-2}
hot_lower <- hot_mid - 2
hot_upper <- hot_mid + 2
fp <- paste0("../figures/Chr", hot_chr, "-")
knitr::opts_chunk$set(fig.path = fp)
```

## Reading intermediate result files

Let's examine the Chr `r hot_chr` files for pleiotropy results.

```{r readRDS-2}
pleio_mixed_list <- readRDS(paste0("_cache/Chr", hot_chr, "_pleio.rds"))
pleio_local_list <- readRDS(paste0("_cache/Chr", hot_chr, "_pleio_local.rds"))
pleio_nonlocal_list <- readRDS(paste0("_cache/Chr", hot_chr, "_pleio_nonlocal.rds"))
med_list <- readRDS(paste0("_cache/Chr", hot_chr, "_mediations.rds"))
```

```{r, load_attie}
load("~/Box Sync/attie/keller2018-chr2-hotspot-chtc/data-to-ignore/Attie_DO378_eQTL_viewer_v1.Rdata")
library(tidyverse)
options(tibble.print_max = 50, tibble.print_min = 20)
```

```{r make_local_annot-2}
local_annot <- dataset.islet.rnaseq$lod.peaks %>%
  filter(chrom == hot_chr, pos <= hot_upper, pos >= hot_lower) %>%
  left_join(y = dataset.islet.rnaseq$annots, by = c("annot.id" = "gene_id")) %>%
  filter(!duplicated(annot.id)) %>%
  filter(lod >= ult) %>%
  mutate(local_tx = chr == hot_chr) %>%
  filter(local_tx)
```

```{r make_nonlocal_annot-2}
nonlocal_annot <- dataset.islet.rnaseq$lod.peaks %>%
  filter(chrom == hot_chr, pos <= hot_upper, pos >= hot_lower) %>%
  left_join(y = dataset.islet.rnaseq$annots, by = c("annot.id" = "gene_id")) %>%
  filter(!duplicated(annot.id)) %>%
  filter(lod >= ult) %>% #genomewide threshold
  mutate(local_tx = chr == hot_chr) %>%
  filter(!local_tx) %>% 
  filter(!is.na(hotspot))
```

```{r, make_local_annot2-2}
local_annot2 <- local_annot %>%
  filter(middle <= hot_upper + 5, middle >= hot_lower - 5) %>%
  arrange(desc(lod))
```

```{r subset_expr_matrix-2}
hot_local <- dataset.islet.rnaseq$expr[, colnames(dataset.islet.rnaseq$expr) %in% local_annot2$annot.id]
hot_nonlocal <- dataset.islet.rnaseq$expr[, colnames(dataset.islet.rnaseq$expr) %in% nonlocal_annot$annot.id]
nnonlocal <- ncol(hot_nonlocal)
nlocal <- ncol(hot_local)
```

```{r tabulate-2, message=FALSE, warning = FALSE, eval = FALSE}
med_tib <- med_list %>%
  purrr::map(bind_rows) %>%
  bind_rows() %>%
  as_tibble() %>%
  rename(lod_without_med = lod_no_med, lod_with_med = lod_med) %>%
  mutate(lod_diff = lod_without_med - lod_with_med, lod_diff_proportion = lod_diff / lod_without_med) %>%
  mutate(local_gene_id = rep(colnames(hot_local), each = nnonlocal), nonlocal_gene_id = rep(colnames(hot_nonlocal), times = nlocal)) %>%
  left_join(local_annot2, by = c("local_gene_id" = "annot.id")) %>%
  left_join(nonlocal_annot, by = c("nonlocal_gene_id" = "annot.id")) %>%
  dplyr::select(symbol.x, symbol.y, lod_without_med, lod_with_med, lod_diff, lod_diff_proportion, local_gene_id, nonlocal_gene_id) %>%
  rename(local_symbol = symbol.x, nonlocal_symbol = symbol.y)
```

```{r make_pleio_tib-2}
hot_annot <- local_annot2 %>%
  bind_rows(nonlocal_annot) %>%
  select(annot.id, symbol)
(pleio_tib <- pleio_mixed_list %>%
  purrr::map(.f = function(x) rename(x, gene1_id = local_gene_id, gene2_id = nonlocal_gene_id)) %>%
  c(pleio_local_list, pleio_nonlocal_list) %>%
  bind_rows() %>%
  rename(pleiotropy_lod = lod) %>%
  inner_join(hot_annot, by = c("gene1_id" = "annot.id")) %>%
  rename(gene1_symbol = symbol) %>%
  inner_join(hot_annot, by = c("gene2_id" = "annot.id")) %>%
  rename(gene2_symbol = symbol) %>%
  select(gene1_symbol, gene2_symbol, pleiotropy_lod)
)  
```

```{r make_pm-2}
pm <- pleio_tib %>%
  tibble_to_matrix(symmetric = TRUE)
```

## heatmaps for full matrix

```{r load_heatmaply}
library(heatmaply)
```

```{r make_scols-2}
scols <- 1L + (rownames(pm) %in% local_annot2$symbol)
```

The vignette for `heatmaply` is quite helpful and includes discussion of its methods for ordering rows.

```{r heat-pleio-2}
heatmaply(pm, symm = TRUE, cexRow = 0.2, cexCol = 0.2, row_side_colors = scols)
```

```{r heat3-2}
(pm > pleio_threshold) %>%
  apply(FUN = as.integer, MARGIN = 2) %>%
  (function(x) {rownames(x) <- colnames(x); return(x)}) %>%
  heatmaply(cexRow = 0.2, cexCol = 0.2, symm = TRUE, row_side_colors = scols)
```

## Heatmaps for matrix containing only nonlocal traits

```{r heat5-2}
pm_nonlocal <- pleio_tib %>%
  filter(gene1_symbol %in% nonlocal_annot$symbol, gene2_symbol %in% nonlocal_annot$symbol) %>%
  tibble_to_matrix(symmetric = TRUE)
pm_nonlocal %>%
  heatmaply(symm = TRUE, cexRow = 0.3, cexCol = 0.3)

```

```{r heat6-2}
(pm_nonlocal > pleio_threshold) %>%
  apply(FUN = as.integer, MARGIN = 2) %>%
  (function(x) {rownames(x) <- colnames(x); return(x)}) %>%
  heatmaply(symm = TRUE, cexRow = 0.3, cexCol = 0.3)
```

# Chr 5

```{r setup_continued-5}
hot_chr <- 5
hot_mid <- 146
```

```{r more_setup-5}
hot_lower <- hot_mid - 2
hot_upper <- hot_mid + 2
fp <- paste0("../figures/Chr", hot_chr, "-")
knitr::opts_chunk$set(fig.path = fp)
```

## Reading intermediate result files

Let's examine the Chr `r hot_chr` files for pleiotropy results.

```{r readRDS-5}
pleio_mixed_list <- readRDS(paste0("_cache/Chr", hot_chr, "_pleio.rds"))
pleio_local_list <- readRDS(paste0("_cache/Chr", hot_chr, "_pleio_local.rds"))
pleio_nonlocal_list <- readRDS(paste0("_cache/Chr", hot_chr, "_pleio_nonlocal.rds"))
med_list <- readRDS(paste0("_cache/Chr", hot_chr, "_mediations.rds"))
```

```{r make_local_annot-5}
local_annot <- dataset.islet.rnaseq$lod.peaks %>%
  filter(chrom == hot_chr, pos <= hot_upper, pos >= hot_lower) %>%
  left_join(y = dataset.islet.rnaseq$annots, by = c("annot.id" = "gene_id")) %>%
  filter(!duplicated(annot.id)) %>%
  filter(lod >= ult) %>%
  mutate(local_tx = chr == hot_chr) %>%
  filter(local_tx)
```

```{r make_nonlocal_annot-5}
nonlocal_annot <- dataset.islet.rnaseq$lod.peaks %>%
  filter(chrom == hot_chr, pos <= hot_upper, pos >= hot_lower) %>%
  left_join(y = dataset.islet.rnaseq$annots, by = c("annot.id" = "gene_id")) %>%
  filter(!duplicated(annot.id)) %>%
  filter(lod >= ult) %>% #genomewide threshold
  mutate(local_tx = chr == hot_chr) %>%
  filter(!local_tx) %>% 
  filter(!is.na(hotspot))
```

```{r, make_local_annot2-5}
local_annot2 <- local_annot %>%
  filter(middle <= hot_upper + 5, middle >= hot_lower - 5) %>%
  arrange(desc(lod))
```

```{r subset_expr_matrix-5}
hot_local <- dataset.islet.rnaseq$expr[, colnames(dataset.islet.rnaseq$expr) %in% local_annot2$annot.id]
hot_nonlocal <- dataset.islet.rnaseq$expr[, colnames(dataset.islet.rnaseq$expr) %in% nonlocal_annot$annot.id]
nnonlocal <- ncol(hot_nonlocal)
nlocal <- ncol(hot_local)
```

```{r tabulate-5, message=FALSE, warning = FALSE, eval = FALSE}
med_tib <- med_list %>%
  purrr::map(bind_rows) %>%
  bind_rows() %>%
  as_tibble() %>%
  rename(lod_without_med = lod_no_med, lod_with_med = lod_med) %>%
  mutate(lod_diff = lod_without_med - lod_with_med, lod_diff_proportion = lod_diff / lod_without_med) %>%
  mutate(local_gene_id = rep(colnames(hot_local), each = nnonlocal), nonlocal_gene_id = rep(colnames(hot_nonlocal), times = nlocal)) %>%
  left_join(local_annot2, by = c("local_gene_id" = "annot.id")) %>%
  left_join(nonlocal_annot, by = c("nonlocal_gene_id" = "annot.id")) %>%
  dplyr::select(symbol.x, symbol.y, lod_without_med, lod_with_med, lod_diff, lod_diff_proportion, local_gene_id, nonlocal_gene_id) %>%
  rename(local_symbol = symbol.x, nonlocal_symbol = symbol.y)
```

```{r make_pleio_tib-5}
hot_annot <- local_annot2 %>%
  bind_rows(nonlocal_annot) %>%
  select(annot.id, symbol)
(pleio_tib <- pleio_mixed_list %>%
  purrr::map(.f = function(x) rename(x, gene1_id = local_gene_id, gene2_id = nonlocal_gene_id)) %>%
  c(pleio_local_list, pleio_nonlocal_list) %>%
  bind_rows() %>%
  rename(pleiotropy_lod = lod) %>%
  inner_join(hot_annot, by = c("gene1_id" = "annot.id")) %>%
  rename(gene1_symbol = symbol) %>%
  inner_join(hot_annot, by = c("gene2_id" = "annot.id")) %>%
  rename(gene2_symbol = symbol) %>%
  select(gene1_symbol, gene2_symbol, pleiotropy_lod)
)  
```

```{r make_pm-5}
pm <- pleio_tib %>%
  tibble_to_matrix(symmetric = TRUE)
```

## heatmaps for full matrix

```{r make_scols-5}
scols <- 1L + (rownames(pm) %in% local_annot2$symbol)
```

```{r heat-pleio-5}
heatmaply(pm, symm = TRUE, cexRow = 0.2, cexCol = 0.2, row_side_colors = scols)
```

```{r heat3-5}
(pm > pleio_threshold) %>%
  apply(FUN = as.integer, MARGIN = 2) %>%
  (function(x) {rownames(x) <- colnames(x); return(x)}) %>%
  heatmaply(cexRow = 0.2, cexCol = 0.2, symm = TRUE, row_side_colors = scols)
```

## Heatmaps for matrix containing only nonlocal traits

```{r heat5-5}
pm_nonlocal <- pleio_tib %>%
  filter(gene1_symbol %in% nonlocal_annot$symbol, gene2_symbol %in% nonlocal_annot$symbol) %>%
  tibble_to_matrix(symmetric = TRUE)
pm_nonlocal %>%
  heatmaply(symm = TRUE, cexRow = 0.3, cexCol = 0.3)

```

```{r heat6-5}
(pm_nonlocal > pleio_threshold) %>%
  apply(FUN = as.integer, MARGIN = 2) %>%
  (function(x) {rownames(x) <- colnames(x); return(x)}) %>%
  heatmaply(symm = TRUE, cexRow = 0.3, cexCol = 0.3)
```

# Chr 7

```{r setup_continued-7}
hot_chr <- 7
hot_mid <- 46
```

```{r more_setup-7}
hot_lower <- hot_mid - 2
hot_upper <- hot_mid + 2
fp <- paste0("../figures/Chr", hot_chr, "-")
knitr::opts_chunk$set(fig.path = fp)
```

## Reading intermediate result files

Let's examine the Chr `r hot_chr` files for pleiotropy results.

```{r readRDS-7}
pleio_mixed_list <- readRDS(paste0("_cache/Chr", hot_chr, "_pleio.rds"))
pleio_local_list <- readRDS(paste0("_cache/Chr", hot_chr, "_pleio_local.rds"))
pleio_nonlocal_list <- readRDS(paste0("_cache/Chr", hot_chr, "_pleio_nonlocal.rds"))
med_list <- readRDS(paste0("_cache/Chr", hot_chr, "_mediations.rds"))
```

```{r make_local_annot-7}
local_annot <- dataset.islet.rnaseq$lod.peaks %>%
  filter(chrom == hot_chr, pos <= hot_upper, pos >= hot_lower) %>%
  left_join(y = dataset.islet.rnaseq$annots, by = c("annot.id" = "gene_id")) %>%
  filter(!duplicated(annot.id)) %>%
  filter(lod >= ult) %>%
  mutate(local_tx = chr == hot_chr) %>%
  filter(local_tx)
```

```{r make_nonlocal_annot-7}
nonlocal_annot <- dataset.islet.rnaseq$lod.peaks %>%
  filter(chrom == hot_chr, pos <= hot_upper, pos >= hot_lower) %>%
  left_join(y = dataset.islet.rnaseq$annots, by = c("annot.id" = "gene_id")) %>%
  filter(!duplicated(annot.id)) %>%
  filter(lod >= ult) %>% #genomewide threshold
  mutate(local_tx = chr == hot_chr) %>%
  filter(!local_tx) %>% 
  filter(!is.na(hotspot))
```

```{r, make_local_annot2-7}
local_annot2 <- local_annot %>%
  filter(middle <= hot_upper + 5, middle >= hot_lower - 5) %>%
  arrange(desc(lod))
```

```{r subset_expr_matrix-7}
hot_local <- dataset.islet.rnaseq$expr[, colnames(dataset.islet.rnaseq$expr) %in% local_annot2$annot.id]
hot_nonlocal <- dataset.islet.rnaseq$expr[, colnames(dataset.islet.rnaseq$expr) %in% nonlocal_annot$annot.id]
nnonlocal <- ncol(hot_nonlocal)
nlocal <- ncol(hot_local)
```

```{r tabulate-7, message=FALSE, warning = FALSE, eval = FALSE}
med_tib <- med_list %>%
  purrr::map(bind_rows) %>%
  bind_rows() %>%
  as_tibble() %>%
  rename(lod_without_med = lod_no_med, lod_with_med = lod_med) %>%
  mutate(lod_diff = lod_without_med - lod_with_med, lod_diff_proportion = lod_diff / lod_without_med) %>%
  mutate(local_gene_id = rep(colnames(hot_local), each = nnonlocal), nonlocal_gene_id = rep(colnames(hot_nonlocal), times = nlocal)) %>%
  left_join(local_annot2, by = c("local_gene_id" = "annot.id")) %>%
  left_join(nonlocal_annot, by = c("nonlocal_gene_id" = "annot.id")) %>%
  dplyr::select(symbol.x, symbol.y, lod_without_med, lod_with_med, lod_diff, lod_diff_proportion, local_gene_id, nonlocal_gene_id) %>%
  rename(local_symbol = symbol.x, nonlocal_symbol = symbol.y)
```

```{r make_pleio_tib-7}
hot_annot <- local_annot2 %>%
  bind_rows(nonlocal_annot) %>%
  select(annot.id, symbol)
(pleio_tib <- pleio_mixed_list %>%
  purrr::map(.f = function(x) rename(x, gene1_id = local_gene_id, gene2_id = nonlocal_gene_id)) %>%
  c(pleio_local_list, pleio_nonlocal_list) %>%
  bind_rows() %>%
  rename(pleiotropy_lod = lod) %>%
  inner_join(hot_annot, by = c("gene1_id" = "annot.id")) %>%
  rename(gene1_symbol = symbol) %>%
  inner_join(hot_annot, by = c("gene2_id" = "annot.id")) %>%
  rename(gene2_symbol = symbol) %>%
  select(gene1_symbol, gene2_symbol, pleiotropy_lod)
)  
```

```{r make_pm-7}
pm <- pleio_tib %>%
  tibble_to_matrix(symmetric = TRUE)
```

## heatmaps for full matrix

```{r make_scols-7}
scols <- 1L + (rownames(pm) %in% local_annot2$symbol)
```

```{r heat-pleio-7}
heatmaply(pm, symm = TRUE, cexRow = 0.2, cexCol = 0.2)
```

```{r heat3-7}
(pm > pleio_threshold) %>%
  apply(FUN = as.integer, MARGIN = 2) %>%
  (function(x) {rownames(x) <- colnames(x); return(x)}) %>%
  heatmaply(cexRow = 0.2, cexCol = 0.2, symm = TRUE)
```

## Heatmaps for matrix containing only nonlocal traits

```{r heat5-7}
pm_nonlocal <- pleio_tib %>%
  filter(gene1_symbol %in% nonlocal_annot$symbol, gene2_symbol %in% nonlocal_annot$symbol) %>%
  tibble_to_matrix(symmetric = TRUE)
pm_nonlocal %>%
  heatmaply(symm = TRUE, cexRow = 0.3, cexCol = 0.3)

```

```{r heat6-7}
(pm_nonlocal > pleio_threshold) %>%
  apply(FUN = as.integer, MARGIN = 2) %>%
  (function(x) {rownames(x) <- colnames(x); return(x)}) %>%
  heatmaply(symm = TRUE, cexRow = 0.3, cexCol = 0.3)
```

# Chr 11

```{r setup_continued-11}
hot_chr <- 11
hot_mid <- 71
```

```{r more_setup-11}
hot_lower <- hot_mid - 2
hot_upper <- hot_mid + 2
fp <- paste0("../figures/Chr", hot_chr, "-")
knitr::opts_chunk$set(fig.path = fp)
```

## Reading intermediate result files

Let's examine the Chr `r hot_chr` files for pleiotropy results.

```{r readRDS-11}
pleio_mixed_list <- readRDS("_cache/Chr11_pleio.rds")
pleio_local_list <- readRDS("_cache/Chr11_pleio_local.rds")
pleio_nonlocal_list <- readRDS("_cache/Chr11_pleio_nonlocal.rds")
med_list <- readRDS("_cache/Chr11_mediations.rds")
```

```{r make_local_annot-11}
local_annot <- dataset.islet.rnaseq$lod.peaks %>%
  filter(chrom == hot_chr, pos <= hot_upper, pos >= hot_lower) %>%
  left_join(y = dataset.islet.rnaseq$annots, by = c("annot.id" = "gene_id")) %>%
  filter(!duplicated(annot.id)) %>%
  filter(lod >= ult) %>%
  mutate(local_tx = chr == hot_chr) %>%
  filter(local_tx)
```

```{r make_nonlocal_annot-11}
nonlocal_annot <- dataset.islet.rnaseq$lod.peaks %>%
  filter(chrom == hot_chr, pos <= hot_upper, pos >= hot_lower) %>%
  left_join(y = dataset.islet.rnaseq$annots, by = c("annot.id" = "gene_id")) %>%
  filter(!duplicated(annot.id)) %>%
  filter(lod >= ult) %>% #genomewide threshold
  mutate(local_tx = chr == hot_chr) %>%
  filter(!local_tx) %>% 
  filter(!is.na(hotspot))
```

```{r, make_local_annot2-11}
local_annot2 <- local_annot %>%
  filter(middle <= hot_upper + 5, middle >= hot_lower - 5) %>%
  arrange(desc(lod))
```

```{r subset_expr_matrix-11}
hot_local <- dataset.islet.rnaseq$expr[, colnames(dataset.islet.rnaseq$expr) %in% local_annot2$annot.id]
hot_nonlocal <- dataset.islet.rnaseq$expr[, colnames(dataset.islet.rnaseq$expr) %in% nonlocal_annot$annot.id]
nnonlocal <- ncol(hot_nonlocal)
nlocal <- ncol(hot_local)
```

```{r tabulate-11, message=FALSE, warning = FALSE, eval = FALSE}
med_tib <- med_list %>%
  purrr::map(bind_rows) %>%
  bind_rows() %>%
  as_tibble() %>%
  rename(lod_without_med = lod_no_med, lod_with_med = lod_med) %>%
  mutate(lod_diff = lod_without_med - lod_with_med, lod_diff_proportion = lod_diff / lod_without_med) %>%
  mutate(local_gene_id = rep(colnames(hot_local), each = nnonlocal), nonlocal_gene_id = rep(colnames(hot_nonlocal), times = nlocal)) %>%
  left_join(local_annot2, by = c("local_gene_id" = "annot.id")) %>%
  left_join(nonlocal_annot, by = c("nonlocal_gene_id" = "annot.id")) %>%
  dplyr::select(symbol.x, symbol.y, lod_without_med, lod_with_med, lod_diff, lod_diff_proportion, local_gene_id, nonlocal_gene_id) %>%
  rename(local_symbol = symbol.x, nonlocal_symbol = symbol.y)
```

```{r make_pleio_tib-11}
hot_annot <- local_annot2 %>%
  bind_rows(nonlocal_annot) %>%
  select(annot.id, symbol)
(pleio_tib <- pleio_mixed_list %>%
  purrr::map(.f = function(x) rename(x, gene1_id = local_gene_id, gene2_id = nonlocal_gene_id)) %>%
  c(pleio_local_list, pleio_nonlocal_list) %>%
  bind_rows() %>%
  rename(pleiotropy_lod = lod) %>%
  inner_join(hot_annot, by = c("gene1_id" = "annot.id")) %>%
  rename(gene1_symbol = symbol) %>%
  inner_join(hot_annot, by = c("gene2_id" = "annot.id")) %>%
  rename(gene2_symbol = symbol) %>%
  select(gene1_symbol, gene2_symbol, pleiotropy_lod)
)  
```

```{r make_pm-11}
pm <- pleio_tib %>%
  tibble_to_matrix(symmetric = TRUE)
```

## heatmaps for full matrix

```{r make_scols-11}
scols <- 1L + (rownames(pm) %in% local_annot2$symbol)
```

```{r heat-pleio-11}
heatmaply(pm, symm = TRUE, cexRow = 0.2, cexCol = 0.2, row_side_colors = scols)
```

```{r heat3-11}
(pm > pleio_threshold) %>%
  apply(FUN = as.integer, MARGIN = 2) %>%
  (function(x) {rownames(x) <- colnames(x); return(x)}) %>%
  heatmaply(cexRow = 0.2, cexCol = 0.2, symm = TRUE, row_side_colors = scols)
```

## Heatmaps for matrix containing only nonlocal traits

```{r heat5-11}
pm_nonlocal <- pleio_tib %>%
  filter(gene1_symbol %in% nonlocal_annot$symbol, gene2_symbol %in% nonlocal_annot$symbol) %>%
  tibble_to_matrix(symmetric = TRUE)
pm_nonlocal %>%
  heatmaply(symm = TRUE, cexRow = 0.3, cexCol = 0.3)

```

```{r heat6-11}
(pm_nonlocal > pleio_threshold) %>%
  apply(FUN = as.integer, MARGIN = 2) %>%
  (function(x) {rownames(x) <- colnames(x); return(x)}) %>%
  heatmaply(symm = TRUE, cexRow = 0.3, cexCol = 0.3)
```

# Chr 13

```{r setup_continued-13}
hot_chr <- 13
hot_mid <- 112.5
```

```{r more_setup-13}
hot_lower <- hot_mid - 2
hot_upper <- hot_mid + 2
fp <- paste0("../figures/Chr", hot_chr, "-")
knitr::opts_chunk$set(fig.path = fp)
```

## Reading intermediate result files

Let's examine the Chr `r hot_chr` files for pleiotropy results.

```{r readRDS-13}
pleio_mixed_list <- readRDS(paste0("_cache/Chr", hot_chr, "_pleio.rds"))
pleio_local_list <- readRDS(paste0("_cache/Chr", hot_chr, "_pleio_local.rds"))
pleio_nonlocal_list <- readRDS(paste0("_cache/Chr", hot_chr, "_pleio_nonlocal.rds"))
med_list <- readRDS(paste0("_cache/Chr", hot_chr, "_mediations.rds"))
```

```{r make_local_annot-13}
local_annot <- dataset.islet.rnaseq$lod.peaks %>%
  filter(chrom == hot_chr, pos <= hot_upper, pos >= hot_lower) %>%
  left_join(y = dataset.islet.rnaseq$annots, by = c("annot.id" = "gene_id")) %>%
  filter(!duplicated(annot.id)) %>%
  filter(lod >= ult) %>%
  mutate(local_tx = chr == hot_chr) %>%
  filter(local_tx)
```

```{r make_nonlocal_annot-13}
nonlocal_annot <- dataset.islet.rnaseq$lod.peaks %>%
  filter(chrom == hot_chr, pos <= hot_upper, pos >= hot_lower) %>%
  left_join(y = dataset.islet.rnaseq$annots, by = c("annot.id" = "gene_id")) %>%
  filter(!duplicated(annot.id)) %>%
  filter(lod >= ult) %>% #genomewide threshold
  mutate(local_tx = chr == hot_chr) %>%
  filter(!local_tx) %>% 
  filter(!is.na(hotspot))
```

```{r, make_local_annot2-13}
local_annot2 <- local_annot %>%
  filter(middle <= hot_upper + 5, middle >= hot_lower - 5) %>%
  arrange(desc(lod))
```

```{r subset_expr_matrix-13}
hot_local <- dataset.islet.rnaseq$expr[, colnames(dataset.islet.rnaseq$expr) %in% local_annot2$annot.id]
hot_nonlocal <- dataset.islet.rnaseq$expr[, colnames(dataset.islet.rnaseq$expr) %in% nonlocal_annot$annot.id]
nnonlocal <- ncol(hot_nonlocal)
nlocal <- ncol(hot_local)
```

```{r tabulate-13, message=FALSE, warning = FALSE, eval = FALSE}
med_tib <- med_list %>%
  purrr::map(bind_rows) %>%
  bind_rows() %>%
  as_tibble() %>%
  rename(lod_without_med = lod_no_med, lod_with_med = lod_med) %>%
  mutate(lod_diff = lod_without_med - lod_with_med, lod_diff_proportion = lod_diff / lod_without_med) %>%
  mutate(local_gene_id = rep(colnames(hot_local), each = nnonlocal), nonlocal_gene_id = rep(colnames(hot_nonlocal), times = nlocal)) %>%
  left_join(local_annot2, by = c("local_gene_id" = "annot.id")) %>%
  left_join(nonlocal_annot, by = c("nonlocal_gene_id" = "annot.id")) %>%
  dplyr::select(symbol.x, symbol.y, lod_without_med, lod_with_med, lod_diff, lod_diff_proportion, local_gene_id, nonlocal_gene_id) %>%
  rename(local_symbol = symbol.x, nonlocal_symbol = symbol.y)
```

```{r make_pleio_tib-13}
hot_annot <- local_annot2 %>%
  bind_rows(nonlocal_annot) %>%
  select(annot.id, symbol)
(pleio_tib <- pleio_mixed_list %>%
  purrr::map(.f = function(x) rename(x, gene1_id = local_gene_id, gene2_id = nonlocal_gene_id)) %>%
  c(pleio_local_list, pleio_nonlocal_list) %>%
  bind_rows() %>%
  rename(pleiotropy_lod = lod) %>%
  inner_join(hot_annot, by = c("gene1_id" = "annot.id")) %>%
  rename(gene1_symbol = symbol) %>%
  inner_join(hot_annot, by = c("gene2_id" = "annot.id")) %>%
  rename(gene2_symbol = symbol) %>%
  select(gene1_symbol, gene2_symbol, pleiotropy_lod)
)  
```

```{r make_pm-13}
pm <- pleio_tib %>%
  tibble_to_matrix(symmetric = TRUE)
```

## heatmaps for full matrix

```{r make_scols-13}
scols <- 1L + (rownames(pm) %in% local_annot2$symbol)
```

```{r heat-pleio-13}
heatmaply(pm, symm = TRUE, cexRow = 0.2, cexCol = 0.2, row_side_colors = scols)
```

```{r heat3-13}
(pm > pleio_threshold) %>%
  apply(FUN = as.integer, MARGIN = 2) %>%
  (function(x) {rownames(x) <- colnames(x); return(x)}) %>%
  heatmaply(cexRow = 0.2, cexCol = 0.2, symm = TRUE, row_side_colors = scols)
```

## Heatmaps for matrix containing only nonlocal traits

```{r heat5-13}
pm_nonlocal <- pleio_tib %>%
  filter(gene1_symbol %in% nonlocal_annot$symbol, gene2_symbol %in% nonlocal_annot$symbol) %>%
  tibble_to_matrix(symmetric = TRUE)
pm_nonlocal %>%
  heatmaply(symm = TRUE, cexRow = 0.3, cexCol = 0.3)
```

```{r heat6-13}
(pm_nonlocal > pleio_threshold) %>%
  apply(FUN = as.integer, MARGIN = 2) %>%
  heatmaply(symm = TRUE, cexRow = 0.3, cexCol = 0.3)
```

### 2019-07-16 analysis/paper/heatmaps-pleiotropy-statistics-subset-locals.Rmd
added Rmd files for heatmaps

--- 2019-07-16 analysis/paper/heatmaps-pleiotropy-statistics-LOD.Rmd
@@
 ---
-title: "Heatmaps of pleiotropy test statistics - all traits above LOD 10"
+title: "Heatmaps of pleiotropy test statistics - all nonlocals and all locals with LOD above 15"
 author: "Frederick Boehm"
-date: "7/15/2019"
+date: "7/16/2019"
 output: 
@@
 pleio_threshold <- 4
-ult <- 10 # univariate lod threshold
+ult <- 15 # univariate lod threshold
 ```
@@
 Below, we produce heatmaps for every hotspot.
-We consider, for every hotspot, all local and nonlocal traits with univariate LOD at least `r ult`.
+We consider, for every hotspot, all local traits with univariate LOD at least `r ult` and all nonlocal traits.
 Note that the choice of `r ult` is arbitrary and limits the sets of traits to those that are rather strongly associated.
-
-In future analyses, we might subset only the local traits... ie, use only the strongest local traits and all of the nonlocals
 
@@
   filter(!duplicated(annot.id)) %>%
-  filter(lod >= ult) %>% #genomewide threshold
   mutate(local_tx = chr == hot_chr) %>%
@@
   filter(!duplicated(annot.id)) %>%
-  filter(lod >= ult) %>% #genomewide threshold
+#  filter(lod >= ult) %>% #genomewide threshold
   mutate(local_tx = chr == hot_chr) %>%

### 2019-07-16 analysis/paper/heatmaps-pleiotropy-statistics.Rmd
added Rmd files for heatmaps

--- 2019-07-16 analysis/paper/heatmaps-pleiotropy-statistics-subset-locals.Rmd
@@
 ---
-title: "Heatmaps of pleiotropy test statistics - all nonlocals and all locals with LOD above 15"
+title: "Heatmaps of pleiotropy test statistics - all traits"
 author: "Frederick Boehm"
-date: "7/16/2019"
+date: "7/15/2019"
 output: 
@@
 pleio_threshold <- 4
-ult <- 15 # univariate lod threshold
 ```
@@
 Below, we produce heatmaps for every hotspot.
-We consider, for every hotspot, all local traits with univariate LOD at least `r ult` and all nonlocal traits.
-Note that the choice of `r ult` is arbitrary and limits the sets of traits to those that are rather strongly associated.
+We consider, for every hotspot, all local and nonlocal traits with univariate LOD at least 7.18, which is the genomewide threshold that Keller et al identified via permutations.
 
@@
   filter(!duplicated(annot.id)) %>%
-  filter(lod >= ult) %>%
+  filter(lod >= 7.18) %>%
   mutate(local_tx = chr == hot_chr) %>%
@@
   filter(!duplicated(annot.id)) %>%
+  filter(lod >= 7.18) %>% #genomewide threshold
   mutate(local_tx = chr == hot_chr) %>%
@@
 
-```{r tabulate-2, message=FALSE, warning = FALSE, eval = FALSE}
+```{r tabulate-2, message=FALSE, warning = FALSE}
 med_tib <- med_list %>%
@@
   rename(pleiotropy_lod = lod) %>%
-  inner_join(hot_annot, by = c("gene1_id" = "annot.id")) %>%
+  left_join(hot_annot, by = c("gene1_id" = "annot.id")) %>%
   rename(gene1_symbol = symbol) %>%
-  inner_join(hot_annot, by = c("gene2_id" = "annot.id")) %>%
+  left_join(hot_annot, by = c("gene2_id" = "annot.id")) %>%
   rename(gene2_symbol = symbol) %>%
@@
 
+```{r heat2-2}
+heatmaply(pm, cexRow = 0.2, cexCol = 0.2, seriate = "mean", symm = TRUE, row_side_colors = scols)
+```
+
 ```{r heat3-2}
@@
   heatmaply(cexRow = 0.2, cexCol = 0.2, symm = TRUE, row_side_colors = scols)
+```
+
+```{r heat4-2}
+(pm > pleio_threshold) %>%
+  apply(FUN = as.integer, MARGIN = 2) %>%
+  (function(x) {rownames(x) <- colnames(x); return(x)}) %>%
+  heatmaply(cexRow = 0.2, cexCol = 0.2, symm = TRUE, seriate = "mean", row_side_colors = scols)
 ```
@@
   filter(!duplicated(annot.id)) %>%
-#  filter(lod >= ult) %>% #genomewide threshold
+  filter(lod >= 7.18) %>% #genomewide threshold
   mutate(local_tx = chr == hot_chr) %>%
@@
 
-```{r tabulate-5, message=FALSE, warning = FALSE, eval = FALSE}
+```{r tabulate-5, message=FALSE, warning = FALSE}
 med_tib <- med_list %>%
@@
 
+```{r heat2-5}
+heatmaply(pm, cexRow = 0.2, cexCol = 0.2, seriate = "mean", symm = TRUE, row_side_colors = scols)
+```
+
 ```{r heat3-5}
@@
   heatmaply(cexRow = 0.2, cexCol = 0.2, symm = TRUE, row_side_colors = scols)
+```
+
+```{r heat4-5}
+(pm > pleio_threshold) %>%
+  apply(FUN = as.integer, MARGIN = 2) %>%
+  (function(x) {rownames(x) <- colnames(x); return(x)}) %>%
+  heatmaply(cexRow = 0.2, cexCol = 0.2, symm = TRUE, seriate = "mean", row_side_colors = scols)
 ```
@@
 
-```{r tabulate-7, message=FALSE, warning = FALSE, eval = FALSE}
+```{r tabulate-7, message=FALSE, warning = FALSE}
 med_tib <- med_list %>%
@@
 
+```{r heat2-7}
+heatmaply(pm, cexRow = 0.2, cexCol = 0.2, seriate = "mean", symm = TRUE)
+```
+
 ```{r heat3-7}
@@
   heatmaply(cexRow = 0.2, cexCol = 0.2, symm = TRUE)
+```
+
+```{r heat4-7}
+(pm > pleio_threshold) %>%
+  apply(FUN = as.integer, MARGIN = 2) %>%
+  (function(x) {rownames(x) <- colnames(x); return(x)}) %>%
+  heatmaply(cexRow = 0.2, cexCol = 0.2, symm = TRUE, seriate = "mean")
 ```
@@
 
-```{r tabulate-11, message=FALSE, warning = FALSE, eval = FALSE}
+```{r tabulate-11, message=FALSE, warning = FALSE}
 med_tib <- med_list %>%
@@
 
+```{r heat2-11}
+heatmaply(pm, cexRow = 0.2, cexCol = 0.2, seriate = "mean", symm = TRUE, row_side_colors = scols)
+```
+
 ```{r heat3-11}
@@
   heatmaply(cexRow = 0.2, cexCol = 0.2, symm = TRUE, row_side_colors = scols)
+```
+
+```{r heat4-11}
+(pm > pleio_threshold) %>%
+  apply(FUN = as.integer, MARGIN = 2) %>%
+  (function(x) {rownames(x) <- colnames(x); return(x)}) %>%
+  heatmaply(cexRow = 0.2, cexCol = 0.2, symm = TRUE, seriate = "mean", row_side_colors = scols)
 ```
@@
 
-```{r tabulate-13, message=FALSE, warning = FALSE, eval = FALSE}
+```{r tabulate-13, message=FALSE, warning = FALSE}
 med_tib <- med_list %>%
@@
 
+```{r heat2-13}
+heatmaply(pm, cexRow = 0.2, cexCol = 0.2, seriate = "mean", symm = TRUE, row_side_colors = scols)
+```
+
 ```{r heat3-13}
@@
   heatmaply(cexRow = 0.2, cexCol = 0.2, symm = TRUE, row_side_colors = scols)
+```
+
+```{r heat4-13}
+(pm > pleio_threshold) %>%
+  apply(FUN = as.integer, MARGIN = 2) %>%
+  (function(x) {rownames(x) <- colnames(x); return(x)}) %>%
+  heatmaply(cexRow = 0.2, cexCol = 0.2, symm = TRUE, seriate = "mean", row_side_colors = scols)
 ```

### 2019-07-16 R/trunc2.R
added trunc2 function

@@ -1,0 +1,12 @@
+#' If the inputted number exceeds `threshold`, return `threshold`; otherwise, return the inputted value.
+#'
+#' @param input numeric vector to be modified
+#' @param threshold threshold value
+#' @export
+#' @return a numeric vector with length equaling that of `input`
+#' @examples
+#' trunc2(input = 1:10, threshold = 5)
+trunc2 <- function(input, threshold){
+  input[input > threshold] <- threshold
+  return(input)
+}

### 2019-07-16 analysis/paper/Chr11.Rmd
updated Rmd files to get modified medations.rds files

@@
                                              addcovar = addcovar
+                                             ) %>%
+        mutate(local_gene_id = colnames(hot_local)[j], 
+               nonlocal_gene_id = colnames(hot_nonlocal)[i]
                )

### 2019-07-16 analysis/paper/Chr2.Rmd
updated Rmd files to get modified medations.rds files

@@
   mutate(lod_diff = lod_without_med - lod_with_med, lod_diff_proportion = lod_diff / lod_without_med) %>%
-  mutate(local_gene_id = rep(colnames(hot_local), each = nnonlocal), nonlocal_gene_id = rep(colnames(hot_nonlocal), times = nlocal)) %>%
   left_join(local_annot2, by = c("local_gene_id" = "annot.id")) %>%

### 2019-07-16 analysis/paper/heatmaps-pleiotropy-statistics-subset-locals.Rmd
updated Rmd files to get modified medations.rds files

@@
 hot_nonlocal <- dataset.islet.rnaseq$expr[, colnames(dataset.islet.rnaseq$expr) %in% nonlocal_annot$annot.id]
-nnonlocal <- ncol(hot_nonlocal)
-nlocal <- ncol(hot_local)
+nnonlocal <- nrow(nonlocal_annot)
+nlocal <- nrow(local_annot2)
 ```
 
-```{r tabulate-2, message=FALSE, warning = FALSE, eval = FALSE}
+```{r tabulate-2, message=FALSE, warning = FALSE}
 med_tib <- med_list %>%
@@
 ```{r heat-pleio-2}
-heatmaply(pm, symm = TRUE, cexRow = 0.2, cexCol = 0.2, row_side_colors = scols)
+heatmaply(pm, symm = TRUE, cexRow = 0.2, cexCol = 0.2, row_side_colors = scols, limits = c(0,4))
 ```
@@
   heatmaply(cexRow = 0.2, cexCol = 0.2, symm = TRUE, row_side_colors = scols)
+```
+
+```{r}
+pm %>%
+  apply(FUN = hotspots::trunc2, MARGIN = 2, threshold = 5) %>%
+  (function(x) {rownames(x) <- colnames(x); return(x)}) %>%
+  heatmaply(cexRow = 0.2, cexCol = 0.2, symm = TRUE, row_side_colors = scols, limits = c(0, 5))
+
 ```

### 2019-07-17 README.Rmd
updated README files

@@
 
-The goal of hotspots is to ...
+The goal of hotspots is to analyze the Keller et al. (2018) expression trait hotspots.
 
@@
 
+## Keller, et al. (2018)'s hotspots table
+
+Here is Keller's Table S1 (from the Supplemental Documents).
+
+| Chr | position (Mb) |  number of transcripts  | candidate mediator  | number of genes with LOD difference > 1.5  |     
+| --- | -----------   |  ---------------------  |  ------------------ |  ----------------------------------------- |   
+| 2    | 165.5        |   147                   |  Hnf4a              |  88                                        | 
+| 5    | 146          |   182                   |  Pdx1               |  77                                        |
+| 7    | 46           |   123                   |  Fam83e             |  96                                        |
+| 11   | 71           |   126                   |  Sat2               |  115                                       |
+| 13   | 112.5        |   104                   |  Il6st              |  82                                        |
+
+Table: From [Keller et al.
+2018 GENETICS](https://www.genetics.org/content/209/1/335), [Supplementary table 1](https://figshare.com/articles/Supplemental_Material_for_Attie_et_al_2018_in_review_/5977459)
+
 ## Example
@@
 
-## Build the manuscript `paper.Rmd`
+## Build the manuscript `paper.Rnw`
 
-To render the manuscript `paper.Rmd`, into html, tex, and pdf, use a shell to enter the following text from the 'hotspots' directory:
+To render the manuscript `paper.Rnw`, into pdf, use a shell to enter the following text from the 'hotspots' directory:
 
@@
 
-If you're not using Docker, and have cloned the repository from Github, you can use the Makefile to render `paper.Rmd`. Just type `make` in the shell:
+If you're not using Docker, and have cloned the repository from Github, you can use the Makefile to render `paper.Rnw`. Just type `make` in the shell (from within the `hotspots` directory):
 

### 2019-07-17 analysis/paper/heatmaps-pleiotropy-statistics-subset-locals.Rmd
updated README files

@@
 hot_mid <- 165.5
+keller_mediator <- "Hnf4a"
 ```
@@
   dplyr::select(symbol.x, symbol.y, lod_without_med, lod_with_med, lod_diff, lod_diff_proportion, local_gene_id, nonlocal_gene_id) %>%
-  rename(local_symbol = symbol.x, nonlocal_symbol = symbol.y)
+  rename(local_symbol = symbol.x, nonlocal_symbol = symbol.y) %>%
+  filter(!is.na(local_symbol))
 ```
@@
   bind_rows(nonlocal_annot) %>%
-  select(annot.id, symbol)
+  dplyr::select(annot.id, symbol)
 (pleio_tib <- pleio_mixed_list %>%
@@
   rename(gene2_symbol = symbol) %>%
-  select(gene1_symbol, gene2_symbol, pleiotropy_lod)
+  dplyr::select(gene1_symbol, gene2_symbol, pleiotropy_lod)
 )  
@@
   tibble_to_matrix(symmetric = TRUE)
+```
+
+```{r}
+med_tib %>%
+  filter(local_symbol == keller_mediator) %>%
+  inner_join(pleio_tib, by = c("nonlocal_symbol" = "gene2_symbol", "local_symbol" = "gene1_symbol")) %>%
+  dplyr::select(1, 2, 9, 5, 6, 3, 4) %>%
+  arrange(pleiotropy_lod, desc(lod_diff)) %>%
+  filter(nonlocal_symbol %in% c("Clrn3", "Zfp541", "Cdhr2", "Sema4g", "Agap2", "Cacnb3", "Tff3"))
+
+```
+
+```{r}
+med_tib %>%
+  filter(local_symbol == "Ttpal") %>%
+  inner_join(pleio_tib, by = c("nonlocal_symbol" = "gene2_symbol", "local_symbol" = "gene1_symbol")) %>%
+  dplyr::select(1, 2, 9, 5, 6, 3, 4) %>%
+  arrange(desc(lod_diff))
+
 ```
@@
 scols <- 1L + (rownames(pm) %in% local_annot2$symbol)
+scols2 <- rownames(pm) %in% (med_tib %>% filter(local_symbol == keller_mediator, lod_diff > 1.5))$nonlocal_symbol
 ```
@@
 ```{r heat-pleio-2}
-heatmaply(pm, symm = TRUE, cexRow = 0.2, cexCol = 0.2, row_side_colors = scols, limits = c(0,4))
+heatmaply(pm, symm = TRUE, cexRow = 0.2, cexCol = 0.2, row_side_colors = tibble(scols, scols2), limits = c(0,4))
 ```
@@
 
-```{r}
-pm %>%
+```{r heat4-2}
+foo <- pm %>%
   apply(FUN = hotspots::trunc2, MARGIN = 2, threshold = 5) %>%
   (function(x) {rownames(x) <- colnames(x); return(x)}) %>%
-  heatmaply(cexRow = 0.2, cexCol = 0.2, symm = TRUE, row_side_colors = scols, limits = c(0, 5))
+  heatmaply(cexRow = 0.2, 
+            cexCol = 0.2, 
+            symm = TRUE, 
+            row_side_colors = tibble(scols, scols2), 
+            #reorderfun = identity,
+            limits = c(0, 5)
 
+            )
+
+```
+
+```{r heat8-2}
+ord <- pm %>%
+  dist() %>%
+  hclust() %>%
+  as.dendrogram() %>%
+  dendextend::seriate_dendrogram(dist(pm)) %>%
+  order.dendrogram()
+## order pm (and sidecols!) according to dendrogram from above
+pm[ord, ord] %>%
+  apply(FUN = hotspots::trunc2, MARGIN = 2, threshold = 5) %>%
+  heatmaply(Rowv = FALSE, 
+            Colv = FALSE, 
+            cexRow = 0.2, 
+            cexCol = 0.2, 
+            symm = TRUE, 
+            row_side_colors = tibble(scols, scols2)[ord, ],
+            limits = c(0, 5)
+            )
 ```
@@
 
+```{r heat7-2}
+pm_nonlocal %>%
+  apply(FUN = trunc2, MARGIN = 2, threshold = 5) %>%
+  (function(x) {rownames(x) <- colnames(x); return(x)}) %>%
+  heatmaply(symm = TRUE, cexRow = 0.3, cexCol = 0.3)
+```
+
 # Chr 5
@@
 hot_mid <- 146
+keller_mediator <- "Pdx1"
 ```
@@
 scols <- 1L + (rownames(pm) %in% local_annot2$symbol)
+scols2 <- rownames(pm) %in% (med_tib %>% filter(local_symbol == keller_mediator, lod_diff > 1.5))$nonlocal_symbol
 ```
 
+The vignette for `heatmaply` is quite helpful and includes discussion of its methods for ordering rows.
+
 ```{r heat-pleio-5}
-heatmaply(pm, symm = TRUE, cexRow = 0.2, cexCol = 0.2, row_side_colors = scols)
+heatmaply(pm, symm = TRUE, cexRow = 0.2, cexCol = 0.2, row_side_colors = tibble(scols, scols2), limits = c(0,4))
 ```
@@
   heatmaply(cexRow = 0.2, cexCol = 0.2, symm = TRUE, row_side_colors = scols)
+```
+
+```{r heat4-5}
+foo <- pm %>%
+  apply(FUN = hotspots::trunc2, MARGIN = 2, threshold = 5) %>%
+  (function(x) {rownames(x) <- colnames(x); return(x)}) %>%
+  heatmaply(cexRow = 0.2, 
+            cexCol = 0.2, 
+            symm = TRUE, 
+            row_side_colors = tibble(scols, scols2), 
+            #reorderfun = identity,
+            limits = c(0, 5)
+
+            )
+
+```
+
+```{r heat8-5}
+ord <- pm %>%
+  dist() %>%
+  hclust() %>%
+  as.dendrogram() %>%
+  dendextend::seriate_dendrogram(dist(pm)) %>%
+  order.dendrogram()
+## order pm (and sidecols!) according to dendrogram from above
+pm[ord, ord] %>%
+  apply(FUN = hotspots::trunc2, MARGIN = 2, threshold = 5) %>%
+  heatmaply(Rowv = FALSE, 
+            Colv = FALSE, 
+            cexRow = 0.2, 
+            cexCol = 0.2, 
+            symm = TRUE, 
+            row_side_colors = tibble(scols, scols2)[ord, ],
+            limits = c(0, 5)
+            )
 ```
@@
 
+```{r heat7-5}
+pm_nonlocal %>%
+  apply(FUN = trunc2, MARGIN = 2, threshold = 5) %>%
+  (function(x) {rownames(x) <- colnames(x); return(x)}) %>%
+  heatmaply(symm = TRUE, cexRow = 0.3, cexCol = 0.3)
+```
+
 # Chr 7
@@
 hot_mid <- 46
+keller_mediator <- "Fam83e"
 ```
@@
 scols <- 1L + (rownames(pm) %in% local_annot2$symbol)
+scols2 <- rownames(pm) %in% (med_tib %>% filter(local_symbol == keller_mediator, lod_diff > 1.5))$nonlocal_symbol
 ```
 
+The vignette for `heatmaply` is quite helpful and includes discussion of its methods for ordering rows.
+
 ```{r heat-pleio-7}
-heatmaply(pm, symm = TRUE, cexRow = 0.2, cexCol = 0.2)
+heatmaply(pm, symm = TRUE, cexRow = 0.2, cexCol = 0.2, row_side_colors = tibble(scols, scols2), limits = c(0,4))
 ```
@@
   (function(x) {rownames(x) <- colnames(x); return(x)}) %>%
-  heatmaply(cexRow = 0.2, cexCol = 0.2, symm = TRUE)
+  heatmaply(cexRow = 0.2, cexCol = 0.2, symm = TRUE, row_side_colors = scols)
+```
+
+```{r heat4-7}
+foo <- pm %>%
+  apply(FUN = hotspots::trunc2, MARGIN = 2, threshold = 5) %>%
+  (function(x) {rownames(x) <- colnames(x); return(x)}) %>%
+  heatmaply(cexRow = 0.2, 
+            cexCol = 0.2, 
+            symm = TRUE, 
+            row_side_colors = tibble(scols, scols2), 
+            #reorderfun = identity,
+            limits = c(0, 5)
+
+            )
+
+```
+
+```{r heat8-7}
+ord <- pm %>%
+  dist() %>%
+  hclust() %>%
+  as.dendrogram() %>%
+  dendextend::seriate_dendrogram(dist(pm)) %>%
+  order.dendrogram()
+## order pm (and sidecols!) according to dendrogram from above
+pm[ord, ord] %>%
+  apply(FUN = hotspots::trunc2, MARGIN = 2, threshold = 5) %>%
+  heatmaply(Rowv = FALSE, 
+            Colv = FALSE, 
+            cexRow = 0.2, 
+            cexCol = 0.2, 
+            symm = TRUE, 
+            row_side_colors = tibble(scols, scols2)[ord, ],
+            limits = c(0, 5)
+            )
 ```
@@
 
+```{r heat7-7}
+pm_nonlocal %>%
+  apply(FUN = trunc2, MARGIN = 2, threshold = 5) %>%
+  (function(x) {rownames(x) <- colnames(x); return(x)}) %>%
+  heatmaply(symm = TRUE, cexRow = 0.3, cexCol = 0.3)
+```
+
 # Chr 11
@@
 hot_mid <- 71
+keller_mediator <- "Sat2"
 ```
@@
 scols <- 1L + (rownames(pm) %in% local_annot2$symbol)
+scols2 <- rownames(pm) %in% (med_tib %>% filter(local_symbol == keller_mediator, lod_diff > 1.5))$nonlocal_symbol
 ```
 
+The vignette for `heatmaply` is quite helpful and includes discussion of its methods for ordering rows.
+
 ```{r heat-pleio-11}
-heatmaply(pm, symm = TRUE, cexRow = 0.2, cexCol = 0.2, row_side_colors = scols)
+heatmaply(pm, symm = TRUE, cexRow = 0.2, cexCol = 0.2, row_side_colors = tibble(scols, scols2), limits = c(0,4))
 ```
@@
   heatmaply(cexRow = 0.2, cexCol = 0.2, symm = TRUE, row_side_colors = scols)
+```
+
+```{r heat4-11}
+foo <- pm %>%
+  apply(FUN = hotspots::trunc2, MARGIN = 2, threshold = 5) %>%
+  (function(x) {rownames(x) <- colnames(x); return(x)}) %>%
+  heatmaply(cexRow = 0.2, 
+            cexCol = 0.2, 
+            symm = TRUE, 
+            row_side_colors = tibble(scols, scols2), 
+            #reorderfun = identity,
+            limits = c(0, 5)
+
+            )
+
+```
+
+```{r heat8-11}
+ord <- pm %>%
+  dist() %>%
+  hclust() %>%
+  as.dendrogram() %>%
+  dendextend::seriate_dendrogram(dist(pm)) %>%
+  order.dendrogram()
+## order pm (and sidecols!) according to dendrogram from above
+pm[ord, ord] %>%
+  apply(FUN = hotspots::trunc2, MARGIN = 2, threshold = 5) %>%
+  heatmaply(Rowv = FALSE, 
+            Colv = FALSE, 
+            cexRow = 0.2, 
+            cexCol = 0.2, 
+            symm = TRUE, 
+            row_side_colors = tibble(scols, scols2)[ord, ],
+            limits = c(0, 5)
+            )
 ```
@@
 
+```{r heat7-11}
+pm_nonlocal %>%
+  apply(FUN = trunc2, MARGIN = 2, threshold = 5) %>%
+  (function(x) {rownames(x) <- colnames(x); return(x)}) %>%
+  heatmaply(symm = TRUE, cexRow = 0.3, cexCol = 0.3)
+```
+
 # Chr 13
@@
 hot_mid <- 112.5
+keller_mediator <- "Il6st"
 ```
@@
 scols <- 1L + (rownames(pm) %in% local_annot2$symbol)
+scols2 <- rownames(pm) %in% (med_tib %>% filter(local_symbol == keller_mediator, lod_diff > 1.5))$nonlocal_symbol
 ```
 
+The vignette for `heatmaply` is quite helpful and includes discussion of its methods for ordering rows.
+
 ```{r heat-pleio-13}
-heatmaply(pm, symm = TRUE, cexRow = 0.2, cexCol = 0.2, row_side_colors = scols)
+heatmaply(pm, symm = TRUE, cexRow = 0.2, cexCol = 0.2, row_side_colors = tibble(scols, scols2), limits = c(0,4))
 ```
@@
   heatmaply(cexRow = 0.2, cexCol = 0.2, symm = TRUE, row_side_colors = scols)
+```
+
+```{r heat4-13}
+foo <- pm %>%
+  apply(FUN = hotspots::trunc2, MARGIN = 2, threshold = 5) %>%
+  (function(x) {rownames(x) <- colnames(x); return(x)}) %>%
+  heatmaply(cexRow = 0.2, 
+            cexCol = 0.2, 
+            symm = TRUE, 
+            row_side_colors = tibble(scols, scols2), 
+            #reorderfun = identity,
+            limits = c(0, 5)
+
+            )
+
+```
+
+```{r heat8-13}
+ord <- pm %>%
+  dist() %>%
+  hclust() %>%
+  as.dendrogram() %>%
+  dendextend::seriate_dendrogram(dist(pm)) %>%
+  order.dendrogram()
+## order pm (and sidecols!) according to dendrogram from above
+pm[ord, ord] %>%
+  apply(FUN = hotspots::trunc2, MARGIN = 2, threshold = 5) %>%
+  heatmaply(Rowv = FALSE, 
+            Colv = FALSE, 
+            cexRow = 0.2, 
+            cexCol = 0.2, 
+            symm = TRUE, 
+            row_side_colors = tibble(scols, scols2)[ord, ],
+            limits = c(0, 5)
+            )
 ```
@@
 ```
+
+```{r heat7-13}
+pm_nonlocal %>%
+  apply(FUN = trunc2, MARGIN = 2, threshold = 5) %>%
+  (function(x) {rownames(x) <- colnames(x); return(x)}) %>%
+  heatmaply(symm = TRUE, cexRow = 0.3, cexCol = 0.3)
+```

### 2019-07-18 analysis/paper/heatmaps-pleiotropy-statistics-subset-locals.Rmd
added ref.label to R code chunks

--- 2019-07-16 analysis/paper/heatmaps-pleiotropy-statistics-subset-locals.Rmd
@@
 ---
-title: "Heatmaps of pleiotropy test statistics - all nonlocals and all locals with LOD above 15"
+title: "Heatmaps of pleiotropy test statistics"
 author: "Frederick Boehm"
-date: "7/16/2019"
+date: "7/17/2019"
 output: 
@@
 The vignette for `heatmaply` is quite helpful and includes discussion of its methods for ordering rows.
+For a discussion of ordering of rows in `heatmaply`, see this issue & its comments: <https://github.com/talgalili/heatmaply/issues/153>
 
-```{r heat-pleio-2}
-heatmaply(pm, symm = TRUE, cexRow = 0.2, cexCol = 0.2, row_side_colors = scols, limits = c(0,4))
+```{r heat8-2}
+ord <- pm %>%
+  dist() %>%
+  hclust() %>%
+  as.dendrogram() %>%
+  dendextend::seriate_dendrogram(dist(pm)) %>%
+  order.dendrogram()
+## order pm (and sidecols!) according to dendrogram from above
 ```
 
-```{r heat3-2}
-(pm > pleio_threshold) %>%
-  apply(FUN = as.integer, MARGIN = 2) %>%
-  (function(x) {rownames(x) <- colnames(x); return(x)}) %>%
-  heatmaply(cexRow = 0.2, cexCol = 0.2, symm = TRUE, row_side_colors = scols)
+```{r make_scols-extra-2}
+scols_extra <- purrr::map(.x = local_annot2$symbol, 
+                                       .f = function(x) {rownames(pm) %in% (med_tib %>% filter(local_symbol == x, lod_diff > 1.5))$nonlocal_symbol}
+) %>% 
+  bind_cols() %>%
+  (function(x){colnames(x) <- local_annot2$symbol; x}) %>%
+  mutate(local = scols)
 ```
 
-```{r}
-pm %>%
-  apply(FUN = hotspots::trunc2, MARGIN = 2, threshold = 5) %>%
-  (function(x) {rownames(x) <- colnames(x); return(x)}) %>%
-  heatmaply(cexRow = 0.2, cexCol = 0.2, symm = TRUE, row_side_colors = scols, limits = c(0, 5))
+Instead of adding a column for *every* local trait, let's only add a column for those local traits that mediate at least, say, 4 traits at the 1.5 LOD difference value.
 
+```{r count_med}
+n_med_tr <- 4
+(n_med_per_local <- med_tib %>%
+  group_by(local_symbol) %>%
+  filter(lod_diff > 1.5) %>%
+  count() %>%
+  filter(n >= n_med_tr)
+  )
+scols <- scols_extra %>%
+  dplyr::select(local, n_med_per_local$local_symbol)
 ```
 
-## Heatmaps for matrix containing only nonlocal traits
-
-```{r heat5-2}
-pm_nonlocal <- pleio_tib %>%
-  filter(gene1_symbol %in% nonlocal_annot$symbol, gene2_symbol %in% nonlocal_annot$symbol) %>%
-  tibble_to_matrix(symmetric = TRUE)
-pm_nonlocal %>%
-  heatmaply(symm = TRUE, cexRow = 0.3, cexCol = 0.3)
-
-```
-
-```{r heat6-2}
-(pm_nonlocal > pleio_threshold) %>%
-  apply(FUN = as.integer, MARGIN = 2) %>%
-  (function(x) {rownames(x) <- colnames(x); return(x)}) %>%
-  heatmaply(symm = TRUE, cexRow = 0.3, cexCol = 0.3)
+```{r heat9-2}
+pm[ord, ord] %>%
+  apply(FUN = hotspots::trunc2, MARGIN = 2, threshold = 5) %>%
+  heatmaply(Rowv = FALSE, 
+            Colv = FALSE, 
+            cexRow = 0.2, 
+            cexCol = 0.2, 
+            symm = TRUE, 
+            row_side_colors = scols[ord, ],
+            limits = c(0, 5)
+            )
 ```
@@
 hot_mid <- 146
+keller_mediator <- "Pdx1"
 ```
 
-```{r more_setup-5}
-hot_lower <- hot_mid - 2
-hot_upper <- hot_mid + 2
-fp <- paste0("../figures/Chr", hot_chr, "-")
-knitr::opts_chunk$set(fig.path = fp)
+```{r more_setup-5, ref.label = "more_setup-2"}
 ```
@@
 
-Let's examine the Chr `r hot_chr` files for pleiotropy results.
-
-```{r readRDS-5}
-pleio_mixed_list <- readRDS(paste0("_cache/Chr", hot_chr, "_pleio.rds"))
-pleio_local_list <- readRDS(paste0("_cache/Chr", hot_chr, "_pleio_local.rds"))
-pleio_nonlocal_list <- readRDS(paste0("_cache/Chr", hot_chr, "_pleio_nonlocal.rds"))
-med_list <- readRDS(paste0("_cache/Chr", hot_chr, "_mediations.rds"))
+```{r readRDS-5, ref.label="readRDS-2"}
 ```
 
-```{r make_local_annot-5}
-local_annot <- dataset.islet.rnaseq$lod.peaks %>%
-  filter(chrom == hot_chr, pos <= hot_upper, pos >= hot_lower) %>%
-  left_join(y = dataset.islet.rnaseq$annots, by = c("annot.id" = "gene_id")) %>%
-  filter(!duplicated(annot.id)) %>%
-  filter(lod >= ult) %>%
-  mutate(local_tx = chr == hot_chr) %>%
-  filter(local_tx)
+```{r make_local_annot-5, ref.label="make_local_annot-2"}
 ```
 
-```{r make_nonlocal_annot-5}
-nonlocal_annot <- dataset.islet.rnaseq$lod.peaks %>%
-  filter(chrom == hot_chr, pos <= hot_upper, pos >= hot_lower) %>%
-  left_join(y = dataset.islet.rnaseq$annots, by = c("annot.id" = "gene_id")) %>%
-  filter(!duplicated(annot.id)) %>%
-#  filter(lod >= ult) %>% #genomewide threshold
-  mutate(local_tx = chr == hot_chr) %>%
-  filter(!local_tx) %>% 
-  filter(!is.na(hotspot))
+```{r make_nonlocal_annot-5, ref.label = "make_nonlocal_annot-2"}
 ```
 
-```{r, make_local_annot2-5}
-local_annot2 <- local_annot %>%
-  filter(middle <= hot_upper + 5, middle >= hot_lower - 5) %>%
-  arrange(desc(lod))
+```{r, make_local_annot2-5, ref.label = "make_local_annot2-2"}
 ```
 
-```{r subset_expr_matrix-5}
-hot_local <- dataset.islet.rnaseq$expr[, colnames(dataset.islet.rnaseq$expr) %in% local_annot2$annot.id]
-hot_nonlocal <- dataset.islet.rnaseq$expr[, colnames(dataset.islet.rnaseq$expr) %in% nonlocal_annot$annot.id]
-nnonlocal <- ncol(hot_nonlocal)
-nlocal <- ncol(hot_local)
+```{r subset_expr_matrix-5, ref.label = "subset_expr_matrix-2"}
 ```
 
-```{r tabulate-5, message=FALSE, warning = FALSE, eval = FALSE}
-med_tib <- med_list %>%
-  purrr::map(bind_rows) %>%
-  bind_rows() %>%
-  as_tibble() %>%
-  rename(lod_without_med = lod_no_med, lod_with_med = lod_med) %>%
-  mutate(lod_diff = lod_without_med - lod_with_med, lod_diff_proportion = lod_diff / lod_without_med) %>%
-  mutate(local_gene_id = rep(colnames(hot_local), each = nnonlocal), nonlocal_gene_id = rep(colnames(hot_nonlocal), times = nlocal)) %>%
-  left_join(local_annot2, by = c("local_gene_id" = "annot.id")) %>%
-  left_join(nonlocal_annot, by = c("nonlocal_gene_id" = "annot.id")) %>%
-  dplyr::select(symbol.x, symbol.y, lod_without_med, lod_with_med, lod_diff, lod_diff_proportion, local_gene_id, nonlocal_gene_id) %>%
-  rename(local_symbol = symbol.x, nonlocal_symbol = symbol.y)
+```{r tabulate-5, ref.label = "tabulate-2", message=FALSE, warning = FALSE}
 ```
 
-```{r make_pleio_tib-5}
-hot_annot <- local_annot2 %>%
-  bind_rows(nonlocal_annot) %>%
-  select(annot.id, symbol)
-(pleio_tib <- pleio_mixed_list %>%
-  purrr::map(.f = function(x) rename(x, gene1_id = local_gene_id, gene2_id = nonlocal_gene_id)) %>%
-  c(pleio_local_list, pleio_nonlocal_list) %>%
-  bind_rows() %>%
-  rename(pleiotropy_lod = lod) %>%
-  inner_join(hot_annot, by = c("gene1_id" = "annot.id")) %>%
-  rename(gene1_symbol = symbol) %>%
-  inner_join(hot_annot, by = c("gene2_id" = "annot.id")) %>%
-  rename(gene2_symbol = symbol) %>%
-  select(gene1_symbol, gene2_symbol, pleiotropy_lod)
-)  
+```{r make_pleio_tib-5, ref.label="make_pleio_tib-2"}
 ```
 
-```{r make_pm-5}
-pm <- pleio_tib %>%
-  tibble_to_matrix(symmetric = TRUE)
+```{r make_pm-5, ref.label = "make_pm-2"}
 ```
@@
 
-```{r make_scols-5}
-scols <- 1L + (rownames(pm) %in% local_annot2$symbol)
+```{r make_scols-5, ref.label="make_scols-2"}
 ```
 
-```{r heat-pleio-5}
-heatmaply(pm, symm = TRUE, cexRow = 0.2, cexCol = 0.2, row_side_colors = scols)
+```{r heat8-5, ref.label="heat8-2"}
 ```
 
-```{r heat3-5}
-(pm > pleio_threshold) %>%
-  apply(FUN = as.integer, MARGIN = 2) %>%
-  (function(x) {rownames(x) <- colnames(x); return(x)}) %>%
-  heatmaply(cexRow = 0.2, cexCol = 0.2, symm = TRUE, row_side_colors = scols)
+```{r ref.label = "make_scols-extra-2"}
 ```
 
-## Heatmaps for matrix containing only nonlocal traits
-
-```{r heat5-5}
-pm_nonlocal <- pleio_tib %>%
-  filter(gene1_symbol %in% nonlocal_annot$symbol, gene2_symbol %in% nonlocal_annot$symbol) %>%
-  tibble_to_matrix(symmetric = TRUE)
-pm_nonlocal %>%
-  heatmaply(symm = TRUE, cexRow = 0.3, cexCol = 0.3)
-
+```{r ref.label = "count_med"}
 ```
 
-```{r heat6-5}
-(pm_nonlocal > pleio_threshold) %>%
-  apply(FUN = as.integer, MARGIN = 2) %>%
-  (function(x) {rownames(x) <- colnames(x); return(x)}) %>%
-  heatmaply(symm = TRUE, cexRow = 0.3, cexCol = 0.3)
+```{r ref.label = "heat9-2"}
 ```
@@
 hot_mid <- 46
+keller_mediator <- "Fam83e"
 ```
 
-```{r more_setup-7}
-hot_lower <- hot_mid - 2
-hot_upper <- hot_mid + 2
-fp <- paste0("../figures/Chr", hot_chr, "-")
-knitr::opts_chunk$set(fig.path = fp)
+```{r ref.label = "more_setup-2"}
 ```
@@
 
-Let's examine the Chr `r hot_chr` files for pleiotropy results.
-
-```{r readRDS-7}
-pleio_mixed_list <- readRDS(paste0("_cache/Chr", hot_chr, "_pleio.rds"))
-pleio_local_list <- readRDS(paste0("_cache/Chr", hot_chr, "_pleio_local.rds"))
-pleio_nonlocal_list <- readRDS(paste0("_cache/Chr", hot_chr, "_pleio_nonlocal.rds"))
-med_list <- readRDS(paste0("_cache/Chr", hot_chr, "_mediations.rds"))
+```{r ref.label="readRDS-2"}
 ```
 
-```{r make_local_annot-7}
-local_annot <- dataset.islet.rnaseq$lod.peaks %>%
-  filter(chrom == hot_chr, pos <= hot_upper, pos >= hot_lower) %>%
-  left_join(y = dataset.islet.rnaseq$annots, by = c("annot.id" = "gene_id")) %>%
-  filter(!duplicated(annot.id)) %>%
-  filter(lod >= ult) %>%
-  mutate(local_tx = chr == hot_chr) %>%
-  filter(local_tx)
+```{r ref.label="make_local_annot-2"}
 ```
 
-```{r make_nonlocal_annot-7}
-nonlocal_annot <- dataset.islet.rnaseq$lod.peaks %>%
-  filter(chrom == hot_chr, pos <= hot_upper, pos >= hot_lower) %>%
-  left_join(y = dataset.islet.rnaseq$annots, by = c("annot.id" = "gene_id")) %>%
-  filter(!duplicated(annot.id)) %>%
-#  filter(lod >= ult) %>% #genomewide threshold
-  mutate(local_tx = chr == hot_chr) %>%
-  filter(!local_tx) %>% 
-  filter(!is.na(hotspot))
+```{r ref.label = "make_nonlocal_annot-2"}
 ```
 
-```{r, make_local_annot2-7}
-local_annot2 <- local_annot %>%
-  filter(middle <= hot_upper + 5, middle >= hot_lower - 5) %>%
-  arrange(desc(lod))
+```{r ref.label = "make_local_annot2-2"}
 ```
 
-```{r subset_expr_matrix-7}
-hot_local <- dataset.islet.rnaseq$expr[, colnames(dataset.islet.rnaseq$expr) %in% local_annot2$annot.id]
-hot_nonlocal <- dataset.islet.rnaseq$expr[, colnames(dataset.islet.rnaseq$expr) %in% nonlocal_annot$annot.id]
-nnonlocal <- ncol(hot_nonlocal)
-nlocal <- ncol(hot_local)
+```{r ref.label = "subset_expr_matrix-2"}
 ```
 
-```{r tabulate-7, message=FALSE, warning = FALSE, eval = FALSE}
-med_tib <- med_list %>%
-  purrr::map(bind_rows) %>%
-  bind_rows() %>%
-  as_tibble() %>%
-  rename(lod_without_med = lod_no_med, lod_with_med = lod_med) %>%
-  mutate(lod_diff = lod_without_med - lod_with_med, lod_diff_proportion = lod_diff / lod_without_med) %>%
-  mutate(local_gene_id = rep(colnames(hot_local), each = nnonlocal), nonlocal_gene_id = rep(colnames(hot_nonlocal), times = nlocal)) %>%
-  left_join(local_annot2, by = c("local_gene_id" = "annot.id")) %>%
-  left_join(nonlocal_annot, by = c("nonlocal_gene_id" = "annot.id")) %>%
-  dplyr::select(symbol.x, symbol.y, lod_without_med, lod_with_med, lod_diff, lod_diff_proportion, local_gene_id, nonlocal_gene_id) %>%
-  rename(local_symbol = symbol.x, nonlocal_symbol = symbol.y)
+```{r ref.label = "tabulate-2", message=FALSE, warning = FALSE}
 ```
 
-```{r make_pleio_tib-7}
-hot_annot <- local_annot2 %>%
-  bind_rows(nonlocal_annot) %>%
-  select(annot.id, symbol)
-(pleio_tib <- pleio_mixed_list %>%
-  purrr::map(.f = function(x) rename(x, gene1_id = local_gene_id, gene2_id = nonlocal_gene_id)) %>%
-  c(pleio_local_list, pleio_nonlocal_list) %>%
-  bind_rows() %>%
-  rename(pleiotropy_lod = lod) %>%
-  inner_join(hot_annot, by = c("gene1_id" = "annot.id")) %>%
-  rename(gene1_symbol = symbol) %>%
-  inner_join(hot_annot, by = c("gene2_id" = "annot.id")) %>%
-  rename(gene2_symbol = symbol) %>%
-  select(gene1_symbol, gene2_symbol, pleiotropy_lod)
-)  
+```{r ref.label="make_pleio_tib-2"}
 ```
 
-```{r make_pm-7}
-pm <- pleio_tib %>%
-  tibble_to_matrix(symmetric = TRUE)
+```{r ref.label = "make_pm-2"}
 ```
@@
 
-```{r make_scols-7}
-scols <- 1L + (rownames(pm) %in% local_annot2$symbol)
+```{r ref.label="make_scols-2"}
 ```
 
-```{r heat-pleio-7}
-heatmaply(pm, symm = TRUE, cexRow = 0.2, cexCol = 0.2)
+```{r ref.label="heat8-2"}
 ```
 
-```{r heat3-7}
-(pm > pleio_threshold) %>%
-  apply(FUN = as.integer, MARGIN = 2) %>%
-  (function(x) {rownames(x) <- colnames(x); return(x)}) %>%
-  heatmaply(cexRow = 0.2, cexCol = 0.2, symm = TRUE)
+```{r ref.label = "make_scols-extra-2"}
 ```
 
-## Heatmaps for matrix containing only nonlocal traits
-
-```{r heat5-7}
-pm_nonlocal <- pleio_tib %>%
-  filter(gene1_symbol %in% nonlocal_annot$symbol, gene2_symbol %in% nonlocal_annot$symbol) %>%
-  tibble_to_matrix(symmetric = TRUE)
-pm_nonlocal %>%
-  heatmaply(symm = TRUE, cexRow = 0.3, cexCol = 0.3)
-
+```{r ref.label = "count_med"}
 ```
 
-```{r heat6-7}
-(pm_nonlocal > pleio_threshold) %>%
-  apply(FUN = as.integer, MARGIN = 2) %>%
-  (function(x) {rownames(x) <- colnames(x); return(x)}) %>%
-  heatmaply(symm = TRUE, cexRow = 0.3, cexCol = 0.3)
+```{r ref.label = "heat9-2"}
 ```
@@
 hot_mid <- 71
+keller_mediator <- "Sat2"
 ```
 
-```{r more_setup-11}
-hot_lower <- hot_mid - 2
-hot_upper <- hot_mid + 2
-fp <- paste0("../figures/Chr", hot_chr, "-")
-knitr::opts_chunk$set(fig.path = fp)
+```{r ref.label = "more_setup-2"}
 ```
@@
 
-Let's examine the Chr `r hot_chr` files for pleiotropy results.
-
-```{r readRDS-11}
-pleio_mixed_list <- readRDS("_cache/Chr11_pleio.rds")
-pleio_local_list <- readRDS("_cache/Chr11_pleio_local.rds")
-pleio_nonlocal_list <- readRDS("_cache/Chr11_pleio_nonlocal.rds")
-med_list <- readRDS("_cache/Chr11_mediations.rds")
+```{r ref.label="readRDS-2"}
 ```
 
-```{r make_local_annot-11}
-local_annot <- dataset.islet.rnaseq$lod.peaks %>%
-  filter(chrom == hot_chr, pos <= hot_upper, pos >= hot_lower) %>%
-  left_join(y = dataset.islet.rnaseq$annots, by = c("annot.id" = "gene_id")) %>%
-  filter(!duplicated(annot.id)) %>%
-  filter(lod >= ult) %>%
-  mutate(local_tx = chr == hot_chr) %>%
-  filter(local_tx)
+```{r ref.label="make_local_annot-2"}
 ```
 
-```{r make_nonlocal_annot-11}
-nonlocal_annot <- dataset.islet.rnaseq$lod.peaks %>%
-  filter(chrom == hot_chr, pos <= hot_upper, pos >= hot_lower) %>%
-  left_join(y = dataset.islet.rnaseq$annots, by = c("annot.id" = "gene_id")) %>%
-  filter(!duplicated(annot.id)) %>%
-#  filter(lod >= ult) %>% #genomewide threshold
-  mutate(local_tx = chr == hot_chr) %>%
-  filter(!local_tx) %>% 
-  filter(!is.na(hotspot))
+```{r ref.label = "make_nonlocal_annot-2"}
 ```
 
-```{r, make_local_annot2-11}
-local_annot2 <- local_annot %>%
-  filter(middle <= hot_upper + 5, middle >= hot_lower - 5) %>%
-  arrange(desc(lod))
+```{r ref.label = "make_local_annot2-2"}
 ```
 
-```{r subset_expr_matrix-11}
-hot_local <- dataset.islet.rnaseq$expr[, colnames(dataset.islet.rnaseq$expr) %in% local_annot2$annot.id]
-hot_nonlocal <- dataset.islet.rnaseq$expr[, colnames(dataset.islet.rnaseq$expr) %in% nonlocal_annot$annot.id]
-nnonlocal <- ncol(hot_nonlocal)
-nlocal <- ncol(hot_local)
+```{r ref.label = "subset_expr_matrix-2"}
 ```
 
-```{r tabulate-11, message=FALSE, warning = FALSE, eval = FALSE}
-med_tib <- med_list %>%
-  purrr::map(bind_rows) %>%
-  bind_rows() %>%
-  as_tibble() %>%
-  rename(lod_without_med = lod_no_med, lod_with_med = lod_med) %>%
-  mutate(lod_diff = lod_without_med - lod_with_med, lod_diff_proportion = lod_diff / lod_without_med) %>%
-  mutate(local_gene_id = rep(colnames(hot_local), each = nnonlocal), nonlocal_gene_id = rep(colnames(hot_nonlocal), times = nlocal)) %>%
-  left_join(local_annot2, by = c("local_gene_id" = "annot.id")) %>%
-  left_join(nonlocal_annot, by = c("nonlocal_gene_id" = "annot.id")) %>%
-  dplyr::select(symbol.x, symbol.y, lod_without_med, lod_with_med, lod_diff, lod_diff_proportion, local_gene_id, nonlocal_gene_id) %>%
-  rename(local_symbol = symbol.x, nonlocal_symbol = symbol.y)
+```{r ref.label = "tabulate-2", message=FALSE, warning = FALSE}
 ```
 
-```{r make_pleio_tib-11}
-hot_annot <- local_annot2 %>%
-  bind_rows(nonlocal_annot) %>%
-  select(annot.id, symbol)
-(pleio_tib <- pleio_mixed_list %>%
-  purrr::map(.f = function(x) rename(x, gene1_id = local_gene_id, gene2_id = nonlocal_gene_id)) %>%
-  c(pleio_local_list, pleio_nonlocal_list) %>%
-  bind_rows() %>%
-  rename(pleiotropy_lod = lod) %>%
-  inner_join(hot_annot, by = c("gene1_id" = "annot.id")) %>%
-  rename(gene1_symbol = symbol) %>%
-  inner_join(hot_annot, by = c("gene2_id" = "annot.id")) %>%
-  rename(gene2_symbol = symbol) %>%
-  select(gene1_symbol, gene2_symbol, pleiotropy_lod)
-)  
+```{r ref.label="make_pleio_tib-2"}
 ```
 
-```{r make_pm-11}
-pm <- pleio_tib %>%
-  tibble_to_matrix(symmetric = TRUE)
+```{r ref.label = "make_pm-2"}
 ```
@@
 
-```{r make_scols-11}
-scols <- 1L + (rownames(pm) %in% local_annot2$symbol)
+```{r ref.label="make_scols-2"}
 ```
 
-```{r heat-pleio-11}
-heatmaply(pm, symm = TRUE, cexRow = 0.2, cexCol = 0.2, row_side_colors = scols)
+```{r ref.label="heat8-2"}
 ```
 
-```{r heat3-11}
-(pm > pleio_threshold) %>%
-  apply(FUN = as.integer, MARGIN = 2) %>%
-  (function(x) {rownames(x) <- colnames(x); return(x)}) %>%
-  heatmaply(cexRow = 0.2, cexCol = 0.2, symm = TRUE, row_side_colors = scols)
+```{r ref.label = "make_scols-extra-2"}
 ```
 
-## Heatmaps for matrix containing only nonlocal traits
-
-```{r heat5-11}
-pm_nonlocal <- pleio_tib %>%
-  filter(gene1_symbol %in% nonlocal_annot$symbol, gene2_symbol %in% nonlocal_annot$symbol) %>%
-  tibble_to_matrix(symmetric = TRUE)
-pm_nonlocal %>%
-  heatmaply(symm = TRUE, cexRow = 0.3, cexCol = 0.3)
-
+```{r ref.label = "count_med"}
 ```
 
-```{r heat6-11}
-(pm_nonlocal > pleio_threshold) %>%
-  apply(FUN = as.integer, MARGIN = 2) %>%
-  (function(x) {rownames(x) <- colnames(x); return(x)}) %>%
-  heatmaply(symm = TRUE, cexRow = 0.3, cexCol = 0.3)
+```{r ref.label = "heat9-2"}
 ```
@@
 hot_mid <- 112.5
+keller_mediator <- "Il6st"
 ```
 
-```{r more_setup-13}
-hot_lower <- hot_mid - 2
-hot_upper <- hot_mid + 2
-fp <- paste0("../figures/Chr", hot_chr, "-")
-knitr::opts_chunk$set(fig.path = fp)
+```{r ref.label = "more_setup-2"}
 ```
@@
 
-Let's examine the Chr `r hot_chr` files for pleiotropy results.
-
-```{r readRDS-13}
-pleio_mixed_list <- readRDS(paste0("_cache/Chr", hot_chr, "_pleio.rds"))
-pleio_local_list <- readRDS(paste0("_cache/Chr", hot_chr, "_pleio_local.rds"))
-pleio_nonlocal_list <- readRDS(paste0("_cache/Chr", hot_chr, "_pleio_nonlocal.rds"))
-med_list <- readRDS(paste0("_cache/Chr", hot_chr, "_mediations.rds"))
+```{r ref.label="readRDS-2"}
 ```
 
-```{r make_local_annot-13}
-local_annot <- dataset.islet.rnaseq$lod.peaks %>%
-  filter(chrom == hot_chr, pos <= hot_upper, pos >= hot_lower) %>%
-  left_join(y = dataset.islet.rnaseq$annots, by = c("annot.id" = "gene_id")) %>%
-  filter(!duplicated(annot.id)) %>%
-  filter(lod >= ult) %>%
-  mutate(local_tx = chr == hot_chr) %>%
-  filter(local_tx)
+```{r ref.label="make_local_annot-2"}
 ```
 
-```{r make_nonlocal_annot-13}
-nonlocal_annot <- dataset.islet.rnaseq$lod.peaks %>%
-  filter(chrom == hot_chr, pos <= hot_upper, pos >= hot_lower) %>%
-  left_join(y = dataset.islet.rnaseq$annots, by = c("annot.id" = "gene_id")) %>%
-  filter(!duplicated(annot.id)) %>%
-#  filter(lod >= ult) %>% #genomewide threshold
-  mutate(local_tx = chr == hot_chr) %>%
-  filter(!local_tx) %>% 
-  filter(!is.na(hotspot))
+```{r ref.label = "make_nonlocal_annot-2"}
 ```
 
-```{r, make_local_annot2-13}
-local_annot2 <- local_annot %>%
-  filter(middle <= hot_upper + 5, middle >= hot_lower - 5) %>%
-  arrange(desc(lod))
+```{r ref.label = "make_local_annot2-2"}
 ```
 
-```{r subset_expr_matrix-13}
-hot_local <- dataset.islet.rnaseq$expr[, colnames(dataset.islet.rnaseq$expr) %in% local_annot2$annot.id]
-hot_nonlocal <- dataset.islet.rnaseq$expr[, colnames(dataset.islet.rnaseq$expr) %in% nonlocal_annot$annot.id]
-nnonlocal <- ncol(hot_nonlocal)
-nlocal <- ncol(hot_local)
+```{r ref.label = "subset_expr_matrix-2"}
 ```
 
-```{r tabulate-13, message=FALSE, warning = FALSE, eval = FALSE}
-med_tib <- med_list %>%
-  purrr::map(bind_rows) %>%
-  bind_rows() %>%
-  as_tibble() %>%
-  rename(lod_without_med = lod_no_med, lod_with_med = lod_med) %>%
-  mutate(lod_diff = lod_without_med - lod_with_med, lod_diff_proportion = lod_diff / lod_without_med) %>%
-  mutate(local_gene_id = rep(colnames(hot_local), each = nnonlocal), nonlocal_gene_id = rep(colnames(hot_nonlocal), times = nlocal)) %>%
-  left_join(local_annot2, by = c("local_gene_id" = "annot.id")) %>%
-  left_join(nonlocal_annot, by = c("nonlocal_gene_id" = "annot.id")) %>%
-  dplyr::select(symbol.x, symbol.y, lod_without_med, lod_with_med, lod_diff, lod_diff_proportion, local_gene_id, nonlocal_gene_id) %>%
-  rename(local_symbol = symbol.x, nonlocal_symbol = symbol.y)
+```{r ref.label = "tabulate-2", message=FALSE, warning = FALSE}
 ```
 
-```{r make_pleio_tib-13}
-hot_annot <- local_annot2 %>%
-  bind_rows(nonlocal_annot) %>%
-  select(annot.id, symbol)
-(pleio_tib <- pleio_mixed_list %>%
-  purrr::map(.f = function(x) rename(x, gene1_id = local_gene_id, gene2_id = nonlocal_gene_id)) %>%
-  c(pleio_local_list, pleio_nonlocal_list) %>%
-  bind_rows() %>%
-  rename(pleiotropy_lod = lod) %>%
-  inner_join(hot_annot, by = c("gene1_id" = "annot.id")) %>%
-  rename(gene1_symbol = symbol) %>%
-  inner_join(hot_annot, by = c("gene2_id" = "annot.id")) %>%
-  rename(gene2_symbol = symbol) %>%
-  select(gene1_symbol, gene2_symbol, pleiotropy_lod)
-)  
+```{r ref.label="make_pleio_tib-2"}
 ```
 
-```{r make_pm-13}
-pm <- pleio_tib %>%
-  tibble_to_matrix(symmetric = TRUE)
+```{r ref.label = "make_pm-2"}
 ```
@@
 
-```{r make_scols-13}
-scols <- 1L + (rownames(pm) %in% local_annot2$symbol)
+```{r ref.label="make_scols-2"}
 ```
 
-```{r heat-pleio-13}
-heatmaply(pm, symm = TRUE, cexRow = 0.2, cexCol = 0.2, row_side_colors = scols)
+```{r ref.label="heat8-2"}
 ```
 
-```{r heat3-13}
-(pm > pleio_threshold) %>%
-  apply(FUN = as.integer, MARGIN = 2) %>%
-  (function(x) {rownames(x) <- colnames(x); return(x)}) %>%
-  heatmaply(cexRow = 0.2, cexCol = 0.2, symm = TRUE, row_side_colors = scols)
+```{r ref.label = "make_scols-extra-2"}
 ```
 
-## Heatmaps for matrix containing only nonlocal traits
-
-```{r heat5-13}
-pm_nonlocal <- pleio_tib %>%
-  filter(gene1_symbol %in% nonlocal_annot$symbol, gene2_symbol %in% nonlocal_annot$symbol) %>%
-  tibble_to_matrix(symmetric = TRUE)
-pm_nonlocal %>%
-  heatmaply(symm = TRUE, cexRow = 0.3, cexCol = 0.3)
+```{r ref.label = "count_med"}
 ```
 
-```{r heat6-13}
-(pm_nonlocal > pleio_threshold) %>%
-  apply(FUN = as.integer, MARGIN = 2) %>%
-  heatmaply(symm = TRUE, cexRow = 0.3, cexCol = 0.3)
+```{r ref.label = "heat9-2"}
 ```

### 2019-07-22 analysis/paper/heatmaps-pleiotropy-statistics-subset-locals.Rmd
created a Rmd file to outline the ms

@@
   filter(chrom == hot_chr, pos <= hot_upper, pos >= hot_lower) %>%
-  left_join(y = dataset.islet.rnaseq$annots, by = c("annot.id" = "gene_id")) %>%
+  inner_join(y = dataset.islet.rnaseq$annots, by = c("annot.id" = "gene_id")) %>%
   filter(!duplicated(annot.id)) %>%
+  filter(lod >= 7.18) %>%
   mutate(local_tx = chr == hot_chr) %>%
@@
 ```{r make_scols-2}
-scols <- 1L + (rownames(pm) %in% local_annot2$symbol)
+scols1 <- 1L + (rownames(pm) %in% local_annot2$symbol)
 scols2 <- rownames(pm) %in% (med_tib %>% filter(local_symbol == keller_mediator, lod_diff > 1.5))$nonlocal_symbol
@@
 ```{r heat8-2}
-ord <- pm %>%
+# order the nonlocal traits first
+## indicator for being a nonlocal trait
+indic_nonlocal <- rownames(pm) %in% nonlocal_annot$symbol
+# define ord_nonlocal
+ord_nonlocal <- pm[indic_nonlocal, indic_nonlocal] %>%
   dist() %>%
@@
   as.dendrogram() %>%
-  dendextend::seriate_dendrogram(dist(pm)) %>%
+  dendextend::seriate_dendrogram(dist(pm[indic_nonlocal, indic_nonlocal])) %>%
   order.dendrogram()
-## order pm (and sidecols!) according to dendrogram from above
+rn_nonlocal <- rownames(pm[indic_nonlocal, indic_nonlocal][ord_nonlocal, ord_nonlocal])
+# define ord_local
+ord_local<- pm[!indic_nonlocal, !indic_nonlocal] %>%
+  dist() %>%
+  hclust() %>%
+  as.dendrogram() %>%
+  dendextend::seriate_dendrogram(dist(pm[!indic_nonlocal, !indic_nonlocal])) %>%
+  order.dendrogram()
+rn_local <- rownames(pm[!indic_nonlocal, !indic_nonlocal][ord_local, ord_local])
+rn <- c(rn_nonlocal, rn_local)
 ```
@@
   (function(x){colnames(x) <- local_annot2$symbol; x}) %>%
-  mutate(local = scols)
+  mutate(local = scols1)
 ```
@@
 ```{r count_med}
-n_med_tr <- 4
+n_med_tr <- 4 # arbitrary threshold
 (n_med_per_local <- med_tib %>%
@@
   )
-scols <- scols_extra %>%
-  dplyr::select(local, n_med_per_local$local_symbol)
+scols_pre <- scols_extra %>%
+  dplyr::select(n_med_per_local$local_symbol)
+# get ordering for columns of scols_pre
+col_order <- scols_pre %>%
+  t() %>%
+  dist() %>%
+  hclust() %>%
+  as.dendrogram() %>%
+  dendextend::seriate_dendrogram(dist(t(scols_pre))) %>%
+  order.dendrogram()
+scols <- scols_pre[ , col_order] %>%
+  mutate(local = scols1)
 ```
@@
 ```{r heat9-2}
+ord <- match(rn, rownames(pm))
+
 pm[ord, ord] %>%
@@
             row_side_colors = scols[ord, ],
+            col_side_colors = scols1[ord],
             limits = c(0, 5)

### 2019-07-22 analysis/paper/figures.Rmd
added figs and renamed Rmd file

--- 2019-07-22 analysis/paper/heatmaps-pleiotropy-statistics-subset-locals.Rmd
@@
 
+```{r scatter-2}
+p <- pleio_mixed_list %>%
+  bind_rows() %>%
+  rename(pleiotropy_lod = lod) %>%
+  inner_join(med_tib, by = c("local_gene_id", "nonlocal_gene_id")) %>%
+  dplyr::select(local_symbol, nonlocal_symbol, pleiotropy_lod, lod_diff, lod_diff_proportion, lod_without_med, lod_with_med, filename, nonlocal_gene_id, local_gene_id) %>%
+  ggplot() + geom_point(mapping = aes(x = pleiotropy_lod, y = lod_diff, colour = local_symbol, nonlocal = nonlocal_symbol), size = 0.05)
+fnsvg <- paste0("Chr", hot_chr, "_scatter.svg")
+fnpdf <- paste0("Chr", hot_chr, "_scatter.pdf")
+ggsave(plot = p, filename = file.path("../figures", fnsvg))
+ggsave(plot = p, filename = file.path("../figures", fnpdf))
+```
+
 ## heatmaps for full matrix
@@
 
+## Scatter: index v QTL peak position
+
+```{r index-v-peak-position-2}
+tibble(symbol = rownames(pm[indic_nonlocal, indic_nonlocal])[ord_nonlocal], index = 1:length(symbol)) %>%
+  left_join(nonlocal_annot, by = "symbol") %>%
+  ggplot() + geom_point(aes(y = pos, x = index))
+```
+
 # Chr 5
@@
 
+```{r scatter-5, ref.label="scatter-2"}
+```
+
 ## heatmaps for full matrix
@@
 ```{r ref.label = "heat9-2"}
+```
+
+## Scatter: index v QTL peak position
+
+```{r ref.label = "index-v-peak-position-2"}
 ```
@@
 
+```{r scatter-7, ref.label="scatter-2"}
+```
+
 ## heatmaps for full matrix
@@
 
+```{r scatter-11, ref.label="scatter-2"}
+```
+
 ## heatmaps for full matrix
@@
 
+```{r scatter-13, ref.label="scatter-2"}
+```
+
 ## heatmaps for full matrix
@@
 ```
+
+## Scatter: index v QTL peak position
+
+```{r ref.label = "index-v-peak-position-2"}
+```

### 2019-07-22 analysis/paper/paper.Rnw
added subsubsection header
fixed Makefile
updated Makefile for compiling paper.pdf

@@
 
+\subsubsection{Two-dimensional, two-QTL scans}
+
 For each hotspot, we tested all pairs of traits, regardless of whether the traits were local or nonlocal. We performed a
@@
 For $n$ mice, $\Sigma$ is a $2n$ by $2n$ covariance matrix, while $V_g$ and $V_e$ are $2$ by $2$ covariance matrices.
-Estimates $\hat V_g$ and $\hat V_e$ are obtained via an expectation-maximization algorithm \citep{zhou2014efficient,gemma2}
+Estimates $\hat V_g$ and $\hat V_e$ are obtained via an expectation-maximization algorithm \citep{zhou2014efficient,gemma2}.
 
@@
 
-We generated scatter plots of pleiotropy test statistics vs. mediation LOD difference proportion values. Each
-point represents a pairing of a local trait with a nonlocal trait. In all five scatter plots, we observe few
-points in the upper right quadrant of the figure. That is, few pairs have high pleiotropy test statistics and
+We generated scatter plots of pleiotropy test statistics vs. mediation LOD difference values.
+Each point represents a pairing of a local trait with a nonlocal trait.
+In all five scatter plots, we observe few
+points in the upper right quadrant of the figure.
+That is, few pairs have high pleiotropy test statistics and
 high mediation LOD difference values.
+
+<<>>=
+include_graphics("../figures/Chr2_scatter.pdf")
+@
 

### 2019-07-23 analysis/paper/figures.Rmd
removed legend from scatter plots

--- 2019-07-22 analysis/paper/heatmaps-pleiotropy-statistics-subset-locals.Rmd
@@
 
+```{r scatter-2}
+p <- pleio_mixed_list %>%
+  bind_rows() %>%
+  rename(pleiotropy_lod = lod) %>%
+  inner_join(med_tib, by = c("local_gene_id", "nonlocal_gene_id")) %>%
+  dplyr::select(local_symbol, nonlocal_symbol, pleiotropy_lod, lod_diff, lod_diff_proportion, lod_without_med, lod_with_med, filename, nonlocal_gene_id, local_gene_id) %>%
+  ggplot() + geom_point(mapping = aes(x = pleiotropy_lod, y = lod_diff, colour = local_symbol, nonlocal = nonlocal_symbol), size = 0.05) + theme(legend.position = "none")
+fnsvg <- paste0("Chr", hot_chr, "_scatter.svg")
+fnpdf <- paste0("Chr", hot_chr, "_scatter.pdf")
+ggsave(plot = p, filename = file.path("../figures", fnsvg))
+ggsave(plot = p, filename = file.path("../figures", fnpdf))
+```
+
 ## heatmaps for full matrix

### 2019-07-23 analysis/paper/figures.Rmd
updated figures

--- 2019-07-22 analysis/paper/heatmaps-pleiotropy-statistics-subset-locals.Rmd
@@
 
+```{r scatter-2}
+p <- pleio_mixed_list %>%
+  bind_rows() %>%
+  rename(pleiotropy_lod = lod) %>%
+  inner_join(med_tib, by = c("local_gene_id", "nonlocal_gene_id")) %>%
+  dplyr::select(local_symbol, nonlocal_symbol, pleiotropy_lod, lod_diff, lod_diff_proportion, lod_without_med, lod_with_med, filename, nonlocal_gene_id, local_gene_id) %>%
+  ggplot() + geom_point(mapping = aes(x = pleiotropy_lod, y = lod_diff, xlab = "Pleiotropy LOD", ylab = "Mediation LOD difference", colour = local_symbol, nonlocal = nonlocal_symbol), size = 0.05) + theme(legend.position = "none")
+fnsvg <- paste0("Chr", hot_chr, "_scatter.svg")
+fnpdf <- paste0("Chr", hot_chr, "_scatter.pdf")
+ggsave(plot = p, filename = file.path("../figures", fnsvg))
+ggsave(plot = p, filename = file.path("../figures", fnpdf))
+```
+
 ## heatmaps for full matrix

### 2019-07-23 analysis/paper/figures.Rmd
added scatter by local plots

--- 2019-07-22 analysis/paper/figures.Rmd
@@
   dplyr::select(local_symbol, nonlocal_symbol, pleiotropy_lod, lod_diff, lod_diff_proportion, lod_without_med, lod_with_med, filename, nonlocal_gene_id, local_gene_id) %>%
-  ggplot() + geom_point(mapping = aes(x = pleiotropy_lod, y = lod_diff, colour = local_symbol, nonlocal = nonlocal_symbol), size = 0.05)
+  ggplot() + geom_point(mapping = aes(x = pleiotropy_lod, y = lod_diff, colour = local_symbol == keller_mediator, nonlocal = nonlocal_symbol), size = 0.05) + theme(legend.position = "none") + xlab("Pleiotropy LOD") + ylab("Mediation LOD difference")
 fnsvg <- paste0("Chr", hot_chr, "_scatter.svg")
@@
 ggsave(plot = p, filename = file.path("../figures", fnpdf))
+```
+
+```{r, scatter_by_local-2}
+p <- pleio_mixed_list %>%
+  bind_rows() %>%
+  rename(pleiotropy_lod = lod) %>%
+  inner_join(med_tib, by = c("local_gene_id", "nonlocal_gene_id")) %>%
+  dplyr::select(local_symbol, nonlocal_symbol, pleiotropy_lod, lod_diff, lod_diff_proportion, lod_without_med, lod_with_med, filename, nonlocal_gene_id, local_gene_id) %>%
+  ggplot() + 
+  geom_point(mapping = aes(x = pleiotropy_lod, y = lod_diff, colour = nonlocal_symbol, nonlocal = nonlocal_symbol), size = 0.1) + 
+  facet_wrap(. ~ local_symbol, nrow = ceiling(nlocal / 2), ncol = 2) +
+  broman::karl_theme(strip.placement = "inside") + 
+  theme(legend.position = "none") +
+  xlab("Pleiotropy LOD") + 
+  ylab("Mediation LOD difference")
+fnsvg <- paste0("Chr", hot_chr, "_scatter-panel.svg")
+fnpdf <- paste0("Chr", hot_chr, "_scatter-panel.pdf")
+ggsave(plot = p, filename = file.path("../figures", fnsvg), height = ceiling(nlocal / 2))
+ggsave(plot = p, filename = file.path("../figures", fnpdf), height = ceiling(nlocal / 2))
 ```
@@
 
+```{r, ref.label = "scatter_by_local-2"}
+```
+
 ## heatmaps for full matrix
@@
 ```{r scatter-7, ref.label="scatter-2"}
+```
+
+```{r, ref.label = "scatter_by_local-2"}
 ```

### 2019-07-23 analysis/paper/paper.Rnw
updated figures
added scatter by local plots

@@
 
-<<>>=
+<<fig.cap="Scatter plot of mediation LOD difference against pleiotropy LOD for all nonlocal-local trait pairs mapping to Chromosome 2 hotspot.">>=
 include_graphics("../figures/Chr2_scatter.pdf")
+@
+
+<<fig.cap = "Chr 2 hotspot scatter plots of mediation LOD difference against pleiotropy LOD, separated by local trait.">>=
+include_graphics("../figures/Chr2_scatter-panel.pdf")
+@
+
+<<fig.cap="Scatter plot of mediation LOD difference against pleiotropy LOD for all nonlocal-local trait pairs mapping to Chromosome 5 hotspot.">>=
+include_graphics("../figures/Chr5_scatter.pdf")
+@
+
+<<fig.cap = "Chr 5 hotspot scatter plots of mediation LOD difference against pleiotropy LOD, separated by local trait.">>=
+include_graphics("../figures/Chr5_scatter-panel.pdf")
+@
+
+<<fig.cap="Scatter plot of mediation LOD difference against pleiotropy LOD for all nonlocal-local trait pairs mapping to Chromosome 7 hotspot.">>=
+include_graphics("../figures/Chr7_scatter.pdf")
+@
+
+<<fig.cap = "Chr 7 hotspot scatter plots of mediation LOD difference against pleiotropy LOD, separated by local trait.">>=
+include_graphics("../figures/Chr7_scatter-panel.pdf")
+@
+
+<<fig.cap="Scatter plot of mediation LOD difference against pleiotropy LOD for all nonlocal-local trait pairs mapping to Chromosome 11 hotspot.">>=
+include_graphics("../figures/Chr11_scatter.pdf")
+@
+
+<<fig.cap = "Chr 11 hotspot scatter plots of mediation LOD difference against pleiotropy LOD, separated by local trait.">>=
+include_graphics("../figures/Chr11_scatter-panel.pdf")
+@
+
+<<fig.cap="Scatter plot of mediation LOD difference against pleiotropy LOD for all nonlocal-local trait pairs mapping to Chromosome 13 hotspot.">>=
+include_graphics("../figures/Chr13_scatter.pdf")
+@
+
+<<fig.cap = "Chr 13 hotspot scatter plots of mediation LOD difference against pleiotropy LOD, separated by local trait.">>=
+include_graphics("../figures/Chr13_scatter-panel.pdf")
 @

### 2019-07-23 analysis/paper/figures.Rmd
updated ms

@@
 ```{r, scatter_by_local-2}
-p <- pleio_mixed_list %>%
+foo <- pleio_mixed_list %>%
   bind_rows() %>%
@@
   inner_join(med_tib, by = c("local_gene_id", "nonlocal_gene_id")) %>%
-  dplyr::select(local_symbol, nonlocal_symbol, pleiotropy_lod, lod_diff, lod_diff_proportion, lod_without_med, lod_with_med, filename, nonlocal_gene_id, local_gene_id) %>%
+  dplyr::select(local_symbol, nonlocal_symbol, pleiotropy_lod, lod_diff, lod_diff_proportion, lod_without_med, lod_with_med, filename, nonlocal_gene_id, local_gene_id)
+p <- foo %>%
   ggplot() + 
@@
   theme(legend.position = "none") +
+  theme(
+    strip.background = element_blank(),
+    strip.text.x = element_blank()
+) + 
   xlab("Pleiotropy LOD") + 
-  ylab("Mediation LOD difference")
+  ylab("Mediation LOD difference") + 
+  geom_text(aes(x = 1, y = 1, label = local_symbol), inherit.aes=FALSE)
+
 fnsvg <- paste0("Chr", hot_chr, "_scatter-panel.svg")
@@
 ggsave(plot = p, filename = file.path("../figures", fnpdf), height = ceiling(nlocal / 2))
+
+# put in multiple figures, if there are many local traits
+n_figs <- 1 + nlocal %/% 18 # 18 is the number of panels per page in pdf
+n_last <- nlocal %% 18 # number of panels on last figure
+max_pleio_lod <- max(foo$pleiotropy_lod)
+max_med_lod_difference <- max(foo$lod_diff) 
+min_med_lod_difference <- min(foo$lod_diff)
+
+for (i in 1:n_figs){
+  n_panels <- 18 * (i < n_figs) + n_last * (i == n_figs)
+  foo %>%
+    filter(local_symbol %in% local_annot2$symbol[((i - 1) * 18 + 1): (18 * i)]) %>%
+    ggplot() + 
+      geom_point(mapping = aes(x = pleiotropy_lod, y = lod_diff, colour = nonlocal_symbol, nonlocal = nonlocal_symbol), size = 0.1) + 
+      facet_wrap(. ~ local_symbol, nrow = ceiling(n_panels / 2), ncol = 2) +
+      broman::karl_theme() + 
+      theme(legend.position = "none") +
+      theme(
+        strip.background = element_blank(),
+        strip.text.x = element_blank()
+    ) + 
+      xlab("Pleiotropy LOD") + xlim(c(0, max_pleio_lod)) + ylim(c(min_med_lod_difference, max_med_lod_difference)) + 
+      ylab("Mediation LOD difference") + 
+      geom_text(vjust = 0.9, hjust = 0.9, size = 3, aes(x = max_pleio_lod / 2, y = 2 * max_med_lod_difference / 3, label = local_symbol), inherit.aes=FALSE)
+  fnsvg <- paste0("Chr", hot_chr, "_scatter-panel", i, ".svg")
+  fnpdf <- paste0("Chr", hot_chr, "_scatter-panel", i, ".pdf")
+  ggsave(filename = file.path("../figures", fnsvg), height = 9)
+  ggsave(filename = file.path("../figures", fnpdf), height = 9)
+}
 ```

### 2019-07-23 analysis/paper/paper.Rnw
updated ms

--- 2019-07-22 analysis/paper/paper.Rnw
@@
 <<>>=
-include_graphics("../figures/Chr2_scatter.pdf")
+hot_chr <- 2
+@
+
+<<fig.cap="Scatter plot of mediation LOD difference against pleiotropy LOD for all nonlocal-local trait pairs mapping to Chromosome 2 hotspot.">>=
+include_graphics(paste0("../figures/Chr", hot_chr, "_scatter.pdf"))
+@
+
+<<define_nfig>>=
+nfig <- c(2, 2, 3, 3, 1) # number of panel figs per hotspot
+names(nfig) <- paste0("Chr", c(2, 5, 7, 11, 13))
+@
+
+<<panel1, fig.cap = paste0("Chr", hot_chr, "hotspot scatter plots of mediation LOD difference against pleiotropy LOD, separated by local trait.")>>=
+include_graphics(paste0("../figures/Chr", hot_chr, "_scatter-panel1.pdf"))
+@
+
+<<panel2, eval = nfig[which(names(nfig) == paste0("Chr", hot_chr))] > 1>>=
+include_graphics(paste0("../figures/Chr", hot_chr, "_scatter-panel2.pdf"))
+@
+
+<<panel3, eval = nfig[which(names(nfig) == paste0("Chr", hot_chr))] > 2>>=
+include_graphics(paste0("../figures/Chr", hot_chr, "_scatter-panel3.pdf"))
+@
+
+<<>>=
+hot_chr <- 5
+@
+
+<<fig.cap="Scatter plot of mediation LOD difference against pleiotropy LOD for all nonlocal-local trait pairs mapping to Chromosome 5 hotspot.">>=
+include_graphics("../figures/Chr5_scatter.pdf")
+@
+
+<<ref.label = "panel1", fig.cap = paste0("Chr", hot_chr, "hotspot scatter plots of mediation LOD difference against pleiotropy LOD, separated by local trait.")>>=
+@
+
+<<ref.label = "panel2", eval = nfig[which(names(nfig) == paste0("Chr", hot_chr))] > 1>>=
+@
+
+<<ref.label = "panel3", eval = nfig[which(names(nfig) == paste0("Chr", hot_chr))] > 2>>=
+@
+
+<<>>=
+hot_chr <- 7
+@
+
+<<fig.cap="Scatter plot of mediation LOD difference against pleiotropy LOD for all nonlocal-local trait pairs mapping to Chromosome 7 hotspot.">>=
+include_graphics("../figures/Chr7_scatter.pdf")
+@
+
+<<ref.label = "panel1", fig.cap = paste0("Chr", hot_chr, "hotspot scatter plots of mediation LOD difference against pleiotropy LOD, separated by local trait.")>>=
+@
+
+<<ref.label = "panel2", eval = nfig[which(names(nfig) == paste0("Chr", hot_chr))] > 1>>=
+@
+
+<<ref.label = "panel3", eval = nfig[which(names(nfig) == paste0("Chr", hot_chr))] > 2>>=
+@
+
+<<>>=
+hot_chr <- 11
+@
+
+<<fig.cap="Scatter plot of mediation LOD difference against pleiotropy LOD for all nonlocal-local trait pairs mapping to Chromosome 11 hotspot.">>=
+include_graphics("../figures/Chr11_scatter.pdf")
+@
+
+<<ref.label = "panel1", fig.cap = paste0("Chr", hot_chr, "hotspot scatter plots of mediation LOD difference against pleiotropy LOD, separated by local trait.")>>=
+@
+
+<<ref.label = "panel2", eval = nfig[which(names(nfig) == paste0("Chr", hot_chr))] > 1>>=
+@
+
+<<ref.label = "panel3", eval = nfig[which(names(nfig) == paste0("Chr", hot_chr))] > 2>>=
+@
+
+<<>>=
+hot_chr <- 13
+@
+
+<<fig.cap="Scatter plot of mediation LOD difference against pleiotropy LOD for all nonlocal-local trait pairs mapping to Chromosome 13 hotspot.">>=
+include_graphics("../figures/Chr13_scatter.pdf")
+@
+
+<<ref.label = "panel1", fig.cap = paste0("Chr", hot_chr, "hotspot scatter plots of mediation LOD difference against pleiotropy LOD, separated by local trait.")>>=
+@
+
+<<ref.label = "panel2", eval = nfig[which(names(nfig) == paste0("Chr", hot_chr))] > 1>>=
+@
+
+<<ref.label = "panel3", eval = nfig[which(names(nfig) == paste0("Chr", hot_chr))] > 2>>=
 @

### 2019-07-24 analysis/paper/paper.Rnw
updated ms again

@@
 %\usepackage{etoolbox} % for messing with page style & title pages
+\usepackage{lscape}
 \usepackage{fancyhdr}
@@
 
-\doublespacing
-
 \begin{document}
@@
 
-<<setup, include=FALSE>>=
+<<setuprnw, include=TRUE>>=
 library(knitr)
@@
 concordance=TRUE,
-echo = TRUE,
+echo = FALSE,
 fig.lp = "fig:",
 warning = FALSE,
-message = FALSE)
-
+message = FALSE,
+comment = NA
+)
+def.chunk.hook  <- knitr::knit_hooks$get("chunk")
+knitr::knit_hooks$set(chunk = function(x, options) {
+  x <- def.chunk.hook(x, options)
+  ifelse(options$size != "normalsize", paste0("\\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
+})
 options(tinytex.verbose = TRUE)
@@
 A genome-wide QTL ``scan'' reveals
-associations between genotypes and phenotypes by considering each position,
+associations between genotypes and phenotypes by considering each position along the genome,
 one at a time, as a candidate QTL for the trait of interest.
@@
 and may contain more than a hundred genes.
-Identification of the causal gene (for a given complex trait) from among those genes near the QTL is
+Identification of a causal gene (for a given complex trait) from among those genes near the QTL is
 challenging and may require costly and time-consuming experiments.
@@
 study of pancreatic islet gene expression in 378 Diversity Outbred mice. They
-defined a hotspot to be a 4-Mb region that affects at least 100 expression
+defined a hotspot to be a 4-Mb region that affects at least 100 nonlocal expression
 traits.
@@
 
+A mediation analysis in systems genetics begins with an identified QTL-expression trait association. In our case, we choose a QTL that affects a nonlocal trait.
+
 For each hotspot and each local trait - nonlocal trait pair, we also performed a
@@
 
+\subsection{Visualizations}
+
+We created scatter plots with the {\tt ggplot2} package \citep{ggplot2} for the R
+statistical computing environment \citep{r}.
+
 \section{Results}
 
-We generated scatter plots of pleiotropy test statistics vs. mediation LOD difference values.
+We generated scatter plots of pleiotropy test statistics vs. mediation LOD difference values for each of the five hotspots (Figure~\ref{}, \ref{}).
 Each point represents a pairing of a local trait with a nonlocal trait.
+The mediation LOD difference value is that observed when the local trait is considered
+as a candidate mediator of the QTL-nonlocal trait association.
+
 In all five scatter plots, we observe few
@@
 high mediation LOD difference values.
+
+<<setup>>=
+library(hotspots)
+pleio_threshold <- 4
+ult <- 15 # univariate lod threshold
+@
+
+<<setup_continued-2>>=
+hot_chr <- 2
+hot_mid <- 165.5
+keller_mediator <- "Hnf4a"
+@
+
+<<more_setup-2>>=
+hot_lower <- hot_mid - 2
+hot_upper <- hot_mid + 2
+fp <- paste0("../figures/Chr", hot_chr, "-")
+knitr::opts_chunk$set(fig.path = fp)
+@
+
+<<readRDS-2>>=
+pleio_mixed_list <- readRDS(paste0("_cache/Chr", hot_chr, "_pleio.rds"))
+pleio_local_list <- readRDS(paste0("_cache/Chr", hot_chr, "_pleio_local.rds"))
+pleio_nonlocal_list <- readRDS(paste0("_cache/Chr", hot_chr, "_pleio_nonlocal.rds"))
+med_list <- readRDS(paste0("_cache/Chr", hot_chr, "_mediations.rds"))
+@
+
+<<load_attie>>=
+load("~/Box Sync/attie/keller2018-chr2-hotspot-chtc/data-to-ignore/Attie_DO378_eQTL_viewer_v1.Rdata")
+library(tidyverse)
+options(tibble.print_max = 50, tibble.print_min = 20)
+@
+
+<<make_local_annot-2>>=
+local_annot <- dataset.islet.rnaseq$lod.peaks %>%
+  filter(chrom == hot_chr, pos <= hot_upper, pos >= hot_lower) %>%
+  left_join(y = dataset.islet.rnaseq$annots, by = c("annot.id" = "gene_id")) %>%
+  filter(!duplicated(annot.id)) %>%
+  filter(lod >= ult) %>%
+  mutate(local_tx = chr == hot_chr) %>%
+  filter(local_tx)
+@
+
+<<make_nonlocal_annot-2>>=
+nonlocal_annot <- dataset.islet.rnaseq$lod.peaks %>%
+  filter(chrom == hot_chr, pos <= hot_upper, pos >= hot_lower) %>%
+  inner_join(y = dataset.islet.rnaseq$annots, by = c("annot.id" = "gene_id")) %>%
+  filter(!duplicated(annot.id)) %>%
+  filter(lod >= 7.18) %>%
+  mutate(local_tx = chr == hot_chr) %>%
+  filter(!local_tx) %>%
+  filter(!is.na(hotspot))
+@
+
+<<make_local_annot2-2>>=
+local_annot2 <- local_annot %>%
+  filter(middle <= hot_upper + 5, middle >= hot_lower - 5) %>%
+  arrange(desc(lod))
+@
+
+<<subset_expr_matrix-2>>=
+hot_local <- dataset.islet.rnaseq$expr[, colnames(dataset.islet.rnaseq$expr) %in% local_annot2$annot.id]
+hot_nonlocal <- dataset.islet.rnaseq$expr[, colnames(dataset.islet.rnaseq$expr) %in% nonlocal_annot$annot.id]
+nnonlocal <- nrow(nonlocal_annot)
+nlocal <- nrow(local_annot2)
+@
+
+<<tabulate-2, message=FALSE, warning = FALSE>>=
+med_tib <- med_list %>%
+  purrr::map(bind_rows) %>%
+  bind_rows() %>%
+  as_tibble() %>%
+  rename(lod_without_med = lod_no_med, lod_with_med = lod_med) %>%
+  mutate(lod_diff = lod_without_med - lod_with_med, lod_diff_proportion = lod_diff / lod_without_med) %>%
+  left_join(local_annot2, by = c("local_gene_id" = "annot.id")) %>%
+  left_join(nonlocal_annot, by = c("nonlocal_gene_id" = "annot.id")) %>%
+  dplyr::select(symbol.x, symbol.y, lod_without_med, lod_with_med, lod_diff, lod_diff_proportion, local_gene_id, nonlocal_gene_id) %>%
+  rename(local_symbol = symbol.x, nonlocal_symbol = symbol.y) %>%
+  filter(!is.na(local_symbol))
+@
+
+<<make_pleio_tib-2>>=
+hot_annot <- local_annot2 %>%
+  bind_rows(nonlocal_annot) %>%
+  dplyr::select(annot.id, symbol)
+pleio_tib <- pleio_mixed_list %>%
+  purrr::map(.f = function(x) rename(x, gene1_id = local_gene_id, gene2_id = nonlocal_gene_id)) %>%
+  c(pleio_local_list, pleio_nonlocal_list) %>%
+  bind_rows() %>%
+  rename(pleiotropy_lod = lod) %>%
+  inner_join(hot_annot, by = c("gene1_id" = "annot.id")) %>%
+  rename(gene1_symbol = symbol) %>%
+  inner_join(hot_annot, by = c("gene2_id" = "annot.id")) %>%
+  rename(gene2_symbol = symbol) %>%
+  dplyr::select(gene1_symbol, gene2_symbol, pleiotropy_lod)
+@
+
+<<make_pm-2>>=
+pm <- pleio_tib %>%
+  hotspots::tibble_to_matrix(symmetric = TRUE)
+@
+
+<<scatter-2>>=
+p <- pleio_mixed_list %>%
+  bind_rows() %>%
+  rename(pleiotropy_lod = lod) %>%
+  inner_join(med_tib, by = c("local_gene_id", "nonlocal_gene_id")) %>%
+  dplyr::select(local_symbol, nonlocal_symbol, pleiotropy_lod, lod_diff, lod_diff_proportion, lod_without_med, lod_with_med, filename, nonlocal_gene_id, local_gene_id) %>%
+  ggplot() + geom_point(mapping = aes(x = pleiotropy_lod, y = lod_diff, colour = local_symbol == keller_mediator, nonlocal = nonlocal_symbol), size = 0.05) + theme(legend.position = "none") + xlab("Pleiotropy LOD") + ylab("Mediation LOD difference")
+fnsvg <- paste0("Chr", hot_chr, "_scatter.svg")
+fnpdf <- paste0("Chr", hot_chr, "_scatter.pdf")
+ggsave(plot = p, filename = file.path("../figures", fnsvg))
+ggsave(plot = p, filename = file.path("../figures", fnpdf))
+p
+@
+
+<<scatter_by_local-2>>=
+foo <- pleio_mixed_list %>%
+  bind_rows() %>%
+  rename(pleiotropy_lod = lod) %>%
+  inner_join(med_tib, by = c("local_gene_id", "nonlocal_gene_id")) %>%
+  dplyr::select(local_symbol, nonlocal_symbol, pleiotropy_lod, lod_diff, lod_diff_proportion, lod_without_med, lod_with_med, filename, nonlocal_gene_id, local_gene_id)
+p <- foo %>%
+  ggplot() +
+  geom_point(mapping = aes(x = pleiotropy_lod, y = lod_diff, colour = nonlocal_symbol, nonlocal = nonlocal_symbol), size = 0.1) +
+  facet_wrap(. ~ local_symbol, nrow = ceiling(nlocal / 2), ncol = 2) +
+  broman::karl_theme(strip.placement = "inside") +
+  theme(legend.position = "none") +
+  theme(
+    strip.background = element_blank(),
+    strip.text.x = element_blank()
+) +
+  xlab("Pleiotropy LOD") +
+  ylab("Mediation LOD difference") +
+  geom_text(aes(x = 1, y = 1, label = local_symbol), inherit.aes=FALSE)
+
+fnsvg <- paste0("Chr", hot_chr, "_scatter-panel.svg")
+fnpdf <- paste0("Chr", hot_chr, "_scatter-panel.pdf")
+ggsave(plot = p, filename = file.path("../figures", fnsvg), height = ceiling(nlocal / 2))
+ggsave(plot = p, filename = file.path("../figures", fnpdf), height = ceiling(nlocal / 2))
+
+# put in multiple figures, if there are many local traits
+n_figs <- 1 + nlocal %/% 18 # 18 is the number of panels per page in pdf
+n_last <- nlocal %% 18 # number of panels on last figure
+max_pleio_lod <- max(foo$pleiotropy_lod)
+max_med_lod_difference <- max(foo$lod_diff)
+min_med_lod_difference <- min(foo$lod_diff)
+
+for (i in 1:n_figs){
+  n_panels <- 18 * (i < n_figs) + n_last * (i == n_figs)
+  foo %>%
+    filter(local_symbol %in% local_annot2$symbol[((i - 1) * 18 + 1): (18 * i)]) %>%
+    ggplot() +
+      geom_point(mapping = aes(x = pleiotropy_lod, y = lod_diff, colour = nonlocal_symbol, nonlocal = nonlocal_symbol), size = 0.1) +
+      facet_wrap(. ~ local_symbol, nrow = ceiling(n_panels / 2), ncol = 2) +
+      broman::karl_theme() +
+      theme(legend.position = "none") +
+      theme(
+        strip.background = element_blank(),
+        strip.text.x = element_blank()
+    ) +
+      xlab("Pleiotropy LOD") + xlim(c(0, max_pleio_lod)) + ylim(c(min_med_lod_difference, max_med_lod_difference)) +
+      ylab("Mediation LOD difference") +
+      geom_text(size = 2, aes(x = max_pleio_lod / 2, y = 2 * max_med_lod_difference / 3, label = local_symbol), inherit.aes=FALSE)
+  fnsvg <- paste0("Chr", hot_chr, "_scatter-panel", i, ".svg")
+  fnpdf <- paste0("Chr", hot_chr, "_scatter-panel", i, ".pdf")
+  ggsave(filename = file.path("../figures", fnsvg), height = ceiling(n_panels / 2))
+  ggsave(filename = file.path("../figures", fnpdf), height = ceiling(n_panels / 2))
+}
+@
+
+<<load_heatmaply>>=
+library(heatmaply)
+@
+
+<<make_scols-2>>=
+scols1 <- 1L + (rownames(pm) %in% local_annot2$symbol)
+scols2 <- rownames(pm) %in% (med_tib %>% filter(local_symbol == keller_mediator, lod_diff > 1.5))$nonlocal_symbol
+@
+
+<<heat8-2>>=
+# order the nonlocal traits first
+## indicator for being a nonlocal trait
+indic_nonlocal <- rownames(pm) %in% nonlocal_annot$symbol
+# define ord_nonlocal
+ord_nonlocal <- pm[indic_nonlocal, indic_nonlocal] %>%
+  dist() %>%
+  hclust() %>%
+  as.dendrogram() %>%
+  dendextend::seriate_dendrogram(dist(pm[indic_nonlocal, indic_nonlocal])) %>%
+  order.dendrogram()
+rn_nonlocal <- rownames(pm[indic_nonlocal, indic_nonlocal][ord_nonlocal, ord_nonlocal])
+# define ord_local
+ord_local<- pm[!indic_nonlocal, !indic_nonlocal] %>%
+  dist() %>%
+  hclust() %>%
+  as.dendrogram() %>%
+  dendextend::seriate_dendrogram(dist(pm[!indic_nonlocal, !indic_nonlocal])) %>%
+  order.dendrogram()
+rn_local <- rownames(pm[!indic_nonlocal, !indic_nonlocal][ord_local, ord_local])
+rn <- c(rn_nonlocal, rn_local)
+@
+
+<<make_scols-extra-2>>=
+scols_extra <- purrr::map(.x = local_annot2$symbol,
+                          .f = function(x) {rownames(pm) %in% (med_tib %>% filter(local_symbol == x, lod_diff > 1.5))$nonlocal_symbol}
+) %>%
+  bind_cols() %>%
+  (function(x){colnames(x) <- local_annot2$symbol; x}) %>%
+  mutate(local = scols1)
+@
+
+<<count_med>>=
+n_med_tr <- 4 # arbitrary threshold
+n_med_per_local <- med_tib %>%
+  group_by(local_symbol) %>%
+  filter(lod_diff > 1.5) %>%
+  count() %>%
+  filter(n >= n_med_tr)
+
+scols_pre <- scols_extra %>%
+  dplyr::select(n_med_per_local$local_symbol)
+# get ordering for columns of scols_pre
+col_order <- scols_pre %>%
+  t() %>%
+  dist() %>%
+  hclust() %>%
+  as.dendrogram() %>%
+  dendextend::seriate_dendrogram(dist(t(scols_pre))) %>%
+  order.dendrogram()
+scols <- scols_pre[ , col_order] %>%
+  mutate(local = scols1)
+@
+
+<<heat9-2>>=
+ord <- match(rn, rownames(pm))
+
+pm[ord, ord] %>%
+  apply(FUN = hotspots::trunc2, MARGIN = 2, threshold = 5) %>%
+  heatmaply(Rowv = FALSE,
+            Colv = FALSE,
+            cexRow = 0.2,
+            cexCol = 0.2,
+            symm = TRUE,
+            row_side_colors = scols[ord, ],
+            col_side_colors = scols1[ord],
+            limits = c(0, 5)
+            )
+@
+
+<<index-v-peak-position-2>>=
+tibble(symbol = rownames(pm[indic_nonlocal, indic_nonlocal])[ord_nonlocal], index = 1:length(symbol)) %>%
+  left_join(nonlocal_annot, by = "symbol") %>%
+  ggplot() + geom_point(aes(y = pos, x = index))
+@
+
+<<setup_continued-5>>=
+hot_chr <- 5
+hot_mid <- 146
+keller_mediator <- "Pdx1"
+@
+
+<<more_setup-5, ref.label = "more_setup-2">>=
+@
+
+<<readRDS-5, ref.label="readRDS-2">>=
+@
+
+<<make_local_annot-5, ref.label="make_local_annot-2">>=
+@
+
+<<make_nonlocal_annot-5, ref.label = "make_nonlocal_annot-2">>=
+@
+
+<<make_local_annot2-5, ref.label = "make_local_annot2-2">>=
+@
+
+<<subset_expr_matrix-5, ref.label = "subset_expr_matrix-2">>=
+@
+
+<<tabulate-5, ref.label = "tabulate-2", message=FALSE, warning = FALSE>>=
+@
+
+<<make_pleio_tib-5, ref.label="make_pleio_tib-2">>=
+@
+
+<<make_pm-5, ref.label = "make_pm-2">>=
+@
+
+<<scatter-5, ref.label="scatter-2">>=
+@
+
+<<ref.label = "scatter_by_local-2">>=
+@
+
+<<make_scols-5, ref.label="make_scols-2">>=
+@
+
+<<heat8-5, ref.label="heat8-2">>=
+@
+
+<<ref.label = "make_scols-extra-2">>=
+@
+
+<<ref.label = "count_med">>=
+@
+
+<<ref.label = "heat9-2">>=
+@
+
+<<ref.label = "index-v-peak-position-2">>=
+@
+
+<<setup_continued-7>>=
+hot_chr <- 7
+hot_mid <- 46
+keller_mediator <- "Fam83e"
+@
+
+<<ref.label = "more_setup-2">>=
+@
+
+<<ref.label="readRDS-2">>=
+@
+
+<<ref.label="make_local_annot-2">>=
+@
+
+<<ref.label = "make_nonlocal_annot-2">>=
+@
+
+<<ref.label = "make_local_annot2-2">>=
+@
+
+<<ref.label = "subset_expr_matrix-2">>=
+@
+
+<<ref.label = "tabulate-2", message=FALSE, warning = FALSE>>=
+@
+
+<<ref.label="make_pleio_tib-2">>=
+@
+
+<<ref.label = "make_pm-2">>=
+@
+
+<<scatter-7, ref.label="scatter-2">>=
+@
+
+<<ref.label = "scatter_by_local-2">>=
+@
+
+<<ref.label="make_scols-2">>=
+@
+
+<<ref.label="heat8-2">>=
+@
+
+<<ref.label = "make_scols-extra-2">>=
+@
+
+<<ref.label = "count_med">>=
+@
+
+<<ref.label = "heat9-2">>=
+@
+
+<<ref.label = "index-v-peak-position-2">>=
+@
+
+<<setup_continued-11>>=
+hot_chr <- 11
+hot_mid <- 71
+keller_mediator <- "Sat2"
+@
+
+<<ref.label = "more_setup-2">>=
+@
+
+<<ref.label="readRDS-2">>=
+@
+
+<<ref.label="make_local_annot-2">>=
+@
+
+<<ref.label = "make_nonlocal_annot-2">>=
+@
+
+<<ref.label = "make_local_annot2-2">>=
+@
+
+<<ref.label = "subset_expr_matrix-2">>=
+@
+
+<<ref.label = "tabulate-2", message=FALSE, warning = FALSE>>=
+@
+
+<<ref.label="make_pleio_tib-2">>=
+@
+
+<<ref.label = "make_pm-2">>=
+@
+
+<<scatter-11, ref.label="scatter-2">>=
+@
+
+<<ref.label = "scatter_by_local-2">>=
+@
+
+<<ref.label="make_scols-2">>=
+@
+
+<<ref.label="heat8-2">>=
+@
+
+<<ref.label = "make_scols-extra-2">>=
+@
+
+<<ref.label = "count_med">>=
+@
+
+<<ref.label = "heat9-2">>=
+@
+
+<<ref.label = "index-v-peak-position-2">>=
+@
+
+<<setup_continued-13>>=
+hot_chr <- 13
+hot_mid <- 112.5
+keller_mediator <- "Il6st"
+@
+
+<<ref.label = "more_setup-2">>=
+@
+
+<<ref.label="readRDS-2">>=
+@
+
+<<ref.label="make_local_annot-2">>=
+@
+
+<<ref.label = "make_nonlocal_annot-2">>=
+@
+
+<<ref.label = "make_local_annot2-2">>=
+@
+
+<<ref.label = "subset_expr_matrix-2">>=
+@
+
+<<ref.label = "tabulate-2", message=FALSE, warning = FALSE>>=
+@
+
+<<ref.label="make_pleio_tib-2">>=
+@
+
+<<ref.label = "make_pm-2">>=
+@
+
+<<scatter-13, ref.label="scatter-2">>=
+@
+
+<<ref.label = "scatter_by_local-2">>=
+@
+
+<<ref.label="make_scols-2">>=
+@
+
+<<ref.label="heat8-2">>=
+@
+
+<<ref.label = "make_scols-extra-2">>=
+@
+
+<<ref.label = "count_med">>=
+@
+
+<<ref.label = "heat9-2">>=
+@
+
+<<ref.label = "index-v-peak-position-2">>=
+@
 
@@
 \begin{singlespacing}
+%\begin{landscape}
+<<colophon, cache = FALSE, echo = TRUE, tidy.opts=list(width.cutoff=50), size = "tiny", results = "markup">>=
+# which R packages and versions?
+foo <- devtools::session_info()
+readr::write_csv(x = foo$packages, path = "packages.csv")
+knitr::kable(tibble(field = names(foo$platform), value = unlist(foo$platform)))
+foo$packages %>%
+  as_tibble() %>%
+  dplyr::select(-4, -5, -10, -11) %>%
+  knitr::kable()
+@
 
-<<colophon, cache = FALSE>>=
-# which R packages and versions?
-devtools::session_info()
-@
+%\end{landscape}
+
+\end{singlespacing}
 
@@
 
-<<git2r, cache = FALSE>>=
+<<git2r, cache = FALSE, echo = TRUE>>=
 # what commit is this file at?
@@
 
-\end{singlespacing}
-
 \end{document}

### 2019-07-24 analysis/paper/paper.Rnw
fixed package name in paper.Rnw

@@
 <<setup>>=
-library(hotspots)
+library(qtl2hotspots)
 pleio_threshold <- 4
@@
 pm <- pleio_tib %>%
-  hotspots::tibble_to_matrix(symmetric = TRUE)
+  tibble_to_matrix(symmetric = TRUE)
 @
@@
 pm[ord, ord] %>%
-  apply(FUN = hotspots::trunc2, MARGIN = 2, threshold = 5) %>%
+  apply(FUN = trunc2, MARGIN = 2, threshold = 5) %>%
   heatmaply(Rowv = FALSE,

### 2019-07-25 analysis/paper/paper.Rnw
fixed figures

@@
 message = FALSE,
-comment = NA
+comment = NA,
+fig.dim = c(7, 7),
+out.width = "\\linewidth"
 )
@@
   filter(n >= n_med_tr)
+n_med_per_local %>%
+  arrange(desc(n)) %>%
+  knitr::kable()
 
@@
 
-<<heat9-2>>=
+<<heat9-2, fig.height = 7, fig.width = 7, out.width = "\\textwidth">>=
 ord <- match(rn, rownames(pm))
 
-pm[ord, ord] %>%
+p <- pm[ord, ord] %>%
   apply(FUN = trunc2, MARGIN = 2, threshold = 5) %>%
@@
             row_side_colors = scols[ord, ],
-            col_side_colors = scols1[ord],
+            col_side_colors = tibble(local = scols1[ord]),
             limits = c(0, 5)
-            )
+            ) %>%
+  hide_guides()
+fnsvg <- paste0("Chr", hot_chr, "_heat", ".svg")
+fnpdf <- paste0("Chr", hot_chr, "_heat", ".pdf")
+plotly::orca(p, file.path("../figures", fnsvg))
+plotly::orca(p, file.path("../figures", fnpdf))
+p
 @
@@
 
-<<ref.label = "scatter_by_local-2">>=
+<<scatter_by_local-5, ref.label = "scatter_by_local-2">>=
 @
@@
 
-<<ref.label = "make_scols-extra-2">>=
+<<make_scols-extra-5, ref.label = "make_scols-extra-2">>=
 @
 
-<<ref.label = "count_med">>=
+<<count_med-5, ref.label = "count_med">>=
 @
 
-<<ref.label = "heat9-2">>=
+<<heat9-5, ref.label = "heat9-2">>=
 @
 
-<<ref.label = "index-v-peak-position-2">>=
+<<index-v-peak-position-5, ref.label = "index-v-peak-position-2">>=
 @
@@
 
-<<ref.label = "more_setup-2">>=
+<<more_setup-7, ref.label = "more_setup-2">>=
 @
 
-<<ref.label="readRDS-2">>=
+<<readRDS-7, ref.label="readRDS-2">>=
 @
 
-<<ref.label="make_local_annot-2">>=
+<<make_local_annot-7, ref.label="make_local_annot-2">>=
 @
 
-<<ref.label = "make_nonlocal_annot-2">>=
+<<make_local_annot2-7, ref.label = "make_local_annot2-2">>=
 @
 
-<<ref.label = "make_local_annot2-2">>=
+<<subset_expr_matrix-7, ref.label = "subset_expr_matrix-2">>=
 @
 
-<<ref.label = "subset_expr_matrix-2">>=
+<<tabulate-7, ref.label = "tabulate-2", message=FALSE, warning = FALSE>>=
 @
 
-<<ref.label = "tabulate-2", message=FALSE, warning = FALSE>>=
+<<make_pleio_tib-7, ref.label="make_pleio_tib-2">>=
 @
 
-<<ref.label="make_pleio_tib-2">>=
-@
-
-<<ref.label = "make_pm-2">>=
+<<make_pm-7, ref.label = "make_pm-2">>=
 @
@@
 
-<<ref.label = "scatter_by_local-2">>=
+<<scatter_by_local-7, ref.label = "scatter_by_local-2">>=
 @
 
-<<ref.label="make_scols-2">>=
+<<make_scols-7, ref.label="make_scols-2">>=
 @
 
-<<ref.label="heat8-2">>=
+<<heat8-7, ref.label="heat8-2">>=
 @
 
-<<ref.label = "make_scols-extra-2">>=
+<<make_scols-extra-7, ref.label = "make_scols-extra-2">>=
 @
 
-<<ref.label = "count_med">>=
+<<count_med-7, ref.label = "count_med">>=
 @
 
-<<ref.label = "heat9-2">>=
+<<heat9-7, ref.label = "heat9-2">>=
 @
 
-<<ref.label = "index-v-peak-position-2">>=
+<<index-v-peak-position-7, ref.label = "index-v-peak-position-2">>=
 @
@@
 
-<<ref.label = "more_setup-2">>=
+<<more_setup-11, ref.label = "more_setup-2">>=
 @
 
-<<ref.label="readRDS-2">>=
+<<readRDS-11, ref.label="readRDS-2">>=
 @
 
-<<ref.label="make_local_annot-2">>=
+<<make_local_annot-11, ref.label="make_local_annot-2">>=
 @
 
-<<ref.label = "make_nonlocal_annot-2">>=
+<<make_local_annot2-11, ref.label = "make_local_annot2-2">>=
 @
 
-<<ref.label = "make_local_annot2-2">>=
+<<subset_expr_matrix-11, ref.label = "subset_expr_matrix-2">>=
 @
 
-<<ref.label = "subset_expr_matrix-2">>=
+<<tabulate-11, ref.label = "tabulate-2", message=FALSE, warning = FALSE>>=
 @
 
-<<ref.label = "tabulate-2", message=FALSE, warning = FALSE>>=
+<<make_pleio_tib-11, ref.label="make_pleio_tib-2">>=
 @
 
-<<ref.label="make_pleio_tib-2">>=
-@
-
-<<ref.label = "make_pm-2">>=
+<<make_pm-11, ref.label = "make_pm-2">>=
 @
@@
 
-<<ref.label = "scatter_by_local-2">>=
+<<scatter_by_local-11, ref.label = "scatter_by_local-2">>=
 @
 
-<<ref.label="make_scols-2">>=
+<<make_scols-11, ref.label="make_scols-2">>=
 @
 
-<<ref.label="heat8-2">>=
+<<heat8-11, ref.label="heat8-2">>=
 @
 
-<<ref.label = "make_scols-extra-2">>=
+<<make_scols_extra-11, ref.label = "make_scols-extra-2">>=
 @
 
-<<ref.label = "count_med">>=
+<<count_med-11, ref.label = "count_med">>=
 @
 
-<<ref.label = "heat9-2">>=
+<<heat9-11, ref.label = "heat9-2">>=
 @
 
-<<ref.label = "index-v-peak-position-2">>=
+<<index-v-peak-position-11, ref.label = "index-v-peak-position-2">>=
 @
@@
 
-<<ref.label = "more_setup-2">>=
+<<more_setup-13, ref.label = "more_setup-2">>=
 @
 
-<<ref.label="readRDS-2">>=
+<<readRDS-13, ref.label="readRDS-2">>=
 @
 
-<<ref.label="make_local_annot-2">>=
+<<make_local_annot-13, ref.label="make_local_annot-2">>=
 @
 
-<<ref.label = "make_nonlocal_annot-2">>=
+<<make_local_annot2-13, ref.label = "make_local_annot2-2">>=
 @
 
-<<ref.label = "make_local_annot2-2">>=
+<<subset_expr_matrix-13, ref.label = "subset_expr_matrix-2">>=
 @
 
-<<ref.label = "subset_expr_matrix-2">>=
+<<tabulate-13, ref.label = "tabulate-2", message=FALSE, warning = FALSE>>=
 @
 
-<<ref.label = "tabulate-2", message=FALSE, warning = FALSE>>=
+<<make_pleio_tib-13, ref.label="make_pleio_tib-2">>=
 @
 
-<<ref.label="make_pleio_tib-2">>=
-@
-
-<<ref.label = "make_pm-2">>=
+<<make_pm-13, ref.label = "make_pm-2">>=
 @
@@
 
-<<ref.label = "scatter_by_local-2">>=
+<<scatter_by_local-13, ref.label = "scatter_by_local-2">>=
 @
 
-<<ref.label="make_scols-2">>=
+<<make_scols-13, ref.label="make_scols-2">>=
 @
 
-<<ref.label="heat8-2">>=
+<<heat8-13, ref.label="heat8-2">>=
 @
 
-<<ref.label = "make_scols-extra-2">>=
+<<make_scols-extra-13, ref.label = "make_scols-extra-2">>=
 @
 
-<<ref.label = "count_med">>=
+<<count_med-13, ref.label = "count_med">>=
 @
 
-<<ref.label = "heat9-2">>=
+<<heat9-13, ref.label = "heat9-2">>=
 @
 
-<<ref.label = "index-v-peak-position-2">>=
+<<index-v-peak-position-13, ref.label = "index-v-peak-position-2">>=
 @
 
-<<>>=
+<<def-2>>=
 hot_chr <- 2
@@
 
-<<>>=
+<<def-5>>=
 hot_chr <- 5
@@
 
-<<>>=
+<<def-7>>=
 hot_chr <- 7
@@
 
-<<>>=
+<<def-11>>=
 hot_chr <- 11
@@
 
-<<>>=
+<<def-13>>=
 hot_chr <- 13

### 2019-07-25 analysis/paper/paper.Rnw
removed cached files in analysis/figures
removed calls to knitr::include_graphics()

@@
 
-We generated scatter plots of pleiotropy test statistics vs. mediation LOD difference values for each of the five hotspots (Figure~\ref{}, \ref{}).
+We generated scatter plots of pleiotropy test statistics vs. mediation LOD difference values for each of the five hotspots (Figure).
 Each point represents a pairing of a local trait with a nonlocal trait.
@@
 <<fig.cap="Scatter plot of mediation LOD difference against pleiotropy LOD for all nonlocal-local trait pairs mapping to Chromosome 2 hotspot.">>=
-include_graphics(paste0("../figures/Chr", hot_chr, "_scatter.pdf"))
+#include_graphics(paste0("../figures/Chr", hot_chr, "_scatter.pdf"))
 @
@@
 <<panel1, fig.cap = paste0("Chr", hot_chr, "hotspot scatter plots of mediation LOD difference against pleiotropy LOD, separated by local trait.")>>=
-include_graphics(paste0("../figures/Chr", hot_chr, "_scatter-panel1.pdf"))
+#include_graphics(paste0("../figures/Chr", hot_chr, "_scatter-panel1.pdf"))
 @
@@
 <<panel2, eval = nfig[which(names(nfig) == paste0("Chr", hot_chr))] > 1>>=
-include_graphics(paste0("../figures/Chr", hot_chr, "_scatter-panel2.pdf"))
+#include_graphics(paste0("../figures/Chr", hot_chr, "_scatter-panel2.pdf"))
 @
@@
 <<panel3, eval = nfig[which(names(nfig) == paste0("Chr", hot_chr))] > 2>>=
-include_graphics(paste0("../figures/Chr", hot_chr, "_scatter-panel3.pdf"))
+#include_graphics(paste0("../figures/Chr", hot_chr, "_scatter-panel3.pdf"))
 @
@@
 <<fig.cap="Scatter plot of mediation LOD difference against pleiotropy LOD for all nonlocal-local trait pairs mapping to Chromosome 5 hotspot.">>=
-include_graphics("../figures/Chr5_scatter.pdf")
+#include_graphics("../figures/Chr5_scatter.pdf")
 @
@@
 <<fig.cap="Scatter plot of mediation LOD difference against pleiotropy LOD for all nonlocal-local trait pairs mapping to Chromosome 7 hotspot.">>=
-include_graphics("../figures/Chr7_scatter.pdf")
+#include_graphics("../figures/Chr7_scatter.pdf")
 @
@@
 <<fig.cap="Scatter plot of mediation LOD difference against pleiotropy LOD for all nonlocal-local trait pairs mapping to Chromosome 11 hotspot.">>=
-include_graphics("../figures/Chr11_scatter.pdf")
+#include_graphics("../figures/Chr11_scatter.pdf")
 @
@@
 <<fig.cap="Scatter plot of mediation LOD difference against pleiotropy LOD for all nonlocal-local trait pairs mapping to Chromosome 13 hotspot.">>=
-include_graphics("../figures/Chr13_scatter.pdf")
+#include_graphics("../figures/Chr13_scatter.pdf")
 @

### 2019-07-25 analysis/paper/paper.Rnw
completed parts of DESCRIPTION

@@
 library(broman)
+library(here)
 @
@@
 hot_upper <- hot_mid + 2
-fp <- paste0("../figures/Chr", hot_chr, "-")
-knitr::opts_chunk$set(fig.path = fp)
+fp <- paste0("Chr", hot_chr, "-")
+knitr::opts_chunk$set(fig.path = here("analysis", "figures", fp))
 @
@@
 <<readRDS-2>>=
-pleio_mixed_list <- readRDS(paste0("_cache/Chr", hot_chr, "_pleio.rds"))
-pleio_local_list <- readRDS(paste0("_cache/Chr", hot_chr, "_pleio_local.rds"))
-pleio_nonlocal_list <- readRDS(paste0("_cache/Chr", hot_chr, "_pleio_nonlocal.rds"))
-med_list <- readRDS(paste0("_cache/Chr", hot_chr, "_mediations.rds"))
+pleio_mixed_list <- readRDS(here("analysis", "data", paste0("Chr", hot_chr, "_pleio.rds")))
+pleio_local_list <- readRDS(here("analysis", "data", paste0("Chr", hot_chr, "_pleio_local.rds")))
+pleio_nonlocal_list <- readRDS(here("analysis", "data", paste0("Chr", hot_chr, "_pleio_nonlocal.rds")))
+med_list <- readRDS(here("analysis", "data", paste0("Chr", hot_chr, "_mediations.rds")))
 @

### 2019-07-26 analysis/paper/paper.Rnw
replaced file.path calls with here::here

@@
 <<readRDS-2>>=
-pleio_mixed_list <- readRDS(here("analysis", "data", paste0("Chr", hot_chr, "_pleio.rds")))
-pleio_local_list <- readRDS(here("analysis", "data", paste0("Chr", hot_chr, "_pleio_local.rds")))
-pleio_nonlocal_list <- readRDS(here("analysis", "data", paste0("Chr", hot_chr, "_pleio_nonlocal.rds")))
-med_list <- readRDS(here("analysis", "data", paste0("Chr", hot_chr, "_mediations.rds")))
+pleio_mixed_list <- readRDS(here("analysis", "data", "derived_data", paste0("Chr", hot_chr, "_pleio.rds")))
+pleio_local_list <- readRDS(here("analysis", "data", "derived_data", paste0("Chr", hot_chr, "_pleio_local.rds")))
+pleio_nonlocal_list <- readRDS(here("analysis", "data", "derived_data", paste0("Chr", hot_chr, "_pleio_nonlocal.rds")))
+med_list <- readRDS(here("analysis", "data", "derived_data", paste0("Chr", hot_chr, "_mediations.rds")))
 @
 
-<<load_attie>>=
-load("~/Box Sync/attie/keller2018-chr2-hotspot-chtc/data-to-ignore/Attie_DO378_eQTL_viewer_v1.Rdata")
+<<load_attie, eval=!file.exists(here("analysis", "data", "raw_data", "Attie_DO378_eQTL_viewer_v1.Rdata"))>>=
+download.file("https://datadryad.org/bitstream/handle/10255/dryad.166774/Attie_DO378_eQTL_viewer_v1.Rdata?sequence=2",
+              destfile = here("analysis", "data", "raw_data", "Attie_DO378_eQTL_viewer_v1.Rdata")
+              )
+load(here("analysis", "data", "raw_data", "Attie_DO378_eQTL_viewer_v1.Rdata"))
+saveRDS(dataset.islet.rnaseq$lod.peaks, here("analysis", "data", "derived_data", "lod_peaks.rds"))
+saveRDS(dataset.islet.rnaseq$annots, here("analysis", "data", "derived_data", "annots.rds"))
+saveRDS(dataset.islet.rnaseq$expr, here("analysis", "data", "derived_data", "expr.rds"))
+@
+
+<<read_attie_files>>=
+lod_peaks <- readRDS(here("analysis", "data", "derived_data", "lod_peaks.rds"))
+annots <- readRDS(here("analysis", "data", "derived_data", "annots.rds"))
+@
+
+<<load_tidyverse>>=
 library(tidyverse)
@@
 <<make_local_annot-2>>=
-local_annot <- dataset.islet.rnaseq$lod.peaks %>%
+local_annot <- lod_peaks %>%
   filter(chrom == hot_chr, pos <= hot_upper, pos >= hot_lower) %>%
-  left_join(y = dataset.islet.rnaseq$annots, by = c("annot.id" = "gene_id")) %>%
+  left_join(y = annots, by = c("annot.id" = "gene_id")) %>%
   filter(!duplicated(annot.id)) %>%
@@
 <<make_nonlocal_annot-2>>=
-nonlocal_annot <- dataset.islet.rnaseq$lod.peaks %>%
+nonlocal_annot <- lod_peaks %>%
   filter(chrom == hot_chr, pos <= hot_upper, pos >= hot_lower) %>%
-  inner_join(y = dataset.islet.rnaseq$annots, by = c("annot.id" = "gene_id")) %>%
+  inner_join(y = annots, by = c("annot.id" = "gene_id")) %>%
   filter(!duplicated(annot.id)) %>%
@@
 <<subset_expr_matrix-2>>=
-hot_local <- dataset.islet.rnaseq$expr[, colnames(dataset.islet.rnaseq$expr) %in% local_annot2$annot.id]
-hot_nonlocal <- dataset.islet.rnaseq$expr[, colnames(dataset.islet.rnaseq$expr) %in% nonlocal_annot$annot.id]
+hot_local <- expr[, colnames(expr) %in% local_annot2$annot.id]
+hot_nonlocal <- expr[, colnames(expr) %in% nonlocal_annot$annot.id]
 nnonlocal <- nrow(nonlocal_annot)
@@
 fnpdf <- paste0("Chr", hot_chr, "_scatter.pdf")
-ggsave(plot = p, filename = file.path("../figures", fnsvg))
-ggsave(plot = p, filename = file.path("../figures", fnpdf))
+ggsave(plot = p, filename = here("analysis", "figures", fnsvg))
+ggsave(plot = p, filename = here("analysis", "figures", fnpdf))
 p
@@
 fnpdf <- paste0("Chr", hot_chr, "_scatter-panel.pdf")
-ggsave(plot = p, filename = file.path("../figures", fnsvg), height = ceiling(nlocal / 2))
-ggsave(plot = p, filename = file.path("../figures", fnpdf), height = ceiling(nlocal / 2))
+ggsave(plot = p, filename = here("analysis", "figures", fnsvg), height = ceiling(nlocal / 2))
+ggsave(plot = p, filename = here("analysis", "figures", fnpdf), height = ceiling(nlocal / 2))
 
@@
   fnpdf <- paste0("Chr", hot_chr, "_scatter-panel", i, ".pdf")
-  ggsave(filename = file.path("../figures", fnsvg), height = ceiling(n_panels / 2))
-  ggsave(filename = file.path("../figures", fnpdf), height = ceiling(n_panels / 2))
+  ggsave(filename = here("analysis", "figures", fnsvg), height = ceiling(n_panels / 2))
+  ggsave(filename = here("analysis", "figures", fnpdf), height = ceiling(n_panels / 2))
 }
@@
 fnpdf <- paste0("Chr", hot_chr, "_heat", ".pdf")
-plotly::orca(p, file.path("../figures", fnsvg))
-plotly::orca(p, file.path("../figures", fnpdf))
+plotly::orca(p, here("analysis", "figures", fnsvg))
+plotly::orca(p, here("analysis", "figures", fnpdf))
 p

### 2019-07-26 README.Rmd
fixed pkg name in README.Rmd & knitted to md

@@
 ```
-# hotspots
+# qtl2hotspots
 
 <!-- badges: start -->
-[![Travis-CI Build Status](https://travis-ci.org/fboehm/hotspots.svg?branch=master)](https://travis-ci.org/fboehm/hotspots)
+[![Travis-CI Build Status](https://travis-ci.org/fboehm/qtl2hotspots.svg?branch=master)](https://travis-ci.org/fboehm/qtl2hotspots)
 
@@
 
-The goal of hotspots is to analyze the Keller et al. (2018) expression trait hotspots.
+The goal of qtl2hotspots is to analyze the Keller et al. (2018) expression trait hotspots.
 
@@
 ```{bash, docker_pull, eval = FALSE}
-docker pull fjboehm/hotspots
+docker pull fjboehm/qtl2hotspots
 ```
@@
 install.packages("devtools")
-devtools::install_github("fboehm/hotspots")
+devtools::install_github("fboehm/qtl2hotspots")
 ```
@@
 ```{r example}
-#library(hotspots)
+#library(qtl2hotspots)
 ## basic example code
@@
 
-Please note that the 'hotspots' project is released with a
+Please note that the 'qtl2hotspots' project is released with a
 [Contributor Code of Conduct](.github/CODE_OF_CONDUCT.md).

### 2019-07-26 analysis/paper/paper.Rnw
added svglite to Imports & turned on 'echo' for Rnw file

@@
 concordance=TRUE,
-echo = FALSE,
+echo = TRUE,
 fig.lp = "fig:",
@@
 the hotspot center.
+
+\todo[inline]{add a "table 1" detailing the keller hotspot data.
+Columns: number of local traits, number of nonlocal traits, center position, keller mediator, number of nonlocal traits mediated by keller mediator}
 
@@
 statistical computing environment \citep{r}.
+We initially plotted all local-nonlocal pairs for each hotspot.
+We highlighted in blue those points that correspond to pairs involving the putative mediator per \citet{keller2018genetic}.
+
+We created heatmaps to examine patterns among pleiotropy test results in each hotspot.
+We represented the pleiotropy test statistics as a symmetric matrix, where each row was one expression
+trait and each column was an expression trait. Each cell, then, contained the pleiotropy test statistic
+from the analysis involving the row's expression trait and the column's expression trait.
+To create a heatmap, we first performed two hierarchical clusterings, one on the submatrix involving only
+nonlocal expression traits and a second on the submatrix involving only local expression traits. We
+calculated the Euclidean distances between all trait pairs before clustering. We ordered the rows and
+columns according to optimal leaf ordering \todo[inline]{what is the OLO citation?} We then combined the
+local and nonlocal traits into a single heatmap. To achieve greater coloring contrasts in the heatmaps, we
+set the color spectrum limits at 0 (dark blue) and 5 (yellow). Test statistics above 5 are represented by
+yellow.
 
@@
 
-<<load_attie, eval=!file.exists(here("analysis", "data", "raw_data", "Attie_DO378_eQTL_viewer_v1.Rdata"))>>=
+<<load_attie, eval=!file.exists(here("analysis", "data", "derived_data", "expr.rds"))>>=
 download.file("https://datadryad.org/bitstream/handle/10255/dryad.166774/Attie_DO378_eQTL_viewer_v1.Rdata?sequence=2",

### 2019-07-26 README.Rmd
fixed README

@@
 
-The goal of qtl2hotspots is to analyze the Keller et al. (2018) expression trait hotspots.
+The goal of `qtl2hotspots` is to analyze the Keller et al. (2018) expression trait hotspots with both mediation methods and pleiotropy tests.
 
@@
 
-To render the manuscript `paper.Rnw`, into pdf, use a shell to enter the following text from the 'hotspots' directory:
+To render the manuscript `paper.Rnw`, into pdf, use a shell to enter the following text from the `qtl2hotspots` directory:
 
@@
 
-If you're not using Docker, and have cloned the repository from Github, you can use the Makefile to render `paper.Rnw`. Just type `make` in the shell (from within the `hotspots` directory):
+If you're not using Docker, and have cloned the repository from Github, you can use the Makefile to render `paper.Rnw`. Just type `make` in the shell (from within the `qtl2hotspots` directory):
 
@@
 
-Please note that the 'qtl2hotspots' project is released with a
+Please note that the `qtl2hotspots` project is released with a
 [Contributor Code of Conduct](.github/CODE_OF_CONDUCT.md).

### 2019-07-26 analysis/paper/paper.Rnw
added R code to read expr file

@@
 annots <- readRDS(here("analysis", "data", "derived_data", "annots.rds"))
+expr <- readRDS(here("analysis", "data", "derived_data", "expr.rds"))
 @

### 2019-07-26 analysis/paper/paper.Rnw
removed orca calls

@@
   hide_guides()
-fnsvg <- paste0("Chr", hot_chr, "_heat", ".svg")
-fnpdf <- paste0("Chr", hot_chr, "_heat", ".pdf")
-plotly::orca(p, here("analysis", "figures", fnsvg))
-plotly::orca(p, here("analysis", "figures", fnpdf))
 p

### 2019-07-26 analysis/paper/paper.Rnw
changed heatmaps to landscape pages

@@
 
-<<heat9-2, fig.height = 7, fig.width = 7, out.width = "\\textwidth">>=
+\begin{landscape}
+
+<<heat9-2, fig.height = 7, fig.width = 10, out.width = "\\textwidth">>=
 ord <- match(rn, rownames(pm))
 
-p <- pm[ord, ord] %>%
+pm[ord, ord] %>%
   apply(FUN = trunc2, MARGIN = 2, threshold = 5) %>%
@@
   hide_guides()
-p
 @
+
+\end{landscape}
 
@@
 
-<<heat9-5, ref.label = "heat9-2">>=
+\begin{landscape}
+
+<<heat9-5, ref.label = "heat9-2", fig.height = 7, fig.width = 10, out.width="\\textwidth">>=
 @
+
+\end{landscape}
 
@@
 
-<<heat9-7, ref.label = "heat9-2">>=
+\begin{landscape}
+
+<<heat9-7, ref.label = "heat9-2", fig.height = 7, fig.width = 10, out.width="\\textwidth">>=
 @
+
+\end{landscape}
 
@@
 
-<<heat9-11, ref.label = "heat9-2">>=
+\begin{landscape}
+
+<<heat9-11, ref.label = "heat9-2", fig.height = 7, fig.width = 10, out.width="\\textwidth">>=
 @
+
+\end{landscape}
 
@@
 
-<<heat9-13, ref.label = "heat9-2">>=
+\begin{landscape}
+
+<<heat9-13, ref.label = "heat9-2", fig.height = 7, fig.width = 10, out.width="\\textwidth">>=
 @
+
+\end{landscape}
 

### 2019-07-27 analysis/paper/paper.Rnw
fixed missing chunk problem in Rnw file

--- 2019-07-26 analysis/paper/paper.Rnw
@@
 %\usepackage{etoolbox} % for messing with page style & title pages
-\usepackage{lscape}
+%\usepackage{lscape}
+\usepackage{pdflscape}
 \usepackage{fancyhdr}
@@
 
-<<heat9-2, fig.height = 7, fig.width = 7, out.width = "\\textwidth">>=
+<<heat9-2>>=
 ord <- match(rn, rownames(pm))
@@
   hide_guides()
-p
+fnhtml <- here("analysis", "figures", paste0("Chr", hot_chr, "hm.html"))
+fnpdf <- here("analysis", "figures", paste0("Chr", hot_chr, "hm.pdf"))
+htmlwidgets::saveWidget(widget = p,
+                        file = fnhtml
+                        )
+webshot::webshot(url = fnhtml,
+                 file = fnpdf,
+                 cliprect = "viewport"
+                 )
 @
+
+\newpage
+\begin{landscape}
+
+<<hm_plot-2>>=
+knitr::include_graphics(path = paste0("../figures/Chr", hot_chr, "hm.pdf"))
+@
+
+\end{landscape}
 
@@
 
+\newpage
+\begin{landscape}
+
+<<hm_plot-5, ref.label = "hm_plot-2">>=
+
+@
+
+\end{landscape}
+
 <<index-v-peak-position-5, ref.label = "index-v-peak-position-2">>=
@@
 <<make_local_annot-7, ref.label="make_local_annot-2">>=
+@
+
+<<make_nonlocal_annot-7, ref.label = "make_nonlocal_annot-2">>=
 @
@@
 
+\newpage
+\begin{landscape}
+
+<<hm_plot-7, ref.label = "hm_plot-2">>=
+
+@
+
+\end{landscape}
+
 <<index-v-peak-position-7, ref.label = "index-v-peak-position-2">>=
@@
 <<make_local_annot-11, ref.label="make_local_annot-2">>=
+@
+
+<<make_nonlocal_annot-11, ref.label = "make_nonlocal_annot-2">>=
 @
@@
 
+\newpage
+\begin{landscape}
+
+<<hm_plot-11, ref.label = "hm_plot-2">>=
+
+@
+
+\end{landscape}
+
 <<index-v-peak-position-11, ref.label = "index-v-peak-position-2">>=
@@
 <<make_local_annot-13, ref.label="make_local_annot-2">>=
+@
+
+<<make_nonlocal_annot-13, ref.label = "make_nonlocal_annot-2">>=
 @
@@
 @
+
+\newpage
+\begin{landscape}
+
+<<hm_plot-13, ref.label = "hm_plot-2">>=
+
+@
+
+\end{landscape}
 

### 2019-07-27 analysis/paper/hotspots-paper.Rmd
created a Rmd file from the paper.Rnw file

---
title: "Expression QTL hotspot dissection with mediation analysis and pleiotropy testing"
author: "Frederick J. Boehm, Brian S. Yandell, and Karl W. Broman"
date: "7/27/2019"
bibliography: research.bib
output: 
  bookdown::html_document2
---

# Introduction

Genetics studies in model organisms like mice can identify
genomic regions that affect complex traits, such as systolic blood pressure and body weight [@sax1923association;@soller1976power;@lander1989mapping;@broman2009guide;@jansen2007quantitative]. 
These genomic regions are called "quantitative trait loci" or "QTL".
A genome-wide QTL "scan" reveals
associations between genotypes and phenotypes by considering each position along the genome,
one at a time, as a candidate QTL for the trait of interest.
A region with strong evidence of association with a complex trait, then, defines a QTL (for that trait).
Because nearby markers have correlated
genotypes, a QTL in a two-parent cross often spans multiple megabases in length
and may contain more than a hundred genes.
Identification of a causal gene (for a given complex trait) from among those genes near the QTL is
challenging and may require costly and time-consuming experiments.
The growing need for greater QTL mapping resolution fueled development over the
last two decades of model organism multiparental populations for high-resolution QTL mapping [@de2017back;@churchill2004collaborative;@svenson2012high;@huang2012multiparent;@shivakumar2018soybean;@huang2011analysis;@kover2009multiparent;@tisne2017identification;@stanley2017genetic].
With experimentalists now measuring tens of thousands of biomolecular traits
in multiparental populations, the systems genetics community needs multivariate
statistical tools to fully examine the large volumes of data [@keller2018genetic;@chick2016defining].
Identifying loci that affect multiple traits can aid in identifying biomolecular interactions and in clarifying complex trait genetic architecture. A QTL that affects more than one trait is called a "pleiotropic" QTL.
A test of pleiotropy vs. separate QTL is one multivariate statistical tool that will inform
complex trait genetics by enabling researchers to identify the number of
unique QTL in a genomic region of interest.

```{r load-broman}
library(broman)
library(here)
```

\section{Methods}

@keller2018genetic identified five expression trait hotspots in their
study of pancreatic islet gene expression in 378 Diversity Outbred mice. They
defined a hotspot to be a 4-Mb region that affects at least 100 nonlocal expression
traits.

For each hotspot, we identified a set of local expression traits and a
set of nonlocal expression traits. All traits demonstrated LOD peaks that met
the genome-wide significance threshold, 7.18 [@keller2018genetic].
The nonlocal expression traits arose from genes
on other chromosomes, while the local traits reflected expression levels for
genes on the same chromosome as the hotspot and within 4 Mb of the hotspot
center. We ignored transcripts for genes on the hotspot chromosome but not within 4 Mb of
the hotspot center.

## Pleiotropy testing

### Two-dimensional, two-QTL scans

For each hotspot, we tested all pairs of traits, regardless of whether the traits were local or nonlocal. We performed a
two-dimensional, two-QTL scan over a genomic region that included the entirety of the
hotspot. We fitted bivariate linear mixed effects models at each ordered pair of markers in a two-dimensional grid [@boehm2019testing].

\begin{equation}
vec(Y) = X vec(B) + vec(G) + vec(E)
\label{eq:mvlmm}
\end{equation}

\begin{equation}
G \sim MN_{n x 2}(0, K, V_g)
\end{equation}

\begin{equation}
E \sim MN_{n x 2}(0, I_n, V_e)
\end{equation}

and $G$ and $E$ are independent.

\begin{equation}
\hat B = (X^T\hat \Sigma^{-1}X)^{-1}X^T\hat\Sigma^{-1}Y
\label{eq:Bhat}
\end{equation}

\begin{equation}
\hat \Sigma = \hat V_g \otimes \hat K + \hat V_e \otimes I_n
\label{eq:Sigmahat}
\end{equation}

\begin{equation}
l_{10}(\hat B, \hat \Sigma) = - \frac{1}{2}\log_{10}\left((Y - X\hat B)^T\hat \Sigma^{-1}(Y - X\hat B)\right) - n\log_{10}(2\pi) - \log_{10}|\hat \Sigma|
\label{eq:loglik}
\end{equation}

For $n$ mice, $\Sigma$ is a $2n$ by $2n$ covariance matrix, while $V_g$ and $V_e$ are $2$ by $2$ covariance matrices. Estimates $\hat V_g$ and $\hat V_e$ are obtained via an expectation-maximization algorithm [@zhou2014efficient;@gemma2].

### Calculating likelihood ratio test statistics}
From the two-dimensional scans, we calculated pleiotropy likelihood ratio test statistics for each
pair of traits [@jiang1995multiple;@boehm2019testing;@boehm2019thesis]. Because of the required
computing time for bootstrap analyses, we didn't obtain p-values for the pleiotropy tests. Instead,
we worked directly with the test statistic values.

\begin{equation}
LOD = \log_{10} \left(\frac{\max_{\lambda_1, \lambda_2}L(B, \Sigma, \lambda_1, \lambda_2)}{\max_{\lambda}L(B, \Sigma, \lambda, \lambda)} \right)
\label{eq:lrt}
\end{equation}

\begin{equation}
ll(\lambda_1, \lambda_2)
\end{equation}

## Mediation analysis

A mediation analysis in systems genetics begins with an identified QTL-expression trait association. In our case, we choose a QTL that affects a nonlocal trait.

For each hotspot and each local trait - nonlocal trait pair, we also performed a
mediation analysis to assess the extent which the local trait mediates the QTL effect on the nonlocal trait. Each mediation analysis involved fitting four linear models (Equations~\ref{model1}, \ref{model2}, \ref{model3}, \ref{model4}) [@chick2016defining].

\begin{equation}
Y = \alpha 1 + WC + E
\label{model1}
\end{equation}

\begin{equation}
Y = XB + WC + E
\label{model2}
\end{equation}

\begin{equation}
Y = \alpha 1 + WC + M\beta + E
\label{model3}
\end{equation}

\begin{equation}
Y = XB + WC + M\beta + E
\label{model4}
\end{equation}

In all mediation analyses, the nonlocal transcripts are the targets, while the local
transcripts serve as candidate mediators of the QTL-target relationship.
In Equations~\ref{model1} through \ref{model4}, $X$ is a matrix of founder allele dosages,
$B$ is a matrix
of founder allele effects, $W$ is a matrix of covariates, $C$ is a matrix of covariate
effects, $M$ is a vector of values for a single candidate mediator, $\beta$ is the effect
of the mediator on the complex trait, 1 is a vector with all entries set to 1, $\alpha$
is the mean trait value, and $E$ is a vector of normally distributed random errors.

## Visualizations

We created scatter plots with the `ggplot2` package [@ggplot2] for the R
statistical computing environment [@r]. We initially plotted all local-nonlocal pairs for each hotspot. We highlighted in blue those points that correspond to pairs involving the putative mediator per [@keller2018genetic].

We created heatmaps to examine patterns among pleiotropy test results in each hotspot.
We represented the pleiotropy test statistics as a symmetric matrix, where each row was one expression
trait and each column was an expression trait. Each cell, then, contained the pleiotropy test statistic
from the analysis involving the row's expression trait and the column's expression trait.
To create a heatmap, we first performed two hierarchical clusterings, one on the submatrix involving only
nonlocal expression traits and a second on the submatrix involving only local expression traits. We
calculated the Euclidean distances between all trait pairs before clustering. We ordered the rows and
columns according to optimal leaf ordering. We then combined the
local and nonlocal traits into a single heatmap. To achieve greater coloring contrasts in the heatmaps, we
set the color spectrum limits at 0 (dark blue) and 5 (yellow). Test statistics above 5 are represented by
yellow.

# Results

We generated scatter plots of pleiotropy test statistics vs. mediation LOD difference values for each of the five hotspots (Figures).
Each point represents a pairing of a local trait with a nonlocal trait.
The mediation LOD difference value is that observed when the local trait is considered
as a candidate mediator of the QTL-nonlocal trait association.

In all five scatter plots, we observe few
points in the upper right quadrant of the figure.
That is, few pairs have high pleiotropy test statistics and
high mediation LOD difference values.

```{r setup}
library(qtl2hotspots)
pleio_threshold <- 4
ult <- 15 # univariate lod threshold
```

```{r setup_continued-2}
hot_chr <- 2
hot_mid <- 165.5
keller_mediator <- "Hnf4a"
```

```{r more_setup-2}
hot_lower <- hot_mid - 2
hot_upper <- hot_mid + 2
fp <- paste0("Chr", hot_chr, "-")
knitr::opts_chunk$set(fig.path = here("analysis", "figures", fp))
```

```{r readRDS-2}
pleio_mixed_list <- readRDS(here("analysis", "data", "derived_data", paste0("Chr", hot_chr, "_pleio.rds")))
pleio_local_list <- readRDS(here("analysis", "data", "derived_data", paste0("Chr", hot_chr, "_pleio_local.rds")))
pleio_nonlocal_list <- readRDS(here("analysis", "data", "derived_data", paste0("Chr", hot_chr, "_pleio_nonlocal.rds")))
med_list <- readRDS(here("analysis", "data", "derived_data", paste0("Chr", hot_chr, "_mediations.rds")))
```

```{r load_attie, eval=!file.exists(here("analysis", "data", "derived_data", "expr.rds"))}
download.file("https://datadryad.org/bitstream/handle/10255/dryad.166774/Attie_DO378_eQTL_viewer_v1.Rdata?sequence=2",
              destfile = here("analysis", "data", "raw_data", "Attie_DO378_eQTL_viewer_v1.Rdata")
              )
load(here("analysis", "data", "raw_data", "Attie_DO378_eQTL_viewer_v1.Rdata"))
saveRDS(dataset.islet.rnaseq$lod.peaks, here("analysis", "data", "derived_data", "lod_peaks.rds"))
saveRDS(dataset.islet.rnaseq$annots, here("analysis", "data", "derived_data", "annots.rds"))
saveRDS(dataset.islet.rnaseq$expr, here("analysis", "data", "derived_data", "expr.rds"))
```

```{r read_attie_files}
lod_peaks <- readRDS(here("analysis", "data", "derived_data", "lod_peaks.rds"))
annots <- readRDS(here("analysis", "data", "derived_data", "annots.rds"))
expr <- readRDS(here("analysis", "data", "derived_data", "expr.rds"))
```

```{r load_tidyverse}
library(tidyverse)
options(tibble.print_max = 50, tibble.print_min = 20)
```

```{r make_local_annot-2}
local_annot <- lod_peaks %>%
  filter(chrom == hot_chr, pos <= hot_upper, pos >= hot_lower) %>%
  left_join(y = annots, by = c("annot.id" = "gene_id")) %>%
  filter(!duplicated(annot.id)) %>%
  filter(lod >= ult) %>%
  mutate(local_tx = chr == hot_chr) %>%
  filter(local_tx)
```

```{r make_nonlocal_annot-2}
nonlocal_annot <- lod_peaks %>%
  filter(chrom == hot_chr, pos <= hot_upper, pos >= hot_lower) %>%
  inner_join(y = annots, by = c("annot.id" = "gene_id")) %>%
  filter(!duplicated(annot.id)) %>%
  filter(lod >= 7.18) %>%
  mutate(local_tx = chr == hot_chr) %>%
  filter(!local_tx) %>%
  filter(!is.na(hotspot))
```

```{r make_local_annot2-2}
local_annot2 <- local_annot %>%
  filter(middle <= hot_upper + 5, middle >= hot_lower - 5) %>%
  arrange(desc(lod))
```

```{r subset_expr_matrix-2}
hot_local <- expr[, colnames(expr) %in% local_annot2$annot.id]
hot_nonlocal <- expr[, colnames(expr) %in% nonlocal_annot$annot.id]
nnonlocal <- nrow(nonlocal_annot)
nlocal <- nrow(local_annot2)
```

```{r tabulate-2, message=FALSE, warning = FALSE}
med_tib <- med_list %>%
  purrr::map(bind_rows) %>%
  bind_rows() %>%
  as_tibble() %>%
  rename(lod_without_med = lod_no_med, lod_with_med = lod_med) %>%
  mutate(lod_diff = lod_without_med - lod_with_med, lod_diff_proportion = lod_diff / lod_without_med) %>%
  left_join(local_annot2, by = c("local_gene_id" = "annot.id")) %>%
  left_join(nonlocal_annot, by = c("nonlocal_gene_id" = "annot.id")) %>%
  dplyr::select(symbol.x, symbol.y, lod_without_med, lod_with_med, lod_diff, lod_diff_proportion, local_gene_id, nonlocal_gene_id) %>%
  rename(local_symbol = symbol.x, nonlocal_symbol = symbol.y) %>%
  filter(!is.na(local_symbol))
```

```{r make_pleio_tib-2}
hot_annot <- local_annot2 %>%
  bind_rows(nonlocal_annot) %>%
  dplyr::select(annot.id, symbol)
pleio_tib <- pleio_mixed_list %>%
  purrr::map(.f = function(x) rename(x, gene1_id = local_gene_id, gene2_id = nonlocal_gene_id)) %>%
  c(pleio_local_list, pleio_nonlocal_list) %>%
  bind_rows() %>%
  rename(pleiotropy_lod = lod) %>%
  inner_join(hot_annot, by = c("gene1_id" = "annot.id")) %>%
  rename(gene1_symbol = symbol) %>%
  inner_join(hot_annot, by = c("gene2_id" = "annot.id")) %>%
  rename(gene2_symbol = symbol) %>%
  dplyr::select(gene1_symbol, gene2_symbol, pleiotropy_lod)
```

```{r make_pm-2}
pm <- pleio_tib %>%
  tibble_to_matrix(symmetric = TRUE)
```

```{r scatter-2}
p <- pleio_mixed_list %>%
  bind_rows() %>%
  rename(pleiotropy_lod = lod) %>%
  inner_join(med_tib, by = c("local_gene_id", "nonlocal_gene_id")) %>%
  dplyr::select(local_symbol, nonlocal_symbol, pleiotropy_lod, lod_diff, lod_diff_proportion, lod_without_med, lod_with_med, filename, nonlocal_gene_id, local_gene_id) %>%
  ggplot() + geom_point(mapping = aes(x = pleiotropy_lod, y = lod_diff, colour = local_symbol == keller_mediator, nonlocal = nonlocal_symbol), size = 0.05) + theme(legend.position = "none") + xlab("Pleiotropy LOD") + ylab("Mediation LOD difference")
fnsvg <- paste0("Chr", hot_chr, "_scatter.svg")
fnpdf <- paste0("Chr", hot_chr, "_scatter.pdf")
ggsave(plot = p, filename = here("analysis", "figures", fnsvg))
ggsave(plot = p, filename = here("analysis", "figures", fnpdf))
p
```

```{r scatter_by_local-2}
foo <- pleio_mixed_list %>%
  bind_rows() %>%
  rename(pleiotropy_lod = lod) %>%
  inner_join(med_tib, by = c("local_gene_id", "nonlocal_gene_id")) %>%
  dplyr::select(local_symbol, nonlocal_symbol, pleiotropy_lod, lod_diff, lod_diff_proportion, lod_without_med, lod_with_med, filename, nonlocal_gene_id, local_gene_id)
foo %>%
  ggplot() +
  geom_point(mapping = aes(x = pleiotropy_lod, y = lod_diff, colour = nonlocal_symbol, nonlocal = nonlocal_symbol), size = 0.1) +
  facet_wrap(. ~ local_symbol, nrow = ceiling(nlocal / 2), ncol = 2) +
  broman::karl_theme(strip.placement = "inside") +
  theme(legend.position = "none") +
  theme(
    strip.background = element_blank(),
    strip.text.x = element_blank()
) +
  xlab("Pleiotropy LOD") +
  ylab("Mediation LOD difference") +
  geom_text(aes(x = 1, y = 1, label = local_symbol), inherit.aes=FALSE)

fnsvg <- paste0("Chr", hot_chr, "_scatter-panel.svg")
fnpdf <- paste0("Chr", hot_chr, "_scatter-panel.pdf")
ggsave(filename = here("analysis", "figures", fnsvg), height = ceiling(nlocal / 2))
ggsave(filename = here("analysis", "figures", fnpdf), height = ceiling(nlocal / 2))

# put in multiple figures, if there are many local traits
n_figs <- 1 + nlocal %/% 18 # 18 is the number of panels per page in pdf
n_last <- nlocal %% 18 # number of panels on last figure
max_pleio_lod <- max(foo$pleiotropy_lod)
max_med_lod_difference <- max(foo$lod_diff)
min_med_lod_difference <- min(foo$lod_diff)

for (i in 1:n_figs){
  n_panels <- 18 * (i < n_figs) + n_last * (i == n_figs)
  foo %>%
    filter(local_symbol %in% local_annot2$symbol[((i - 1) * 18 + 1): (18 * i)]) %>%
    ggplot() +
      geom_point(mapping = aes(x = pleiotropy_lod, y = lod_diff, colour = nonlocal_symbol, nonlocal = nonlocal_symbol), size = 0.1) +
      facet_wrap(. ~ local_symbol, nrow = ceiling(n_panels / 2), ncol = 2) +
      broman::karl_theme() +
      theme(legend.position = "none") +
      theme(
        strip.background = element_blank(),
        strip.text.x = element_blank()
    ) +
      xlab("Pleiotropy LOD") + xlim(c(0, max_pleio_lod)) + ylim(c(min_med_lod_difference, max_med_lod_difference)) +
      ylab("Mediation LOD difference") +
      geom_text(size = 2, aes(x = max_pleio_lod / 2, y = 2 * max_med_lod_difference / 3, label = local_symbol), inherit.aes=FALSE)
  fnsvg <- paste0("Chr", hot_chr, "_scatter-panel", i, ".svg")
  fnpdf <- paste0("Chr", hot_chr, "_scatter-panel", i, ".pdf")
  ggsave(filename = here("analysis", "figures", fnsvg), height = ceiling(n_panels / 2))
  ggsave(filename = here("analysis", "figures", fnpdf), height = ceiling(n_panels / 2))
}
```

```{r load_heatmaply}
library(heatmaply)
```

```{r make_scols-2}
scols1 <- 1L + (rownames(pm) %in% local_annot2$symbol)
scols2 <- rownames(pm) %in% (med_tib %>% filter(local_symbol == keller_mediator, lod_diff > 1.5))$nonlocal_symbol
```

```{r heat8-2}
# order the nonlocal traits first
## indicator for being a nonlocal trait
indic_nonlocal <- rownames(pm) %in% nonlocal_annot$symbol
# define ord_nonlocal
ord_nonlocal <- pm[indic_nonlocal, indic_nonlocal] %>%
  dist() %>%
  hclust() %>%
  as.dendrogram() %>%
  dendextend::seriate_dendrogram(dist(pm[indic_nonlocal, indic_nonlocal])) %>%
  order.dendrogram()
rn_nonlocal <- rownames(pm[indic_nonlocal, indic_nonlocal][ord_nonlocal, ord_nonlocal])
# define ord_local
ord_local<- pm[!indic_nonlocal, !indic_nonlocal] %>%
  dist() %>%
  hclust() %>%
  as.dendrogram() %>%
  dendextend::seriate_dendrogram(dist(pm[!indic_nonlocal, !indic_nonlocal])) %>%
  order.dendrogram()
rn_local <- rownames(pm[!indic_nonlocal, !indic_nonlocal][ord_local, ord_local])
rn <- c(rn_nonlocal, rn_local)
```

```{r make_scols-extra-2}
scols_extra <- purrr::map(.x = local_annot2$symbol,
                          .f = function(x) {rownames(pm) %in% (med_tib %>% filter(local_symbol == x, lod_diff > 1.5))$nonlocal_symbol}
) %>%
  bind_cols() %>%
  (function(x){colnames(x) <- local_annot2$symbol; x}) %>%
  mutate(local = scols1)
```

```{r count_med}
n_med_tr <- 4 # arbitrary threshold
n_med_per_local <- med_tib %>%
  group_by(local_symbol) %>%
  filter(lod_diff > 1.5) %>%
  count() %>%
  filter(n >= n_med_tr)
n_med_per_local %>%
  arrange(desc(n)) %>%
  knitr::kable()

scols_pre <- scols_extra %>%
  dplyr::select(n_med_per_local$local_symbol)
# get ordering for columns of scols_pre
col_order <- scols_pre %>%
  t() %>%
  dist() %>%
  hclust() %>%
  as.dendrogram() %>%
  dendextend::seriate_dendrogram(dist(t(scols_pre))) %>%
  order.dendrogram()
scols <- scols_pre[ , col_order] %>%
  mutate(local = scols1)
```

```{r heat9-2}
ord <- match(rn, rownames(pm))

p <- pm[ord, ord] %>%
  apply(FUN = trunc2, MARGIN = 2, threshold = 5) %>%
  heatmaply(Rowv = FALSE,
            Colv = FALSE,
            cexRow = 0.2,
            cexCol = 0.2,
            symm = TRUE,
            row_side_colors = scols[ord, ],
            col_side_colors = tibble(local = scols1[ord]),
            limits = c(0, 5)
            ) %>%
  hide_guides()
fnhtml <- here("analysis", "figures", paste0("Chr", hot_chr, "hm.html"))
fnpdf <- here("analysis", "figures", paste0("Chr", hot_chr, "hm.pdf"))
htmlwidgets::saveWidget(widget = p,
                        file = fnhtml
                        )
webshot::webshot(url = fnhtml,
                 file = fnpdf,
                 cliprect = "viewport"
                 )
```

```{r index-v-peak-position-2}
tibble(symbol = rownames(pm[indic_nonlocal, indic_nonlocal])[ord_nonlocal], index = 1:length(symbol)) %>%
  left_join(nonlocal_annot, by = "symbol") %>%
  ggplot() + geom_point(aes(y = pos, x = index))
```

```{r setup_continued-5}
hot_chr <- 5
hot_mid <- 146
keller_mediator <- "Pdx1"
```

```{r more_setup-5, ref.label = "more_setup-2"}
```

```{r readRDS-5, ref.label="readRDS-2"}
```

```{r make_local_annot-5, ref.label="make_local_annot-2"}
```

```{r make_nonlocal_annot-5, ref.label = "make_nonlocal_annot-2"}
```

```{r make_local_annot2-5, ref.label = "make_local_annot2-2"}
```

```{r subset_expr_matrix-5, ref.label = "subset_expr_matrix-2"}
```

```{r tabulate-5, ref.label = "tabulate-2", message=FALSE, warning = FALSE}
```

```{r make_pleio_tib-5, ref.label="make_pleio_tib-2"}
```

```{r make_pm-5, ref.label = "make_pm-2"}
```

```{r scatter-5, ref.label="scatter-2"}
```

```{r scatter_by_local-5, ref.label = "scatter_by_local-2"}
```

```{r make_scols-5, ref.label="make_scols-2"}
```

```{r heat8-5, ref.label="heat8-2"}
```

```{r make_scols-extra-5, ref.label = "make_scols-extra-2"}
```

```{r count_med-5, ref.label = "count_med"}
```

```{r heat9-5, ref.label = "heat9-2"}
```

```{r hm_plot-5, ref.label = "hm_plot-2"}
```

```{r index-v-peak-position-5, ref.label = "index-v-peak-position-2"}
```

```{r setup_continued-7}
hot_chr <- 7
hot_mid <- 46
keller_mediator <- "Fam83e"
```

```{r more_setup-7, ref.label = "more_setup-2"}
```

```{r readRDS-7, ref.label="readRDS-2"}
```

```{r make_local_annot-7, ref.label="make_local_annot-2"}
```

```{r make_nonlocal_annot-7, ref.label = "make_nonlocal_annot-2"}
```

```{r make_local_annot2-7, ref.label = "make_local_annot2-2"}
```

```{r subset_expr_matrix-7, ref.label = "subset_expr_matrix-2"}
```

```{r tabulate-7, ref.label = "tabulate-2", message=FALSE, warning = FALSE}
```

```{r make_pleio_tib-7, ref.label="make_pleio_tib-2"}
```

```{r make_pm-7, ref.label = "make_pm-2"}
```

```{r scatter-7, ref.label="scatter-2"}
```

```{r scatter_by_local-7, ref.label = "scatter_by_local-2"}
```

```{r make_scols-7, ref.label="make_scols-2"}
```

```{r heat8-7, ref.label="heat8-2"}
```

```{r make_scols-extra-7, ref.label = "make_scols-extra-2"}
```

```{r count_med-7, ref.label = "count_med"}
```

```{r heat9-7, ref.label = "heat9-2"}
```

```{r hm_plot-7, ref.label = "hm_plot-2"}

```

```{r index-v-peak-position-7, ref.label = "index-v-peak-position-2"}
```

```{r setup_continued-11}
hot_chr <- 11
hot_mid <- 71
keller_mediator <- "Sat2"
```

```{r more_setup-11, ref.label = "more_setup-2"}
```

```{r readRDS-11, ref.label="readRDS-2"}
```

```{r make_local_annot-11, ref.label="make_local_annot-2"}
```

```{r make_nonlocal_annot-11, ref.label = "make_nonlocal_annot-2"}
```

```{r make_local_annot2-11, ref.label = "make_local_annot2-2"}
```

```{r subset_expr_matrix-11, ref.label = "subset_expr_matrix-2"}
```

```{r tabulate-11, ref.label = "tabulate-2", message=FALSE, warning = FALSE}
```

```{r make_pleio_tib-11, ref.label="make_pleio_tib-2"}
```

```{r make_pm-11, ref.label = "make_pm-2"}
```

```{r scatter-11, ref.label="scatter-2"}
```

```{r scatter_by_local-11, ref.label = "scatter_by_local-2"}
```

```{r make_scols-11, ref.label="make_scols-2"}
```

```{r heat8-11, ref.label="heat8-2"}
```

```{r make_scols_extra-11, ref.label = "make_scols-extra-2"}
```

```{r count_med-11, ref.label = "count_med"}
```

```{r heat9-11, ref.label = "heat9-2"}
```

```{r hm_plot-11, ref.label = "hm_plot-2"}

```

```{r index-v-peak-position-11, ref.label = "index-v-peak-position-2"}
```

```{r setup_continued-13}
hot_chr <- 13
hot_mid <- 112.5
keller_mediator <- "Il6st"
```

```{r more_setup-13, ref.label = "more_setup-2"}
```

```{r readRDS-13, ref.label="readRDS-2"}
```

```{r make_local_annot-13, ref.label="make_local_annot-2"}
```

```{r make_nonlocal_annot-13, ref.label = "make_nonlocal_annot-2"}
```

```{r make_local_annot2-13, ref.label = "make_local_annot2-2"}
```

```{r subset_expr_matrix-13, ref.label = "subset_expr_matrix-2"}
```

```{r tabulate-13, ref.label = "tabulate-2", message=FALSE, warning = FALSE}
```

```{r make_pleio_tib-13, ref.label="make_pleio_tib-2"}
```

```{r make_pm-13, ref.label = "make_pm-2"}
```

```{r scatter-13, ref.label="scatter-2"}
```

```{r scatter_by_local-13, ref.label = "scatter_by_local-2"}
```

```{r make_scols-13, ref.label="make_scols-2"}
```

```{r heat8-13, ref.label="heat8-2"}
```

```{r make_scols-extra-13, ref.label = "make_scols-extra-2"}
```

```{r count_med-13, ref.label = "count_med"}
```

```{r heat9-13, ref.label = "heat9-2"}
```

```{r hm_plot-13, ref.label = "hm_plot-2"}

```

```{r index-v-peak-position-13, ref.label = "index-v-peak-position-2"}
```

```{r def-2}
hot_chr <- 2
```

```{r fig.cap="Scatter plot of mediation LOD difference against pleiotropy LOD for all nonlocal-local trait pairs mapping to Chromosome 2 hotspot."}
#include_graphics(paste0("../figures/Chr", hot_chr, "_scatter.pdf"))
```

```{r define_nfig}
nfig <- c(2, 2, 3, 3, 1) # number of panel figs per hotspot
names(nfig) <- paste0("Chr", c(2, 5, 7, 11, 13))
```

```{r panel1, fig.cap = paste0("Chr", hot_chr, "hotspot scatter plots of mediation LOD difference against pleiotropy LOD, separated by local trait.")}
include_graphics(paste0("../figures/Chr", hot_chr, "_scatter-panel1.pdf"))
```

```{r panel2, eval = nfig[which(names(nfig) == paste0("Chr", hot_chr))] > 1}
include_graphics(paste0("../figures/Chr", hot_chr, "_scatter-panel2.pdf"))
```

```{r panel3, eval = nfig[which(names(nfig) == paste0("Chr", hot_chr))] > 2}
include_graphics(paste0("../figures/Chr", hot_chr, "_scatter-panel3.pdf"))
```

```{r def-5}
hot_chr <- 5
```

```{r fig.cap="Scatter plot of mediation LOD difference against pleiotropy LOD for all nonlocal-local trait pairs mapping to Chromosome 5 hotspot."}
#include_graphics("../figures/Chr5_scatter.pdf")
```

```{r ref.label = "panel1", fig.cap = paste0("Chr", hot_chr, "hotspot scatter plots of mediation LOD difference against pleiotropy LOD, separated by local trait.")}
```

```{r ref.label = "panel2", eval = nfig[which(names(nfig) == paste0("Chr", hot_chr))] > 1}
```

```{r ref.label = "panel3", eval = nfig[which(names(nfig) == paste0("Chr", hot_chr))] > 2}
```

```{r def-7}
hot_chr <- 7
```

```{r fig.cap="Scatter plot of mediation LOD difference against pleiotropy LOD for all nonlocal-local trait pairs mapping to Chromosome 7 hotspot."}
#include_graphics("../figures/Chr7_scatter.pdf")
```

```{r ref.label = "panel1", fig.cap = paste0("Chr", hot_chr, "hotspot scatter plots of mediation LOD difference against pleiotropy LOD, separated by local trait.")}
```

```{r ref.label = "panel2", eval = nfig[which(names(nfig) == paste0("Chr", hot_chr))] > 1}
```

```{r ref.label = "panel3", eval = nfig[which(names(nfig) == paste0("Chr", hot_chr))] > 2}
```

```{r def-11}
hot_chr <- 11
```

```{r fig.cap="Scatter plot of mediation LOD difference against pleiotropy LOD for all nonlocal-local trait pairs mapping to Chromosome 11 hotspot."}
#include_graphics("../figures/Chr11_scatter.pdf")
```

```{r ref.label = "panel1", fig.cap = paste0("Chr", hot_chr, "hotspot scatter plots of mediation LOD difference against pleiotropy LOD, separated by local trait.")}
```

```{r ref.label = "panel2", eval = nfig[which(names(nfig) == paste0("Chr", hot_chr))] > 1}
```

```{r ref.label = "panel3", eval = nfig[which(names(nfig) == paste0("Chr", hot_chr))] > 2}
```

```{r def-13}
hot_chr <- 13
```

```{r fig.cap="Scatter plot of mediation LOD difference against pleiotropy LOD for all nonlocal-local trait pairs mapping to Chromosome 13 hotspot."}
#include_graphics("../figures/Chr13_scatter.pdf")
```

```{r ref.label = "panel1", fig.cap = paste0("Chr", hot_chr, "hotspot scatter plots of mediation LOD difference against pleiotropy LOD, separated by local trait.")}
```

```{r ref.label = "panel2", eval = nfig[which(names(nfig) == paste0("Chr", hot_chr))] > 1}
```

```{r ref.label = "panel3", eval = nfig[which(names(nfig) == paste0("Chr", hot_chr))] > 2}
```

\section{Discussion}

Pleiotropy tests complement mediation analyses in the following manner. When considering a
collection of potential mediators for a nonlocal trait, mediation analyses clearly identify the
mediator when there is, in fact, a mediator among the candidates. However, when there is no
mediator, mediation analyses provide little information about the genetics of the complex trait.
Pleiotropy testing is particularly useful in this setting, since a pleiotropy test always gives
an answer about the number of underlying QTL. A drawback of pleiotropy testing is that the scientific question that it addresses - do two traits share a single QTL - informs, but doesn't fully resolve, the genetics of the nonlocal complex traits under study.

\printbibliography

```{r install_here}
install.packages("here", repos = "https://cloud.r-project.org")
```

\subsection{Colophon}

This report was generated on \Sexpr{Sys.time()} using the following computational environment and dependencies:

\begin{singlespacing}
%\begin{landscape}
```{r colophon, cache = FALSE, echo = TRUE, tidy.opts=list(width.cutoff=50), size = "tiny", results = "markup"}
# which R packages and versions?
foo <- devtools::session_info()
readr::write_csv(x = foo$packages, path = "packages.csv")
knitr::kable(tibble(field = names(foo$platform), value = unlist(foo$platform)))
foo$packages %>%
  as_tibble() %>%
  dplyr::select(-4, -5, -10, -11) %>%
  knitr::kable()
```

%\end{landscape}

\end{singlespacing}

The current Git commit details are:

```{r git2r, cache = FALSE, echo = TRUE}
# what commit is this file at?
git2r::repository(here::here())
```

### 2019-07-27 analysis/paper/hotspots-paper.Rmd
simplified R code dealing with multipanel figs

@@
 output: 
-  bookdown::html_document2
+  - bookdown::html_document2
+  - github_document:
+      toc: true
+      toc_depth: 2
 ---
@@
 
-```{r define_nfig}
-nfig <- c(2, 2, 3, 3, 1) # number of panel figs per hotspot
-names(nfig) <- paste0("Chr", c(2, 5, 7, 11, 13))
-```
-
-```{r panel1, fig.cap = paste0("Chr", hot_chr, "hotspot scatter plots of mediation LOD difference against pleiotropy LOD, separated by local trait.")}
-include_graphics(paste0("../figures/Chr", hot_chr, "_scatter-panel1.pdf"))
-```
-
-```{r panel2, eval = nfig[which(names(nfig) == paste0("Chr", hot_chr))] > 1}
-include_graphics(paste0("../figures/Chr", hot_chr, "_scatter-panel2.pdf"))
-```
-
-```{r panel3, eval = nfig[which(names(nfig) == paste0("Chr", hot_chr))] > 2}
-include_graphics(paste0("../figures/Chr", hot_chr, "_scatter-panel3.pdf"))
-```
-
-```{r def-5}
-hot_chr <- 5
-```
-
-```{r fig.cap="Scatter plot of mediation LOD difference against pleiotropy LOD for all nonlocal-local trait pairs mapping to Chromosome 5 hotspot."}
-#include_graphics("../figures/Chr5_scatter.pdf")
-```
-
-```{r ref.label = "panel1", fig.cap = paste0("Chr", hot_chr, "hotspot scatter plots of mediation LOD difference against pleiotropy LOD, separated by local trait.")}
-```
-
-```{r ref.label = "panel2", eval = nfig[which(names(nfig) == paste0("Chr", hot_chr))] > 1}
-```
-
-```{r ref.label = "panel3", eval = nfig[which(names(nfig) == paste0("Chr", hot_chr))] > 2}
-```
-
-```{r def-7}
-hot_chr <- 7
-```
-
-```{r fig.cap="Scatter plot of mediation LOD difference against pleiotropy LOD for all nonlocal-local trait pairs mapping to Chromosome 7 hotspot."}
-#include_graphics("../figures/Chr7_scatter.pdf")
-```
-
-```{r ref.label = "panel1", fig.cap = paste0("Chr", hot_chr, "hotspot scatter plots of mediation LOD difference against pleiotropy LOD, separated by local trait.")}
-```
-
-```{r ref.label = "panel2", eval = nfig[which(names(nfig) == paste0("Chr", hot_chr))] > 1}
-```
-
-```{r ref.label = "panel3", eval = nfig[which(names(nfig) == paste0("Chr", hot_chr))] > 2}
-```
-
-```{r def-11}
-hot_chr <- 11
-```
-
-```{r fig.cap="Scatter plot of mediation LOD difference against pleiotropy LOD for all nonlocal-local trait pairs mapping to Chromosome 11 hotspot."}
-#include_graphics("../figures/Chr11_scatter.pdf")
-```
-
-```{r ref.label = "panel1", fig.cap = paste0("Chr", hot_chr, "hotspot scatter plots of mediation LOD difference against pleiotropy LOD, separated by local trait.")}
-```
-
-```{r ref.label = "panel2", eval = nfig[which(names(nfig) == paste0("Chr", hot_chr))] > 1}
-```
-
-```{r ref.label = "panel3", eval = nfig[which(names(nfig) == paste0("Chr", hot_chr))] > 2}
-```
-
-```{r def-13}
-hot_chr <- 13
-```
-
-```{r fig.cap="Scatter plot of mediation LOD difference against pleiotropy LOD for all nonlocal-local trait pairs mapping to Chromosome 13 hotspot."}
-#include_graphics("../figures/Chr13_scatter.pdf")
-```
-
-```{r ref.label = "panel1", fig.cap = paste0("Chr", hot_chr, "hotspot scatter plots of mediation LOD difference against pleiotropy LOD, separated by local trait.")}
-```
-
-```{r ref.label = "panel2", eval = nfig[which(names(nfig) == paste0("Chr", hot_chr))] > 1}
-```
-
-```{r ref.label = "panel3", eval = nfig[which(names(nfig) == paste0("Chr", hot_chr))] > 2}
-```
-
-\section{Discussion}
+# Discussion
 
@@
 
-\printbibliography
+# Colophon
 
-```{r install_here}
-install.packages("here", repos = "https://cloud.r-project.org")
-```
+This report was generated on ` r Sys.time()` using the following computational environment and dependencies:
 
-\subsection{Colophon}
-
-This report was generated on \Sexpr{Sys.time()} using the following computational environment and dependencies:
-
-\begin{singlespacing}
-%\begin{landscape}
 ```{r colophon, cache = FALSE, echo = TRUE, tidy.opts=list(width.cutoff=50), size = "tiny", results = "markup"}
 # which R packages and versions?
-foo <- devtools::session_info()
-readr::write_csv(x = foo$packages, path = "packages.csv")
-knitr::kable(tibble(field = names(foo$platform), value = unlist(foo$platform)))
-foo$packages %>%
-  as_tibble() %>%
-  dplyr::select(-4, -5, -10, -11) %>%
-  knitr::kable()
+devtools::session_info()
 ```
-
-%\end{landscape}
-
-\end{singlespacing}
 

### 2019-07-27 analysis/paper/hotspots-paper.Rmd
updated Rmd; replaced latex eqns with pandoc eqns

@@
 
-\begin{equation}
+$$
 vec(Y) = X vec(B) + vec(G) + vec(E)
 \label{eq:mvlmm}
-\end{equation}
+$$
 
-\begin{equation}
+$$
 G \sim MN_{n x 2}(0, K, V_g)
-\end{equation}
+$$
 
-\begin{equation}
+$$
 E \sim MN_{n x 2}(0, I_n, V_e)
-\end{equation}
+$$
 
@@
 
-\begin{equation}
+$$
 \hat B = (X^T\hat \Sigma^{-1}X)^{-1}X^T\hat\Sigma^{-1}Y
 \label{eq:Bhat}
-\end{equation}
+$$
 
-\begin{equation}
+$$
 \hat \Sigma = \hat V_g \otimes \hat K + \hat V_e \otimes I_n
 \label{eq:Sigmahat}
-\end{equation}
+$$
 
-\begin{equation}
+$$
 l_{10}(\hat B, \hat \Sigma) = - \frac{1}{2}\log_{10}\left((Y - X\hat B)^T\hat \Sigma^{-1}(Y - X\hat B)\right) - n\log_{10}(2\pi) - \log_{10}|\hat \Sigma|
 \label{eq:loglik}
-\end{equation}
+$$
 
@@
 
-\begin{equation}
+$$
 LOD = \log_{10} \left(\frac{\max_{\lambda_1, \lambda_2}L(B, \Sigma, \lambda_1, \lambda_2)}{\max_{\lambda}L(B, \Sigma, \lambda, \lambda)} \right)
 \label{eq:lrt}
-\end{equation}
+$$
 
-\begin{equation}
+$$
 ll(\lambda_1, \lambda_2)
-\end{equation}
+$$
 
@@
 
-\begin{equation}
+$$
 Y = \alpha 1 + WC + E
 \label{model1}
-\end{equation}
+$$
 
-\begin{equation}
+$$
 Y = XB + WC + E
 \label{model2}
-\end{equation}
+$$
 
-\begin{equation}
+$$
 Y = \alpha 1 + WC + M\beta + E
 \label{model3}
-\end{equation}
+$$
 
-\begin{equation}
+$$
 Y = XB + WC + M\beta + E
 \label{model4}
-\end{equation}
+$$
 
@@
   mutate(local = scols1)
+fnhtml <- here("analysis", "figures", paste0("Chr", hot_chr, "hm.html"))
 ```
 
-```{r heat9-2}
+```{r heat9-2, eval = !file.exists(fnhtml)}
 ord <- match(rn, rownames(pm))
@@
   hide_guides()
-fnhtml <- here("analysis", "figures", paste0("Chr", hot_chr, "hm.html"))
 fnpdf <- here("analysis", "figures", paste0("Chr", hot_chr, "hm.pdf"))
@@
                  )
+```
+
+```{r heat10-2}
+webshot::webshot(fnhtml)
 ```
@@
 
-```{r heat9-5, ref.label = "heat9-2"}
+```{r heat9-5, ref.label = "heat9-2", eval = !file.exists(fnhtml)}
 ```
 
-```{r hm_plot-5, ref.label = "hm_plot-2"}
+```{r heat10-5, ref.label="heat10-2"}
 ```
@@
 
-```{r heat9-7, ref.label = "heat9-2"}
+```{r heat9-7, ref.label = "heat9-2", eval = !file.exists(fnhtml)}
 ```
 
-```{r hm_plot-7, ref.label = "hm_plot-2"}
-
+```{r heat10-7, ref.label="heat10-2"}
 ```
@@
 
-```{r heat9-11, ref.label = "heat9-2"}
+```{r heat9-11, ref.label = "heat9-2", eval = !file.exists(fnhtml)}
 ```
 
-```{r hm_plot-11, ref.label = "hm_plot-2"}
-
+```{r heat10-11, ref.label="heat10-2"}
 ```
@@
 
-```{r heat9-13, ref.label = "heat9-2"}
+```{r heat9-13, ref.label = "heat9-2", eval = !file.exists(fnhtml)}
 ```
 
-```{r hm_plot-13, ref.label = "hm_plot-2"}
-
+```{r heat10-13, ref.label="heat10-2"}
 ```
@@
 ```{r index-v-peak-position-13, ref.label = "index-v-peak-position-2"}
-```
-
-```{r def-2}
-hot_chr <- 2
-```
-
-```{r fig.cap="Scatter plot of mediation LOD difference against pleiotropy LOD for all nonlocal-local trait pairs mapping to Chromosome 2 hotspot."}
-#include_graphics(paste0("../figures/Chr", hot_chr, "_scatter.pdf"))
 ```

### 2019-07-27 analysis/paper/hotspots-paper.Rmd
reordered output formats; added ggplotly calls

@@
 output: 
-  - bookdown::html_document2
   - github_document:
@@
       toc_depth: 2
+  - bookdown::html_document2
 ---
@@
 ggsave(plot = p, filename = here("analysis", "figures", fnpdf))
-p
+plotly::ggplotly(p)
 ```
 
-```{r scatter_by_local-2}
+```{r scatter_by_local-2, fig.height = ceiling(nlocal / 2)}
 foo <- pleio_mixed_list %>%
@@
   dplyr::select(local_symbol, nonlocal_symbol, pleiotropy_lod, lod_diff, lod_diff_proportion, lod_without_med, lod_with_med, filename, nonlocal_gene_id, local_gene_id)
-foo %>%
+p <- foo %>%
   ggplot() +
@@
 fnpdf <- paste0("Chr", hot_chr, "_scatter-panel.pdf")
-ggsave(filename = here("analysis", "figures", fnsvg), height = ceiling(nlocal / 2))
-ggsave(filename = here("analysis", "figures", fnpdf), height = ceiling(nlocal / 2))
+ggsave(p, filename = here("analysis", "figures", fnsvg), height = ceiling(nlocal / 2))
+ggsave(p, filename = here("analysis", "figures", fnpdf), height = ceiling(nlocal / 2))
+plotly::ggplotly(p)
+```
 
+```{r multipanel_scatter-2, eval = FALSE}
 # put in multiple figures, if there are many local traits
@@
 
-```{r heat9-2, eval = !file.exists(fnhtml)}
+```{r heat9-2, eval = TRUE}
 ord <- match(rn, rownames(pm))
 
-p <- pm[ord, ord] %>%
+(p <- pm[ord, ord] %>%
   apply(FUN = trunc2, MARGIN = 2, threshold = 5) %>%
@@
   hide_guides()
+)
 fnpdf <- here("analysis", "figures", paste0("Chr", hot_chr, "hm.pdf"))
@@
                         )
+```
+
+```{r webshot-2, eval = FALSE}
 webshot::webshot(url = fnhtml,
@@
 
-```{r heat10-2}
+```{r heat10-2, eval = FALSE}
 webshot::webshot(fnhtml)
@@
 
+```{r multipanel_scatter-5, ref.label = "multipanel_scatter-2", eval = FALSE}
+```
+
 ```{r make_scols-5, ref.label="make_scols-2"}
@@
 
-```{r heat9-5, ref.label = "heat9-2", eval = !file.exists(fnhtml)}
+```{r heat9-5, ref.label = "heat9-2", eval = TRUE}
 ```
@@
 
+```{r multipanel_scatter-7, ref.label = "multipanel_scatter-2", eval = FALSE}
+```
+
 ```{r make_scols-7, ref.label="make_scols-2"}
@@
 
-```{r heat9-7, ref.label = "heat9-2", eval = !file.exists(fnhtml)}
+```{r heat9-7, ref.label = "heat9-2", eval = TRUE}
 ```
@@
 
+```{r multipanel_scatter-11, ref.label = "multipanel_scatter-2", eval = FALSE}
+```
+
 ```{r make_scols-11, ref.label="make_scols-2"}
@@
 
-```{r heat9-11, ref.label = "heat9-2", eval = !file.exists(fnhtml)}
+```{r heat9-11, ref.label = "heat9-2", eval = TRUE}
 ```
@@
 
+```{r multipanel_scatter-13, ref.label = "multipanel_scatter-2", eval = FALSE}
+```
+
 ```{r make_scols-13, ref.label="make_scols-2"}
@@
 
-```{r heat9-13, ref.label = "heat9-2", eval = !file.exists(fnhtml)}
+```{r heat9-13, ref.label = "heat9-2", eval = TRUE}
 ```

### 2019-07-27 analysis/paper/hotspots-paper.Rmd
moved eval chunk argument to an "if" conditional statement

@@
 
-```{r load_attie, eval=!file.exists(here("analysis", "data", "derived_data", "expr.rds"))}
+```{r load_attie}
+if (!file.exists(here("analysis", "data", "derived_data", "expr.rds"))){
   download.file("https://datadryad.org/bitstream/handle/10255/dryad.166774/Attie_DO378_eQTL_viewer_v1.Rdata?sequence=2",
@@
   saveRDS(dataset.islet.rnaseq$expr, here("analysis", "data", "derived_data", "expr.rds"))
+}
 ```
@@
   dplyr::select(local_symbol, nonlocal_symbol, pleiotropy_lod, lod_diff, lod_diff_proportion, lod_without_med, lod_with_med, filename, nonlocal_gene_id, local_gene_id) %>%
-  ggplot() + geom_point(mapping = aes(x = pleiotropy_lod, y = lod_diff, colour = local_symbol == keller_mediator, nonlocal = nonlocal_symbol), size = 0.05) + theme(legend.position = "none") + xlab("Pleiotropy LOD") + ylab("Mediation LOD difference")
+  ggplot() + geom_point(mapping = aes(x = pleiotropy_lod, y = lod_diff, colour = local_symbol == keller_mediator, nonlocal = nonlocal_symbol, local = local_symbol), size = 0.05) + theme(legend.position = "none") + xlab("Pleiotropy LOD") + ylab("Mediation LOD difference")
 fnsvg <- paste0("Chr", hot_chr, "_scatter.svg")
@@
   ggplot() +
-  geom_point(mapping = aes(x = pleiotropy_lod, y = lod_diff, colour = nonlocal_symbol, nonlocal = nonlocal_symbol), size = 0.1) +
+  geom_point(mapping = aes(x = pleiotropy_lod, y = lod_diff, colour = nonlocal_symbol), size = 0.1) +
   facet_wrap(. ~ local_symbol, nrow = ceiling(nlocal / 2), ncol = 2) +
@@
     ggplot() +
-      geom_point(mapping = aes(x = pleiotropy_lod, y = lod_diff, colour = nonlocal_symbol, nonlocal = nonlocal_symbol), size = 0.1) +
+      geom_point(mapping = aes(x = pleiotropy_lod, y = lod_diff, colour = nonlocal_symbol), size = 0.1) +
       facet_wrap(. ~ local_symbol, nrow = ceiling(n_panels / 2), ncol = 2) +

### 2019-07-28 analysis/paper/hotspot-one-chromosome.Rmd
created child Rmd file

# Chr `r hot_chr`

```{r}
hot_lower <- hot_mid - 2
hot_upper <- hot_mid + 2
fp <- paste0("Chr", hot_chr, "-")
knitr::opts_chunk$set(fig.path = here("analysis", "figures", fp))
```

```{r}
pleio_mixed_list <- readRDS(here("analysis", "data", "derived_data", paste0("Chr", hot_chr, "_pleio.rds")))
pleio_local_list <- readRDS(here("analysis", "data", "derived_data", paste0("Chr", hot_chr, "_pleio_local.rds")))
pleio_nonlocal_list <- readRDS(here("analysis", "data", "derived_data", paste0("Chr", hot_chr, "_pleio_nonlocal.rds")))
med_list <- readRDS(here("analysis", "data", "derived_data", paste0("Chr", hot_chr, "_mediations.rds")))
```

```{r}
if (!file.exists(here("analysis", "data", "derived_data", "expr.rds"))){
  download.file("https://datadryad.org/bitstream/handle/10255/dryad.166774/Attie_DO378_eQTL_viewer_v1.Rdata?sequence=2",
              destfile = here("analysis", "data", "raw_data", "Attie_DO378_eQTL_viewer_v1.Rdata")
              )
  load(here("analysis", "data", "raw_data", "Attie_DO378_eQTL_viewer_v1.Rdata"))
  saveRDS(dataset.islet.rnaseq$lod.peaks, here("analysis", "data", "derived_data", "lod_peaks.rds"))
  saveRDS(dataset.islet.rnaseq$annots, here("analysis", "data", "derived_data", "annots.rds"))
  saveRDS(dataset.islet.rnaseq$expr, here("analysis", "data", "derived_data", "expr.rds"))
}
```

```{r}
lod_peaks <- readRDS(here("analysis", "data", "derived_data", "lod_peaks.rds"))
annots <- readRDS(here("analysis", "data", "derived_data", "annots.rds"))
expr <- readRDS(here("analysis", "data", "derived_data", "expr.rds"))
```

```{r}
local_annot <- lod_peaks %>%
  filter(chrom == hot_chr, pos <= hot_upper, pos >= hot_lower) %>%
  left_join(y = annots, by = c("annot.id" = "gene_id")) %>%
  filter(!duplicated(annot.id)) %>%
  filter(lod >= ult) %>%
  mutate(local_tx = chr == hot_chr) %>%
  filter(local_tx)
```

```{r}
nonlocal_annot <- lod_peaks %>%
  filter(chrom == hot_chr, pos <= hot_upper, pos >= hot_lower) %>%
  inner_join(y = annots, by = c("annot.id" = "gene_id")) %>%
  filter(!duplicated(annot.id)) %>%
  filter(lod >= 7.18) %>%
  mutate(local_tx = chr == hot_chr) %>%
  filter(!local_tx) %>%
  filter(!is.na(hotspot))
```

```{r}
local_annot2 <- local_annot %>%
  filter(middle <= hot_upper + 1, middle >= hot_lower - 1) %>% 
  # extend only 1 mb beyond hotspot end points (total interval length, for local traits, = 6 Mb)
  arrange(desc(lod))
```

```{r}
hot_local <- expr[, colnames(expr) %in% local_annot2$annot.id]
hot_nonlocal <- expr[, colnames(expr) %in% nonlocal_annot$annot.id]
nnonlocal <- nrow(nonlocal_annot)
nlocal <- nrow(local_annot2)
```

```{r, message=FALSE, warning = FALSE}
med_tib <- med_list %>%
  purrr::map(bind_rows) %>%
  bind_rows() %>%
  as_tibble() %>%
  rename(lod_without_med = lod_no_med, lod_with_med = lod_med) %>%
  mutate(lod_diff = lod_without_med - lod_with_med, lod_diff_proportion = lod_diff / lod_without_med) %>%
  left_join(local_annot2, by = c("local_gene_id" = "annot.id")) %>%
  left_join(nonlocal_annot, by = c("nonlocal_gene_id" = "annot.id")) %>%
  dplyr::select(symbol.x, symbol.y, lod_without_med, lod_with_med, lod_diff, lod_diff_proportion, local_gene_id, nonlocal_gene_id) %>%
  rename(local_symbol = symbol.x, nonlocal_symbol = symbol.y) %>%
  filter(!is.na(local_symbol))
```

```{r}
hot_annot <- local_annot2 %>%
  bind_rows(nonlocal_annot) %>%
  dplyr::select(annot.id, symbol)
pleio_tib <- pleio_mixed_list %>%
  purrr::map(.f = function(x) rename(x, gene1_id = local_gene_id, gene2_id = nonlocal_gene_id)) %>%
  c(pleio_local_list, pleio_nonlocal_list) %>%
  bind_rows() %>%
  rename(pleiotropy_lod = lod) %>%
  inner_join(hot_annot, by = c("gene1_id" = "annot.id")) %>%
  rename(gene1_symbol = symbol) %>%
  inner_join(hot_annot, by = c("gene2_id" = "annot.id")) %>%
  rename(gene2_symbol = symbol) %>%
  dplyr::select(gene1_symbol, gene2_symbol, pleiotropy_lod)
```

```{r}
pm <- pleio_tib %>%
  tibble_to_matrix(symmetric = TRUE)
```

```{r}
p <- pleio_mixed_list %>%
  bind_rows() %>%
  rename(pleiotropy_lod = lod) %>%
  inner_join(med_tib, by = c("local_gene_id", "nonlocal_gene_id")) %>%
  dplyr::select(local_symbol, nonlocal_symbol, pleiotropy_lod, lod_diff, lod_diff_proportion, lod_without_med, lod_with_med, filename, nonlocal_gene_id, local_gene_id) %>%
  ggplot() + geom_point(mapping = aes(x = pleiotropy_lod, y = lod_diff, colour = local_symbol == keller_mediator, nonlocal = nonlocal_symbol, local = local_symbol), size = 0.05) + theme(legend.position = "none") + xlab("Pleiotropy LOD") + ylab("Mediation LOD difference")
fnsvg <- paste0("Chr", hot_chr, "_scatter.svg")
fnpdf <- paste0("Chr", hot_chr, "_scatter.pdf")
ggsave(plot = p, filename = here("analysis", "figures", fnsvg))
ggsave(plot = p, filename = here("analysis", "figures", fnpdf))
plotly::ggplotly(p)
```

```{r, fig.height = ceiling(nlocal / 2)}
foo <- pleio_mixed_list %>%
  bind_rows() %>%
  rename(pleiotropy_lod = lod) %>%
  inner_join(med_tib, by = c("local_gene_id", "nonlocal_gene_id")) %>%
  dplyr::select(local_symbol, nonlocal_symbol, pleiotropy_lod, lod_diff, lod_diff_proportion, lod_without_med, lod_with_med, filename, nonlocal_gene_id, local_gene_id)
p <- foo %>%
  ggplot() +
  geom_point(mapping = aes(x = pleiotropy_lod, y = lod_diff, colour = nonlocal_symbol), size = 0.1) +
  facet_wrap(. ~ local_symbol, nrow = ceiling(nlocal / 2), ncol = 2) +
  broman::karl_theme(strip.placement = "inside") +
  theme(legend.position = "none") +
  theme(
    strip.background = element_blank(),
    strip.text.x = element_blank()
) +
  xlab("Pleiotropy LOD") +
  ylab("Mediation LOD difference") +
  geom_text(aes(x = 1, y = 1, label = local_symbol), inherit.aes=FALSE)

fnsvg <- paste0("Chr", hot_chr, "_scatter-panel.svg")
fnpdf <- paste0("Chr", hot_chr, "_scatter-panel.pdf")
ggsave(p, filename = here("analysis", "figures", fnsvg), height = ceiling(nlocal / 2))
ggsave(p, filename = here("analysis", "figures", fnpdf), height = ceiling(nlocal / 2))
plotly::ggplotly(p)
```

```{r, eval = FALSE}
# put in multiple figures, if there are many local traits
n_figs <- 1 + nlocal %/% 18 # 18 is the number of panels per page in pdf
n_last <- nlocal %% 18 # number of panels on last figure
max_pleio_lod <- max(foo$pleiotropy_lod)
max_med_lod_difference <- max(foo$lod_diff)
min_med_lod_difference <- min(foo$lod_diff)

for (i in 1:n_figs){
  n_panels <- 18 * (i < n_figs) + n_last * (i == n_figs)
  foo %>%
    filter(local_symbol %in% local_annot2$symbol[((i - 1) * 18 + 1): (18 * i)]) %>%
    ggplot() +
      geom_point(mapping = aes(x = pleiotropy_lod, y = lod_diff, colour = nonlocal_symbol), size = 0.1) +
      facet_wrap(. ~ local_symbol, nrow = ceiling(n_panels / 2), ncol = 2) +
      broman::karl_theme() +
      theme(legend.position = "none") +
      theme(
        strip.background = element_blank(),
        strip.text.x = element_blank()
    ) +
      xlab("Pleiotropy LOD") + xlim(c(0, max_pleio_lod)) + ylim(c(min_med_lod_difference, max_med_lod_difference)) +
      ylab("Mediation LOD difference") +
      geom_text(size = 2, aes(x = max_pleio_lod / 2, y = 2 * max_med_lod_difference / 3, label = local_symbol), inherit.aes=FALSE)
  fnsvg <- paste0("Chr", hot_chr, "_scatter-panel", i, ".svg")
  fnpdf <- paste0("Chr", hot_chr, "_scatter-panel", i, ".pdf")
  ggsave(filename = here("analysis", "figures", fnsvg), height = ceiling(n_panels / 2))
  ggsave(filename = here("analysis", "figures", fnpdf), height = ceiling(n_panels / 2))
}
```

```{r}
scols1 <- 1L + (rownames(pm) %in% local_annot2$symbol)
scols2 <- rownames(pm) %in% (med_tib %>% filter(local_symbol == keller_mediator, lod_diff > 1.5))$nonlocal_symbol
```

```{r}
# order the nonlocal traits first
## indicator for being a nonlocal trait
indic_nonlocal <- rownames(pm) %in% nonlocal_annot$symbol
# define ord_nonlocal
ord_nonlocal <- pm[indic_nonlocal, indic_nonlocal] %>%
  dist() %>%
  hclust() %>%
  as.dendrogram() %>%
  dendextend::seriate_dendrogram(dist(pm[indic_nonlocal, indic_nonlocal])) %>%
  order.dendrogram()
rn_nonlocal <- rownames(pm[indic_nonlocal, indic_nonlocal][ord_nonlocal, ord_nonlocal])
# define ord_local
ord_local<- pm[!indic_nonlocal, !indic_nonlocal] %>%
  dist() %>%
  hclust() %>%
  as.dendrogram() %>%
  dendextend::seriate_dendrogram(dist(pm[!indic_nonlocal, !indic_nonlocal])) %>%
  order.dendrogram()
rn_local <- rownames(pm[!indic_nonlocal, !indic_nonlocal][ord_local, ord_local])
rn <- c(rn_nonlocal, rn_local)
```

```{r}
scols_extra <- purrr::map(.x = local_annot2$symbol,
                          .f = function(x) {rownames(pm) %in% (med_tib %>% filter(local_symbol == x, lod_diff > 1.5))$nonlocal_symbol}
) %>%
  bind_cols() %>%
  (function(x){colnames(x) <- local_annot2$symbol; x}) %>%
  mutate(local = scols1)
```

```{r}
n_med_tr <- 4 # arbitrary threshold for number of mediated traits per local
# only those local traits with n_med_tr or more mediated nonlocal traits 
# appear as annotation columns in the heatmap
n_med_per_local <- med_tib %>%
  group_by(local_symbol) %>%
  filter(lod_diff > 1.5) %>%
  count() %>%
  filter(n >= n_med_tr)
n_med_per_local %>%
  arrange(desc(n)) %>%
  knitr::kable()

scols_pre <- scols_extra %>%
  dplyr::select(n_med_per_local$local_symbol)
# get ordering for columns of scols_pre
col_order <- scols_pre %>%
  t() %>%
  dist() %>%
  hclust() %>%
  as.dendrogram() %>%
  dendextend::seriate_dendrogram(dist(t(scols_pre))) %>%
  order.dendrogram()
scols <- scols_pre[ , col_order] %>%
  mutate(local = scols1)
fnhtml <- here("analysis", "figures", paste0("Chr", hot_chr, "hm.html"))
```

```{r}
ord <- match(rn, rownames(pm))

(p <- pm[ord, ord] %>%
  apply(FUN = trunc2, MARGIN = 2, threshold = 5) %>%
  heatmaply(Rowv = FALSE,
            Colv = FALSE,
            cexRow = 0.2,
            cexCol = 0.2,
            symm = TRUE,
            row_side_colors = scols[ord, ],
            col_side_colors = tibble(local = scols1[ord]),
            limits = c(0, 5)
            ) %>%
  hide_guides()
)
fnpdf <- here("analysis", "figures", paste0("Chr", hot_chr, "hm.pdf"))
htmlwidgets::saveWidget(widget = p, 
                        file = fnhtml
                        )
```

```{r eval = FALSE}
webshot::webshot(url = fnhtml,
                 file = fnpdf,
                 cliprect = "viewport"
                 )
```

```{r}
tibble(symbol = rownames(pm[indic_nonlocal, indic_nonlocal])[ord_nonlocal], index = 1:length(symbol)) %>%
  left_join(nonlocal_annot, by = "symbol") %>%
  ggplot() + geom_point(aes(y = pos, x = index))
```

### 2019-07-28 analysis/paper/hotspots-paper-Chr2-only.Rmd
added files for dift sections, to be assembled in hotspots-paper.Rmd

--- 2019-07-27 analysis/paper/hotspots-paper.Rmd
@@
 output: 
-  - bookdown::html_document2
-  - github_document:
-      toc: true
-      toc_depth: 2
+  bookdown::html_document2:
+    pandoc_args: 
+      - --filter
+      - /usr/local/Cellar/pandoc-crossref/0.3.4.1/bin/pandoc-crossref
 ---
-
-# Introduction
-
-Genetics studies in model organisms like mice can identify
-genomic regions that affect complex traits, such as systolic blood pressure and body weight [@sax1923association;@soller1976power;@lander1989mapping;@broman2009guide;@jansen2007quantitative]. 
-These genomic regions are called "quantitative trait loci" or "QTL".
-A genome-wide QTL "scan" reveals
-associations between genotypes and phenotypes by considering each position along the genome,
-one at a time, as a candidate QTL for the trait of interest.
-A region with strong evidence of association with a complex trait, then, defines a QTL (for that trait).
-Because nearby markers have correlated
-genotypes, a QTL in a two-parent cross often spans multiple megabases in length
-and may contain more than a hundred genes.
-Identification of a causal gene (for a given complex trait) from among those genes near the QTL is
-challenging and may require costly and time-consuming experiments.
-The growing need for greater QTL mapping resolution fueled development over the
-last two decades of model organism multiparental populations for high-resolution QTL mapping [@de2017back;@churchill2004collaborative;@svenson2012high;@huang2012multiparent;@shivakumar2018soybean;@huang2011analysis;@kover2009multiparent;@tisne2017identification;@stanley2017genetic].
-With experimentalists now measuring tens of thousands of biomolecular traits
-in multiparental populations, the systems genetics community needs multivariate
-statistical tools to fully examine the large volumes of data [@keller2018genetic;@chick2016defining].
-Identifying loci that affect multiple traits can aid in identifying biomolecular interactions and in clarifying complex trait genetic architecture.
-A QTL that affects more than one trait is called a "pleiotropic" QTL.
-A test of pleiotropy vs. separate QTL is one multivariate statistical tool that will inform
-complex trait genetics by enabling researchers to identify the number of
-unique QTL in a genomic region of interest.
 
@@
 
-\section{Methods}
+# Methods
 
@@
 vec(Y) = X vec(B) + vec(G) + vec(E)
-\label{eq:mvlmm}
 $$
@@
 \hat B = (X^T\hat \Sigma^{-1}X)^{-1}X^T\hat\Sigma^{-1}Y
-\label{eq:Bhat}
 $$
@@
 \hat \Sigma = \hat V_g \otimes \hat K + \hat V_e \otimes I_n
-\label{eq:Sigmahat}
 $$
@@
 l_{10}(\hat B, \hat \Sigma) = - \frac{1}{2}\log_{10}\left((Y - X\hat B)^T\hat \Sigma^{-1}(Y - X\hat B)\right) - n\log_{10}(2\pi) - \log_{10}|\hat \Sigma|
-\label{eq:loglik}
 $$
@@
 LOD = \log_{10} \left(\frac{\max_{\lambda_1, \lambda_2}L(B, \Sigma, \lambda_1, \lambda_2)}{\max_{\lambda}L(B, \Sigma, \lambda, \lambda)} \right)
-\label{eq:lrt}
 $$
@@
 
+```{r load_tidyverse}
+library(tidyverse)
+options(tibble.print_max = 50, tibble.print_min = 20)
+```
+
+```{r load_heatmaply}
+library(heatmaply)
+```
+
 ```{r setup_continued-2}
@@
 
-```{r more_setup-2}
-hot_lower <- hot_mid - 2
-hot_upper <- hot_mid + 2
-fp <- paste0("Chr", hot_chr, "-")
-knitr::opts_chunk$set(fig.path = here("analysis", "figures", fp))
-```
-
-```{r readRDS-2}
-pleio_mixed_list <- readRDS(here("analysis", "data", "derived_data", paste0("Chr", hot_chr, "_pleio.rds")))
-pleio_local_list <- readRDS(here("analysis", "data", "derived_data", paste0("Chr", hot_chr, "_pleio_local.rds")))
-pleio_nonlocal_list <- readRDS(here("analysis", "data", "derived_data", paste0("Chr", hot_chr, "_pleio_nonlocal.rds")))
-med_list <- readRDS(here("analysis", "data", "derived_data", paste0("Chr", hot_chr, "_mediations.rds")))
-```
-
-```{r load_attie, eval=!file.exists(here("analysis", "data", "derived_data", "expr.rds"))}
-download.file("https://datadryad.org/bitstream/handle/10255/dryad.166774/Attie_DO378_eQTL_viewer_v1.Rdata?sequence=2",
-              destfile = here("analysis", "data", "raw_data", "Attie_DO378_eQTL_viewer_v1.Rdata")
-              )
-load(here("analysis", "data", "raw_data", "Attie_DO378_eQTL_viewer_v1.Rdata"))
-saveRDS(dataset.islet.rnaseq$lod.peaks, here("analysis", "data", "derived_data", "lod_peaks.rds"))
-saveRDS(dataset.islet.rnaseq$annots, here("analysis", "data", "derived_data", "annots.rds"))
-saveRDS(dataset.islet.rnaseq$expr, here("analysis", "data", "derived_data", "expr.rds"))
-```
-
-```{r read_attie_files}
-lod_peaks <- readRDS(here("analysis", "data", "derived_data", "lod_peaks.rds"))
-annots <- readRDS(here("analysis", "data", "derived_data", "annots.rds"))
-expr <- readRDS(here("analysis", "data", "derived_data", "expr.rds"))
-```
-
-```{r load_tidyverse}
-library(tidyverse)
-options(tibble.print_max = 50, tibble.print_min = 20)
-```
-
-```{r make_local_annot-2}
-local_annot <- lod_peaks %>%
-  filter(chrom == hot_chr, pos <= hot_upper, pos >= hot_lower) %>%
-  left_join(y = annots, by = c("annot.id" = "gene_id")) %>%
-  filter(!duplicated(annot.id)) %>%
-  filter(lod >= ult) %>%
-  mutate(local_tx = chr == hot_chr) %>%
-  filter(local_tx)
-```
-
-```{r make_nonlocal_annot-2}
-nonlocal_annot <- lod_peaks %>%
-  filter(chrom == hot_chr, pos <= hot_upper, pos >= hot_lower) %>%
-  inner_join(y = annots, by = c("annot.id" = "gene_id")) %>%
-  filter(!duplicated(annot.id)) %>%
-  filter(lod >= 7.18) %>%
-  mutate(local_tx = chr == hot_chr) %>%
-  filter(!local_tx) %>%
-  filter(!is.na(hotspot))
-```
-
-```{r make_local_annot2-2}
-local_annot2 <- local_annot %>%
-  filter(middle <= hot_upper + 5, middle >= hot_lower - 5) %>%
-  arrange(desc(lod))
-```
-
-```{r subset_expr_matrix-2}
-hot_local <- expr[, colnames(expr) %in% local_annot2$annot.id]
-hot_nonlocal <- expr[, colnames(expr) %in% nonlocal_annot$annot.id]
-nnonlocal <- nrow(nonlocal_annot)
-nlocal <- nrow(local_annot2)
-```
-
-```{r tabulate-2, message=FALSE, warning = FALSE}
-med_tib <- med_list %>%
-  purrr::map(bind_rows) %>%
-  bind_rows() %>%
-  as_tibble() %>%
-  rename(lod_without_med = lod_no_med, lod_with_med = lod_med) %>%
-  mutate(lod_diff = lod_without_med - lod_with_med, lod_diff_proportion = lod_diff / lod_without_med) %>%
-  left_join(local_annot2, by = c("local_gene_id" = "annot.id")) %>%
-  left_join(nonlocal_annot, by = c("nonlocal_gene_id" = "annot.id")) %>%
-  dplyr::select(symbol.x, symbol.y, lod_without_med, lod_with_med, lod_diff, lod_diff_proportion, local_gene_id, nonlocal_gene_id) %>%
-  rename(local_symbol = symbol.x, nonlocal_symbol = symbol.y) %>%
-  filter(!is.na(local_symbol))
-```
-
-```{r make_pleio_tib-2}
-hot_annot <- local_annot2 %>%
-  bind_rows(nonlocal_annot) %>%
-  dplyr::select(annot.id, symbol)
-pleio_tib <- pleio_mixed_list %>%
-  purrr::map(.f = function(x) rename(x, gene1_id = local_gene_id, gene2_id = nonlocal_gene_id)) %>%
-  c(pleio_local_list, pleio_nonlocal_list) %>%
-  bind_rows() %>%
-  rename(pleiotropy_lod = lod) %>%
-  inner_join(hot_annot, by = c("gene1_id" = "annot.id")) %>%
-  rename(gene1_symbol = symbol) %>%
-  inner_join(hot_annot, by = c("gene2_id" = "annot.id")) %>%
-  rename(gene2_symbol = symbol) %>%
-  dplyr::select(gene1_symbol, gene2_symbol, pleiotropy_lod)
-```
-
-```{r make_pm-2}
-pm <- pleio_tib %>%
-  tibble_to_matrix(symmetric = TRUE)
-```
-
-```{r scatter-2}
-p <- pleio_mixed_list %>%
-  bind_rows() %>%
-  rename(pleiotropy_lod = lod) %>%
-  inner_join(med_tib, by = c("local_gene_id", "nonlocal_gene_id")) %>%
-  dplyr::select(local_symbol, nonlocal_symbol, pleiotropy_lod, lod_diff, lod_diff_proportion, lod_without_med, lod_with_med, filename, nonlocal_gene_id, local_gene_id) %>%
-  ggplot() + geom_point(mapping = aes(x = pleiotropy_lod, y = lod_diff, colour = local_symbol == keller_mediator, nonlocal = nonlocal_symbol), size = 0.05) + theme(legend.position = "none") + xlab("Pleiotropy LOD") + ylab("Mediation LOD difference")
-fnsvg <- paste0("Chr", hot_chr, "_scatter.svg")
-fnpdf <- paste0("Chr", hot_chr, "_scatter.pdf")
-ggsave(plot = p, filename = here("analysis", "figures", fnsvg))
-ggsave(plot = p, filename = here("analysis", "figures", fnpdf))
-p
-```
-
-```{r scatter_by_local-2}
-foo <- pleio_mixed_list %>%
-  bind_rows() %>%
-  rename(pleiotropy_lod = lod) %>%
-  inner_join(med_tib, by = c("local_gene_id", "nonlocal_gene_id")) %>%
-  dplyr::select(local_symbol, nonlocal_symbol, pleiotropy_lod, lod_diff, lod_diff_proportion, lod_without_med, lod_with_med, filename, nonlocal_gene_id, local_gene_id)
-foo %>%
-  ggplot() +
-  geom_point(mapping = aes(x = pleiotropy_lod, y = lod_diff, colour = nonlocal_symbol, nonlocal = nonlocal_symbol), size = 0.1) +
-  facet_wrap(. ~ local_symbol, nrow = ceiling(nlocal / 2), ncol = 2) +
-  broman::karl_theme(strip.placement = "inside") +
-  theme(legend.position = "none") +
-  theme(
-    strip.background = element_blank(),
-    strip.text.x = element_blank()
-) +
-  xlab("Pleiotropy LOD") +
-  ylab("Mediation LOD difference") +
-  geom_text(aes(x = 1, y = 1, label = local_symbol), inherit.aes=FALSE)
-
-fnsvg <- paste0("Chr", hot_chr, "_scatter-panel.svg")
-fnpdf <- paste0("Chr", hot_chr, "_scatter-panel.pdf")
-ggsave(filename = here("analysis", "figures", fnsvg), height = ceiling(nlocal / 2))
-ggsave(filename = here("analysis", "figures", fnpdf), height = ceiling(nlocal / 2))
-
-# put in multiple figures, if there are many local traits
-n_figs <- 1 + nlocal %/% 18 # 18 is the number of panels per page in pdf
-n_last <- nlocal %% 18 # number of panels on last figure
-max_pleio_lod <- max(foo$pleiotropy_lod)
-max_med_lod_difference <- max(foo$lod_diff)
-min_med_lod_difference <- min(foo$lod_diff)
-
-for (i in 1:n_figs){
-  n_panels <- 18 * (i < n_figs) + n_last * (i == n_figs)
-  foo %>%
-    filter(local_symbol %in% local_annot2$symbol[((i - 1) * 18 + 1): (18 * i)]) %>%
-    ggplot() +
-      geom_point(mapping = aes(x = pleiotropy_lod, y = lod_diff, colour = nonlocal_symbol, nonlocal = nonlocal_symbol), size = 0.1) +
-      facet_wrap(. ~ local_symbol, nrow = ceiling(n_panels / 2), ncol = 2) +
-      broman::karl_theme() +
-      theme(legend.position = "none") +
-      theme(
-        strip.background = element_blank(),
-        strip.text.x = element_blank()
-    ) +
-      xlab("Pleiotropy LOD") + xlim(c(0, max_pleio_lod)) + ylim(c(min_med_lod_difference, max_med_lod_difference)) +
-      ylab("Mediation LOD difference") +
-      geom_text(size = 2, aes(x = max_pleio_lod / 2, y = 2 * max_med_lod_difference / 3, label = local_symbol), inherit.aes=FALSE)
-  fnsvg <- paste0("Chr", hot_chr, "_scatter-panel", i, ".svg")
-  fnpdf <- paste0("Chr", hot_chr, "_scatter-panel", i, ".pdf")
-  ggsave(filename = here("analysis", "figures", fnsvg), height = ceiling(n_panels / 2))
-  ggsave(filename = here("analysis", "figures", fnpdf), height = ceiling(n_panels / 2))
-}
-```
-
-```{r load_heatmaply}
-library(heatmaply)
-```
-
-```{r make_scols-2}
-scols1 <- 1L + (rownames(pm) %in% local_annot2$symbol)
-scols2 <- rownames(pm) %in% (med_tib %>% filter(local_symbol == keller_mediator, lod_diff > 1.5))$nonlocal_symbol
-```
-
-```{r heat8-2}
-# order the nonlocal traits first
-## indicator for being a nonlocal trait
-indic_nonlocal <- rownames(pm) %in% nonlocal_annot$symbol
-# define ord_nonlocal
-ord_nonlocal <- pm[indic_nonlocal, indic_nonlocal] %>%
-  dist() %>%
-  hclust() %>%
-  as.dendrogram() %>%
-  dendextend::seriate_dendrogram(dist(pm[indic_nonlocal, indic_nonlocal])) %>%
-  order.dendrogram()
-rn_nonlocal <- rownames(pm[indic_nonlocal, indic_nonlocal][ord_nonlocal, ord_nonlocal])
-# define ord_local
-ord_local<- pm[!indic_nonlocal, !indic_nonlocal] %>%
-  dist() %>%
-  hclust() %>%
-  as.dendrogram() %>%
-  dendextend::seriate_dendrogram(dist(pm[!indic_nonlocal, !indic_nonlocal])) %>%
-  order.dendrogram()
-rn_local <- rownames(pm[!indic_nonlocal, !indic_nonlocal][ord_local, ord_local])
-rn <- c(rn_nonlocal, rn_local)
-```
-
-```{r make_scols-extra-2}
-scols_extra <- purrr::map(.x = local_annot2$symbol,
-                          .f = function(x) {rownames(pm) %in% (med_tib %>% filter(local_symbol == x, lod_diff > 1.5))$nonlocal_symbol}
-) %>%
-  bind_cols() %>%
-  (function(x){colnames(x) <- local_annot2$symbol; x}) %>%
-  mutate(local = scols1)
-```
-
-```{r count_med}
-n_med_tr <- 4 # arbitrary threshold
-n_med_per_local <- med_tib %>%
-  group_by(local_symbol) %>%
-  filter(lod_diff > 1.5) %>%
-  count() %>%
-  filter(n >= n_med_tr)
-n_med_per_local %>%
-  arrange(desc(n)) %>%
-  knitr::kable()
-
-scols_pre <- scols_extra %>%
-  dplyr::select(n_med_per_local$local_symbol)
-# get ordering for columns of scols_pre
-col_order <- scols_pre %>%
-  t() %>%
-  dist() %>%
-  hclust() %>%
-  as.dendrogram() %>%
-  dendextend::seriate_dendrogram(dist(t(scols_pre))) %>%
-  order.dendrogram()
-scols <- scols_pre[ , col_order] %>%
-  mutate(local = scols1)
-fnhtml <- here("analysis", "figures", paste0("Chr", hot_chr, "hm.html"))
-```
-
-```{r heat9-2, eval = !file.exists(fnhtml)}
-ord <- match(rn, rownames(pm))
-
-p <- pm[ord, ord] %>%
-  apply(FUN = trunc2, MARGIN = 2, threshold = 5) %>%
-  heatmaply(Rowv = FALSE,
-            Colv = FALSE,
-            cexRow = 0.2,
-            cexCol = 0.2,
-            symm = TRUE,
-            row_side_colors = scols[ord, ],
-            col_side_colors = tibble(local = scols1[ord]),
-            limits = c(0, 5)
-            ) %>%
-  hide_guides()
-fnpdf <- here("analysis", "figures", paste0("Chr", hot_chr, "hm.pdf"))
-htmlwidgets::saveWidget(widget = p,
-                        file = fnhtml
-                        )
-webshot::webshot(url = fnhtml,
-                 file = fnpdf,
-                 cliprect = "viewport"
-                 )
-```
-
-```{r heat10-2}
-webshot::webshot(fnhtml)
-```
-
-```{r index-v-peak-position-2}
-tibble(symbol = rownames(pm[indic_nonlocal, indic_nonlocal])[ord_nonlocal], index = 1:length(symbol)) %>%
-  left_join(nonlocal_annot, by = "symbol") %>%
-  ggplot() + geom_point(aes(y = pos, x = index))
-```
-
-```{r setup_continued-5}
-hot_chr <- 5
-hot_mid <- 146
-keller_mediator <- "Pdx1"
-```
-
-```{r more_setup-5, ref.label = "more_setup-2"}
-```
-
-```{r readRDS-5, ref.label="readRDS-2"}
-```
-
-```{r make_local_annot-5, ref.label="make_local_annot-2"}
-```
-
-```{r make_nonlocal_annot-5, ref.label = "make_nonlocal_annot-2"}
-```
-
-```{r make_local_annot2-5, ref.label = "make_local_annot2-2"}
-```
-
-```{r subset_expr_matrix-5, ref.label = "subset_expr_matrix-2"}
-```
-
-```{r tabulate-5, ref.label = "tabulate-2", message=FALSE, warning = FALSE}
-```
-
-```{r make_pleio_tib-5, ref.label="make_pleio_tib-2"}
-```
-
-```{r make_pm-5, ref.label = "make_pm-2"}
-```
-
-```{r scatter-5, ref.label="scatter-2"}
-```
-
-```{r scatter_by_local-5, ref.label = "scatter_by_local-2"}
-```
-
-```{r make_scols-5, ref.label="make_scols-2"}
-```
-
-```{r heat8-5, ref.label="heat8-2"}
-```
-
-```{r make_scols-extra-5, ref.label = "make_scols-extra-2"}
-```
-
-```{r count_med-5, ref.label = "count_med"}
-```
-
-```{r heat9-5, ref.label = "heat9-2", eval = !file.exists(fnhtml)}
-```
-
-```{r heat10-5, ref.label="heat10-2"}
-```
-
-```{r index-v-peak-position-5, ref.label = "index-v-peak-position-2"}
-```
-
-```{r setup_continued-7}
-hot_chr <- 7
-hot_mid <- 46
-keller_mediator <- "Fam83e"
-```
-
-```{r more_setup-7, ref.label = "more_setup-2"}
-```
-
-```{r readRDS-7, ref.label="readRDS-2"}
-```
-
-```{r make_local_annot-7, ref.label="make_local_annot-2"}
-```
-
-```{r make_nonlocal_annot-7, ref.label = "make_nonlocal_annot-2"}
-```
-
-```{r make_local_annot2-7, ref.label = "make_local_annot2-2"}
-```
-
-```{r subset_expr_matrix-7, ref.label = "subset_expr_matrix-2"}
-```
-
-```{r tabulate-7, ref.label = "tabulate-2", message=FALSE, warning = FALSE}
-```
-
-```{r make_pleio_tib-7, ref.label="make_pleio_tib-2"}
-```
-
-```{r make_pm-7, ref.label = "make_pm-2"}
-```
-
-```{r scatter-7, ref.label="scatter-2"}
-```
-
-```{r scatter_by_local-7, ref.label = "scatter_by_local-2"}
-```
-
-```{r make_scols-7, ref.label="make_scols-2"}
-```
-
-```{r heat8-7, ref.label="heat8-2"}
-```
-
-```{r make_scols-extra-7, ref.label = "make_scols-extra-2"}
-```
-
-```{r count_med-7, ref.label = "count_med"}
-```
-
-```{r heat9-7, ref.label = "heat9-2", eval = !file.exists(fnhtml)}
-```
-
-```{r heat10-7, ref.label="heat10-2"}
-```
-
-```{r index-v-peak-position-7, ref.label = "index-v-peak-position-2"}
-```
-
-```{r setup_continued-11}
-hot_chr <- 11
-hot_mid <- 71
-keller_mediator <- "Sat2"
-```
-
-```{r more_setup-11, ref.label = "more_setup-2"}
-```
-
-```{r readRDS-11, ref.label="readRDS-2"}
-```
-
-```{r make_local_annot-11, ref.label="make_local_annot-2"}
-```
-
-```{r make_nonlocal_annot-11, ref.label = "make_nonlocal_annot-2"}
-```
-
-```{r make_local_annot2-11, ref.label = "make_local_annot2-2"}
-```
-
-```{r subset_expr_matrix-11, ref.label = "subset_expr_matrix-2"}
-```
-
-```{r tabulate-11, ref.label = "tabulate-2", message=FALSE, warning = FALSE}
-```
-
-```{r make_pleio_tib-11, ref.label="make_pleio_tib-2"}
-```
-
-```{r make_pm-11, ref.label = "make_pm-2"}
-```
-
-```{r scatter-11, ref.label="scatter-2"}
-```
-
-```{r scatter_by_local-11, ref.label = "scatter_by_local-2"}
-```
-
-```{r make_scols-11, ref.label="make_scols-2"}
-```
-
-```{r heat8-11, ref.label="heat8-2"}
-```
-
-```{r make_scols_extra-11, ref.label = "make_scols-extra-2"}
-```
-
-```{r count_med-11, ref.label = "count_med"}
-```
-
-```{r heat9-11, ref.label = "heat9-2", eval = !file.exists(fnhtml)}
-```
-
-```{r heat10-11, ref.label="heat10-2"}
-```
-
-```{r index-v-peak-position-11, ref.label = "index-v-peak-position-2"}
-```
-
-```{r setup_continued-13}
-hot_chr <- 13
-hot_mid <- 112.5
-keller_mediator <- "Il6st"
-```
-
-```{r more_setup-13, ref.label = "more_setup-2"}
-```
-
-```{r readRDS-13, ref.label="readRDS-2"}
-```
-
-```{r make_local_annot-13, ref.label="make_local_annot-2"}
-```
-
-```{r make_nonlocal_annot-13, ref.label = "make_nonlocal_annot-2"}
-```
-
-```{r make_local_annot2-13, ref.label = "make_local_annot2-2"}
-```
-
-```{r subset_expr_matrix-13, ref.label = "subset_expr_matrix-2"}
-```
-
-```{r tabulate-13, ref.label = "tabulate-2", message=FALSE, warning = FALSE}
-```
-
-```{r make_pleio_tib-13, ref.label="make_pleio_tib-2"}
-```
-
-```{r make_pm-13, ref.label = "make_pm-2"}
-```
-
-```{r scatter-13, ref.label="scatter-2"}
-```
-
-```{r scatter_by_local-13, ref.label = "scatter_by_local-2"}
-```
-
-```{r make_scols-13, ref.label="make_scols-2"}
-```
-
-```{r heat8-13, ref.label="heat8-2"}
-```
-
-```{r make_scols-extra-13, ref.label = "make_scols-extra-2"}
-```
-
-```{r count_med-13, ref.label = "count_med"}
-```
-
-```{r heat9-13, ref.label = "heat9-2", eval = !file.exists(fnhtml)}
-```
-
-```{r heat10-13, ref.label="heat10-2"}
-```
-
-```{r index-v-peak-position-13, ref.label = "index-v-peak-position-2"}
+```{r child="hotspot-one-chromosome.Rmd"}
 ```

### 2019-07-28 analysis/paper/hotspots-paper.Rmd
added files for dift sections, to be assembled in hotspots-paper.Rmd

--- 2019-07-27 analysis/paper/hotspots-paper.Rmd
@@
 bibliography: research.bib
+numberSections: true
+autoEqnLabels: true
 output: 
-  - bookdown::html_document2
-  - github_document:
-      toc: true
-      toc_depth: 2
+  bookdown::html_document2:
+    pandoc_args: 
+      - --filter
+      - /usr/local/Cellar/pandoc-crossref/0.3.4.1/bin/pandoc-crossref
 ---
 
-# Introduction
-
-Genetics studies in model organisms like mice can identify
-genomic regions that affect complex traits, such as systolic blood pressure and body weight [@sax1923association;@soller1976power;@lander1989mapping;@broman2009guide;@jansen2007quantitative]. 
-These genomic regions are called "quantitative trait loci" or "QTL".
-A genome-wide QTL "scan" reveals
-associations between genotypes and phenotypes by considering each position along the genome,
-one at a time, as a candidate QTL for the trait of interest.
-A region with strong evidence of association with a complex trait, then, defines a QTL (for that trait).
-Because nearby markers have correlated
-genotypes, a QTL in a two-parent cross often spans multiple megabases in length
-and may contain more than a hundred genes.
-Identification of a causal gene (for a given complex trait) from among those genes near the QTL is
-challenging and may require costly and time-consuming experiments.
-The growing need for greater QTL mapping resolution fueled development over the
-last two decades of model organism multiparental populations for high-resolution QTL mapping [@de2017back;@churchill2004collaborative;@svenson2012high;@huang2012multiparent;@shivakumar2018soybean;@huang2011analysis;@kover2009multiparent;@tisne2017identification;@stanley2017genetic].
-With experimentalists now measuring tens of thousands of biomolecular traits
-in multiparental populations, the systems genetics community needs multivariate
-statistical tools to fully examine the large volumes of data [@keller2018genetic;@chick2016defining].
-Identifying loci that affect multiple traits can aid in identifying biomolecular interactions and in clarifying complex trait genetic architecture.
-A QTL that affects more than one trait is called a "pleiotropic" QTL.
-A test of pleiotropy vs. separate QTL is one multivariate statistical tool that will inform
-complex trait genetics by enabling researchers to identify the number of
-unique QTL in a genomic region of interest.
+```{r child="intro.Rmd"}
+```
 
@@
 
-\section{Methods}
-
-@keller2018genetic identified five expression trait hotspots in their
-study of pancreatic islet gene expression in 378 Diversity Outbred mice. They
-defined a hotspot to be a 4-Mb region that affects at least 100 nonlocal expression
-traits.
-
-For each hotspot, we identified a set of local expression traits and a
-set of nonlocal expression traits. All traits demonstrated LOD peaks that met
-the genome-wide significance threshold, 7.18 [@keller2018genetic].
-The nonlocal expression traits arose from genes
-on other chromosomes, while the local traits reflected expression levels for
-genes on the same chromosome as the hotspot and within 4 Mb of the hotspot
-center. We ignored transcripts for genes on the hotspot chromosome but not within 4 Mb of
-the hotspot center.
-
-## Pleiotropy testing
-
-### Two-dimensional, two-QTL scans
-
-For each hotspot, we tested all pairs of traits, regardless of whether the traits were local or nonlocal. We performed a
-two-dimensional, two-QTL scan over a genomic region that included the entirety of the
-hotspot. We fitted bivariate linear mixed effects models at each ordered pair of markers in a two-dimensional grid [@boehm2019testing].
-
-$$
-vec(Y) = X vec(B) + vec(G) + vec(E)
-\label{eq:mvlmm}
-$$
-
-$$
-G \sim MN_{n x 2}(0, K, V_g)
-$$
-
-$$
-E \sim MN_{n x 2}(0, I_n, V_e)
-$$
-
-and $G$ and $E$ are independent.
-
-$$
-\hat B = (X^T\hat \Sigma^{-1}X)^{-1}X^T\hat\Sigma^{-1}Y
-\label{eq:Bhat}
-$$
-
-$$
-\hat \Sigma = \hat V_g \otimes \hat K + \hat V_e \otimes I_n
-\label{eq:Sigmahat}
-$$
-
-$$
-l_{10}(\hat B, \hat \Sigma) = - \frac{1}{2}\log_{10}\left((Y - X\hat B)^T\hat \Sigma^{-1}(Y - X\hat B)\right) - n\log_{10}(2\pi) - \log_{10}|\hat \Sigma|
-\label{eq:loglik}
-$$
-
-For $n$ mice, $\Sigma$ is a $2n$ by $2n$ covariance matrix, while $V_g$ and $V_e$ are $2$ by $2$ covariance matrices.
-Estimates $\hat V_g$ and $\hat V_e$ are obtained via an expectation-maximization algorithm [@zhou2014efficient;@gemma2].
-
-### Calculating likelihood ratio test statistics}
-From the two-dimensional scans, we calculated pleiotropy likelihood ratio test statistics for each
-pair of traits [@jiang1995multiple;@boehm2019testing;@boehm2019thesis]. Because of the required
-computing time for bootstrap analyses, we didn't obtain p-values for the pleiotropy tests. Instead,
-we worked directly with the test statistic values.
-
-$$
-LOD = \log_{10} \left(\frac{\max_{\lambda_1, \lambda_2}L(B, \Sigma, \lambda_1, \lambda_2)}{\max_{\lambda}L(B, \Sigma, \lambda, \lambda)} \right)
-\label{eq:lrt}
-$$
-
-$$
-ll(\lambda_1, \lambda_2)
-$$
-
-## Mediation analysis
-
-A mediation analysis in systems genetics begins with an identified QTL-expression trait association. In our case, we choose a QTL that affects a nonlocal trait.
-
-For each hotspot and each local trait - nonlocal trait pair, we also performed a
-mediation analysis to assess the extent which the local trait mediates the QTL effect on the nonlocal trait.
-Each mediation analysis involved fitting four linear models (Equations~\ref{model1}, \ref{model2}, \ref{model3}, \ref{model4}) [@chick2016defining].
-
-$$
-Y = \alpha 1 + WC + E
-\label{model1}
-$$
-
-$$
-Y = XB + WC + E
-\label{model2}
-$$
-
-$$
-Y = \alpha 1 + WC + M\beta + E
-\label{model3}
-$$
-
-$$
-Y = XB + WC + M\beta + E
-\label{model4}
-$$
-
-In all mediation analyses, the nonlocal transcripts are the targets, while the local
-transcripts serve as candidate mediators of the QTL-target relationship.
-In Equations~\ref{model1} through \ref{model4}, $X$ is a matrix of founder allele dosages,
-$B$ is a matrix
-of founder allele effects, $W$ is a matrix of covariates, $C$ is a matrix of covariate
-effects, $M$ is a vector of values for a single candidate mediator, $\beta$ is the effect
-of the mediator on the complex trait, 1 is a vector with all entries set to 1, $\alpha$
-is the mean trait value, and $E$ is a vector of normally distributed random errors.
-
-## Visualizations
-
-We created scatter plots with the `ggplot2` package [@ggplot2] for the R
-statistical computing environment [@r].
-We initially plotted all local-nonlocal pairs for each hotspot.
-We highlighted in blue those points that correspond to pairs involving the putative mediator per [@keller2018genetic].
-
-We created heatmaps to examine patterns among pleiotropy test results in each hotspot.
-We represented the pleiotropy test statistics as a symmetric matrix, where each row was one expression
-trait and each column was an expression trait. Each cell, then, contained the pleiotropy test statistic
-from the analysis involving the row's expression trait and the column's expression trait.
-To create a heatmap, we first performed two hierarchical clusterings, one on the submatrix involving only
-nonlocal expression traits and a second on the submatrix involving only local expression traits. We
-calculated the Euclidean distances between all trait pairs before clustering. We ordered the rows and
-columns according to optimal leaf ordering. We then combined the
-local and nonlocal traits into a single heatmap. To achieve greater coloring contrasts in the heatmaps, we
-set the color spectrum limits at 0 (dark blue) and 5 (yellow). Test statistics above 5 are represented by
-yellow.
+```{r child="methods.Rmd"}
+```
 
@@
 
-```{r more_setup-2}
-hot_lower <- hot_mid - 2
-hot_upper <- hot_mid + 2
-fp <- paste0("Chr", hot_chr, "-")
-knitr::opts_chunk$set(fig.path = here("analysis", "figures", fp))
-```
-
-```{r readRDS-2}
-pleio_mixed_list <- readRDS(here("analysis", "data", "derived_data", paste0("Chr", hot_chr, "_pleio.rds")))
-pleio_local_list <- readRDS(here("analysis", "data", "derived_data", paste0("Chr", hot_chr, "_pleio_local.rds")))
-pleio_nonlocal_list <- readRDS(here("analysis", "data", "derived_data", paste0("Chr", hot_chr, "_pleio_nonlocal.rds")))
-med_list <- readRDS(here("analysis", "data", "derived_data", paste0("Chr", hot_chr, "_mediations.rds")))
-```
-
-```{r load_attie, eval=!file.exists(here("analysis", "data", "derived_data", "expr.rds"))}
-download.file("https://datadryad.org/bitstream/handle/10255/dryad.166774/Attie_DO378_eQTL_viewer_v1.Rdata?sequence=2",
-              destfile = here("analysis", "data", "raw_data", "Attie_DO378_eQTL_viewer_v1.Rdata")
-              )
-load(here("analysis", "data", "raw_data", "Attie_DO378_eQTL_viewer_v1.Rdata"))
-saveRDS(dataset.islet.rnaseq$lod.peaks, here("analysis", "data", "derived_data", "lod_peaks.rds"))
-saveRDS(dataset.islet.rnaseq$annots, here("analysis", "data", "derived_data", "annots.rds"))
-saveRDS(dataset.islet.rnaseq$expr, here("analysis", "data", "derived_data", "expr.rds"))
-```
-
-```{r read_attie_files}
-lod_peaks <- readRDS(here("analysis", "data", "derived_data", "lod_peaks.rds"))
-annots <- readRDS(here("analysis", "data", "derived_data", "annots.rds"))
-expr <- readRDS(here("analysis", "data", "derived_data", "expr.rds"))
-```
-
-```{r load_tidyverse}
-library(tidyverse)
-options(tibble.print_max = 50, tibble.print_min = 20)
-```
-
-```{r make_local_annot-2}
-local_annot <- lod_peaks %>%
-  filter(chrom == hot_chr, pos <= hot_upper, pos >= hot_lower) %>%
-  left_join(y = annots, by = c("annot.id" = "gene_id")) %>%
-  filter(!duplicated(annot.id)) %>%
-  filter(lod >= ult) %>%
-  mutate(local_tx = chr == hot_chr) %>%
-  filter(local_tx)
-```
-
-```{r make_nonlocal_annot-2}
-nonlocal_annot <- lod_peaks %>%
-  filter(chrom == hot_chr, pos <= hot_upper, pos >= hot_lower) %>%
-  inner_join(y = annots, by = c("annot.id" = "gene_id")) %>%
-  filter(!duplicated(annot.id)) %>%
-  filter(lod >= 7.18) %>%
-  mutate(local_tx = chr == hot_chr) %>%
-  filter(!local_tx) %>%
-  filter(!is.na(hotspot))
-```
-
-```{r make_local_annot2-2}
-local_annot2 <- local_annot %>%
-  filter(middle <= hot_upper + 5, middle >= hot_lower - 5) %>%
-  arrange(desc(lod))
-```
-
-```{r subset_expr_matrix-2}
-hot_local <- expr[, colnames(expr) %in% local_annot2$annot.id]
-hot_nonlocal <- expr[, colnames(expr) %in% nonlocal_annot$annot.id]
-nnonlocal <- nrow(nonlocal_annot)
-nlocal <- nrow(local_annot2)
-```
-
-```{r tabulate-2, message=FALSE, warning = FALSE}
-med_tib <- med_list %>%
-  purrr::map(bind_rows) %>%
-  bind_rows() %>%
-  as_tibble() %>%
-  rename(lod_without_med = lod_no_med, lod_with_med = lod_med) %>%
-  mutate(lod_diff = lod_without_med - lod_with_med, lod_diff_proportion = lod_diff / lod_without_med) %>%
-  left_join(local_annot2, by = c("local_gene_id" = "annot.id")) %>%
-  left_join(nonlocal_annot, by = c("nonlocal_gene_id" = "annot.id")) %>%
-  dplyr::select(symbol.x, symbol.y, lod_without_med, lod_with_med, lod_diff, lod_diff_proportion, local_gene_id, nonlocal_gene_id) %>%
-  rename(local_symbol = symbol.x, nonlocal_symbol = symbol.y) %>%
-  filter(!is.na(local_symbol))
-```
-
-```{r make_pleio_tib-2}
-hot_annot <- local_annot2 %>%
-  bind_rows(nonlocal_annot) %>%
-  dplyr::select(annot.id, symbol)
-pleio_tib <- pleio_mixed_list %>%
-  purrr::map(.f = function(x) rename(x, gene1_id = local_gene_id, gene2_id = nonlocal_gene_id)) %>%
-  c(pleio_local_list, pleio_nonlocal_list) %>%
-  bind_rows() %>%
-  rename(pleiotropy_lod = lod) %>%
-  inner_join(hot_annot, by = c("gene1_id" = "annot.id")) %>%
-  rename(gene1_symbol = symbol) %>%
-  inner_join(hot_annot, by = c("gene2_id" = "annot.id")) %>%
-  rename(gene2_symbol = symbol) %>%
-  dplyr::select(gene1_symbol, gene2_symbol, pleiotropy_lod)
-```
-
-```{r make_pm-2}
-pm <- pleio_tib %>%
-  tibble_to_matrix(symmetric = TRUE)
-```
-
-```{r scatter-2}
-p <- pleio_mixed_list %>%
-  bind_rows() %>%
-  rename(pleiotropy_lod = lod) %>%
-  inner_join(med_tib, by = c("local_gene_id", "nonlocal_gene_id")) %>%
-  dplyr::select(local_symbol, nonlocal_symbol, pleiotropy_lod, lod_diff, lod_diff_proportion, lod_without_med, lod_with_med, filename, nonlocal_gene_id, local_gene_id) %>%
-  ggplot() + geom_point(mapping = aes(x = pleiotropy_lod, y = lod_diff, colour = local_symbol == keller_mediator, nonlocal = nonlocal_symbol), size = 0.05) + theme(legend.position = "none") + xlab("Pleiotropy LOD") + ylab("Mediation LOD difference")
-fnsvg <- paste0("Chr", hot_chr, "_scatter.svg")
-fnpdf <- paste0("Chr", hot_chr, "_scatter.pdf")
-ggsave(plot = p, filename = here("analysis", "figures", fnsvg))
-ggsave(plot = p, filename = here("analysis", "figures", fnpdf))
-p
-```
-
-```{r scatter_by_local-2}
-foo <- pleio_mixed_list %>%
-  bind_rows() %>%
-  rename(pleiotropy_lod = lod) %>%
-  inner_join(med_tib, by = c("local_gene_id", "nonlocal_gene_id")) %>%
-  dplyr::select(local_symbol, nonlocal_symbol, pleiotropy_lod, lod_diff, lod_diff_proportion, lod_without_med, lod_with_med, filename, nonlocal_gene_id, local_gene_id)
-foo %>%
-  ggplot() +
-  geom_point(mapping = aes(x = pleiotropy_lod, y = lod_diff, colour = nonlocal_symbol, nonlocal = nonlocal_symbol), size = 0.1) +
-  facet_wrap(. ~ local_symbol, nrow = ceiling(nlocal / 2), ncol = 2) +
-  broman::karl_theme(strip.placement = "inside") +
-  theme(legend.position = "none") +
-  theme(
-    strip.background = element_blank(),
-    strip.text.x = element_blank()
-) +
-  xlab("Pleiotropy LOD") +
-  ylab("Mediation LOD difference") +
-  geom_text(aes(x = 1, y = 1, label = local_symbol), inherit.aes=FALSE)
-
-fnsvg <- paste0("Chr", hot_chr, "_scatter-panel.svg")
-fnpdf <- paste0("Chr", hot_chr, "_scatter-panel.pdf")
-ggsave(filename = here("analysis", "figures", fnsvg), height = ceiling(nlocal / 2))
-ggsave(filename = here("analysis", "figures", fnpdf), height = ceiling(nlocal / 2))
-
-# put in multiple figures, if there are many local traits
-n_figs <- 1 + nlocal %/% 18 # 18 is the number of panels per page in pdf
-n_last <- nlocal %% 18 # number of panels on last figure
-max_pleio_lod <- max(foo$pleiotropy_lod)
-max_med_lod_difference <- max(foo$lod_diff)
-min_med_lod_difference <- min(foo$lod_diff)
-
-for (i in 1:n_figs){
-  n_panels <- 18 * (i < n_figs) + n_last * (i == n_figs)
-  foo %>%
-    filter(local_symbol %in% local_annot2$symbol[((i - 1) * 18 + 1): (18 * i)]) %>%
-    ggplot() +
-      geom_point(mapping = aes(x = pleiotropy_lod, y = lod_diff, colour = nonlocal_symbol, nonlocal = nonlocal_symbol), size = 0.1) +
-      facet_wrap(. ~ local_symbol, nrow = ceiling(n_panels / 2), ncol = 2) +
-      broman::karl_theme() +
-      theme(legend.position = "none") +
-      theme(
-        strip.background = element_blank(),
-        strip.text.x = element_blank()
-    ) +
-      xlab("Pleiotropy LOD") + xlim(c(0, max_pleio_lod)) + ylim(c(min_med_lod_difference, max_med_lod_difference)) +
-      ylab("Mediation LOD difference") +
-      geom_text(size = 2, aes(x = max_pleio_lod / 2, y = 2 * max_med_lod_difference / 3, label = local_symbol), inherit.aes=FALSE)
-  fnsvg <- paste0("Chr", hot_chr, "_scatter-panel", i, ".svg")
-  fnpdf <- paste0("Chr", hot_chr, "_scatter-panel", i, ".pdf")
-  ggsave(filename = here("analysis", "figures", fnsvg), height = ceiling(n_panels / 2))
-  ggsave(filename = here("analysis", "figures", fnpdf), height = ceiling(n_panels / 2))
-}
-```
-
-```{r load_heatmaply}
-library(heatmaply)
-```
-
-```{r make_scols-2}
-scols1 <- 1L + (rownames(pm) %in% local_annot2$symbol)
-scols2 <- rownames(pm) %in% (med_tib %>% filter(local_symbol == keller_mediator, lod_diff > 1.5))$nonlocal_symbol
-```
-
-```{r heat8-2}
-# order the nonlocal traits first
-## indicator for being a nonlocal trait
-indic_nonlocal <- rownames(pm) %in% nonlocal_annot$symbol
-# define ord_nonlocal
-ord_nonlocal <- pm[indic_nonlocal, indic_nonlocal] %>%
-  dist() %>%
-  hclust() %>%
-  as.dendrogram() %>%
-  dendextend::seriate_dendrogram(dist(pm[indic_nonlocal, indic_nonlocal])) %>%
-  order.dendrogram()
-rn_nonlocal <- rownames(pm[indic_nonlocal, indic_nonlocal][ord_nonlocal, ord_nonlocal])
-# define ord_local
-ord_local<- pm[!indic_nonlocal, !indic_nonlocal] %>%
-  dist() %>%
-  hclust() %>%
-  as.dendrogram() %>%
-  dendextend::seriate_dendrogram(dist(pm[!indic_nonlocal, !indic_nonlocal])) %>%
-  order.dendrogram()
-rn_local <- rownames(pm[!indic_nonlocal, !indic_nonlocal][ord_local, ord_local])
-rn <- c(rn_nonlocal, rn_local)
-```
-
-```{r make_scols-extra-2}
-scols_extra <- purrr::map(.x = local_annot2$symbol,
-                          .f = function(x) {rownames(pm) %in% (med_tib %>% filter(local_symbol == x, lod_diff > 1.5))$nonlocal_symbol}
-) %>%
-  bind_cols() %>%
-  (function(x){colnames(x) <- local_annot2$symbol; x}) %>%
-  mutate(local = scols1)
-```
-
-```{r count_med}
-n_med_tr <- 4 # arbitrary threshold
-n_med_per_local <- med_tib %>%
-  group_by(local_symbol) %>%
-  filter(lod_diff > 1.5) %>%
-  count() %>%
-  filter(n >= n_med_tr)
-n_med_per_local %>%
-  arrange(desc(n)) %>%
-  knitr::kable()
-
-scols_pre <- scols_extra %>%
-  dplyr::select(n_med_per_local$local_symbol)
-# get ordering for columns of scols_pre
-col_order <- scols_pre %>%
-  t() %>%
-  dist() %>%
-  hclust() %>%
-  as.dendrogram() %>%
-  dendextend::seriate_dendrogram(dist(t(scols_pre))) %>%
-  order.dendrogram()
-scols <- scols_pre[ , col_order] %>%
-  mutate(local = scols1)
-fnhtml <- here("analysis", "figures", paste0("Chr", hot_chr, "hm.html"))
-```
-
-```{r heat9-2, eval = !file.exists(fnhtml)}
-ord <- match(rn, rownames(pm))
-
-p <- pm[ord, ord] %>%
-  apply(FUN = trunc2, MARGIN = 2, threshold = 5) %>%
-  heatmaply(Rowv = FALSE,
-            Colv = FALSE,
-            cexRow = 0.2,
-            cexCol = 0.2,
-            symm = TRUE,
-            row_side_colors = scols[ord, ],
-            col_side_colors = tibble(local = scols1[ord]),
-            limits = c(0, 5)
-            ) %>%
-  hide_guides()
-fnpdf <- here("analysis", "figures", paste0("Chr", hot_chr, "hm.pdf"))
-htmlwidgets::saveWidget(widget = p,
-                        file = fnhtml
-                        )
-webshot::webshot(url = fnhtml,
-                 file = fnpdf,
-                 cliprect = "viewport"
-                 )
-```
-
-```{r heat10-2}
-webshot::webshot(fnhtml)
-```
-
-```{r index-v-peak-position-2}
-tibble(symbol = rownames(pm[indic_nonlocal, indic_nonlocal])[ord_nonlocal], index = 1:length(symbol)) %>%
-  left_join(nonlocal_annot, by = "symbol") %>%
-  ggplot() + geom_point(aes(y = pos, x = index))
+```{r child="hotspot-one-chromosome.Rmd"}
 ```
@@
 
-```{r more_setup-5, ref.label = "more_setup-2"}
-```
-
-```{r readRDS-5, ref.label="readRDS-2"}
-```
-
-```{r make_local_annot-5, ref.label="make_local_annot-2"}
-```
-
-```{r make_nonlocal_annot-5, ref.label = "make_nonlocal_annot-2"}
-```
-
-```{r make_local_annot2-5, ref.label = "make_local_annot2-2"}
-```
-
-```{r subset_expr_matrix-5, ref.label = "subset_expr_matrix-2"}
-```
-
-```{r tabulate-5, ref.label = "tabulate-2", message=FALSE, warning = FALSE}
-```
-
-```{r make_pleio_tib-5, ref.label="make_pleio_tib-2"}
-```
-
-```{r make_pm-5, ref.label = "make_pm-2"}
-```
-
-```{r scatter-5, ref.label="scatter-2"}
-```
-
-```{r scatter_by_local-5, ref.label = "scatter_by_local-2"}
-```
-
-```{r make_scols-5, ref.label="make_scols-2"}
-```
-
-```{r heat8-5, ref.label="heat8-2"}
-```
-
-```{r make_scols-extra-5, ref.label = "make_scols-extra-2"}
-```
-
-```{r count_med-5, ref.label = "count_med"}
-```
-
-```{r heat9-5, ref.label = "heat9-2", eval = !file.exists(fnhtml)}
-```
-
-```{r heat10-5, ref.label="heat10-2"}
-```
-
-```{r index-v-peak-position-5, ref.label = "index-v-peak-position-2"}
+```{r child="hotspot-one-chromosome.Rmd"}
 ```
@@
 
-```{r more_setup-7, ref.label = "more_setup-2"}
-```
-
-```{r readRDS-7, ref.label="readRDS-2"}
-```
-
-```{r make_local_annot-7, ref.label="make_local_annot-2"}
-```
-
-```{r make_nonlocal_annot-7, ref.label = "make_nonlocal_annot-2"}
-```
-
-```{r make_local_annot2-7, ref.label = "make_local_annot2-2"}
-```
-
-```{r subset_expr_matrix-7, ref.label = "subset_expr_matrix-2"}
-```
-
-```{r tabulate-7, ref.label = "tabulate-2", message=FALSE, warning = FALSE}
-```
-
-```{r make_pleio_tib-7, ref.label="make_pleio_tib-2"}
-```
-
-```{r make_pm-7, ref.label = "make_pm-2"}
-```
-
-```{r scatter-7, ref.label="scatter-2"}
-```
-
-```{r scatter_by_local-7, ref.label = "scatter_by_local-2"}
-```
-
-```{r make_scols-7, ref.label="make_scols-2"}
-```
-
-```{r heat8-7, ref.label="heat8-2"}
-```
-
-```{r make_scols-extra-7, ref.label = "make_scols-extra-2"}
-```
-
-```{r count_med-7, ref.label = "count_med"}
-```
-
-```{r heat9-7, ref.label = "heat9-2", eval = !file.exists(fnhtml)}
-```
-
-```{r heat10-7, ref.label="heat10-2"}
-```
-
-```{r index-v-peak-position-7, ref.label = "index-v-peak-position-2"}
+```{r child="hotspot-one-chromosome.Rmd"}
 ```
@@
 
-```{r more_setup-11, ref.label = "more_setup-2"}
-```
-
-```{r readRDS-11, ref.label="readRDS-2"}
-```
-
-```{r make_local_annot-11, ref.label="make_local_annot-2"}
-```
-
-```{r make_nonlocal_annot-11, ref.label = "make_nonlocal_annot-2"}
-```
-
-```{r make_local_annot2-11, ref.label = "make_local_annot2-2"}
-```
-
-```{r subset_expr_matrix-11, ref.label = "subset_expr_matrix-2"}
-```
-
-```{r tabulate-11, ref.label = "tabulate-2", message=FALSE, warning = FALSE}
-```
-
-```{r make_pleio_tib-11, ref.label="make_pleio_tib-2"}
-```
-
-```{r make_pm-11, ref.label = "make_pm-2"}
-```
-
-```{r scatter-11, ref.label="scatter-2"}
-```
-
-```{r scatter_by_local-11, ref.label = "scatter_by_local-2"}
-```
-
-```{r make_scols-11, ref.label="make_scols-2"}
-```
-
-```{r heat8-11, ref.label="heat8-2"}
-```
-
-```{r make_scols_extra-11, ref.label = "make_scols-extra-2"}
-```
-
-```{r count_med-11, ref.label = "count_med"}
-```
-
-```{r heat9-11, ref.label = "heat9-2", eval = !file.exists(fnhtml)}
-```
-
-```{r heat10-11, ref.label="heat10-2"}
-```
-
-```{r index-v-peak-position-11, ref.label = "index-v-peak-position-2"}
+```{r child="hotspot-one-chromosome.Rmd"}
 ```
@@
 
-```{r more_setup-13, ref.label = "more_setup-2"}
-```
-
-```{r readRDS-13, ref.label="readRDS-2"}
-```
-
-```{r make_local_annot-13, ref.label="make_local_annot-2"}
-```
-
-```{r make_nonlocal_annot-13, ref.label = "make_nonlocal_annot-2"}
-```
-
-```{r make_local_annot2-13, ref.label = "make_local_annot2-2"}
-```
-
-```{r subset_expr_matrix-13, ref.label = "subset_expr_matrix-2"}
-```
-
-```{r tabulate-13, ref.label = "tabulate-2", message=FALSE, warning = FALSE}
-```
-
-```{r make_pleio_tib-13, ref.label="make_pleio_tib-2"}
-```
-
-```{r make_pm-13, ref.label = "make_pm-2"}
-```
-
-```{r scatter-13, ref.label="scatter-2"}
-```
-
-```{r scatter_by_local-13, ref.label = "scatter_by_local-2"}
-```
-
-```{r make_scols-13, ref.label="make_scols-2"}
-```
-
-```{r heat8-13, ref.label="heat8-2"}
-```
-
-```{r make_scols-extra-13, ref.label = "make_scols-extra-2"}
-```
-
-```{r count_med-13, ref.label = "count_med"}
-```
-
-```{r heat9-13, ref.label = "heat9-2", eval = !file.exists(fnhtml)}
-```
-
-```{r heat10-13, ref.label="heat10-2"}
-```
-
-```{r index-v-peak-position-13, ref.label = "index-v-peak-position-2"}
+```{r child="hotspot-one-chromosome.Rmd"}
 ```

### 2019-07-28 analysis/paper/intro.Rmd
added files for dift sections, to be assembled in hotspots-paper.Rmd

# Introduction

Genetics studies in model organisms like mice can identify
genomic regions that affect complex traits, such as systolic blood pressure and body weight [@sax1923association;@soller1976power;@lander1989mapping;@broman2009guide;@jansen2007quantitative]. 
These genomic regions are called "quantitative trait loci" or "QTL".
A genome-wide QTL "scan" reveals
associations between genotypes and phenotypes by considering each position along the genome,
one at a time, as a candidate QTL for the trait of interest.
A region with strong evidence of association with a complex trait, then, defines a QTL (for that trait).
Because nearby markers have correlated
genotypes, a QTL in a two-parent cross often spans multiple megabases in length
and may contain more than a hundred genes.
Identification of a causal gene (for a given complex trait) from among those genes near the QTL is
challenging and may require costly and time-consuming experiments.
The growing need for greater QTL mapping resolution fueled development over the
last two decades of model organism multiparental populations for high-resolution QTL mapping [@de2017back;@churchill2004collaborative;@svenson2012high;@huang2012multiparent;@shivakumar2018soybean;@huang2011analysis;@kover2009multiparent;@tisne2017identification;@stanley2017genetic].
With experimentalists now measuring tens of thousands of biomolecular traits
in multiparental populations, the systems genetics community needs multivariate
statistical tools to fully examine the large volumes of data [@keller2018genetic;@chick2016defining].
Identifying loci that affect multiple traits can aid in identifying biomolecular interactions and in clarifying complex trait genetic architecture. A QTL that affects more than one trait is called a "pleiotropic" QTL.
A test of pleiotropy vs. separate QTL is one multivariate statistical tool that will inform
complex trait genetics by enabling researchers to identify the number of
unique QTL in a genomic region of interest.

### 2019-07-28 analysis/paper/methods.Rmd
added files for dift sections, to be assembled in hotspots-paper.Rmd

--- 2019-07-28 analysis/paper/hotspots-paper-Chr2-only.Rmd
@@
----
-title: "Expression QTL hotspot dissection with mediation analysis and pleiotropy testing"
-author: "Frederick J. Boehm, Brian S. Yandell, and Karl W. Broman"
-date: "7/27/2019"
-bibliography: research.bib
-output: 
-  bookdown::html_document2:
-    pandoc_args: 
-      - --filter
-      - /usr/local/Cellar/pandoc-crossref/0.3.4.1/bin/pandoc-crossref
----
-
-```{r load-broman}
-library(broman)
-library(here)
-```
-
 # Methods
@@
 vec(Y) = X vec(B) + vec(G) + vec(E)
-$$
+$${eq:mvlmm}
 
@@
 
-### Calculating likelihood ratio test statistics}
+### Calculating likelihood ratio test statistics
+
 From the two-dimensional scans, we calculated pleiotropy likelihood ratio test statistics for each
@@
 LOD = \log_{10} \left(\frac{\max_{\lambda_1, \lambda_2}L(B, \Sigma, \lambda_1, \lambda_2)}{\max_{\lambda}L(B, \Sigma, \lambda, \lambda)} \right)
-$$
+$${eq:lod}
 
@@
 Y = \alpha 1 + WC + E
-\label{model1}
-$$
+$${eq:model1}
 
@@
 Y = XB + WC + E
-\label{model2}
-$$
+$${eq:model2}
 
@@
 Y = \alpha 1 + WC + M\beta + E
-\label{model3}
-$$
+$${eq:model3}
 
@@
 Y = XB + WC + M\beta + E
-\label{model4}
-$$
+$${eq:model4}
 
-In all mediation analyses, the nonlocal transcripts are the targets, while the local
+In all of our mediation analyses, the nonlocal transcripts are the targets, while the local
 transcripts serve as candidate mediators of the QTL-target relationship.
@@
 yellow.
-
-# Results
-
-We generated scatter plots of pleiotropy test statistics vs. mediation LOD difference values for each of the five hotspots (Figures).
-Each point represents a pairing of a local trait with a nonlocal trait.
-The mediation LOD difference value is that observed when the local trait is considered
-as a candidate mediator of the QTL-nonlocal trait association.
-
-In all five scatter plots, we observe few
-points in the upper right quadrant of the figure.
-That is, few pairs have high pleiotropy test statistics and
-high mediation LOD difference values.
-
-```{r setup}
-library(qtl2hotspots)
-pleio_threshold <- 4
-ult <- 15 # univariate lod threshold
-```
-
-```{r load_tidyverse}
-library(tidyverse)
-options(tibble.print_max = 50, tibble.print_min = 20)
-```
-
-```{r load_heatmaply}
-library(heatmaply)
-```
-
-```{r setup_continued-2}
-hot_chr <- 2
-hot_mid <- 165.5
-keller_mediator <- "Hnf4a"
-```
-
-```{r child="hotspot-one-chromosome.Rmd"}
-```
-
-# Discussion
-
-Pleiotropy tests complement mediation analyses in the following manner. When considering a
-collection of potential mediators for a nonlocal trait, mediation analyses clearly identify the
-mediator when there is, in fact, a mediator among the candidates. However, when there is no
-mediator, mediation analyses provide little information about the genetics of the complex trait.
-Pleiotropy testing is particularly useful in this setting, since a pleiotropy test always gives
-an answer about the number of underlying QTL.
-A drawback of pleiotropy testing is that the scientific question that it addresses - do two traits share a single QTL - informs, but doesn't fully resolve, the genetics of the nonlocal complex traits under study.
-
-# Colophon
-
-This report was generated on ` r Sys.time()` using the following computational environment and dependencies:
-
-```{r colophon, cache = FALSE, echo = TRUE, tidy.opts=list(width.cutoff=50), size = "tiny", results = "markup"}
-# which R packages and versions?
-devtools::session_info()
-```
-
-The current Git commit details are:
-
-```{r git2r, cache = FALSE, echo = TRUE}
-# what commit is this file at?
-git2r::repository(here::here())
-```

### 2019-07-29 analysis/paper/hotspots-paper-Chr2-only.Rmd
increased travis_wait to 100

@@
 bibliography: research.bib
+numberSections: true
 output: 
@@
 Y = \alpha 1 + WC + E
-\label{model1}
 $$

### 2019-07-29 analysis/paper/hotspot-one-chromosome.Rmd
fixed equation labels

@@
 ```{r}
-tibble(symbol = rownames(pm[indic_nonlocal, indic_nonlocal])[ord_nonlocal], index = 1:length(symbol)) %>%
-  left_join(nonlocal_annot, by = "symbol") %>%
-  ggplot() + geom_point(aes(y = pos, x = index))
+bar <- foo %>%
+  filter(local_symbol == keller_mediator)
+
+p <- tibble(nonlocal_symbol = rownames(pm[indic_nonlocal, indic_nonlocal])[ord_nonlocal], index = 1:length(nonlocal_symbol)) %>%
+  inner_join(bar, by = "nonlocal_symbol") %>% 
+  left_join(nonlocal_annot, by = c("nonlocal_symbol" = "symbol")) %>%
+  ggplot() + geom_point(aes(y = pos, x = index, color = lod_diff > 1.5, nonlocal = nonlocal_symbol, lod_diff = lod_diff))
+
+ggplotly(p)
 ```

### 2019-07-29 analysis/paper/hotspots-paper.Rmd
fixed equation labels

@@
 numberSections: true
-autoEqnLabels: true
+linkReferences: true
+link-citations: true
 output: 
@@
 
+```{r, tmem189-scatter, fig.cap="Colors correspond to mediation by Tmem189"}
+bar <- foo %>%
+  filter(local_symbol == "Tmem189")
+
+p <- tibble(nonlocal_symbol = rownames(pm[indic_nonlocal, indic_nonlocal])[ord_nonlocal], index = 1:length(nonlocal_symbol)) %>%
+  inner_join(bar, by = "nonlocal_symbol") %>% 
+  left_join(nonlocal_annot, by = c("nonlocal_symbol" = "symbol")) %>%
+  ggplot() + geom_point(aes(y = pos, x = index, color = lod_diff > 1.5, nonlocal = nonlocal_symbol, lod_diff = lod_diff))
+
+ggplotly(p)
+```
+
+## Chr 2 hotspot results
+
 ```{r setup_continued-5}
@@
 
-```{r child="hotspot-one-chromosome.Rmd"}
+```{r child="hotspot-one-chromosome.Rmd", eval = FALSE}
 ```

### 2019-07-29 analysis/paper/intro.Rmd
fixed equation labels

@@
 Genetics studies in model organisms like mice can identify
-genomic regions that affect complex traits, such as systolic blood pressure and body weight [@sax1923association;@soller1976power;@lander1989mapping;@broman2009guide;@jansen2007quantitative]. 
+genomic regions that affect complex quantitative traits, such as systolic blood pressure, body weight, and, more recently, biomolecular traits such as gene expression levels and protein concentrations [@sax1923association;@soller1976power;@lander1989mapping;@broman2009guide;@jansen2007quantitative;@chick2016defining].
 These genomic regions are called "quantitative trait loci" or "QTL".
@@
 one at a time, as a candidate QTL for the trait of interest.
-A region with strong evidence of association with a complex trait, then, defines a QTL (for that trait).
+A region with strong evidence of association with a complex quantitative trait, then, defines a QTL (for that trait).
 Because nearby markers have correlated
@@
 last two decades of model organism multiparental populations for high-resolution QTL mapping [@de2017back;@churchill2004collaborative;@svenson2012high;@huang2012multiparent;@shivakumar2018soybean;@huang2011analysis;@kover2009multiparent;@tisne2017identification;@stanley2017genetic].
-With experimentalists now measuring tens of thousands of biomolecular traits
-in multiparental populations, the systems genetics community needs multivariate
-statistical tools to fully examine the large volumes of data [@keller2018genetic;@chick2016defining].
-Identifying loci that affect multiple traits can aid in identifying biomolecular interactions and in clarifying complex trait genetic architecture.
-A QTL that affects more than one trait is called a "pleiotropic" QTL.
-A test of pleiotropy vs. separate QTL is one multivariate statistical tool that will inform
-complex trait genetics by enabling researchers to identify the number of
-unique QTL in a genomic region of interest.
+
+Recent systems genetics studies have sought to identify QTL for thousands of biomolecular traits [@keller2018genetic;@chick2016defining].

### 2019-07-29 analysis/paper/methods.Rmd
fixed equation labels

@@
 
+$$
+vec(Y) = X vec(B) + vec(G) + vec(E)
+$${#eq:mvlmm}
+
 For each hotspot, we tested all pairs of traits, regardless of whether the traits were local or nonlocal. We performed a
 two-dimensional, two-QTL scan over a genomic region that included the entirety of the
-hotspot. We fitted bivariate linear mixed effects models at each ordered pair of markers in a two-dimensional grid [@boehm2019testing].
-
-$$
-vec(Y) = X vec(B) + vec(G) + vec(E)
-$${eq:mvlmm}
+hotspot. We fitted bivariate linear mixed effects models (Equations [-@eq:mvlmm;-@eq:g;-@eq:e]) at each ordered pair of markers in a two-dimensional grid [@boehm2019testing].
 
@@
 G \sim MN_{n x 2}(0, K, V_g)
-$$
+$${#eq:g}
 
@@
 E \sim MN_{n x 2}(0, I_n, V_e)
-$$
+$${#eq:e}
 
@@
 LOD = \log_{10} \left(\frac{\max_{\lambda_1, \lambda_2}L(B, \Sigma, \lambda_1, \lambda_2)}{\max_{\lambda}L(B, \Sigma, \lambda, \lambda)} \right)
-$${eq:lod}
+$${#eq:lod}
 
@@
 Y = \alpha 1 + WC + E
-$${eq:model1}
+$${#eq:model1}
 
@@
 Y = XB + WC + E
-$${eq:model2}
+$${#eq:model2}
 
@@
 Y = \alpha 1 + WC + M\beta + E
-$${eq:model3}
+$${#eq:model3}
 
@@
 Y = XB + WC + M\beta + E
-$${eq:model4}
+$${#eq:model4}
 
 In all of our mediation analyses, the nonlocal transcripts are the targets, while the local
-transcripts serve as candidate mediators of the QTL-target relationship.
+transcripts serve as candidate mediators of the QTL-target relationship. We only considered as potential mediators those local traits with strong univariate LODs.
 In Equations~\ref{model1} through \ref{model4}, $X$ is a matrix of founder allele dosages,

### 2019-07-29 analysis/paper/hotspot-one-chromosome.Rmd
fixed interval defining local traits

@@
 local_annot2 <- local_annot %>%
-  filter(middle <= hot_upper + 1, middle >= hot_lower - 1) %>% 
-  # extend only 1 mb beyond hotspot end points (total interval length, for local traits, = 6 Mb)
+  filter(middle <= hot_upper, middle >= hot_lower) %>% 
   arrange(desc(lod))

### 2019-07-29 analysis/paper/hotspot-one-chromosome.Rmd
moved code chunks to hotspots-paper.Rmd from hotspot-one-chromosome.Rmd

@@
 med_list <- readRDS(here("analysis", "data", "derived_data", paste0("Chr", hot_chr, "_mediations.rds")))
-```
-
-```{r}
-if (!file.exists(here("analysis", "data", "derived_data", "expr.rds"))){
-  download.file("https://datadryad.org/bitstream/handle/10255/dryad.166774/Attie_DO378_eQTL_viewer_v1.Rdata?sequence=2",
-              destfile = here("analysis", "data", "raw_data", "Attie_DO378_eQTL_viewer_v1.Rdata")
-              )
-  load(here("analysis", "data", "raw_data", "Attie_DO378_eQTL_viewer_v1.Rdata"))
-  saveRDS(dataset.islet.rnaseq$lod.peaks, here("analysis", "data", "derived_data", "lod_peaks.rds"))
-  saveRDS(dataset.islet.rnaseq$annots, here("analysis", "data", "derived_data", "annots.rds"))
-  saveRDS(dataset.islet.rnaseq$expr, here("analysis", "data", "derived_data", "expr.rds"))
-}
-```
-
-```{r}
-lod_peaks <- readRDS(here("analysis", "data", "derived_data", "lod_peaks.rds"))
-annots <- readRDS(here("analysis", "data", "derived_data", "annots.rds"))
-expr <- readRDS(here("analysis", "data", "derived_data", "expr.rds"))
 ```
@@
 local_annot2 <- local_annot %>%
-  filter(middle <= hot_upper, middle >= hot_lower) %>% 
+  filter(middle <= hot_upper + 1, middle >= hot_lower - 1) %>% 
   arrange(desc(lod))

### 2019-07-29 analysis/paper/hotspots-paper.Rmd
moved code chunks to hotspots-paper.Rmd from hotspot-one-chromosome.Rmd

@@
 
+```{r}
+if (!file.exists(here("analysis", "data", "derived_data", "expr.rds"))){
+  download.file("https://datadryad.org/bitstream/handle/10255/dryad.166774/Attie_DO378_eQTL_viewer_v1.Rdata?sequence=2",
+              destfile = here("analysis", "data", "raw_data", "Attie_DO378_eQTL_viewer_v1.Rdata")
+              )
+  load(here("analysis", "data", "raw_data", "Attie_DO378_eQTL_viewer_v1.Rdata"))
+  saveRDS(dataset.islet.rnaseq$lod.peaks, here("analysis", "data", "derived_data", "lod_peaks.rds"))
+  saveRDS(dataset.islet.rnaseq$annots, here("analysis", "data", "derived_data", "annots.rds"))
+  saveRDS(dataset.islet.rnaseq$expr, here("analysis", "data", "derived_data", "expr.rds"))
+}
+```
+
+```{r}
+lod_peaks <- readRDS(here("analysis", "data", "derived_data", "lod_peaks.rds"))
+annots <- readRDS(here("analysis", "data", "derived_data", "annots.rds"))
+expr <- readRDS(here("analysis", "data", "derived_data", "expr.rds"))
+```
+
 ```{r setup_continued-2}
@@
 ```{r child="hotspot-one-chromosome.Rmd"}
-```
-
-```{r, tmem189-scatter, fig.cap="Colors correspond to mediation by Tmem189"}
-bar <- foo %>%
-  filter(local_symbol == "Tmem189")
-
-p <- tibble(nonlocal_symbol = rownames(pm[indic_nonlocal, indic_nonlocal])[ord_nonlocal], index = 1:length(nonlocal_symbol)) %>%
-  inner_join(bar, by = "nonlocal_symbol") %>% 
-  left_join(nonlocal_annot, by = c("nonlocal_symbol" = "symbol")) %>%
-  ggplot() + geom_point(aes(y = pos, x = index, color = lod_diff > 1.5, nonlocal = nonlocal_symbol, lod_diff = lod_diff))
-
-ggplotly(p)
 ```

### 2019-07-29 analysis/paper/hotspots-paper.Rmd
added line to install cabal, which can be used to install pandoc-crossref

@@
       - --filter
-      - /usr/local/Cellar/pandoc-crossref/0.3.4.1/bin/pandoc-crossref
+      - pandoc-crossref
 ---

### 2019-07-30 analysis/paper/hotspots-paper.Rmd
added Chr2 hotspot results

@@
 
+The scatter plot of pleiotropy test statistics and mediation LOD differences reveals that few local-nonlocal 
+trait pairs have both a high value of the pleiotropy test statistic and a high value of the mediation LOD 
+difference. Many points have low values of both, while some points demonstrate a high value of only mediation 
+LOD difference or pleiotropy test statistic.
+
+Many points have negative values of mediation LOD difference. These correspond to pairs that, when conditioning 
+on the putative mediator, the LOD for the nonlocal trait increased.
+
+The points are color-coded to identify those pairs that involve the local trait *Hnf4a*, which 
+@keller2018genetic identified as a key mediator of traits at this hotspot. Blue points involve *Hnf4a*, while 
+red points do not. We see that many blue point have large mediation LOD differences and small pleiotropy test 
+statistics. In fact, most of the points with mediation LOD difference above 15 involve *Hnf4a* and have small 
+pleiotropy test statistics.
+
 ```{r setup_continued-5}

### 2019-07-30 analysis/paper/methods.Rmd
added Chr2 hotspot results

@@
 
+### Heatmaps
+
 We created heatmaps to examine patterns among pleiotropy test results in each hotspot.
 We represented the pleiotropy test statistics as a symmetric matrix, where each row was one expression
-trait and each column was an expression trait. Each cell, then, contained the pleiotropy test statistic
+trait and each column was an expression trait. 
+Excepting the cells on the diagonal, each cell, then, contained the pleiotropy test statistic
 from the analysis involving the row's expression trait and the column's expression trait.
-To create a heatmap, we first performed two hierarchical clusterings, one on the submatrix involving only
-nonlocal expression traits and a second on the submatrix involving only local expression traits. We
-calculated the Euclidean distances between all trait pairs before clustering. We ordered the rows and
-columns according to optimal leaf ordering. We then combined the
-local and nonlocal traits into a single heatmap. To achieve greater coloring contrasts in the heatmaps, we
-set the color spectrum limits at 0 (dark blue) and 5 (yellow). Test statistics above 5 are represented by
-yellow.
+To create a heatmap, we first performed two hierarchical clusterings of rows. 
+One hierarchical clustering involved only the rows and columns containing local traits, while the other contained only 
+rows and columns containing nonlocal traits.
+We calculated the Euclidean distances between all row pairs (in the respective submatrices) before clustering.
+We ordered the rows according to optimal leaf ordering [@hahsler2008getting;@dendextend].
+We then combined the
+local and nonlocal traits into a single heatmap for each hotspot. 
+To achieve coloring contrasts in the heatmaps, we
+set the color spectrum limits at 0 (dark blue) and 5 (yellow). Pleiotropy test statistics above 5, too, are 
+represented by
+yellow. We annotated rows by a collection of columns on the right-hand side of the heatmap. This binary matrix 
+consists of columns that summarize mediation results. Each column represents a local trait. If a row 
+corresponds to a nonlocal trait that is mediated by the local trait, then the cell is colored dark green; 
+otherwise, the cell is light green. The last annotation column and the annotation row above the heatmap both 
+designate local traits with dark blue and nonlocal traits with light blue.

### 2019-07-30 analysis/paper/methods.Rmd
updated methods section on heatmaps

@@
 
-We created heatmaps to examine patterns among pleiotropy test results in each hotspot.
+We created heatmaps to examine patterns among pleiotropy test results in each hotspot. Our analysis used both 
+local and nonlocal traits.
 We represented the pleiotropy test statistics as a symmetric matrix, where each row was one expression
@@
 rows and columns containing nonlocal traits.
-We calculated the Euclidean distances between all row pairs (in the respective submatrices) before clustering.
-We ordered the rows according to optimal leaf ordering [@hahsler2008getting;@dendextend].
+We calculated the Euclidean distances between all row pairs (in the respective 
+submatrices) before clustering. Hierarchical clustering was performed using the complete 
+linkage method. Initially, each row is assigned to its own cluster. An iterative 
+algorithm then combines the two most similar clusters at each step. Distances between 
+clusters were computed with the Lance-Williams dissimilarity update formula for complete 
+linkage. Lastly, we arranged the rows according to optimal leaf ordering [@hahsler2008getting;@dendextend]. 
 We then combined the

### 2019-07-30 analysis/paper/hotspots-paper.Rmd
made a new Makefile-docker.makefile for use with Docker

@@
 bibliography: research.bib
-numberSections: true
 linkReferences: true
 link-citations: true
-output: 
-  bookdown::html_document2:
-    pandoc_args: 
-      - --filter
-      - pandoc-crossref
+output: bookdown::html_document2
 ---
@@
 
+## Chr 5 results
+
 ```{r setup_continued-7}

### 2019-07-30 analysis/paper/methods.Rmd
made a new Makefile-docker.makefile for use with Docker

@@
 clusters were computed with the Lance-Williams dissimilarity update formula for complete 
-linkage. Lastly, we arranged the rows according to optimal leaf ordering [@hahsler2008getting;@dendextend]. 
+linkage [@lance1967general]. Lastly, we arranged the rows according to optimal leaf ordering [@hahsler2008getting;@dendextend]. 
 We then combined the
@@
 consists of columns that summarize mediation results. Each column represents a local trait. If a row 
-corresponds to a nonlocal trait that is mediated by the local trait, then the cell is colored dark green; 
-otherwise, the cell is light green. The last annotation column and the annotation row above the heatmap both 
+corresponds to a nonlocal trait that is mediated by the local trait (*i.e.*, with mediation LOD difference greater than 1.5), then the cell is colored dark green; 
+otherwise, the cell is light green.
+We didn't examine the possibility that a local trait might mediate the relationship between a QTL and another local trait; thus, all local trait rows are colored light green.
+The last annotation column and the annotation row above the heatmap both
 designate local traits with dark blue and nonlocal traits with light blue.

### 2019-07-30 analysis/paper/hotspot-one-chromosome.Rmd
changed params to output_options

@@
   ylab("Mediation LOD difference") +
-  geom_text(aes(x = 1, y = 1, label = local_symbol), inherit.aes=FALSE)
+  geom_text(aes(x = max(foo$pleiotropy_lod), y = 1, label = local_symbol), inherit.aes=FALSE)
 

### 2019-07-31 README.Rmd
added tibble of scan regions

@@
 ```{r example}
-#library(qtl2hotspots)
-## basic example code
+map <- readRDS(here::here("analysis", "data", "derived_data", "map.rds"))
+hot_chr <- c(2, 5, 7, 11, 13)
+start_index <- c(3742, 3498, 927, 1868, 2814) # from s1 values in condor submit files
+scan_length <- c(247, 269, 242, 217, 226) # in number of markers, from condor submit files
+stop_index <- start_index + scan_length - 1 # index of the last marker in the interval
+start_position <- c(map$`2`[start_index[1]], 
+                    map$`5`[start_index[2]], 
+                    map$`7`[start_index[3]],
+                    map$`11`[start_index[4]],
+                    map$`13`[start_index[5]]
+)
+stop_position <- c(map$`2`[stop_index[1]], 
+                    map$`5`[stop_index[2]], 
+                    map$`7`[stop_index[3]],
+                    map$`11`[stop_index[4]],
+                    map$`13`[stop_index[5]]
+)
+tibble::tibble(hot_chr, start_index, scan_length, stop_index, start_position, stop_position)
 ```

### 2019-07-31 analysis/paper/hotspots-paper.Rmd
updated paper text

@@
   saveRDS(dataset.islet.rnaseq$expr, here("analysis", "data", "derived_data", "expr.rds"))
+  saveRDS(K, here("analysis", "data", "derived_data", "kinship.rds"))
+  saveRDS(map, here("analysis", "data", "derived_data", "map.rds"))
+  saveRDS(genoprobs, here("analysis", "data", "derived_data", "genoprobs.rds"))
+  saveRDS(dataset.islet.rnaseq$covar, here("analysis", "data", "derived_data", "covar.rds"))
 }
@@
 pleiotropy test statistics.
+
+An unanticipated observation is that some points have mediation LOD differences of nearly 10 
+and pleiotropy test statistics of nearly 10. With such large pleiotropy test statistics, one 
+has reasonable evidence each point representing two separate QTL. It seems biologically 
+unlikely that a nonlocal trait and a mediating local trait would not share a pleiotropic QTL.
+
+In looking closer at these six points, we find that five of the six have *Tmem189* as the 
+local trait.
+This raises questions about the *Tmem189* trait.
+One known issue with the pleiotropy test is its sensitivity to start and end points for the two-dimensional scan.
+If the scan interval is too small, the test can return misleading results.
+If *Tmem189* were to have a very broad univariate peak, such that the peak extended beyond the scan interval, then we might observe misleading results in tests involving *Tmem189*.
+
+```{r scan1}
+# read oddball points from interactive plot
+tr <- c("Tmem189",
+        "Myo15b",
+        "Sephs2",
+        "Bcmo1",
+        "Tmprss4",
+        "Cacnb3",
+        "Gm12929",
+        "Eya2"
+        )
+tr_tib <- hot_annot %>%
+  filter(symbol %in% tr) %>%
+  arrange(annot.id)
+K <- readRDS(here("analysis", "data", "derived_data", "kinship.rds"))
+map <- readRDS(here("analysis", "data", "derived_data", "map.rds"))
+covar <- readRDS(here("analysis", "data", "derived_data", "covar.rds"))
+aprobs <- readRDS(here("analysis", "data", "derived_data", "genoprobs.rds"))
+expr_sm <- expr[ , colnames(expr) %in% tr_tib$annot.id, drop = FALSE]
+s1 <- qtl2::scan1(genoprobs = aprobs,
+            pheno = expr_sm,
+            kinship = K,
+            addcovar = covar, 
+            cores = 0
+)
+# zoom in on Chr2 after 150 Mb
+# get length of map$`1`
+purrr::map(.x = 1:8, .f = function(x) qtl2::plot_scan1(s1, map, lodcolumn = x, chr = 2, main = paste0(tr_tib[x, ])))
+qtl2::find_peaks(scan1_output = s1, map = map, threshold = 7) %>% 
+  as_tibble() %>%
+  filter(chr == 2) %>%
+  left_join(hot_annot, by = c("lodcolumn" = "annot.id"))
+```
 

### 2019-07-31 analysis/paper/hotspots-paper.Rmd
set eval to FALSE for code chunk about Tmem189

@@
 
-```{r scan1}
+```{r scan1-Chr2, eval = FALSE}
 # read oddball points from interactive plot

### 2019-07-31 README.Rmd
fixed README table

--- 2019-07-26 README.Rmd
@@
 
+```{r load_tidyverse}
+library(tidyverse)
+```
+
 ```{r example}
-#library(qtl2hotspots)
-## basic example code
+map <- readRDS(here::here("analysis", "data", "derived_data", "map.rds"))
+hot_chr <- c(2, 5, 7, 11, 13)
+start_index <- c(3742, 3498, 927, 1868, 2814) # from s1 values in condor submit files
+scan_length <- c(247, 269, 242, 217, 226) # in number of markers, from condor submit files
+stop_index <- start_index + scan_length - 1 # index of the last marker in the interval
+start_position <- c(map$`2`[start_index[1]], 
+                    map$`5`[start_index[2]], 
+                    map$`7`[start_index[3]],
+                    map$`11`[start_index[4]],
+                    map$`13`[start_index[5]]
+)
+stop_position <- c(map$`2`[stop_index[1]], 
+                    map$`5`[stop_index[2]], 
+                    map$`7`[stop_index[3]],
+                    map$`11`[stop_index[4]],
+                    map$`13`[stop_index[5]]
+)
+tibble::tibble(hot_chr, start_index, scan_length, stop_index, start_position, stop_position) %>%
+  knitr::kable()
 ```

### 2019-07-31 analysis/paper/hotspots-paper.Rmd
added back github_document output format
switched eval -> TRUE for Chr5 & 7 hotspots
set outputs to do html_document2 AFTER github_document; also set github_document with html_preview: false

@@
 link-citations: true
-output: bookdown::html_document2
+output: 
+  - github_document:
+      html_preview: false
+      toc: true
+      toc_depth: 3
+  - bookdown::html_document2  
 ---
@@
 
-```{r child="hotspot-one-chromosome.Rmd", eval = FALSE}
+```{r child="hotspot-one-chromosome.Rmd"}
 ```

### 2019-08-01 analysis/paper/hotspots-paper.Rmd
adding results text

@@
 
+The Chr 5 scatter plot of mediation LOD difference values against pleiotropy test statistics 
+reveals a block-diagonal pattern of blue coloring among the nonlocal traits. We observe the 
+set of nonlocal traits clustered into two blocks. Within each block, any two traits share a 
+QTL, ie, are pleiotropic. Yet, the off-diagonal cells, ie, those nonlocal trait pairs that 
+include one trait from each of the two blocks, tend to have greater pleiotropy test 
+statistics. Still, not all off-diagonal blocks have large pleiotropy test statistics. The 
+mostly dark blue matrix on the diagonal (from column 22 to 99) features no cells with pleiotropy test statistics above 3.2.
+While we didn't determine p-values for these analyses, prior analyses suggest that pleiotropy test statistics below 3.2 are consistent with p-values above 0.05, and, thus, are consistent with a single pleiotropic QTL.
+
+Columns 100 to 119 (and rows 100 to 119) are a second (symmetric) block with only small 
+pleiotropy test statistics.
+Columns 120 to 202 contains cells for the third (symmetric) block.
+Column 203, the last column, contains a trait *Gm* that, based on pleiotropy test statistics, doesn't cluster near other traits.
+
 ```{r setup_continued-7}
@@
 
-```{r child="hotspot-one-chromosome.Rmd", eval = FALSE}
+```{r child="hotspot-one-chromosome.Rmd"}
 ```
+
+Our analysis of the Chromosome 13 hotspot presents results much like those for the Chromosome 2 
+hotspot, with the Chromosome 13 local transcript *Il6st* playing a role like that of *Hnf4a* in the 
+Chr 2 hotspot. In examining the scatter plot (Figure ) of mediation LOD differences against 
+pleiotropy test statistics for all local-nonlocal pairs, we see, as in Figure for the Chr 2 
+hotspot, a single local trait *Il6st* has strong mediation effects on a collection of nonlocal traits.
+*Il6st*, first identified by @keller2018genetic, accounts for most of the nonlocal traits that are affected by the hotspot.
+
+The heatmap for pleiotropy statistics for pairwise tests involving nonlocal traits resembles the 
+heatmap for the Chr2 hotspot. Nearly all nonlocal traits have small test statistics when paired 
+with each other. This is visualized as a dark blue, symmetric square along the heatmap's diagonal 
+with only a few outlier nonlocal traits that demonstrate distinct patterns in pleiotropy test 
+statistics.
 
@@
 A drawback of pleiotropy testing is that the scientific question that it addresses - do two traits share a single QTL - informs, but doesn't fully resolve, the genetics of the nonlocal complex traits under study.
+
+- lod difference threshold for declaration of "mediation". 
+  - esp with regards to Chr 5 hotspot, where the keller mediator has a lot of traits with lod differences under 3, but above the 1.5 threshold.
+
+- look at outputs of `lod_int()`, ie, the support intervals for all traits in a hotspot. Could plot interval against cluster order, for example.
 

### 2019-08-01 analysis/paper/hotspots-paper.Rmd
added chunk names to main document

@@
 
+Simple hotspots: Chr 2, Chr 13
+
+Complex hotspots: Chr 5
+
+Undetermined: Chr 7, Chr 11
+
 We generated scatter plots of pleiotropy test statistics vs. mediation LOD difference values for each of the five hotspots (Figures).
@@
 
-```{r}
+```{r saveRDS}
 if (!file.exists(here("analysis", "data", "derived_data", "expr.rds"))){
@@
 
-```{r}
+```{r readRDS}
 lod_peaks <- readRDS(here("analysis", "data", "derived_data", "lod_peaks.rds"))
@@
 
+### Chr 2 scatter plots
+
 The scatter plot of pleiotropy test statistics and mediation LOD differences reveals that few local-nonlocal 
@@
 
-An unanticipated observation is that some points have mediation LOD differences of nearly 10 
-and pleiotropy test statistics of nearly 10. With such large pleiotropy test statistics, one 
-has reasonable evidence each point representing two separate QTL. It seems biologically 
-unlikely that a nonlocal trait and a mediating local trait would not share a pleiotropic QTL.
+An unanticipated observation is that some points have mediation LOD differences of
 
-In looking closer at these six points, we find that five of the six have *Tmem189* as the 
-local trait.
-This raises questions about the *Tmem189* trait.
-One known issue with the pleiotropy test is its sensitivity to start and end points for the two-dimensional scan.
-If the scan interval is too small, the test can return misleading results.
-If *Tmem189* were to have a very broad univariate peak, such that the peak extended beyond the scan interval, then we might observe misleading results in tests involving *Tmem189*.
+### Chr 2 heatmap
 
-```{r scan1-Chr2, eval = FALSE}
-# read oddball points from interactive plot
-tr <- c("Tmem189",
-        "Myo15b",
-        "Sephs2",
-        "Bcmo1",
-        "Tmprss4",
-        "Cacnb3",
-        "Gm12929",
-        "Eya2"
-        )
-tr_tib <- hot_annot %>%
-  filter(symbol %in% tr) %>%
-  arrange(annot.id)
-K <- readRDS(here("analysis", "data", "derived_data", "kinship.rds"))
-map <- readRDS(here("analysis", "data", "derived_data", "map.rds"))
-covar <- readRDS(here("analysis", "data", "derived_data", "covar.rds"))
-aprobs <- readRDS(here("analysis", "data", "derived_data", "genoprobs.rds"))
-expr_sm <- expr[ , colnames(expr) %in% tr_tib$annot.id, drop = FALSE]
-s1 <- qtl2::scan1(genoprobs = aprobs,
-            pheno = expr_sm,
-            kinship = K,
-            addcovar = covar, 
-            cores = 0
-)
-# zoom in on Chr2 after 150 Mb
-# get length of map$`1`
-purrr::map(.x = 1:8, .f = function(x) qtl2::plot_scan1(s1, map, lodcolumn = x, chr = 2, main = paste0(tr_tib[x, ])))
-qtl2::find_peaks(scan1_output = s1, map = map, threshold = 7) %>% 
-  as_tibble() %>%
-  filter(chr == 2) %>%
-  left_join(hot_annot, by = c("lodcolumn" = "annot.id"))
-```
+### Chr 2 index v position plot
 
@@
 
-The Chr 5 scatter plot of mediation LOD difference values against pleiotropy test statistics 
+### Chr 5 scatter plots
+
+### Chr 5 heatmap
+
+The Chr 5 heatmap of pleiotropy test statistics 
 reveals a block-diagonal pattern of blue coloring among the nonlocal traits. We observe the 
@@
 
+### Chr 5 index v position plot
+
 ```{r setup_continued-7}
@@
 ```
+
+## Chr 7 results
+
+### Chr 7 scatter plots
+
+### Chr 7 heatmap
+
+### Chr 7 index v position plot
 
@@
 
+## Chr 11 results
+
+### Chr 11 scatter plots
+
+### Chr 11 heatmap
+
+### Chr 11 index v position plot
+
 ```{r setup_continued-13}
@@
 ```
+
+## Chr 13 results
+
+### Chr 13 scatter plots
+
+### Chr 13 heatmap
+
+### Chr 13 index v position plot
 

### 2019-08-01 analysis/paper/hotspots-paper.Rmd
added code_folding to Rmd

@@
 output: 
-  - github_document:
-      html_preview: false
-      toc: true
-      toc_depth: 3
-  - bookdown::html_document2  
+  bookdown::html_document2:
+  code_folding: hide  
 ---
@@
 
-Undetermined: Chr 7, Chr 11
+Undetermined (right now): Chr 7, Chr 11
 

### 2019-08-02 analysis/paper/hotspot-one-chromosome.Rmd
fixed definition of local traits for a hotspot
started using knitr::knit_expand instead of child = ...

@@
-# Chr `r hot_chr`
+# Chr {{hot_chr}}
 
-```{r}
+```{r, def-{{hot_chr}}}
 hot_lower <- hot_mid - 2
@@
 
-```{r}
+```{r, readRDS-{{hot_chr}}}
 pleio_mixed_list <- readRDS(here("analysis", "data", "derived_data", paste0("Chr", hot_chr, "_pleio.rds")))
@@
 
-```{r}
+```{r, define_local_annot-{{hot_chr}}}
 local_annot <- lod_peaks %>%
@@
 local_annot2 <- local_annot %>%
-  filter(middle <= hot_upper + 1, middle >= hot_lower - 1) %>% 
+  filter(middle <= hot_upper, middle >= hot_lower) %>% 
   arrange(desc(lod))
@@
 
-```{r}
+```{r, scatter-{{hot_chr}}}
 p <- pleio_mixed_list %>%

### 2019-08-02 analysis/paper/hotspots-paper.Rmd
started using knitr::knit_expand instead of child = ...

@@
 output: 
-  bookdown::html_document2:
+  - bookdown::html_document2:
+    code_folding: hide
+  - bookdown::pdf_document2:
     code_folding: hide  
@@
 library(here)
+```
+
+```{r child="methods.Rmd"}
 ```
@@
 
-```{r child="hotspot-one-chromosome.Rmd"}
+```{r}
+knitr::knit_expand(file = here::here("analysis", "paper", "hotspot-one-chromosome.Rmd"))
 ```

### 2019-08-02 analysis/paper/hotspot-one-chromosome.Rmd
fixed Rmd files to work with knit_expand

@@
 ```{r, def-{{hot_chr}}}
-hot_lower <- hot_mid - 2
-hot_upper <- hot_mid + 2
+hot_lower <- {{hot_mid}} - 2
+hot_upper <- {{hot_mid}} + 2
 fp <- paste0("Chr", hot_chr, "-")
@@
 
-```{r}
+```{r, define_nonlocal_annot-{{hot_chr}}}
 nonlocal_annot <- lod_peaks %>%
@@
 
-```{r}
+```{r, local_annot2-{{hot_chr}}}
 local_annot2 <- local_annot %>%
@@
 
-```{r}
+```{r hot_expr-{{hot_chr}}}
 hot_local <- expr[, colnames(expr) %in% local_annot2$annot.id]
@@
 
-```{r, message=FALSE, warning = FALSE}
+```{r, med_tib-{{hot_chr}}, message=FALSE, warning = FALSE}
 med_tib <- med_list %>%
@@
 
-```{r}
+```{r, hot_annot-{{hot_chr}}}
 hot_annot <- local_annot2 %>%
@@
 
-```{r}
+```{r pm-{{hot_chr}}}
 pm <- pleio_tib %>%
@@
   dplyr::select(local_symbol, nonlocal_symbol, pleiotropy_lod, lod_diff, lod_diff_proportion, lod_without_med, lod_with_med, filename, nonlocal_gene_id, local_gene_id) %>%
-  ggplot() + geom_point(mapping = aes(x = pleiotropy_lod, y = lod_diff, colour = local_symbol == keller_mediator, nonlocal = nonlocal_symbol, local = local_symbol), size = 0.05) + theme(legend.position = "none") + xlab("Pleiotropy LOD") + ylab("Mediation LOD difference")
+  ggplot() + geom_point(mapping = aes(x = pleiotropy_lod, y = lod_diff, colour = local_symbol == {{keller_mediator}}, nonlocal = nonlocal_symbol, local = local_symbol), size = 0.05) + theme(legend.position = "none") + xlab("Pleiotropy LOD") + ylab("Mediation LOD difference")
 fnsvg <- paste0("Chr", hot_chr, "_scatter.svg")
@@
 
-```{r, fig.height = ceiling(nlocal / 2)}
+```{r, scatter_by_local-{{hot_chr}}, fig.height = ceiling(nlocal / 2)}
 foo <- pleio_mixed_list %>%
@@
 
-```{r}
+```{r, scols-{{hot_chr}}}
 scols1 <- 1L + (rownames(pm) %in% local_annot2$symbol)
-scols2 <- rownames(pm) %in% (med_tib %>% filter(local_symbol == keller_mediator, lod_diff > 1.5))$nonlocal_symbol
+scols2 <- rownames(pm) %in% (med_tib %>% filter(local_symbol == {{keller_mediator}}, lod_diff > 1.5))$nonlocal_symbol
 ```
 
-```{r}
+```{r, ordering-{{hot_chr}}}
 # order the nonlocal traits first
@@
 
-```{r}
+```{r, scols_extra-{{hot_chr}}}
 scols_extra <- purrr::map(.x = local_annot2$symbol,
@@
 
-```{r}
+```{r, n_med-{{hot_chr}}}
 n_med_tr <- 4 # arbitrary threshold for number of mediated traits per local
@@
 
-```{r}
+```{r, heat-{{hot_chr}}}
 ord <- match(rn, rownames(pm))
@@
 
-```{r eval = FALSE}
+```{r, webshot-{{hot_chr}}, eval = FALSE}
 webshot::webshot(url = fnhtml,
@@
 
-```{r}
+```{r, index_v_position-{{hot_chr}}}
 bar <- foo %>%
-  filter(local_symbol == keller_mediator)
+  filter(local_symbol == {{keller_mediator}})
 

### 2019-08-02 analysis/paper/hotspots-paper.Rmd
fixed Rmd files to work with knit_expand

@@
 keller_mediator <- "Hnf4a"
+out <- knitr::knit_expand(file = here::here("analysis", "paper", "hotspot-one-chromosome.Rmd"))
 ```
 
-```{r}
-knitr::knit_expand(file = here::here("analysis", "paper", "hotspot-one-chromosome.Rmd"))
-```
+`r knit(text = out)`
 
@@
 keller_mediator <- "Pdx1"
+out <- knitr::knit_expand(file = here::here("analysis", "paper", "hotspot-one-chromosome.Rmd"))
 ```
 
-```{r child="hotspot-one-chromosome.Rmd"}
-```
+`r knit(text = out)`
 
@@
 keller_mediator <- "Fam83e"
+out <- knitr::knit_expand(file = here::here("analysis", "paper", "hotspot-one-chromosome.Rmd"))
 ```
 
-```{r child="hotspot-one-chromosome.Rmd"}
-```
+`r knit(text = out)`
 
@@
 keller_mediator <- "Sat2"
+out <- knitr::knit_expand(file = here::here("analysis", "paper", "hotspot-one-chromosome.Rmd"))
 ```
 
-```{r child="hotspot-one-chromosome.Rmd"}
-```
+`r knit(text = out)`
 
@@
 keller_mediator <- "Il6st"
+out <- knitr::knit_expand(file = here::here("analysis", "paper", "hotspot-one-chromosome.Rmd"))
 ```
 
-```{r child="hotspot-one-chromosome.Rmd"}
-```
+`r knit(text = out)`
 

### 2019-08-02 analysis/paper/hotspot-one-chromosome.Rmd
replaced hot_chr with {{hot_chr}}

@@
 hot_upper <- {{hot_mid}} + 2
-fp <- paste0("Chr", hot_chr, "-")
+fp <- paste0("Chr", {{hot_chr}}, "-")
 knitr::opts_chunk$set(fig.path = here("analysis", "figures", fp))
@@
 ```{r, readRDS-{{hot_chr}}}
-pleio_mixed_list <- readRDS(here("analysis", "data", "derived_data", paste0("Chr", hot_chr, "_pleio.rds")))
-pleio_local_list <- readRDS(here("analysis", "data", "derived_data", paste0("Chr", hot_chr, "_pleio_local.rds")))
-pleio_nonlocal_list <- readRDS(here("analysis", "data", "derived_data", paste0("Chr", hot_chr, "_pleio_nonlocal.rds")))
-med_list <- readRDS(here("analysis", "data", "derived_data", paste0("Chr", hot_chr, "_mediations.rds")))
+pleio_mixed_list <- readRDS(here("analysis", "data", "derived_data", paste0("Chr", {{hot_chr}}, "_pleio.rds")))
+pleio_local_list <- readRDS(here("analysis", "data", "derived_data", paste0("Chr", {{hot_chr}}, "_pleio_local.rds")))
+pleio_nonlocal_list <- readRDS(here("analysis", "data", "derived_data", paste0("Chr", {{hot_chr}}, "_pleio_nonlocal.rds")))
+med_list <- readRDS(here("analysis", "data", "derived_data", paste0("Chr", {{hot_chr}}, "_mediations.rds")))
 ```
@@
 local_annot <- lod_peaks %>%
-  filter(chrom == hot_chr, pos <= hot_upper, pos >= hot_lower) %>%
+  filter(chrom == {{hot_chr}}, pos <= hot_upper, pos >= hot_lower) %>%
   left_join(y = annots, by = c("annot.id" = "gene_id")) %>%
@@
   filter(lod >= ult) %>%
-  mutate(local_tx = chr == hot_chr) %>%
+  mutate(local_tx = chr == {{hot_chr}}) %>%
   filter(local_tx)
@@
 nonlocal_annot <- lod_peaks %>%
-  filter(chrom == hot_chr, pos <= hot_upper, pos >= hot_lower) %>%
+  filter(chrom == {{hot_chr}}, pos <= hot_upper, pos >= hot_lower) %>%
   inner_join(y = annots, by = c("annot.id" = "gene_id")) %>%
@@
   filter(lod >= 7.18) %>%
-  mutate(local_tx = chr == hot_chr) %>%
+  mutate(local_tx = chr == {{hot_chr}}) %>%
   filter(!local_tx) %>%
@@
   ggplot() + geom_point(mapping = aes(x = pleiotropy_lod, y = lod_diff, colour = local_symbol == {{keller_mediator}}, nonlocal = nonlocal_symbol, local = local_symbol), size = 0.05) + theme(legend.position = "none") + xlab("Pleiotropy LOD") + ylab("Mediation LOD difference")
-fnsvg <- paste0("Chr", hot_chr, "_scatter.svg")
-fnpdf <- paste0("Chr", hot_chr, "_scatter.pdf")
+fnsvg <- paste0("Chr", {{hot_chr}}, "_scatter.svg")
+fnpdf <- paste0("Chr", {{hot_chr}}, "_scatter.pdf")
 ggsave(plot = p, filename = here("analysis", "figures", fnsvg))
@@
 
-fnsvg <- paste0("Chr", hot_chr, "_scatter-panel.svg")
-fnpdf <- paste0("Chr", hot_chr, "_scatter-panel.pdf")
+fnsvg <- paste0("Chr", {{hot_chr}}, "_scatter-panel.svg")
+fnpdf <- paste0("Chr", {{hot_chr}}, "_scatter-panel.pdf")
 ggsave(p, filename = here("analysis", "figures", fnsvg), height = ceiling(nlocal / 2))
@@
       geom_text(size = 2, aes(x = max_pleio_lod / 2, y = 2 * max_med_lod_difference / 3, label = local_symbol), inherit.aes=FALSE)
-  fnsvg <- paste0("Chr", hot_chr, "_scatter-panel", i, ".svg")
-  fnpdf <- paste0("Chr", hot_chr, "_scatter-panel", i, ".pdf")
+  fnsvg <- paste0("Chr", {{hot_chr}}, "_scatter-panel", i, ".svg")
+  fnpdf <- paste0("Chr", {{hot_chr}}, "_scatter-panel", i, ".pdf")
   ggsave(filename = here("analysis", "figures", fnsvg), height = ceiling(n_panels / 2))
@@
   mutate(local = scols1)
-fnhtml <- here("analysis", "figures", paste0("Chr", hot_chr, "hm.html"))
+fnhtml <- here("analysis", "figures", paste0("Chr", {{hot_chr}}, "hm.html"))
 ```
@@
 )
-fnpdf <- here("analysis", "figures", paste0("Chr", hot_chr, "hm.pdf"))
+fnpdf <- here("analysis", "figures", paste0("Chr", {{hot_chr}}, "hm.pdf"))
 htmlwidgets::saveWidget(widget = p, 

### 2019-08-02 analysis/paper/hotspots-paper.Rmd
replaced hot_chr with {{hot_chr}}

@@
 output: 
-  - bookdown::html_document2:
-    code_folding: hide
-  - bookdown::pdf_document2:
+  bookdown::html_document2:
     code_folding: hide
@@
 ```{r setup}
+library(knitr)
 library(qtl2hotspots)
@@
 
-```{r setup_continued-2}
-hot_chr <- 2
-hot_mid <- 165.5
-keller_mediator <- "Hnf4a"
-out <- knitr::knit_expand(file = here::here("analysis", "paper", "hotspot-one-chromosome.Rmd"))
+```{r, load_purrr}
+library(purrr)
 ```
 
-`r knit(text = out)`
+```{r setup_continued}
+hot_chr <- c(2, 5, 7, 11, 13)
+hot_mid <- c(165.5, 146, 46, 71, 112.5)
+keller_mediator <- c("Hnf4a", "Pdx1", "Fam83e", "Sat2", "Il6st")
+out <- pmap(tibble(hot_chr, hot_mid, keller_mediator), function(hot_chr, hot_mid, keller_mediator) knitr::knit_expand(text = readLines(here::here("analysis", "paper", "hotspot-one-chromosome.Rmd"))))
+```
 
 ## Chr 2 hotspot results
+
+`r knitr::knit(out[[1]])`
 
@@
 keller_mediator <- "Pdx1"
-out <- knitr::knit_expand(file = here::here("analysis", "paper", "hotspot-one-chromosome.Rmd"))
+out <- knitr::knit_expand(file = "hotspot-one-chromosome.Rmd")
 ```
-
-`r knit(text = out)`
 

### 2019-08-02 analysis/paper/hotspots-paper.Rmd
trying to fix call to pmap

--- 2019-08-02 analysis/paper/hotspots-paper.Rmd
@@
 author: "Frederick J. Boehm, Brian S. Yandell, and Karl W. Broman"
-date: "7/27/2019"
+date: "August 1, 2019"
 bibliography: research.bib
@@
 
-```{r setup_continued-2}
-hot_chr <- 2
-hot_mid <- 165.5
-keller_mediator <- "Hnf4a"
+```{r, load_purrr}
+library(purrr)
 ```
 
-```{r}
-knitr::knit_expand(file = here::here("analysis", "paper", "hotspot-one-chromosome.Rmd"))
+```{r setup_continued}
+hot_chr <- c(2, 5, 7, 11, 13)
+hot_mid <- c(165.5, 146, 46, 71, 112.5)
+keller_mediator <- c("Hnf4a", "Pdx1", "Fam83e", "Sat2", "Il6st")
+out <- pmap(tibble(hot_chr, hot_mid, keller_mediator), .f = function(hot_chr, hot_mid, keller_mediator) knitr::knit_expand(text = readLines(here::here("analysis", "paper", "hotspot-one-chromosome.Rmd"))))
+# note that ... argument must be a list, ie, it must be a single argument, not multiple args
 ```
@@
 ## Chr 2 hotspot results
+
+`r knitr::knit(text = out[[1]])`
 
@@
 
-```{r setup_continued-5}
-hot_chr <- 5
-hot_mid <- 146
-keller_mediator <- "Pdx1"
-```
-
-```{r child="hotspot-one-chromosome.Rmd"}
-```
-
 ## Chr 5 results
@@
 
-```{r setup_continued-7}
-hot_chr <- 7
-hot_mid <- 46
-keller_mediator <- "Fam83e"
-```
-
-```{r child="hotspot-one-chromosome.Rmd"}
-```
-
 ## Chr 7 results
@@
 
-```{r setup_continued-11}
-hot_chr <- 11
-hot_mid <- 71
-keller_mediator <- "Sat2"
-```
-
-```{r child="hotspot-one-chromosome.Rmd"}
-```
-
 ## Chr 11 results
@@
 ### Chr 11 index v position plot
-
-```{r setup_continued-13}
-hot_chr <- 13
-hot_mid <- 112.5
-keller_mediator <- "Il6st"
-```
-
-```{r child="hotspot-one-chromosome.Rmd"}
-```
 

### 2019-08-02 analysis/paper/hotspots-paper.Rmd
added quotes around {{keller_mediator}} to fix compilation

@@
 keller_mediator <- c("Hnf4a", "Pdx1", "Fam83e", "Sat2", "Il6st")
-out <- pmap(tibble(hot_chr, hot_mid, keller_mediator), .f = function(hot_chr, hot_mid, keller_mediator) knitr::knit_expand(text = readLines(here::here("analysis", "paper", "hotspot-one-chromosome.Rmd"))))
+inputs <- tibble(hot_chr, hot_mid, keller_mediator)
+out <- pmap_dfc(inputs, .f = function(hot_chr, hot_mid, keller_mediator) knitr::knit_expand(text = readLines(here::here("analysis", "paper", "hotspot-one-chromosome.Rmd"))))
 # note that ... argument must be a list, ie, it must be a single argument, not multiple args

### 2019-08-02 analysis/paper/methods.Rmd
added results

@@
 
-$$
+\begin{equation}
 vec(Y) = X vec(B) + vec(G) + vec(E)
-$${#eq:mvlmm}
+(\#eq:mvlmm)
+\end{equation}
 
@@
 two-dimensional, two-QTL scan over a genomic region that included the entirety of the
-hotspot. We fitted bivariate linear mixed effects models (Equations [-@eq:mvlmm;-@eq:g;-@eq:e]) at each ordered pair of markers in a two-dimensional grid [@boehm2019testing].
+hotspot. We fitted bivariate linear mixed effects models (Equations \@ref(), \@ref(), and \@ref()) at each ordered pair of markers in a two-dimensional grid [@boehm2019testing].
 
-$$
+\begin{equation}
 G \sim MN_{n x 2}(0, K, V_g)
-$${#eq:g}
+(\#eq:g)
+\end{equation}
 
-$$
+\begin{equation}
 E \sim MN_{n x 2}(0, I_n, V_e)
-$${#eq:e}
+(\#eq:e)
+\end{equation}
 
@@
 
-$$
+\begin{equation}
 \hat B = (X^T\hat \Sigma^{-1}X)^{-1}X^T\hat\Sigma^{-1}Y
-$$
+(\#eq:Bhat)
+\end{equation}
 
-$$
+\begin{equation}
 \hat \Sigma = \hat V_g \otimes \hat K + \hat V_e \otimes I_n
-$$
+(\#eq:Sigmahat)
+\end{equation}
 
-$$
+\begin{equation}
 l_{10}(\hat B, \hat \Sigma) = - \frac{1}{2}\log_{10}\left((Y - X\hat B)^T\hat \Sigma^{-1}(Y - X\hat B)\right) - n\log_{10}(2\pi) - \log_{10}|\hat \Sigma|
-$$
+(\#eq:loglik)
+\end{equation}
 
@@
 
-$$
+\begin{equation}
 LOD = \log_{10} \left(\frac{\max_{\lambda_1, \lambda_2}L(B, \Sigma, \lambda_1, \lambda_2)}{\max_{\lambda}L(B, \Sigma, \lambda, \lambda)} \right)
-$${#eq:lod}
+\end{equation}{#eq:lod}
 
-$$
+\begin{equation}
 ll(\lambda_1, \lambda_2)
-$$
+\end{equation}
 
@@
 mediation analysis to assess the extent which the local trait mediates the QTL effect on the nonlocal trait.
-Each mediation analysis involved fitting four linear models (Equations~\ref{model1}, \ref{model2}, \ref{model3}, \ref{model4}) [@chick2016defining].
+Each mediation analysis involved fitting four linear models (Equations \@ref(eq:model1), \@ref(eq:model2), \@ref(eq:model3), \@ref(eq:model4)) [@chick2016defining].
 
-$$
+\begin{equation}
 Y = \alpha 1 + WC + E
-$${#eq:model1}
+(\#eq:model1)
+\end{equation}
 
-$$
+\begin{equation}
 Y = XB + WC + E
-$${#eq:model2}
+(\#eq:model2)
+\end{equation}
 
-$$
+\begin{equation}
 Y = \alpha 1 + WC + M\beta + E
-$${#eq:model3}
+(\#eq:model3)
+\end{equation}
 
-$$
+\begin{equation}
 Y = XB + WC + M\beta + E
-$${#eq:model4}
+(\#eq:model4)
+\end{equation}
 

### 2019-08-02 analysis/paper/results.Rmd
added results

## Results

We studied five expression trait hotspots in pancreatic islets [@keller2018genetic]. Scatter plots of 
pleiotropy test statistic against mediation LOD difference for local - nonlocal trait pairs reveal 
some similarities for the five hotspots. All five scatter plots have few, if any, points in the upper 
right quadrant. This tells us that local-nonlocal trait pairs may have either high pleiotropy test 
statistics or high mediation LOD differences, but not both.

Examinating the ranges of pleiotropy test statistics within a hotspot also informs our conclusions. 
Some hotspots have maximum pleiotropy test statistics under 6 (Figure \@ref(fig:scatter-7)), while other hotspots have local-nonlocal pairs with pleiotropy test statistics over 15 (Figure \@ref(fig:scatter-2)).

### 2019-08-02 analysis/paper/results.Rmd
created discussion.Rmd

@@
 
-We studied five expression trait hotspots in pancreatic islets [@keller2018genetic]. Scatter plots of 
+We studied five expression trait hotspots in pancreatic islets from 378 Diversity Outbred mice [@keller2018genetic]. Scatter plots of 
 pleiotropy test statistic against mediation LOD difference for local - nonlocal trait pairs reveal 
@@
 Examinating the ranges of pleiotropy test statistics within a hotspot also informs our conclusions. 
-Some hotspots have maximum pleiotropy test statistics under 6 (Figure \@ref(fig:scatter-7)), while other hotspots have local-nonlocal pairs with pleiotropy test statistics over 15 (Figure \@ref(fig:scatter-2)).
+Some hotspots have maximum pleiotropy test statistics under 6 (Figure \@ref(fig:scatter-7)), while 
+other hotspots have local-nonlocal pairs with pleiotropy test statistics over 15 (Figure 
+\@ref(fig:scatter-2)).

### 2019-08-02 analysis/paper/discussion.Rmd
started Discussion section

## Discussion

We sought to answer the question of whether a single local trait is primarily responsible for each of 
five hotspots in pancreatic islets from Diversity Outbred mice. Our primary tools for probing this 
question are linear regression-based mediation analysis and pleiotropy testing. 
We

### 2019-08-02 analysis/paper/results.Rmd
started Discussion section

@@
 \@ref(fig:scatter-2)).
+
+Some hotspot scatter plots show points with both high pleiotropy test statistics and high mediation 
+LOD differences. Closer investigation of these points reveals that, for a given hotspot, they 
+include a single local trait. For example, multiple pairs with local trait *Tmem130* have high 
+mediation LOD differences and high pleiotropy test statistics. This may indicate that our QTL scan 
+region was too narrow to encompass the entirety of the *Tmem130* peak.

### 2019-08-02 analysis/paper/discussion.Rmd
moved discussion text from paper to discussion.Rmd

@@
 We
+
+Pleiotropy tests complement mediation analyses in the following manner. When considering a
+collection of potential mediators for a nonlocal trait, mediation analyses clearly identify the
+mediator when there is, in fact, a mediator among the candidates. However, when there is no
+mediator, mediation analyses provide little information about the genetics of the complex trait.
+Pleiotropy testing is particularly useful in this setting, since a pleiotropy test always gives
+an answer about the number of underlying QTL.
+A drawback of pleiotropy testing is that the scientific question that it addresses - do two traits share a single QTL - informs, but doesn't fully resolve, the genetics of the nonlocal complex traits under study.
+
+- lod difference threshold for declaration of "mediation". 
+  - esp with regards to Chr 5 hotspot, where the keller mediator has a lot of traits with lod differences under 3, but above the 1.5 threshold.
+
+- look at outputs of `lod_int()`, ie, the support intervals for all traits in a hotspot. Could plot interval against cluster order, for example.

### 2019-08-02 analysis/paper/hotspots-paper.Rmd
removed duplicate call of methods.Rmd
added calls to knit
moved discussion to its own Rmd file

@@
 library(here)
-```
-
-```{r child="methods.Rmd"}
 ```
@@
 
+`r knitr::knit(text = out[[2]])`
+
 ### Chr 5 heatmap
@@
 
+`r knitr::knit(text = out[[3]])`
+
 ### Chr 7 heatmap
@@
 
+`r knitr::knit(text = out[[4]])`
+
 ### Chr 11 heatmap
@@
 ### Chr 13 scatter plots
+
+`r knitr::knit(text = out[[5]])`
 
@@
 
-# Discussion
-
-Pleiotropy tests complement mediation analyses in the following manner. When considering a
-collection of potential mediators for a nonlocal trait, mediation analyses clearly identify the
-mediator when there is, in fact, a mediator among the candidates. However, when there is no
-mediator, mediation analyses provide little information about the genetics of the complex trait.
-Pleiotropy testing is particularly useful in this setting, since a pleiotropy test always gives
-an answer about the number of underlying QTL.
-A drawback of pleiotropy testing is that the scientific question that it addresses - do two traits share a single QTL - informs, but doesn't fully resolve, the genetics of the nonlocal complex traits under study.
-
-- lod difference threshold for declaration of "mediation". 
-  - esp with regards to Chr 5 hotspot, where the keller mediator has a lot of traits with lod differences under 3, but above the 1.5 threshold.
-
-- look at outputs of `lod_int()`, ie, the support intervals for all traits in a hotspot. Could plot interval against cluster order, for example.
+```{r read_discussion, child = "discussion.Rmd"}
+```
 

### 2019-08-02 analysis/paper/results.Rmd
added text about heatmaps.

@@
 region was too narrow to encompass the entirety of the *Tmem130* peak.
+
+The heatmaps reveal one of two patterns among the nonlocal traits. The first possible pattern is a 
+single, mostly blue, square along the diagonal that includes most of the nonlocal traits. The second 
+possible pattern involves more than one dark blue square. Chr 5 hotspot is an example of this second 
+pattern. Its block-diagonal structure contains off-diagonal blocks that are a mixture of colors.

### 2019-08-02 analysis/paper/hotspots-paper.Rmd
updated intro

@@
     code_folding: hide
+  bookdown::pdf_document2:
+    code_folding: none
 ---

### 2019-08-02 analysis/paper/intro.Rmd
updated intro
adding results text
carriage return added
started Discussion section

@@
 Genetics studies in model organisms like mice can identify
-genomic regions that affect complex quantitative traits, such as systolic blood pressure, body weight, and, more recently, biomolecular traits such as gene expression levels and protein concentrations [@sax1923association;@soller1976power;@lander1989mapping;@broman2009guide;@jansen2007quantitative;@chick2016defining].
+genomic regions that affect complex quantitative traits, such as systolic blood pressure, body 
+weight, and, more recently, biomolecular traits such as gene expression levels and protein 
+concentrations [@sax1923association;@soller1976power;@lander1989mapping;@broman2009guide;@jansen2007quantitative;@chick2016defining]. 
 These genomic regions are called "quantitative trait loci" or "QTL".
@@
 and may contain more than a hundred genes.
-Identification of a causal gene (for a given complex trait) from among those genes near the QTL is
+Identification of a causal gene (for a given complex trait) from 
+among those genes near the QTL is
 challenging and may require costly and time-consuming experiments.
@@
 Recent systems genetics studies have sought to identify QTL for thousands of biomolecular traits [@keller2018genetic;@chick2016defining].
+
+@boehm2019thesis argued for complementary roles for mediation analysis and pleiotropy testing in hotspot 
+dissection. Here, we clarify this argument by presenting a study of five expression trait 
+hotspots in a publicly available data set. We consider each hotspot one at a time. At each hotspot, we identify 
+a set of local traits that have QTL within the hotspot. We also identify a set of nonlocal traits with QTL 
+arising from the hotspot. We perform pleiotropy tests and mediation analyses for all local-nonlocal trait pairs. 
+Graphical summaries of our results enable us to conclude that at least one of the five hotspots is due to multiple QTL, 
+while the evidence at the four other hotspots is consistent with a single local trait affecting most or all of the 
+nonlocal associated traits. The Chr 5 hotspot
+
+## INTRO IDEAS
+
+- why is it useful to study expression hotspots? 
+- why are these two methods - pleiotropy tests and mediation analyses useful for studying hotspots? 
+- who else has studied hotspots? And what did they do?

### 2019-08-02 analysis/paper/hotspots-paper.Rmd
moved results section to results.Rmd

--- 2019-08-02 analysis/paper/hotspots-paper.Rmd
@@
     code_folding: hide
+  bookdown::pdf_document2:
+    fig_caption: true
 ---
 
+```{r child = "abstract.Rmd"}
+```
+
 ```{r child="intro.Rmd"}
+```
+
+```{r presetup}
+eval_code <- FALSE
+knitr::opts_chunk$set(comment=NA, fig.width=6, fig.height=6, eval = eval_code)
 ```
@@
 # Results
-
-Simple hotspots: Chr 2, Chr 13
-
-Complex hotspots: Chr 5
-
-Undetermined (right now): Chr 7, Chr 11
-
-We generated scatter plots of pleiotropy test statistics vs. mediation LOD difference values for each of the five hotspots (Figures).
-Each point represents a pairing of a local trait with a nonlocal trait.
-The mediation LOD difference value is that observed when the local trait is considered
-as a candidate mediator of the QTL-nonlocal trait association.
-
-In all five scatter plots, we observe few
-points in the upper right quadrant of the figure.
-That is, few pairs have high pleiotropy test statistics and
-high mediation LOD difference values.
 
@@
 
-## Chr 2 hotspot results
+### Chr 2 hotspot results
 
-`r knitr::knit(text = out[[1]])`
+`r if (eval_code) knitr::knit(text = out[[1]])`
 
-### Chr 2 scatter plots
+### Chr 5 hotspot results
 
-The scatter plot of pleiotropy test statistics and mediation LOD differences reveals that few local-nonlocal 
-trait pairs have both a high value of the pleiotropy test statistic and a high value of the mediation LOD 
-difference. Many points have low values of both, while some points demonstrate a high value of only mediation 
-LOD difference or pleiotropy test statistic.
-
-Many points have negative values of mediation LOD difference. These correspond to pairs that, when conditioning 
-on the putative mediator, the LOD for the nonlocal trait increased.
-
-The points are color-coded to identify those pairs that involve the local trait *Hnf4a*, which 
-@keller2018genetic identified as a key mediator of traits at this hotspot. Blue points involve *Hnf4a*, while 
-red points do not. We see that many blue point have large mediation LOD differences and small pleiotropy test 
-statistics. In fact, most of the points with mediation LOD difference above 15 involve *Hnf4a* and have small 
-pleiotropy test statistics.
-
-An unanticipated observation is that some points have mediation LOD differences of
-
-### Chr 2 heatmap
-
-### Chr 2 index v position plot
-
-## Chr 5 results
-
-### Chr 5 scatter plots
-
-`r knitr::knit(text = out[[2]])`
-
-### Chr 5 heatmap
-
-The Chr 5 heatmap of pleiotropy test statistics 
-reveals a block-diagonal pattern of blue coloring among the nonlocal traits. We observe the 
-set of nonlocal traits clustered into two blocks. Within each block, any two traits share a 
-QTL, ie, are pleiotropic. Yet, the off-diagonal cells, ie, those nonlocal trait pairs that 
-include one trait from each of the two blocks, tend to have greater pleiotropy test 
-statistics. Still, not all off-diagonal blocks have large pleiotropy test statistics. The 
-mostly dark blue matrix on the diagonal (from column 22 to 99) features no cells with pleiotropy test statistics above 3.2.
-While we didn't determine p-values for these analyses, prior analyses suggest that pleiotropy test statistics below 3.2 are consistent with p-values above 0.05, and, thus, are consistent with a single pleiotropic QTL.
-
-Columns 100 to 119 (and rows 100 to 119) are a second (symmetric) block with only small 
-pleiotropy test statistics.
-Columns 120 to 202 contains cells for the third (symmetric) block.
-Column 203, the last column, contains a trait *Gm* that, based on pleiotropy test statistics, doesn't cluster near other traits.
+`r if (eval_code) knitr::knit(text = out[[2]])`
 
@@
 
-## Chr 7 results
+### Chr 7 results
 
-### Chr 7 scatter plots
+`r if (eval_code) knitr::knit(text = out[[3]])`
 
-`r knitr::knit(text = out[[3]])`
+### Chr 11 results
 
-### Chr 7 heatmap
+`r if (eval_code) knitr::knit(text = out[[4]])`
 
-### Chr 7 index v position plot
+### Chr 13 results
 
-## Chr 11 results
+`r if (eval_code) knitr::knit(text = out[[5]])`
 
-### Chr 11 scatter plots
-
-`r knitr::knit(text = out[[4]])`
-
-### Chr 11 heatmap
-
-### Chr 11 index v position plot
-
-## Chr 13 results
-
-### Chr 13 scatter plots
-
-`r knitr::knit(text = out[[5]])`
-
-### Chr 13 heatmap
-
-### Chr 13 index v position plot
-
-Our analysis of the Chromosome 13 hotspot presents results much like those for the Chromosome 2 
-hotspot, with the Chromosome 13 local transcript *Il6st* playing a role like that of *Hnf4a* in the 
-Chr 2 hotspot. In examining the scatter plot (Figure ) of mediation LOD differences against 
-pleiotropy test statistics for all local-nonlocal pairs, we see, as in Figure for the Chr 2 
-hotspot, a single local trait *Il6st* has strong mediation effects on a collection of nonlocal traits.
-*Il6st*, first identified by @keller2018genetic, accounts for most of the nonlocal traits that are affected by the hotspot.
-
-The heatmap for pleiotropy statistics for pairwise tests involving nonlocal traits resembles the 
-heatmap for the Chr2 hotspot. Nearly all nonlocal traits have small test statistics when paired 
-with each other. This is visualized as a dark blue, symmetric square along the heatmap's diagonal 
-with only a few outlier nonlocal traits that demonstrate distinct patterns in pleiotropy test 
-statistics.
+```{r read_discussion, child = "results.Rmd"}
+```
 

### 2019-08-02 analysis/paper/results.Rmd
moved results section to results.Rmd

@@
 statistics or high mediation LOD differences, but not both.
+
+Complex hotspots: Chr 5; Simple hotspots: Chr 2, Chr 13
+
+Undetermined (right now): Chr 7, Chr 11
+
+We generated scatter plots of pleiotropy test statistics vs. mediation LOD difference values for each of the five hotspots (Figures).
+Each point represents a pairing of a local trait with a nonlocal trait.
+The mediation LOD difference value is that observed when the local trait is considered
+as a candidate mediator of the QTL-nonlocal trait association.
+
+In all five scatter plots, we observe few
+points in the upper right quadrant of the figure.
+That is, few pairs have high pleiotropy test statistics and
+high mediation LOD difference values.
 
@@
 pattern. Its block-diagonal structure contains off-diagonal blocks that are a mixture of colors.
+
+The scatter plot of pleiotropy test statistics and mediation LOD differences reveals that few local-nonlocal 
+trait pairs have both a high value of the pleiotropy test statistic and a high value of the mediation LOD 
+difference. Many points have low values of both, while some points demonstrate a high value of only mediation 
+LOD difference or pleiotropy test statistic.
+
+Many points have negative values of mediation LOD difference. These correspond to pairs that, when conditioning 
+on the putative mediator, the LOD for the nonlocal trait increased.
+
+The points are color-coded to identify those pairs that involve the local trait *Hnf4a*, which 
+@keller2018genetic identified as a key mediator of traits at this hotspot. Blue points involve *Hnf4a*, while 
+red points do not. We see that many blue point have large mediation LOD differences and small pleiotropy test 
+statistics. In fact, most of the points with mediation LOD difference above 15 involve *Hnf4a* and have small 
+pleiotropy test statistics.
+
+An unanticipated observation is that some points have mediation LOD differences of
+
+The Chr 5 heatmap of pleiotropy test statistics 
+reveals a block-diagonal pattern of blue coloring among the nonlocal traits. We observe the 
+set of nonlocal traits clustered into two blocks. Within each block, any two traits share a 
+QTL, ie, are pleiotropic. Yet, the off-diagonal cells, ie, those nonlocal trait pairs that 
+include one trait from each of the two blocks, tend to have greater pleiotropy test 
+statistics. Still, not all off-diagonal blocks have large pleiotropy test statistics. The 
+mostly dark blue matrix on the diagonal (from column 22 to 99) features no cells with pleiotropy test statistics above 3.2.
+While we didn't determine p-values for these analyses, prior analyses suggest that pleiotropy test statistics below 3.2 are consistent with p-values above 0.05, and, thus, are consistent with a single pleiotropic QTL.
+
+Columns 100 to 119 (and rows 100 to 119) are a second (symmetric) block with only small 
+pleiotropy test statistics.
+Columns 120 to 202 contains cells for the third (symmetric) block.
+Column 203, the last column, contains a trait *Gm* that, based on pleiotropy test statistics, doesn't cluster near other traits.
+
+Our analysis of the Chromosome 13 hotspot presents results much like those for the Chromosome 2 
+hotspot, with the Chromosome 13 local transcript *Il6st* playing a role like that of *Hnf4a* in the 
+Chr 2 hotspot. In examining the scatter plot (Figure ) of mediation LOD differences against 
+pleiotropy test statistics for all local-nonlocal pairs, we see, as in Figure for the Chr 2 
+hotspot, a single local trait *Il6st* has strong mediation effects on a collection of nonlocal traits.
+*Il6st*, first identified by @keller2018genetic, accounts for most of the nonlocal traits that are affected by the hotspot.
+
+The heatmap for pleiotropy statistics for pairwise tests involving nonlocal traits resembles the 
+heatmap for the Chr2 hotspot. Nearly all nonlocal traits have small test statistics when paired 
+with each other. This is visualized as a dark blue, symmetric square along the heatmap's diagonal 
+with only a few outlier nonlocal traits that demonstrate distinct patterns in pleiotropy test 
+statistics.

### 2019-08-02 analysis/paper/hotspots-paper.Rmd
fixed code chunks with child arg

@@
 output: 
+  bookdown::pdf_document2:
+    fig_caption: true
   bookdown::html_document2: 
     code_folding: hide
-  bookdown::pdf_document2:
-    fig_caption: true
 ---
 
-```{r child = "abstract.Rmd"}
+```{r child = "abstract.Rmd", eval = TRUE}
 ```
 
-```{r child="intro.Rmd"}
+```{r child="intro.Rmd", eval = TRUE}
 ```
@@
 
-```{r child="methods.Rmd"}
+```{r child="methods.Rmd", eval = TRUE}
 ```
@@
 
-### Chr 5 index v position plot
-
 ### Chr 7 results
@@
 
-```{r read_discussion, child = "results.Rmd"}
+```{r read_results, child = "results.Rmd", eval = TRUE}
 ```
 
-```{r read_discussion, child = "discussion.Rmd"}
+```{r read_discussion, child = "discussion.Rmd", eval = TRUE}
 ```

### 2019-08-02 analysis/paper/methods.Rmd
added Tian et al 2016 citation

@@
 # Methods
+
+We used the publicly available data from @keller2018genetic to demonstrate our strategy. @keller2018genetic
+measured genome-wide gene expression levels with RNA sequencing of pancreatic islets in 378 Diversity Outbred mice.
+From the RNA sequencing counts, they obtained normalized gene expression levels for 21,771 transcripts, which they shared publicly via DataDryad ().
 
@@
 For each hotspot, we identified a set of local expression traits and a
-set of nonlocal expression traits. All traits demonstrated LOD peaks that met
-the genome-wide significance threshold, 7.18 [@keller2018genetic].
-The nonlocal expression traits arose from genes
+set of nonlocal expression traits. We focused on those local 
+traits that strongly associated with QTL within the hotspot.
+We set the univariate LOD threshold for local traits at 15.
+We limited consideration to those local traits for which the middle of the gene lay within the hotspot interval.
+All nonlocal traits demonstrated univariate LOD peaks that met
+the genome-wide significance threshold, 7.18 [@keller2018genetic]. The nonlocal expression traits arose from genes
 on other chromosomes, while the local traits reflected expression levels for
-genes on the same chromosome as the hotspot and within 4 Mb of the hotspot
-center. We ignored transcripts for genes on the hotspot chromosome but not within 4 Mb of
-the hotspot center.
+genes on the same chromosome as the hotspot . We ignored transcripts for genes on the hotspot chromosome but outside the hotspot interval.
 
@@
 
-A mediation analysis in systems genetics begins with an identified QTL-expression trait association. In our case, we choose a QTL that affects a nonlocal trait.
+A mediation analysis in systems genetics begins with an identified QTL-expression trait 
+association [@chick2016defining]. In our case, we choose a QTL that affects a nonlocal trait, then ask if any of 
+the nearby local traits lie in the causal pathway from QTL to nonlocal trait.
 
-For each hotspot and each local trait - nonlocal trait pair, we also performed a
-mediation analysis to assess the extent which the local trait mediates the QTL effect on the nonlocal trait.
-Each mediation analysis involved fitting four linear models (Equations \@ref(eq:model1), \@ref(eq:model2), \@ref(eq:model3), \@ref(eq:model4)) [@chick2016defining].
+involved fitting four linear models (Equations \@ref(eq:model1), \@ref(eq:model2), \@ref(eq:model3), \@ref(eq:model4)) [@chick2016defining].
 

### 2019-08-03 analysis/paper/methods.Rmd
updated text

@@
 
-From the two-dimensional scans, we calculated pleiotropy likelihood ratio test statistics for each
+From the two-dimensional scans, we calculated pleiotropy likelihood ratio test (LOD) statistics for each
 pair of traits [@jiang1995multiple;@boehm2019testing;@boehm2019thesis]. Because of the required
@@
 LOD = \log_{10} \left(\frac{\max_{\lambda_1, \lambda_2}L(B, \Sigma, \lambda_1, \lambda_2)}{\max_{\lambda}L(B, \Sigma, \lambda, \lambda)} \right)
-\end{equation}{#eq:lod}
-
-\begin{equation}
-ll(\lambda_1, \lambda_2)
+(\#eq:lod)
 \end{equation}
@@
 A mediation analysis in systems genetics begins with an identified QTL-expression trait 
-association [@chick2016defining]. In our case, we choose a QTL that affects a nonlocal trait, then ask if any of 
-the nearby local traits lie in the causal pathway from QTL to nonlocal trait.
+association [@chick2016defining]. In our case, we choose a QTL that affects a nonlocal expression trait, then ask if any of 
+the nearby local expression traits lie in the causal pathway from QTL to nonlocal trait.
 
-involved fitting four linear models (Equations \@ref(eq:model1), \@ref(eq:model2), \@ref(eq:model3), \@ref(eq:model4)) [@chick2016defining].
+A single mediation analysis involving one nonlocal trait and one local trait requires fitting four linear models (Equations \@ref(eq:model1), \@ref(eq:model2), \@ref(eq:model3), \@ref(eq:model4)) [@chick2016defining].
 
@@
 In all of our mediation analyses, the nonlocal transcripts are the targets, while the local
-transcripts serve as candidate mediators of the QTL-target relationship. We only considered as potential mediators those local traits with strong univariate LODs.
+transcripts serve as candidate mediators of the QTL-target relationship. 
+We only considered as potential mediators those local traits with strong univariate LODs.
 In Equations~\ref{model1} through \ref{model4}, $X$ is a matrix of founder allele dosages,
@@
 of founder allele effects, $W$ is a matrix of covariates, $C$ is a matrix of covariate
-effects, $M$ is a vector of values for a single candidate mediator, $\beta$ is the effect
-of the mediator on the complex trait, 1 is a vector with all entries set to 1, $\alpha$
-is the mean trait value, and $E$ is a vector of normally distributed random errors.
+effects, $M$ is a vector of values for a single candidate mediator (a local expression trait, in our case), $\beta$ is the effect
+of the mediator on the complex trait (after adjusting for other covariates in the model), 1 is a vector with all entries set to 1, $\alpha$
+is a parameter that corresponds to the mean trait value (after adjusting for other covariates in the model), and $E$ is a vector of normally distributed random errors.
+
+We included four covariates in our models: three wave indicators and sex.
+The wave indicators are binary vectors to account for wave membership.
+@keller2018genetic phenotyped the mice in four waves of nearly 100 mice per wave.
 
@@
 We highlighted in blue those points that correspond to pairs involving the putative mediator per [@keller2018genetic].
-
-### Heatmaps
 

### 2019-08-03 analysis/paper/results.Rmd
updated text

@@
-## Results
+# Results
 
@@
 
-Complex hotspots: Chr 5; Simple hotspots: Chr 2, Chr 13
+Complex hotspots: Chr 5; Simple hotspots: Chr 2, Chr 13, Chr 7
 
-Undetermined (right now): Chr 7, Chr 11
+Undetermined (right now): Chr 11
 
@@
 
-Examinating the ranges of pleiotropy test statistics within a hotspot also informs our conclusions. 
+Examining the ranges of pleiotropy test statistics within a hotspot also informs our conclusions. 
 Some hotspots have maximum pleiotropy test statistics under 6 (Figure \@ref(fig:scatter-7)), while 

### 2019-08-03 analysis/paper/Rmd/abstract.Rmd
organized analysis/paper dir

# Abstract

We present our analysis of five expression trait hotspots in pancreatic islets from 378
Diversity Outbred mice.

### 2019-08-03 analysis/paper/Rmd/hotspot-one-chromosome.Rmd
organized analysis/paper dir

--- 2019-08-02 analysis/paper/hotspot-one-chromosome.Rmd
@@
   dplyr::select(local_symbol, nonlocal_symbol, pleiotropy_lod, lod_diff, lod_diff_proportion, lod_without_med, lod_with_med, filename, nonlocal_gene_id, local_gene_id) %>%
-  ggplot() + geom_point(mapping = aes(x = pleiotropy_lod, y = lod_diff, colour = local_symbol == {{keller_mediator}}, nonlocal = nonlocal_symbol, local = local_symbol), size = 0.05) + theme(legend.position = "none") + xlab("Pleiotropy LOD") + ylab("Mediation LOD difference")
+  ggplot() + geom_point(mapping = aes(x = pleiotropy_lod, y = lod_diff, colour = local_symbol == "{{keller_mediator}}", nonlocal = nonlocal_symbol, local = local_symbol), size = 0.05) + theme(legend.position = "none") + xlab("Pleiotropy LOD") + ylab("Mediation LOD difference")
 fnsvg <- paste0("Chr", {{hot_chr}}, "_scatter.svg")
@@
 scols1 <- 1L + (rownames(pm) %in% local_annot2$symbol)
-scols2 <- rownames(pm) %in% (med_tib %>% filter(local_symbol == {{keller_mediator}}, lod_diff > 1.5))$nonlocal_symbol
+scols2 <- rownames(pm) %in% (med_tib %>% filter(local_symbol == "{{keller_mediator}}", lod_diff > 1.5))$nonlocal_symbol
 ```
@@
 bar <- foo %>%
-  filter(local_symbol == {{keller_mediator}})
+  filter(local_symbol == "{{keller_mediator}}")
 

### 2019-08-03 analysis/paper/Rnw/paper-thesis.Rnw
organized analysis/paper dir

--- 2019-06-17 analysis/paper/paper.Rnw
@@
 \usepackage[utf8]{inputenc}
-%\usepackage{lineno}
-%\linenumbers
+\usepackage{lineno}
+\linenumbers
 
@@
 \renewcommand{\headrule}{}
-\title{The title}
-\author{An Author}
-\patchcmd{\titlepage}{empty}{fancy}{}{}
-\patchcmd{\chapter}{plain}{fancy}{}{}
-\DeclareCaptionType{equ}[][]
-%https://stackoverflow.com/questions/149479/adding-a-caption-to-an-equation-in-latex
-%\captionsetup[equ]{labelformat=empty}
 
-\DeclareFloatingEnvironment[fileext=frm,placement={!ht},name=Frame]{myfloat}
-\captionsetup[myfloat]{labelfont=bf}
-\newenvironment{frameenv}[1]
-    {\begin{myfloat}[tb]
-    \begin{mdframed}[roundcorner=10pt,backgroundcolor=blue!10]
-    \caption{#1}
-    %\begin{multicols*}{2}
-    }
-    {%\end{multicols*}
-    \end{mdframed}\end{myfloat}
-    }
-
-\title{}
+\title{Expression QTL hotspot dissection}
 \author{Frederick J. Boehm, Brian S. Yandell, and Karl W. Broman}
@@
 
-%% from vignette tex file
-\urlstyle{same}  % don't use monospace font for urls
-\usepackage{color}
-\usepackage{fancyvrb}
-\newcommand{\VerbBar}{|}
-\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
-\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
-% Add ',fontsize=\small' for more characters per line
-\usepackage{framed}
-\definecolor{shadecolor}{RGB}{248,248,248}
-\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
-\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
-\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
-\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\BuiltInTok}[1]{#1}
-\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
-\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
-\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
-\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
-\newcommand{\ExtensionTok}[1]{#1}
-\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\ImportTok}[1]{#1}
-\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
-\newcommand{\NormalTok}[1]{#1}
-\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
-\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
-\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
-\newcommand{\RegionMarkerTok}[1]{#1}
-\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\usepackage{graphicx,grffile}
-\makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+\begin{document}
 
-%% end from vignette tex file
-
-\begin{document}
+\maketitle
 
@@
 opts_chunk$set(
-concordance=TRUE, echo = FALSE, fig.lp = "fig:", warning = FALSE, message = FALSE, results = "hide"
-)
+concordance=TRUE,
+echo = TRUE,
+fig.lp = "../figures/fig-",
+warning = FALSE,
+message = FALSE)
 @
@@
 \doublespacing
-\begin{titlepage}\thispagestyle{empty}
 \tableofcontents
-%\listoffigures
-%\listoftables
+\listoffigures
+\listoftables
 %\pagestyle{headings} % page numbers appear at bottom of page, centered!
@@
 
-One then fits a linear mixed effects model and calculates the log$_{10}$ likelihood at each position.
-As in two-parent crosses (Section~\ref{sec:qtl-two-parent}), one summarizes evidence for a QTL by plotting LOD scores across the genome.
-Because multiple hypothesis tests are performed, one typically wants to control family-wise error rate.
-One strategy to do this involves a permutation test, much like that in two-parent crosses \citep{churchill1994empirical}.
+One then fits a linear mixed effects model and calculates the log$_{10}$
+likelihood at each position. As in two-parent crosses, one summarizes evidence for a QTL by plotting LOD scores across
+the genome. Because multiple hypothesis tests are performed, one typically
+wants to control family-wise error rate. One strategy to do this involves a
+permutation test, much like that in two-parent crosses \citep{churchill1994empirical}.
 
@@
 relationship. Failure to measure C1, C2, or C3 leads to violations of Assumption 1,
-Assumption 2, or Assumption 3, respectively (Frame~\ref{frame1}).
-
-\begin{figure}
-\includegraphics{figs/confounding.pdf}
-\caption{Example causal diagram showing relationships between random variables in an expression QTL study.}\label{fig:confounding}
-\end{figure}
+Assumption 2, or Assumption 3, respectively.
 
 In studies of Diversity Outbred mice, relatedness is a possible confounder,
-yet the linear models (Equations~\ref{model1},~\ref{model2},~\ref{model3},~\ref{model4}
-in Frame~\ref{frame3}) fail to account for the complex relatedness patterns
+yet the linear models (Equations~\ref{model1},~\ref{model2},~\ref{model3},~\ref{model4}) fail to account for the complex relatedness patterns
 among Diversity Outbred mice.
@@
 other local genes (Table \ref{tab:annot}).
-Our 13 local genes are the only genes that met three criteria (Frame~\ref{frame2}).
+Our 13 local genes are the only genes that met three criteria.
 The 147 nonlocal traits that we studied all had LOD peak heights above 7.18 and QTL positions within 2 Mb of the center of the hotspot (at 165.5 Mb).
 
-\begin{frameenv}{Local gene inclusion criteria}\label{frame2}
 \begin{enumerate}
-    \item QTL peak with LOD $>$ 40
+    \item QTL peak with LOD $>$ 7.18
     \item QTL peak position within 2 Mb of hotspot center (165.5 Mb)

### 2019-08-03 analysis/paper/Rnw/paper.Rnw
organized analysis/paper dir

--- 2019-07-27 analysis/paper/paper.Rnw
@@
 <<panel1, fig.cap = paste0("Chr", hot_chr, "hotspot scatter plots of mediation LOD difference against pleiotropy LOD, separated by local trait.")>>=
-#include_graphics(paste0("../figures/Chr", hot_chr, "_scatter-panel1.pdf"))
+include_graphics(paste0("../figures/Chr", hot_chr, "_scatter-panel1.pdf"))
 @
@@
 <<panel2, eval = nfig[which(names(nfig) == paste0("Chr", hot_chr))] > 1>>=
-#include_graphics(paste0("../figures/Chr", hot_chr, "_scatter-panel2.pdf"))
+include_graphics(paste0("../figures/Chr", hot_chr, "_scatter-panel2.pdf"))
 @
@@
 <<panel3, eval = nfig[which(names(nfig) == paste0("Chr", hot_chr))] > 2>>=
-#include_graphics(paste0("../figures/Chr", hot_chr, "_scatter-panel3.pdf"))
+include_graphics(paste0("../figures/Chr", hot_chr, "_scatter-panel3.pdf"))
 @

### 2019-08-03 analysis/paper/preliminary/heatmaps-dendro-check.Rmd
organized analysis/paper dir

--- 2019-07-16 analysis/paper/Chr11-heat.Rmd
@@
 ---
-title: "Heatmaps of pleiotropy test statistics"
+title: "Checking heatmaps and dendrograms"
 author: "Frederick Boehm"
@@
 library(hotspots)
+hot_chr <- 11
+hot_mid <- 71
+keller_mediator <- "Sat2"
+```
+
+```{r more_setup}
+hot_lower <- hot_mid - 2
+hot_upper <- hot_mid + 2
+fp <- paste0("../figures/Chr", hot_chr, "-")
+knitr::opts_chunk$set(fig.path = fp)
 ```
@@
 Other questions that I want to consider include the *method* for dendrogram creation. I also want to think about coloring schemes in heatmaps.
-
-# Chr 11
-
-```{r setup_continued}
-hot_chr <- 11
-hot_mid <- 71
-```
-
-```{r more_setup}
-hot_lower <- hot_mid - 2
-hot_upper <- hot_mid + 2
-fp <- paste0("../figures/Chr", hot_chr, "-")
-knitr::opts_chunk$set(fig.path = fp)
-```
 
@@
 In the matrix of nonlocal - nonlocal pairs of pleiotropy test statistics, the fact that the maximum is only `r round(max(pm_nonlocal, na.rm = TRUE))` tells us that the heatmap of binary values is not going to be informative, since all or nearly all entries are less than any reasonable threshold.
-
-Regarding possible transformations of the rows of the matrix, I don't really like transformations here.
-Transformations aim to preserve, in some sense, the distribution of values - test statistics, here - and I don't think that we should do that.
-I don't really care if a given row comes from a normal distribution, for example.
-I'm inclined to prefer heatmaps of the raw values.

### 2019-08-03 analysis/paper/preliminary/hotspots-paper-Chr2-only.Rmd
organized analysis/paper dir

--- 2019-07-28 analysis/paper/hotspots-paper-Chr2-only.Rmd
@@
 bibliography: research.bib
+link-citations: true
+numberSections: true
+linkReferences: true
 output: 
@@
 
-# Methods
-
-@keller2018genetic identified five expression trait hotspots in their
-study of pancreatic islet gene expression in 378 Diversity Outbred mice. They
-defined a hotspot to be a 4-Mb region that affects at least 100 nonlocal expression
-traits.
-
-For each hotspot, we identified a set of local expression traits and a
-set of nonlocal expression traits. All traits demonstrated LOD peaks that met
-the genome-wide significance threshold, 7.18 [@keller2018genetic].
-The nonlocal expression traits arose from genes
-on other chromosomes, while the local traits reflected expression levels for
-genes on the same chromosome as the hotspot and within 4 Mb of the hotspot
-center. We ignored transcripts for genes on the hotspot chromosome but not within 4 Mb of
-the hotspot center.
-
-## Pleiotropy testing
-
-### Two-dimensional, two-QTL scans
-
-For each hotspot, we tested all pairs of traits, regardless of whether the traits were local or nonlocal. We performed a
-two-dimensional, two-QTL scan over a genomic region that included the entirety of the
-hotspot. We fitted bivariate linear mixed effects models at each ordered pair of markers in a two-dimensional grid [@boehm2019testing].
-
-$$
-vec(Y) = X vec(B) + vec(G) + vec(E)
-$$
-
-$$
-G \sim MN_{n x 2}(0, K, V_g)
-$$
-
-$$
-E \sim MN_{n x 2}(0, I_n, V_e)
-$$
-
-and $G$ and $E$ are independent.
-
-$$
-\hat B = (X^T\hat \Sigma^{-1}X)^{-1}X^T\hat\Sigma^{-1}Y
-$$
-
-$$
-\hat \Sigma = \hat V_g \otimes \hat K + \hat V_e \otimes I_n
-$$
-
-$$
-l_{10}(\hat B, \hat \Sigma) = - \frac{1}{2}\log_{10}\left((Y - X\hat B)^T\hat \Sigma^{-1}(Y - X\hat B)\right) - n\log_{10}(2\pi) - \log_{10}|\hat \Sigma|
-$$
-
-For $n$ mice, $\Sigma$ is a $2n$ by $2n$ covariance matrix, while $V_g$ and $V_e$ are $2$ by $2$ covariance matrices.
-Estimates $\hat V_g$ and $\hat V_e$ are obtained via an expectation-maximization algorithm [@zhou2014efficient;@gemma2].
-
-### Calculating likelihood ratio test statistics}
-From the two-dimensional scans, we calculated pleiotropy likelihood ratio test statistics for each
-pair of traits [@jiang1995multiple;@boehm2019testing;@boehm2019thesis]. Because of the required
-computing time for bootstrap analyses, we didn't obtain p-values for the pleiotropy tests. Instead,
-we worked directly with the test statistic values.
-
-$$
-LOD = \log_{10} \left(\frac{\max_{\lambda_1, \lambda_2}L(B, \Sigma, \lambda_1, \lambda_2)}{\max_{\lambda}L(B, \Sigma, \lambda, \lambda)} \right)
-$$
-
-$$
-ll(\lambda_1, \lambda_2)
-$$
-
-## Mediation analysis
-
-A mediation analysis in systems genetics begins with an identified QTL-expression trait association. In our case, we choose a QTL that affects a nonlocal trait.
-
-For each hotspot and each local trait - nonlocal trait pair, we also performed a
-mediation analysis to assess the extent which the local trait mediates the QTL effect on the nonlocal trait.
-Each mediation analysis involved fitting four linear models (Equations~\ref{model1}, \ref{model2}, \ref{model3}, \ref{model4}) [@chick2016defining].
-
-$$
-Y = \alpha 1 + WC + E
-\label{model1}
-$$
-
-$$
-Y = XB + WC + E
-\label{model2}
-$$
-
-$$
-Y = \alpha 1 + WC + M\beta + E
-\label{model3}
-$$
-
-$$
-Y = XB + WC + M\beta + E
-\label{model4}
-$$
-
-In all mediation analyses, the nonlocal transcripts are the targets, while the local
-transcripts serve as candidate mediators of the QTL-target relationship.
-In Equations~\ref{model1} through \ref{model4}, $X$ is a matrix of founder allele dosages,
-$B$ is a matrix
-of founder allele effects, $W$ is a matrix of covariates, $C$ is a matrix of covariate
-effects, $M$ is a vector of values for a single candidate mediator, $\beta$ is the effect
-of the mediator on the complex trait, 1 is a vector with all entries set to 1, $\alpha$
-is the mean trait value, and $E$ is a vector of normally distributed random errors.
-
-## Visualizations
-
-We created scatter plots with the `ggplot2` package [@ggplot2] for the R
-statistical computing environment [@r].
-We initially plotted all local-nonlocal pairs for each hotspot.
-We highlighted in blue those points that correspond to pairs involving the putative mediator per [@keller2018genetic].
-
-We created heatmaps to examine patterns among pleiotropy test results in each hotspot.
-We represented the pleiotropy test statistics as a symmetric matrix, where each row was one expression
-trait and each column was an expression trait. Each cell, then, contained the pleiotropy test statistic
-from the analysis involving the row's expression trait and the column's expression trait.
-To create a heatmap, we first performed two hierarchical clusterings, one on the submatrix involving only
-nonlocal expression traits and a second on the submatrix involving only local expression traits. We
-calculated the Euclidean distances between all trait pairs before clustering. We ordered the rows and
-columns according to optimal leaf ordering. We then combined the
-local and nonlocal traits into a single heatmap. To achieve greater coloring contrasts in the heatmaps, we
-set the color spectrum limits at 0 (dark blue) and 5 (yellow). Test statistics above 5 are represented by
-yellow.
+```{r child="methods.Rmd"}
+```
 

### 2019-08-03 analysis/paper/preliminary/outline.Rmd
organized analysis/paper dir

---
title: "Manuscript Outline"
author: "Frederick Boehm"
date: "7/22/2019"
output: html_document
---

## Figures list

1. pleiotropy v lod difference (one plot per hotspot)
1. pleiotropy v lod difference (multipanel figures: one panel per local trait per hotspot)
1. heatmaps of pleiotropy test statistics with mediation results (one plot per hotspot)

## Tables list

1. summary table with one row per hotspot, like Keller's supp table.
1. mediation counts per local trait (one table per hotspot). Each row is a local trait that mediates at least 4 nonlocal traits.

### 2019-08-03 analysis/paper/Rmd/hotspots-paper.Rmd
moved Makefile to Rmd dir

--- 2019-08-02 analysis/paper/hotspots-paper.Rmd
@@
 author: "Frederick J. Boehm, Brian S. Yandell, and Karl W. Broman"
-date: "August 1, 2019"
+date: "August 3, 2019"
 bibliography: research.bib
@@
     fig_caption: true
+    keep_tex: true
   bookdown::html_document2: 
     code_folding: hide
+    self_contained: no
 ---
@@
 ```{r presetup}
-eval_code <- FALSE
-knitr::opts_chunk$set(comment=NA, fig.width=6, fig.height=6, eval = eval_code)
+eval_code <- TRUE
+knitr::opts_chunk$set(comment=NA, fig.width=6, fig.height=6, eval = eval_code, results = "hide")
 ```
@@
 ```
-
-# Results
 

### 2019-08-03 analysis/paper/Rmd/hotspots-paper.Rmd
added dryad citation

@@
 inputs <- tibble(hot_chr, hot_mid, keller_mediator)
-out <- pmap_dfc(inputs, .f = function(hot_chr, hot_mid, keller_mediator) knitr::knit_expand(text = readLines(here::here("analysis", "paper", "hotspot-one-chromosome.Rmd"))))
+out <- pmap_dfc(inputs, .f = function(hot_chr, hot_mid, keller_mediator) knitr::knit_expand(text = readLines(here::here("analysis", "paper", "Rmd", "hotspot-one-chromosome.Rmd"))))
 # note that ... argument must be a list, ie, it must be a single argument, not multiple args

### 2019-08-03 analysis/paper/Rmd/methods.Rmd
organized analysis/paper dir
added dryad citation

--- 2019-08-03 analysis/paper/methods.Rmd
@@
 measured genome-wide gene expression levels with RNA sequencing of pancreatic islets in 378 Diversity Outbred mice.
-From the RNA sequencing counts, they obtained normalized gene expression levels for 21,771 transcripts, which they shared publicly via DataDryad ().
+From the RNA sequencing counts, they obtained normalized gene expression levels for 21,771 transcripts, which they shared publicly via DataDryad [@dryad].
+@keller2018genetic also included founder allele probabilities at 69,005 markers across the genome, covariates, and leave-one-chromosome-out estimated kinship matrices.
+@keller2018genetic calculated leave-one-chromosome-out kinship matrices for every chromosome [@broman2019rqtl2; @yang2014advantages].
 
@@
 consists of columns that summarize mediation results. Each column represents a local trait. If a row 
-corresponds to a nonlocal trait that is mediated by the local trait (*i.e.*, with mediation LOD difference greater than 1.5), then the cell is colored dark green; 
-otherwise, the cell is light green.
-We didn't examine the possibility that a local trait might mediate the relationship between a QTL and another local trait; thus, all local trait rows are colored light green.
-The last annotation column and the annotation row above the heatmap both
+corresponds to a nonlocal trait that is mediated by the local trait (*i.e.*, with mediation LOD 
+difference greater than 1.5), then the cell is colored dark green; 
+otherwise, the cell is light green. We didn't examine the possibility that a local trait might 
+mediate the relationship between a QTL and another local trait; thus, all local trait rows are 
+colored light green. The last annotation column and the annotation row above the heatmap both 
 designate local traits with dark blue and nonlocal traits with light blue.

### 2019-08-04 analysis/paper/Rmd/hotspot-one-chromosome.Rmd
added fig.cap captions for plots. Also, changed one chunk name to avoid having underscores in figure reference.

@@
 
-```{r, scatter-{{hot_chr}}}
+```{r, scatter-{{hot_chr}}, fig.cap="Pleiotropy test statistic vs. mediation LOD difference for Chr {{hot_chr}} hotspot."}
 p <- pleio_mixed_list %>%
@@
 
-```{r, scatter_by_local-{{hot_chr}}, fig.height = ceiling(nlocal / 2)}
+```{r, scatter_by_local-{{hot_chr}}, fig.height = ceiling(nlocal / 2), fig.cap="Pleiotropy test statistic vs. mediation LOD difference by local expression trait for Chr {{hot_chr}} hotspot."}
 foo <- pleio_mixed_list %>%
@@
 
-```{r, heat-{{hot_chr}}}
+```{r, heat-{{hot_chr}}, fig.cap="Heatmap for pleiotropy test statsitics for Chr {{hot_chr}}."}
 ord <- match(rn, rownames(pm))
@@
 
-```{r, index_v_position-{{hot_chr}}}
+```{r, index-v-position-{{hot_chr}}, fig.cap="QTL peak position vs. dendrogram index for Chr {{hot_chr}}."}
 bar <- foo %>%

### 2019-08-04 analysis/paper/Rmd/methods.Rmd
fixed figure refs

@@
 
-@keller2018genetic identified five expression trait hotspots in their
+@keller2018genetic performed univariate QTL mapping for all expression traits. 
+They identified five expression trait hotspots in their
 study of pancreatic islet gene expression in 378 Diversity Outbred mice. They
@@
 set of nonlocal expression traits. We focused on those local 
-traits that strongly associated with QTL within the hotspot.
-We set the univariate LOD threshold for local traits at 15.
-We limited consideration to those local traits for which the middle of the gene lay within the hotspot interval.
+traits that strongly associated with QTL within the hotspot. We set the univariate LOD threshold for 
+local traits at 15. We limited consideration to those local traits for which the middle of the gene lay 
+within the hotspot interval. 
 All nonlocal traits demonstrated univariate LOD peaks that met
-the genome-wide significance threshold, 7.18 [@keller2018genetic]. The nonlocal expression traits arose from genes
+the genome-wide significance threshold, 7.18 [@keller2018genetic]. The nonlocal expression traits arose 
+from genes
 on other chromosomes, while the local traits reflected expression levels for
-genes on the same chromosome as the hotspot . We ignored transcripts for genes on the hotspot chromosome but outside the hotspot interval.
+genes on the same chromosome as the hotspot . We ignored transcripts for genes on the hotspot chromosome 
+but outside the hotspot interval.
 
@@
 
-For each hotspot, we tested all pairs of traits, regardless of whether the traits were local or nonlocal. We performed a
+For each hotspot, we tested all pairs of traits, regardless of whether the traits were local or nonlocal. 
+We performed a
 two-dimensional, two-QTL scan over a genomic region that included the entirety of the
@@
 We only considered as potential mediators those local traits with strong univariate LODs.
-In Equations~\ref{model1} through \ref{model4}, $X$ is a matrix of founder allele dosages,
+In Equations~\ref{eq:model1} through \ref{eq:model4}, $X$ is a matrix of founder allele dosages,
 $B$ is a matrix

### 2019-08-04 analysis/paper/Rmd/hotspots-paper.Rmd
fixed chunk name by removing underscores

@@
 output: 
+  bookdown::html_document2: 
+    code_folding: hide
+    self_contained: no
   bookdown::pdf_document2:
@@
     keep_tex: true
-  bookdown::html_document2: 
-    code_folding: hide
-    self_contained: no
 ---
@@
 eval_code <- TRUE
-knitr::opts_chunk$set(comment=NA, fig.width=6, fig.height=6, eval = eval_code, results = "hide")
+knitr::opts_chunk$set(comment=NA, fig.width=6, fig.height=6, eval = eval_code)
 ```

### 2019-08-04 analysis/paper/Rmd/hotspots-paper.Rmd
minor edits to paper

@@
     self_contained: no
+    keep_md: true
   bookdown::pdf_document2:

### 2019-08-05 analysis/paper/Rmd/hotspot-one-chromosome-no-curly-braces.Rmd
removed webshot call

# Chr

```{r definitions}
hot_chr <- 2
keller_mediator <- "Hnf4a"
hot_mid <- 165.5
```

```{r, def-hot_chr}
hot_lower <- hot_mid - 2
hot_upper <- hot_mid + 2
fp <- paste0("Chr", hot_chr, "-")
knitr::opts_chunk$set(fig.path = here("analysis", "figures", fp))
```

```{r, readRDS-hot_chr}
pleio_mixed_list <- readRDS(here("analysis", "data", "derived_data", paste0("Chr", hot_chr, "_pleio.rds")))
pleio_local_list <- readRDS(here("analysis", "data", "derived_data", paste0("Chr", hot_chr, "_pleio_local.rds")))
pleio_nonlocal_list <- readRDS(here("analysis", "data", "derived_data", paste0("Chr", hot_chr, "_pleio_nonlocal.rds")))
med_list <- readRDS(here("analysis", "data", "derived_data", paste0("Chr", hot_chr, "_mediations.rds")))
```

```{r, define_local_annot-hot_chr}
local_annot <- lod_peaks %>%
  filter(chrom == hot_chr, pos <= hot_upper, pos >= hot_lower) %>%
  left_join(y = annots, by = c("annot.id" = "gene_id")) %>%
  filter(!duplicated(annot.id)) %>%
  filter(lod >= ult) %>%
  mutate(local_tx = chr == hot_chr) %>%
  filter(local_tx)
```

```{r, define_nonlocal_annot-hot_chr}
nonlocal_annot <- lod_peaks %>%
  filter(chrom == hot_chr, pos <= hot_upper, pos >= hot_lower) %>%
  inner_join(y = annots, by = c("annot.id" = "gene_id")) %>%
  filter(!duplicated(annot.id)) %>%
  filter(lod >= 7.18) %>%
  mutate(local_tx = chr == hot_chr) %>%
  filter(!local_tx) %>%
  filter(!is.na(hotspot))
```

```{r, local_annot2-hot_chr}
local_annot2 <- local_annot %>%
  filter(middle <= hot_upper, middle >= hot_lower) %>% 
  arrange(desc(lod))
```

```{r hot_expr-hot_chr}
hot_local <- expr[, colnames(expr) %in% local_annot2$annot.id]
hot_nonlocal <- expr[, colnames(expr) %in% nonlocal_annot$annot.id]
nnonlocal <- nrow(nonlocal_annot)
nlocal <- nrow(local_annot2)
```

```{r, med_tib-hot_chr, message=FALSE, warning = FALSE}
med_tib <- med_list %>%
  purrr::map(bind_rows) %>%
  bind_rows() %>%
  as_tibble() %>%
  rename(lod_without_med = lod_no_med, lod_with_med = lod_med) %>%
  mutate(lod_diff = lod_without_med - lod_with_med, lod_diff_proportion = lod_diff / lod_without_med) %>%
  left_join(local_annot2, by = c("local_gene_id" = "annot.id")) %>%
  left_join(nonlocal_annot, by = c("nonlocal_gene_id" = "annot.id")) %>%
  dplyr::select(symbol.x, symbol.y, lod_without_med, lod_with_med, lod_diff, lod_diff_proportion, local_gene_id, nonlocal_gene_id) %>%
  rename(local_symbol = symbol.x, nonlocal_symbol = symbol.y) %>%
  filter(!is.na(local_symbol))
```

```{r, hot_annot-hot_chr}
hot_annot <- local_annot2 %>%
  bind_rows(nonlocal_annot) %>%
  dplyr::select(annot.id, symbol)
pleio_tib <- pleio_mixed_list %>%
  purrr::map(.f = function(x) rename(x, gene1_id = local_gene_id, gene2_id = nonlocal_gene_id)) %>%
  c(pleio_local_list, pleio_nonlocal_list) %>%
  bind_rows() %>%
  rename(pleiotropy_lod = lod) %>%
  inner_join(hot_annot, by = c("gene1_id" = "annot.id")) %>%
  rename(gene1_symbol = symbol) %>%
  inner_join(hot_annot, by = c("gene2_id" = "annot.id")) %>%
  rename(gene2_symbol = symbol) %>%
  dplyr::select(gene1_symbol, gene2_symbol, pleiotropy_lod)
```

```{r pm-hot_chr}
pm <- pleio_tib %>%
  tibble_to_matrix(symmetric = TRUE)
```

```{r, scatter-hot_chr, fig.cap="Pleiotropy test statistic vs. mediation LOD difference for Chr hot_chr hotspot."}
p <- pleio_mixed_list %>%
  bind_rows() %>%
  rename(pleiotropy_lod = lod) %>%
  inner_join(med_tib, by = c("local_gene_id", "nonlocal_gene_id")) %>%
  dplyr::select(local_symbol, nonlocal_symbol, pleiotropy_lod, lod_diff, lod_diff_proportion, lod_without_med, lod_with_med, filename, nonlocal_gene_id, local_gene_id) %>%
  ggplot() + geom_point(mapping = aes(x = pleiotropy_lod, y = lod_diff, colour = local_symbol == "keller_mediator", nonlocal = nonlocal_symbol, local = local_symbol), size = 0.05) + theme(legend.position = "none") + xlab("Pleiotropy LOD") + ylab("Mediation LOD difference")
fnsvg <- paste0("Chr", hot_chr, "_scatter.svg")
fnpdf <- paste0("Chr", hot_chr, "_scatter.pdf")
ggsave(plot = p, filename = here("analysis", "figures", fnsvg))
ggsave(plot = p, filename = here("analysis", "figures", fnpdf))
plotly::ggplotly(p)
```

```{r, scatter-by-local-hot_chr, fig.height = ceiling(nlocal / 2), fig.cap="Pleiotropy test statistic vs. mediation LOD difference by local expression trait for Chr hot_chr hotspot."}
foo <- pleio_mixed_list %>%
  bind_rows() %>%
  rename(pleiotropy_lod = lod) %>%
  inner_join(med_tib, by = c("local_gene_id", "nonlocal_gene_id")) %>%
  dplyr::select(local_symbol, nonlocal_symbol, pleiotropy_lod, lod_diff, lod_diff_proportion, lod_without_med, lod_with_med, filename, nonlocal_gene_id, local_gene_id)
p <- foo %>%
  ggplot() +
  geom_point(mapping = aes(x = pleiotropy_lod, y = lod_diff, colour = nonlocal_symbol), size = 0.1) +
  facet_wrap(. ~ local_symbol, nrow = ceiling(nlocal / 2), ncol = 2) +
  broman::karl_theme(strip.placement = "inside") +
  theme(legend.position = "none") +
  theme(
    strip.background = element_blank(),
    strip.text.x = element_blank()
) +
  xlab("Pleiotropy LOD") +
  ylab("Mediation LOD difference") +
  geom_text(aes(x = max(foo$pleiotropy_lod), y = 1, label = local_symbol), inherit.aes=FALSE)

fnsvg <- paste0("Chr", hot_chr, "_scatter-panel.svg")
fnpdf <- paste0("Chr", hot_chr, "_scatter-panel.pdf")
ggsave(p, filename = here("analysis", "figures", fnsvg), height = ceiling(nlocal / 2))
ggsave(p, filename = here("analysis", "figures", fnpdf), height = ceiling(nlocal / 2))
plotly::ggplotly(p)
```

```{r, scols-hot_chr}
scols1 <- 1L + (rownames(pm) %in% local_annot2$symbol)
scols2 <- rownames(pm) %in% (med_tib %>% filter(local_symbol == "keller_mediator", lod_diff > 1.5))$nonlocal_symbol
```

```{r, ordering-hot_chr}
```

```{r, scols_extra-hot_chr}
scols_extra <- purrr::map(.x = local_annot2$symbol,
                          .f = function(x) {rownames(pm) %in% (med_tib %>% filter(local_symbol == x, lod_diff > 1.5))$nonlocal_symbol}
) %>%
  bind_cols() %>%
  (function(x){colnames(x) <- local_annot2$symbol; x}) %>%
  mutate(local = scols1)
```

```{r, n_med-hot_chr}
n_med_tr <- 4 # arbitrary threshold for number of mediated traits per local
# only those local traits with n_med_tr or more mediated nonlocal traits 
# appear as annotation columns in the heatmap
n_med_per_local <- med_tib %>%
  group_by(local_symbol) %>%
  filter(lod_diff > 1.5) %>%
  count() %>%
  filter(n >= n_med_tr)
n_med_per_local %>%
  arrange(desc(n)) %>%
  knitr::kable()

scols_pre <- scols_extra %>%
  dplyr::select(n_med_per_local$local_symbol)
# get ordering for columns of scols_pre
col_order <- scols_pre %>%
  t() %>%
  dist() %>%
  hclust() %>%
  as.dendrogram() %>%
  dendextend::seriate_dendrogram(dist(t(scols_pre))) %>%
  order.dendrogram()
scols <- scols_pre[ , col_order] %>%
  mutate(local = scols1)
fnhtml <- here("analysis", "figures", paste0("Chr", hot_chr, "hm.html"))
```

```{r, heat-hot_chr, fig.cap="Heatmap for pleiotropy test statsitics for Chr hot_chr."}
ord <- match(rn, rownames(pm))

(p <- pm[ord, ord] %>%
  apply(FUN = trunc2, MARGIN = 2, threshold = 5) %>%
  heatmaply(Rowv = FALSE,
            Colv = FALSE,
            cexRow = 0.2,
            cexCol = 0.2,
            symm = TRUE,
            row_side_colors = scols[ord, ],
            col_side_colors = tibble(local = scols1[ord]),
            limits = c(0, 5)
            ) %>%
  hide_guides()
)
fnpdf <- here("analysis", "figures", paste0("Chr", hot_chr, "hm.pdf"))
htmlwidgets::saveWidget(widget = p, 
                        file = fnhtml
                        )
```

```{r, webshot-hot_chr, eval = FALSE}
webshot::webshot(url = fnhtml,
                 file = fnpdf,
                 cliprect = "viewport"
                 )
```

```{r, index-v-position-hot_chr, fig.cap="QTL peak position vs. dendrogram index for Chr hot_chr."}
bar <- foo %>%
  filter(local_symbol == keller_mediator)

p <- tibble(nonlocal_symbol = rownames(pm[indic_nonlocal, ])[ord_nonlocal], index = 1:length(nonlocal_symbol)) %>%
  inner_join(bar, by = "nonlocal_symbol") %>% 
  left_join(nonlocal_annot, by = c("nonlocal_symbol" = "symbol")) %>%
  ggplot() + geom_point(aes(y = pos, x = index, color = lod_diff > 1.5, nonlocal = nonlocal_symbol, lod_diff = lod_diff))

ggplotly(p)
```

### 2019-08-05 analysis/paper/Rmd/intro.Rmd
removed webshot call

--- 2019-08-02 analysis/paper/intro.Rmd
@@
 concentrations [@sax1923association;@soller1976power;@lander1989mapping;@broman2009guide;@jansen2007quantitative;@chick2016defining]. 
-These genomic regions are called "quantitative trait loci" or "QTL".
+These genomic regions that affect variability in one or more complex traits are called "quantitative trait loci" or "QTL".
 A genome-wide QTL "scan" reveals
@@
 one at a time, as a candidate QTL for the trait of interest.
-A region with strong evidence of association with a complex quantitative trait, then, defines a QTL (for that trait).
+A genomic region with strong evidence of association with a complex quantitative trait, then, defines a QTL.
 Because nearby markers have correlated
@@
 
-@boehm2019thesis argued for complementary roles for mediation analysis and pleiotropy testing in hotspot 
+Other researchers have identified and examined QTL hotspots in model organism genetics studies. @tian2016dissection 
+presented quantitative methods for hotspot dissection in F$_2$ intercross mice.
+
+@boehm2019thesis and @boehm2019testing argued for complementary roles for mediation analysis and pleiotropy testing in hotspot 
 dissection. Here, we clarify this argument by presenting a study of five expression trait 
-hotspots in a publicly available data set. We consider each hotspot one at a time. At each hotspot, we identify 
+hotspots in a publicly available data set. At each hotspot, we identify 
 a set of local traits that have QTL within the hotspot. We also identify a set of nonlocal traits with QTL 
@@
 
+I
+
 ## INTRO IDEAS

### 2019-08-05 analysis/paper/Rmd/methods.Rmd
removed webshot call

@@
 
-### Two-dimensional, two-QTL scans
-
 \begin{equation}
@@
 two-dimensional, two-QTL scan over a genomic region that included the entirety of the
-hotspot. We fitted bivariate linear mixed effects models (Equations \@ref(), \@ref(), and \@ref()) at each ordered pair of markers in a two-dimensional grid [@boehm2019testing].
+hotspot. We fitted bivariate linear mixed effects models (Equations \@ref(eq:mvlmm), \@ref(eq:g), and \@ref(eq:e)) at each ordered pair of markers in a two-dimensional grid [@boehm2019testing].
 
@@
 Estimates $\hat V_g$ and $\hat V_e$ are obtained via an expectation-maximization algorithm [@zhou2014efficient;@gemma2].
-
-### Calculating likelihood ratio test statistics
 
@@
 statistical computing environment [@r].
-We initially plotted all local-nonlocal pairs for each hotspot.
-We highlighted in blue those points that correspond to pairs involving the putative mediator per [@keller2018genetic].
+We plotted all local-nonlocal pairs for each hotspot.
+We highlighted in blue those points that correspond to pairs involving the putative mediator that @keller2018genetic identified.
 

### 2019-08-05 analysis/paper/Rmd/hotspot-one-chromosome.Rmd
fixed chunk name by removing underscores
removed calls to save figures

@@
   ggplot() + geom_point(mapping = aes(x = pleiotropy_lod, y = lod_diff, colour = local_symbol == "{{keller_mediator}}", nonlocal = nonlocal_symbol, local = local_symbol), size = 0.05) + theme(legend.position = "none") + xlab("Pleiotropy LOD") + ylab("Mediation LOD difference")
-fnsvg <- paste0("Chr", {{hot_chr}}, "_scatter.svg")
-fnpdf <- paste0("Chr", {{hot_chr}}, "_scatter.pdf")
-ggsave(plot = p, filename = here("analysis", "figures", fnsvg))
-ggsave(plot = p, filename = here("analysis", "figures", fnpdf))
 plotly::ggplotly(p)
@@
 
-```{r, scatter_by_local-{{hot_chr}}, fig.height = ceiling(nlocal / 2), fig.cap="Pleiotropy test statistic vs. mediation LOD difference by local expression trait for Chr {{hot_chr}} hotspot."}
+```{r, scatter-by-local-{{hot_chr}}, fig.height = ceiling(nlocal / 2), fig.cap="Pleiotropy test statistic vs. mediation LOD difference by local expression trait for Chr {{hot_chr}} hotspot."}
 foo <- pleio_mixed_list %>%
@@
   geom_text(aes(x = max(foo$pleiotropy_lod), y = 1, label = local_symbol), inherit.aes=FALSE)
-
-fnsvg <- paste0("Chr", {{hot_chr}}, "_scatter-panel.svg")
-fnpdf <- paste0("Chr", {{hot_chr}}, "_scatter-panel.pdf")
-ggsave(p, filename = here("analysis", "figures", fnsvg), height = ceiling(nlocal / 2))
-ggsave(p, filename = here("analysis", "figures", fnpdf), height = ceiling(nlocal / 2))
 plotly::ggplotly(p)
-```
-
-```{r, eval = FALSE}
-# put in multiple figures, if there are many local traits
-n_figs <- 1 + nlocal %/% 18 # 18 is the number of panels per page in pdf
-n_last <- nlocal %% 18 # number of panels on last figure
-max_pleio_lod <- max(foo$pleiotropy_lod)
-max_med_lod_difference <- max(foo$lod_diff)
-min_med_lod_difference <- min(foo$lod_diff)
-
-for (i in 1:n_figs){
-  n_panels <- 18 * (i < n_figs) + n_last * (i == n_figs)
-  foo %>%
-    filter(local_symbol %in% local_annot2$symbol[((i - 1) * 18 + 1): (18 * i)]) %>%
-    ggplot() +
-      geom_point(mapping = aes(x = pleiotropy_lod, y = lod_diff, colour = nonlocal_symbol), size = 0.1) +
-      facet_wrap(. ~ local_symbol, nrow = ceiling(n_panels / 2), ncol = 2) +
-      broman::karl_theme() +
-      theme(legend.position = "none") +
-      theme(
-        strip.background = element_blank(),
-        strip.text.x = element_blank()
-    ) +
-      xlab("Pleiotropy LOD") + xlim(c(0, max_pleio_lod)) + ylim(c(min_med_lod_difference, max_med_lod_difference)) +
-      ylab("Mediation LOD difference") +
-      geom_text(size = 2, aes(x = max_pleio_lod / 2, y = 2 * max_med_lod_difference / 3, label = local_symbol), inherit.aes=FALSE)
-  fnsvg <- paste0("Chr", {{hot_chr}}, "_scatter-panel", i, ".svg")
-  fnpdf <- paste0("Chr", {{hot_chr}}, "_scatter-panel", i, ".pdf")
-  ggsave(filename = here("analysis", "figures", fnsvg), height = ceiling(n_panels / 2))
-  ggsave(filename = here("analysis", "figures", fnpdf), height = ceiling(n_panels / 2))
-}
 ```
@@
 # define ord_nonlocal
-ord_nonlocal <- pm[indic_nonlocal, indic_nonlocal] %>%
+ord_nonlocal <- pm[indic_nonlocal, ] %>%
   dist() %>%
@@
   as.dendrogram() %>%
-  dendextend::seriate_dendrogram(dist(pm[indic_nonlocal, indic_nonlocal])) %>%
+  dendextend::seriate_dendrogram(dist(pm[indic_nonlocal, ])) %>%
   order.dendrogram()
-rn_nonlocal <- rownames(pm[indic_nonlocal, indic_nonlocal][ord_nonlocal, ord_nonlocal])
+rn_nonlocal <- rownames(pm[indic_nonlocal, ][ord_nonlocal, ])
 # define ord_local
-ord_local<- pm[!indic_nonlocal, !indic_nonlocal] %>%
+ord_local<- pm[!indic_nonlocal, ] %>%
   dist() %>%
@@
   as.dendrogram() %>%
-  dendextend::seriate_dendrogram(dist(pm[!indic_nonlocal, !indic_nonlocal])) %>%
+  dendextend::seriate_dendrogram(dist(pm[!indic_nonlocal, ])) %>%
   order.dendrogram()
-rn_local <- rownames(pm[!indic_nonlocal, !indic_nonlocal][ord_local, ord_local])
+rn_local <- rownames(pm[!indic_nonlocal, ][ord_local, ])
 rn <- c(rn_nonlocal, rn_local)
@@
 
-```{r, n_med-{{hot_chr}}}
-n_med_tr <- 4 # arbitrary threshold for number of mediated traits per local
+```{r, n-med-{{hot_chr}}}
+n_med_tr <- 10 # arbitrary threshold for number of mediated traits per local
 # only those local traits with n_med_tr or more mediated nonlocal traits 
@@
   arrange(desc(n)) %>%
-  knitr::kable()
+  knitr::kable(caption = "Number of mediated traits per local trait for Chr {{hot_chr}} hotspot.")
 
@@
   mutate(local = scols1)
-fnhtml <- here("analysis", "figures", paste0("Chr", {{hot_chr}}, "hm.html"))
 ```
 
-```{r, heat-{{hot_chr}}, fig.cap="Heatmap for pleiotropy test statsitics for Chr {{hot_chr}}."}
+```{r, heat-{{hot_chr}}, fig.cap="Heatmap for pleiotropy test statistics for Chr {{hot_chr}}."}
 ord <- match(rn, rownames(pm))
 
-(p <- pm[ord, ord] %>%
+(p <- pm[ord, ord] %>% # order both cols and rows by 'ord'
   apply(FUN = trunc2, MARGIN = 2, threshold = 5) %>%
@@
 )
-fnpdf <- here("analysis", "figures", paste0("Chr", {{hot_chr}}, "hm.pdf"))
-htmlwidgets::saveWidget(widget = p, 
-                        file = fnhtml
-                        )
-```
-
-```{r, webshot-{{hot_chr}}, eval = FALSE}
-webshot::webshot(url = fnhtml,
-                 file = fnpdf,
-                 cliprect = "viewport"
-                 )
 ```
@@
 
-p <- tibble(nonlocal_symbol = rownames(pm[indic_nonlocal, indic_nonlocal])[ord_nonlocal], index = 1:length(nonlocal_symbol)) %>%
+p <- tibble(nonlocal_symbol = rownames(pm[indic_nonlocal, ])[ord_nonlocal], index = 1:length(nonlocal_symbol)) %>%
   inner_join(bar, by = "nonlocal_symbol") %>% 

### 2019-08-06 R/bin_peak_position.R
added code for bin_peak_position function

@@ -1,0 +1,10 @@
+#' Assign a peak position to a bin number
+#'
+#' @param peak_position one peak position (numeric)
+#' @param breaks vector defining the boundaries of bins
+#' @return an integer, the bin number
+#' @export
+#'
+bin_peak_position <- function(peak_position, breaks){
+  return(sum(peak_position > breaks) + 1)
+}

### 2019-08-07 R/bin_peak_position.R
added tests for function bin_peak_position()

@@ -4,7 +4,14 @@
 #' @param breaks vector defining the boundaries of bins
 #' @return an integer, the bin number
 #' @export
+#' @examples
+#' bin_peak_position(5, 1:10)
+#' bin_peak_position(1, 2)
 #'
 bin_peak_position <- function(peak_position, breaks){
+  if (length(peak_position) > 1) stop("peak_position must be length 1")
+  if (!is.numeric(peak_position)) stop("peak_position must be numeric")
+  if (length(breaks) < 1) stop("breaks must have length at least 1")
+  if (!is.numeric(breaks)) stop("breaks must be numeric")
   return(sum(peak_position > breaks) + 1)
 }

### 2019-08-07 tests/testthat/test-bin_peak_position.R

@@ -1,0 +1,17 @@
+library(hotspots)
+library(testthat)
+
+
+test_that("bin assignment is correct in simple cases", {
+  expect_equal(bin_peak_position(1, 2), 1)
+  expect_equal(bin_peak_position(2, 1), 2)
+  expect_equal(bin_peak_position(5, 1:10), 5)
+})
+
+
+test_that("errors occur when input is bad", {
+  expect_error(bin_peak_position(1:2, 2))
+  expect_error(bin_peak_position("a", 2))
+  expect_error(bin_peak_position(1, NULL))
+  expect_error(bin_peak_position(1, "a"))
+})

### 2019-08-08 analysis/paper/Rmd/hotspot-one-chromosome.Rmd
updated paper.Rmd
added doc to binning functions

@@
-# Chr {{hot_chr}}
+## Chr {{hot_chr}}
 
@@
 hot_upper <- {{hot_mid}} + 2
-fp <- paste0("Chr", {{hot_chr}}, "-")
-knitr::opts_chunk$set(fig.path = here("analysis", "figures", fp))
 ```
@@
 ```{r, scols-{{hot_chr}}}
-scols1 <- 1L + (rownames(pm) %in% local_annot2$symbol)
-scols2 <- rownames(pm) %in% (med_tib %>% filter(local_symbol == "{{keller_mediator}}", lod_diff > 1.5))$nonlocal_symbol
+scols1 <- tibble(local = 1L + (rownames(pm) %in% local_annot2$symbol), 
+                 symbol = rownames(pm), 
+                 mediated_by_keller_mediator = rownames(pm) %in% (med_tib %>% filter(local_symbol == "{{keller_mediator}}", lod_diff > 1.5))$nonlocal_symbol
+                 )
+
 ```
@@
 ```{r, scols_extra-{{hot_chr}}}
-scols_extra <- purrr::map(.x = local_annot2$symbol,
-                          .f = function(x) {rownames(pm) %in% (med_tib %>% filter(local_symbol == x, lod_diff > 1.5))$nonlocal_symbol}
+scols_extra <- local_annot2$symbol %>%
+  as.list() %>%
+  (function(x) {names(x) <- unlist(x); x}) %>%
+  purrr::map(.f = function(x) {rownames(pm) %in% (med_tib %>% filter(local_symbol == x, lod_diff > 1.5))$nonlocal_symbol}
 ) %>%
   bind_cols() %>%
-  (function(x){colnames(x) <- local_annot2$symbol; x}) %>%
-  mutate(local = scols1)
+  mutate(symbol = rownames(pm)) %>%
+left_join(scols1, by = "symbol")
+```
+
+```{r, scols-peak-position-{{hot_chr}}}
+annot_all <- local_annot2 %>%
+  bind_rows(nonlocal_annot)
+
+binning <- annot_all %>%
+  dplyr::select(symbol, pos, lod) %>%
+  mutate(bin = unlist(pos) %>%
+           map(.f = function(x) {x %>% 
+               bin_peak_position(breaks = hot_lower + (1:9) * 0.4) %>%
+               binarize_bin_number(total_number_of_bins = 10L)
+             })
+  )
+```
+
+```{r, scan1-{{hot_chr}}}
+fn <- here::here("analysis", "data", "derived_data", "Chr{{hot_chr}}_scan1_out.rds")
+if (file.exists(fn)) {
+  s1out <- readRDS(fn)
+} else {
+  s1out <- qtl2::scan1(genoprobs = probs, 
+                       pheno = cbind(hot_local, hot_nonlocal), 
+                       kinship = K[["{{hot_chr}}"]], 
+                       addcovar = covar, 
+                       reml = TRUE
+                       )
+  saveRDS(s1out, fn)
+}
+```
+
+```{r allele-effects-{{hot_chr}}}
+fn <- here::here("analysis", "data", "derived_data", "Chr{{hot_chr}}_scan1coef_out.rds")
+if (file.exists(fn)) {
+  s1coef_out <- readRDS(fn)
+} else {
+s1coef_out <- purrr::map(.x = tibble(cbind(hot_local, hot_nonlocal)), 
+           .f = function(pheno) {
+             qtl2::scan1coef(genoprobs = probs[["{{hot_chr}}"]], 
+                             pheno = pheno, 
+                             kinship = K[["{{hot_chr}}"]], 
+                             addcovar = covar, 
+                             reml = TRUE, 
+                             )
+           }
+                )
+saveRDS(s1coef_out, fn)
+}
+```
+
+```{r, get-allele-effects-{{hot_chr}}}
+# add index values for peak position
+local_annot2 %>%
+  mutate(peak_index = match(table = names(map[["{{hot_chr}}"]]), x = marker.id)) %>%
+  mutate(peak_effects = qtl2pleio::get_effects(peak_index, allele_effects_matrix = s1coef_out, map = map[["{{hot_chr}}"]]))
+
+nonlocal_annot %>%
+  mutate(peak_index = match(table = names(map[["{{hot_chr}}"]]), x = marker.id))
+
 ```
@@
   knitr::kable(caption = "Number of mediated traits per local trait for Chr {{hot_chr}} hotspot.")
+```
 
+```{r define-scols-pre-{{hot_chr}}}
 scols_pre <- scols_extra %>%
@@
   apply(FUN = trunc2, MARGIN = 2, threshold = 5) %>%
-  heatmaply(Rowv = FALSE,
+  heatmaply::heatmapr(Rowv = FALSE,
             Colv = FALSE,
-            cexRow = 0.2,
-            cexCol = 0.2,
+            plot_method = "plotly"
+            ) %>%
+    heatmaply::heatmaply_heatmapr2(
+            fontsize_col = 2,
+            fontsize_row = 2,
             symm = TRUE,
-            row_side_colors = scols[ord, ],
-            col_side_colors = tibble(local = scols1[ord]),
+            row_side_colors = list(scols[ord, ], tibble(scols1$local[ord])),
+            row_side_palette = list(heatmaply::cool_warm, heatmaply::Blues), 
+            col_side_colors = tibble(local = scols1$local[ord]),
             limits = c(0, 5)

### 2019-08-08 analysis/paper/Rmd/hotspots-paper-2.Rmd
added doc to binning functions

--- 2019-08-04 analysis/paper/Rmd/hotspots-paper.Rmd
@@
     code_folding: hide
-    self_contained: no
-  bookdown::pdf_document2:
-    fig_caption: true
-    keep_tex: true
 ---
@@
 library(qtl2hotspots)
-pleio_threshold <- 4
 ult <- 15 # univariate lod threshold
@@
 expr <- readRDS(here("analysis", "data", "derived_data", "expr.rds"))
+probs <- readRDS(here("analysis", "data", "derived_data", "genoprobs.rds"))
 ```
@@
 
-### Chr 2 hotspot results
-
 `r if (eval_code) knitr::knit(text = out[[1]])`
-
-### Chr 5 hotspot results
-
-`r if (eval_code) knitr::knit(text = out[[2]])`
-
-### Chr 7 results
-
-`r if (eval_code) knitr::knit(text = out[[3]])`
-
-### Chr 11 results
-
-`r if (eval_code) knitr::knit(text = out[[4]])`
-
-### Chr 13 results
-
-`r if (eval_code) knitr::knit(text = out[[5]])`
 

### 2019-08-08 R/bin_peak_position.R

@@ -15,3 +15,27 @@ bin_peak_position <- function(peak_position, breaks){
   if (!is.numeric(breaks)) stop("breaks must be numeric")
   return(sum(peak_position > breaks) + 1)
 }
+
+#' Binarize bin number
+#'
+#' @param bin_number a positive integer
+#' @param total_number_of_bins a positive integer, at least as big as bin_number
+#' @return a binary vector with exactly one nonzero entry
+#' @export
+#' @examples
+#' binarize_bin_number(3, 10)
+#' binarize_bin_number(10L, 10L)
+#'
+binarize_bin_number <- function(bin_number, total_number_of_bins){
+  if (bin_number > total_number_of_bins) stop("bin_number must be less than or equal to bin_number")
+  if (bin_number < 1) stop("bin_number must be at least 1")
+  if (total_number_of_bins < 1) stop("total_number_of_bins must be at least 1")
+  if (!is.numeric(bin_number)) stop("bin_number must be an integer")
+  if (!is.numeric(total_number_of_bins)) stop("total_number_of_bins must be an integer")
+  if (length(bin_number) != 1) stop("bin_number must have length 1")
+  if (length(total_number_of_bins) != 1) stop("total_number_of_bins must have length 1")
+  # end checks
+  out <- rep(0, total_number_of_bins)
+  out[bin_number] <- 1
+  return(out)
+}

### 2019-08-08 tests/testthat/test-binarize_bin_number.R

@@ -1,0 +1,10 @@
+library(hotspots)
+library(testthat)
+
+
+test_that("binary vector is correct in simple cases", {
+  expect_equal(sum(binarize_bin_number(1, 2)), 1)
+  expect_length(binarize_bin_number(1, 2), 2)
+  expect_equal(class(binarize_bin_number(1, 2)), "numeric")
+})
+

### 2019-08-09 R/listcol_to_cols.R
documented & exported listcol_to_list

@@ -1,0 +1,19 @@
+#' Convert a tibble with one list-column of vectors (of same length) to a tibble without list-column, where each entry of vector (within list-column) is a separate column.
+#'
+#' @param tibble a tibble containing one list-column
+#' @param list_column name of list-column, enclosed with dplyr::quo()
+#' @param another_column name of another column, enclosed with dplyr::quo()
+#' @return a tibble with the list-column replaced with multiple columns. New columns have backticked numbers as names, like `1`
+#' @examples
+#' tibble(a = 1:3, b = c(list(1:2), list(3:4), list(5:6))) %>%
+#' listcol_to_cols(list_column = dplyr::quo(b), another_column = dplyr::quo(a))
+#' @export
+listcol_to_cols <- function(tibble, list_column, another_column){
+  tibble %>%
+    tidyr::unnest() %>%
+    dplyr::group_by(!! another_column) %>%
+    dplyr::mutate(col=seq_along(!! another_column)) %>%
+    tidyr::spread(key = col, value = !! list_column)
+}
+# see: https://cran.r-project.org/web/packages/dplyr/vignettes/programming.html
+# see: https://stackoverflow.com/questions/50881440/split-a-list-column-into-multiple-columns-in-r

### 2019-08-09 analysis/paper/Rmd/hotspot-one-chromosome.Rmd
R coding revisions

@@
 pm <- pleio_tib %>%
+  arrange(gene1_symbol, gene2_symbol) %>%
   tibble_to_matrix(symmetric = TRUE)
@@
                  )
-
 ```
@@
 rn_local <- rownames(pm[!indic_nonlocal, ][ord_local, ])
-rn <- c(rn_nonlocal, rn_local)
+rn <- c(rn_nonlocal, rn_local) # order for dendrogram
 ```
@@
 ```{r, scols-peak-position-{{hot_chr}}}
-annot_all <- local_annot2 %>%
-  bind_rows(nonlocal_annot)
 
-binning <- annot_all %>%
+binning <- local_nonlocal_annot %>%
   dplyr::select(symbol, pos, lod) %>%
@@
              })
-  )
+  ) %>%
+  listcol_to_cols(list_column = quo(bin), another_column = quo(symbol))
 ```
@@
                        addcovar = covar, 
-                       reml = TRUE
+                       reml = TRUE, 
+                       cores = 7
                        )
@@
 } else {
-s1coef_out <- purrr::map(.x = tibble(cbind(hot_local, hot_nonlocal)), 
+s1coef_out <- purrr::map(.x = as_tibble(cbind(hot_local, hot_nonlocal)), 
            .f = function(pheno) {
-             qtl2::scan1coef(genoprobs = probs[["{{hot_chr}}"]], 
-                             pheno = pheno, 
+             ph <- as.matrix(pheno)
+             rownames(ph) <- rownames(expr)
+             qtl2::scan1coef(genoprobs = list(probs[["{{hot_chr}}"]]), 
+                             pheno = ph, 
                              kinship = K[["{{hot_chr}}"]], 
@@
                              reml = TRUE, 
+                             cores = 7
                              )
@@
 # add index values for peak position
-local_annot2 %>%
+local_nonlocal_annot <- dplyr::bind_rows(local_annot2, nonlocal_annot) %>%
   mutate(peak_index = match(table = names(map[["{{hot_chr}}"]]), x = marker.id)) %>%
-  mutate(peak_effects = qtl2pleio::get_effects(peak_index, allele_effects_matrix = s1coef_out, map = map[["{{hot_chr}}"]]))
-
-nonlocal_annot %>%
-  mutate(peak_index = match(table = names(map[["{{hot_chr}}"]]), x = marker.id))
-
+  mutate(peak_effects = purrr::map2(.f = qtl2pleio::get_effects, .x = peak_index, .y = s1coef_out, map = map[["{{hot_chr}}"]])) %>%
+  listcol_to_cols(list_column = quo(peak_effects), another_column = quo(symbol))
 ```
@@
 scols <- scols_pre[ , col_order] %>%
-  mutate(local = scols1)
+  mutate(local = scols1$local)
+```
+
+```{r, prepare-annotations-{{hot_chr}}}
+# need 5 annotation object: 1.
+local v nonlocal, 2.
+mediation LOD difference for all local traits, 3.
+mediation LOD difference proportion for all local traits, 4.
+8 allele effects per trait, 5.
+peak position per trait, encoded as binned position
+# 2,3,4 are numeric; 1 & 5 are binary
+## local_nonlocal_annot already has the allele effect and local v nonlocal; binning has the peak position. Need to join these two on symbol
+### med_tib has lod diff & lod diff proportion for all local - nonlocal pairs
+fake_tibble_mediation <- tidyr::crossing(symbol = local_annot2$symbol, 
+                                         symbol2 = local_annot2$symbol) %>%
+  mutate(dat = NA) %>%
+  spread(key = symbol, value = dat) %>%
+  rename(symbol = symbol2)
+
+lod_diffs_tibble <- med_tib %>%
+  select( - c(lod_without_med, 
+              lod_with_med, 
+              lod_diff_proportion, 
+              local_gene_id, 
+              nonlocal_gene_id)
+          ) %>%
+  tidyr::spread(key = local_symbol, 
+                value = lod_diff) %>% 
+  dplyr::rename(symbol = nonlocal_symbol) %>%
+  dplyr::bind_rows(fake_tibble_mediation)
+lod_diff_proportions_tibble <- med_tib %>%
+  select( - c(lod_without_med, 
+              lod_with_med, 
+              lod_diff, 
+              local_gene_id, 
+              nonlocal_gene_id)
+          ) %>%
+  tidyr::spread(key = local_symbol, 
+                value = lod_diff_proportion) %>% # 147 rows by 32 columns
+  rename(symbol = nonlocal_symbol) %>%
+  bind_rows(fake_tibble_mediation)
+allele_effects_tibble <- local_nonlocal_annot %>%
+  dplyr::select(symbol, 18:25)
+peak_positions <- binning %>%
+  select(symbol, pos)
+local_tibble <- local_nonlocal_annot %>%
+  select(symbol, local_tx) %>% 
+  ungroup()
+
 ```
@@
 ```{r, heat-{{hot_chr}}, fig.cap="Heatmap for pleiotropy test statistics for Chr {{hot_chr}}."}
-ord <- match(rn, rownames(pm))
+ord <- match(x = rn, table = rownames(pm))
 
-(p <- pm[ord, ord] %>% # order both cols and rows by 'ord'
+pm[ord, ord] %>% # order both cols and rows by 'ord'
   apply(FUN = trunc2, MARGIN = 2, threshold = 5) %>%
@@
   hide_guides()
-)
 ```

### 2019-08-09 analysis/paper/Rmd/hotspots-paper.Rmd
removed unused assignment
updated paper.Rmd
R coding revisions

@@
 expr <- readRDS(here("analysis", "data", "derived_data", "expr.rds"))
-```
-
-```{r, load_purrr}
-library(purrr)
+probs <- readRDS(here("analysis", "data", "derived_data", "genoprobs.rds"))
+K <- readRDS(here("analysis", "data", "derived_data", "kinship.rds"))
+covar <- readRDS(here("analysis", "data", "derived_data", "covar.rds"))
+map <- readRDS(here("analysis", "data", "derived_data", "map.rds"))
 ```
@@
 
-### Chr 2 hotspot results
-
 `r if (eval_code) knitr::knit(text = out[[1]])`
-
-### Chr 5 hotspot results
 
@@
 
-### Chr 7 results
-
 `r if (eval_code) knitr::knit(text = out[[3]])`
 
-### Chr 11 results
-
 `r if (eval_code) knitr::knit(text = out[[4]])`
-
-### Chr 13 results
 

### 2019-08-10 analysis/paper/Rmd/hotspot-one-chromosome.Rmd
organized annotations

@@
 
-```{r, scols_extra-{{hot_chr}}}
-scols_extra <- local_annot2$symbol %>%
-  as.list() %>%
-  (function(x) {names(x) <- unlist(x); x}) %>%
-  purrr::map(.f = function(x) {rownames(pm) %in% (med_tib %>% filter(local_symbol == x, lod_diff > 1.5))$nonlocal_symbol}
-) %>%
-  bind_cols() %>%
-  mutate(symbol = rownames(pm)) %>%
-left_join(scols1, by = "symbol")
-```
-
 ```{r, scols-peak-position-{{hot_chr}}}
@@
 
-```{r, n-med-{{hot_chr}}}
+```{r, n-med-{{hot_chr}}, eval = FALSE}
 n_med_tr <- 10 # arbitrary threshold for number of mediated traits per local
@@
   knitr::kable(caption = "Number of mediated traits per local trait for Chr {{hot_chr}} hotspot.")
-```
-
-```{r define-scols-pre-{{hot_chr}}}
-scols_pre <- scols_extra %>%
-  dplyr::select(n_med_per_local$local_symbol)
-# get ordering for columns of scols_pre
-col_order <- scols_pre %>%
-  t() %>%
-  dist() %>%
-  hclust() %>%
-  as.dendrogram() %>%
-  dendextend::seriate_dendrogram(dist(t(scols_pre))) %>%
-  order.dendrogram()
-scols <- scols_pre[ , col_order] %>%
-  mutate(local = scols1$local)
 ```
@@
   rename(symbol = symbol2)
+# to add to mediation tibbles
 
+# first med tibble
 lod_diffs_tibble <- med_tib %>%
@@
   dplyr::bind_rows(fake_tibble_mediation)
+# second med tibble
 lod_diff_proportions_tibble <- med_tib %>%
@@
   bind_rows(fake_tibble_mediation)
+# effects tibble
 allele_effects_tibble <- local_nonlocal_annot %>%
   dplyr::select(symbol, 18:25)
+# peak positions tibble
 peak_positions <- binning %>%
   select(symbol, pos)
+# local indicator
 local_tibble <- local_nonlocal_annot %>%
@@
   ungroup()
+# univariate lods
+lod_tibble <- local_nonlocal_annot %>%
+  select(symbol, lod) %>%
+  ungroup()
+```
 
+```{r, order-annotations-{{hot_chr}}}
+# all annots 
+annots_ord <- list(lod_diffs_tibble, 
+                   lod_diff_proportions_tibble, 
+                   allele_effects_tibble,
+                   peak_positions, 
+                   local_tibble,
+                   lod_tibble
+                    ) %>% 
+  purrr::map(.f = function(df, rn){
+    out <- df[match(x = rn, table = df$symbol), ]
+    return(out)
+  }, 
+  rn = rn)
 ```
@@
   apply(FUN = trunc2, MARGIN = 2, threshold = 5) %>%
-  heatmaply::heatmapr(Rowv = FALSE,
+  heatmaply:::heatmaply(Rowv = FALSE,
             Colv = FALSE,
-            plot_method = "plotly"
-            ) %>%
-    heatmaply::heatmaply_heatmapr2(
+            plot_method = "plotly",
+            symm = TRUE,
+            row_side_colors = as.matrix(annots_ord[[3]][, 2:3]),
             fontsize_col = 2,
             fontsize_row = 2,
-            symm = TRUE,
-            row_side_colors = list(scols[ord, ], tibble(scols1$local[ord])),
-            row_side_palette = list(heatmaply::cool_warm, heatmaply::Blues), 
-            col_side_colors = tibble(local = scols1$local[ord]),
-            limits = c(0, 5)
             ) %>%

### 2019-08-10 analysis/paper/Rmd/hotspot-one-chromosome.Rmd
replaced heatmaply call with pheatmap

@@
   apply(FUN = trunc2, MARGIN = 2, threshold = 5) %>%
-  heatmaply:::heatmaply(Rowv = FALSE,
-            Colv = FALSE,
-            plot_method = "plotly",
-            symm = TRUE,
-            row_side_colors = as.matrix(annots_ord[[3]][, 2:3]),
-            fontsize_col = 2,
-            fontsize_row = 2,
-            ) %>%
-  hide_guides()
+  pheatmap::pheatmap(cluster_rows = FALSE,
+                     cluster_cols = FALSE,
+                     fontsize = 2,
+                     annotation_row = as.data.frame(annfoo)
+                     )
 ```

### 2019-08-10 analysis/paper/preliminary/comparing-heatmaps.Rmd
added Rmd file for comparing heatmap functions in R

--- 2019-08-09 analysis/paper/Rmd/hotspots-paper.Rmd
@@
 ---
-title: "Expression QTL hotspot dissection with mediation analysis and pleiotropy testing"
-author: "Frederick J. Boehm, Brian S. Yandell, and Karl W. Broman"
-date: "August 3, 2019"
-bibliography: research.bib
-linkReferences: true
-link-citations: true 
-output: 
-  bookdown::html_document2: 
-    code_folding: hide
-    self_contained: no
-    keep_md: true
-  bookdown::pdf_document2:
-    fig_caption: true
-    keep_tex: true
+title: "Comparing heatmap functions in R"
+author: "Frederick Boehm"
+date: "8/10/2019"
+output: html_document
 ---
 
-```{r child = "abstract.Rmd", eval = TRUE}
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = TRUE)
 ```
 
-```{r child="intro.Rmd", eval = TRUE}
-```
+The goal here is to compare heatmaps that are produced by R functions. I'm particularly excited by interactive heatmaps, like those using D3 javascript.
 
@@
 
-`r if (eval_code) knitr::knit(text = out[[2]])`
+### `iheatmapr` R package
 
-`r if (eval_code) knitr::knit(text = out[[3]])`
-
-`r if (eval_code) knitr::knit(text = out[[4]])`
-
-`r if (eval_code) knitr::knit(text = out[[5]])`
-
-```{r read_results, child = "results.Rmd", eval = TRUE}
-```
-
-```{r read_discussion, child = "discussion.Rmd", eval = TRUE}
-```
-
-# Colophon
-
-This report was generated on ` r Sys.time()` using the following computational environment and dependencies:
-
-```{r colophon, cache = FALSE, echo = TRUE, tidy.opts=list(width.cutoff=50), size = "tiny", results = "markup"}
-# which R packages and versions?
-devtools::session_info()
-```
-
-The current Git commit details are:
-
-```{r git2r, cache = FALSE, echo = TRUE}
-# what commit is this file at?
-git2r::repository(here::here())
-```
+I installed the `iheatmapr` package from Github: `ropensci/iheatmapr`.

### 2019-08-11 analysis/paper/Rmd/hotspot-one-chromosome.Rmd
added iheatmapr heatmaps to hotspots-paper.Rmd

@@
 
-```{r, scols-peak-position-{{hot_chr}}}
-
-binning <- local_nonlocal_annot %>%
-  dplyr::select(symbol, pos, lod) %>%
-  mutate(bin = unlist(pos) %>%
-           map(.f = function(x) {x %>% 
-               bin_peak_position(breaks = hot_lower + (1:9) * 0.4) %>%
-               binarize_bin_number(total_number_of_bins = 10L)
-             })
-  ) %>%
-  listcol_to_cols(list_column = quo(bin), another_column = quo(symbol))
-```
-
 ```{r, scan1-{{hot_chr}}}
@@
   mutate(peak_effects = purrr::map2(.f = qtl2pleio::get_effects, .x = peak_index, .y = s1coef_out, map = map[["{{hot_chr}}"]])) %>%
-  listcol_to_cols(list_column = quo(peak_effects), another_column = quo(symbol))
+  qtl2hotspots::listcol_to_cols(list_column = quo(peak_effects), another_column = quo(symbol))
 ```
 
-```{r, n-med-{{hot_chr}}, eval = FALSE}
+```{r, scols-peak-position-{{hot_chr}}}
+
+binning <- local_nonlocal_annot %>%
+  dplyr::select(symbol, pos, lod) %>%
+  mutate(bin = unlist(pos) %>%
+           map(.f = function(x) {x %>% 
+               bin_peak_position(breaks = hot_lower + (1:9) * 0.4) %>%
+               binarize_bin_number(total_number_of_bins = 10L)
+             })
+  ) %>%
+  listcol_to_cols(list_column = quo(bin), another_column = quo(symbol))
+```
+
+```{r, n-med-{{hot_chr}}}
 n_med_tr <- 10 # arbitrary threshold for number of mediated traits per local
@@
                      cluster_cols = FALSE,
-                     fontsize = 2,
-                     annotation_row = as.data.frame(annfoo)
+                     fontsize = 2
                      )
@@
 
-ggplotly(p)
+plotly::ggplotly(p)
 ```
+
+```{r iheatmapr-{{hot_chr}}}
+ann3 <- annots_ord[[3]] %>%
+  ungroup() %>%
+  select(- symbol) %>%
+  as.matrix() %>%
+  (function(x){rownames(x) <- annots_ord[[3]]$symbol; return(x)})
+colnames(ann3) <- LETTERS[1:8]
+ann2 <- annots_ord[[2]] %>%
+  select(-symbol) %>%
+  as.matrix() %>%
+  (function(x){rownames(x) <- annots_ord[[2]]$symbol; return(x)})
+ann1 <- annots_ord[[1]] %>%
+  select(-symbol) %>%
+  as.matrix() %>%
+  (function(x){rownames(x) <- annots_ord[[1]]$symbol; return(x)})
+ann5 <- annots_ord[[5]] %>%
+  select(-symbol) %>%
+  as.matrix() %>%
+  (function(x){rownames(x) <- annots_ord[[5]]$symbol; return(x)})
+ann6 <- annots_ord[[6]] %>%
+  select(-symbol) %>%
+  as.matrix() %>%
+  (function(x){rownames(x) <- annots_ord[[6]]$symbol; return(x)})
+ann4 <- annots_ord[[4]] %>%
+  ungroup() %>%
+  select(-symbol) %>%
+  as.matrix() %>%
+  (function(x){rownames(x) <- annots_ord[[4]]$symbol; return(x)})
+mymat <- pm[ord, ord] %>% # order both cols and rows by 'ord'
+  apply(FUN = trunc2, MARGIN = 2, threshold = 5)
+iheatmapr::main_heatmap(mymat, show_colorbar = FALSE, layout = list(font = list(size = 2))) %>%
+  iheatmapr::add_col_labels() %>%
+  iheatmapr::add_row_labels() %>%
+  iheatmapr::add_col_annotation(ann5) %>%
+  iheatmapr::add_main_heatmap(ann3, name = "signal3", size = 0.2, show_colorbar = FALSE) %>%
+  iheatmapr::add_main_heatmap(ann2, name = "signal2", size = 0.5, show_colorbar = FALSE) %>%
+  iheatmapr::add_main_heatmap(ann1, name = "signal1", size = 0.5, show_colorbar = FALSE) %>%
+  iheatmapr::add_row_annotation(ann4) %>%
+  iheatmapr::add_row_annotation(ann5, side = "left") %>%
+  iheatmapr::add_row_annotation(ann6, side = "left")
+
+```

### 2019-08-11 analysis/paper/preliminary/comparing-heatmaps.Rmd
added iheatmapr heatmaps to hotspots-paper.Rmd

@@
 
+# Goals
+
 The goal here is to compare heatmaps that are produced by R functions. I'm particularly excited by interactive heatmaps, like those using D3 javascript.
+
+It's particularly important that I be able to easily annotate, with multiple objects, the rows (and/or columns).
 
@@
 
-```{r child="methods.Rmd", eval = TRUE}
-```
-
 ```{r setup}
@@
 options(tibble.print_max = 50, tibble.print_min = 20)
-```
-
-```{r load_heatmaply}
-library(heatmaply)
-```
-
-```{r saveRDS}
-if (!file.exists(here("analysis", "data", "derived_data", "expr.rds"))){
-  download.file("https://datadryad.org/bitstream/handle/10255/dryad.166774/Attie_DO378_eQTL_viewer_v1.Rdata?sequence=2",
-              destfile = here("analysis", "data", "raw_data", "Attie_DO378_eQTL_viewer_v1.Rdata")
-              )
-  load(here("analysis", "data", "raw_data", "Attie_DO378_eQTL_viewer_v1.Rdata"))
-  saveRDS(dataset.islet.rnaseq$lod.peaks, here("analysis", "data", "derived_data", "lod_peaks.rds"))
-  saveRDS(dataset.islet.rnaseq$annots, here("analysis", "data", "derived_data", "annots.rds"))
-  saveRDS(dataset.islet.rnaseq$expr, here("analysis", "data", "derived_data", "expr.rds"))
-  saveRDS(K, here("analysis", "data", "derived_data", "kinship.rds"))
-  saveRDS(map, here("analysis", "data", "derived_data", "map.rds"))
-  saveRDS(genoprobs, here("analysis", "data", "derived_data", "genoprobs.rds"))
-  saveRDS(dataset.islet.rnaseq$covar, here("analysis", "data", "derived_data", "covar.rds"))
-}
 ```
@@
 
-`r if (eval_code) knitr::knit(text = out[[1]])`
+`r knitr::knit(text = out[[1]])`
 
@@
 I installed the `iheatmapr` package from Github: `ropensci/iheatmapr`.
+
+```{r, eval = FALSE}
+devtools::install_github("ropensci/iheatmapr")
+```
+
+```{r iheatmapr-{{hot_chr}}}
+ann3 <- annots_ord[[3]] %>%
+  ungroup() %>%
+  select(- symbol) %>%
+  as.matrix() %>%
+  (function(x){rownames(x) <- annots_ord[[3]]$symbol; return(x)})
+colnames(ann3) <- LETTERS[1:8]
+ann2 <- annots_ord[[2]] %>%
+  select(-symbol) %>%
+  as.matrix() %>%
+  (function(x){rownames(x) <- annots_ord[[2]]$symbol; return(x)})
+ann1 <- annots_ord[[1]] %>%
+  select(-symbol) %>%
+  as.matrix() %>%
+  (function(x){rownames(x) <- annots_ord[[1]]$symbol; return(x)})
+ann5 <- annots_ord[[5]] %>%
+  select(-symbol) %>%
+  as.matrix() %>%
+  (function(x){rownames(x) <- annots_ord[[5]]$symbol; return(x)})
+ann6 <- annots_ord[[6]] %>%
+  select(-symbol) %>%
+  as.matrix() %>%
+  (function(x){rownames(x) <- annots_ord[[6]]$symbol; return(x)})
+ann4 <- annots_ord[[4]] %>%
+  ungroup() %>%
+  select(-symbol) %>%
+  as.matrix() %>%
+  (function(x){rownames(x) <- annots_ord[[4]]$symbol; return(x)})
+mymat <- pm[ord, ord] %>% # order both cols and rows by 'ord'
+  apply(FUN = trunc2, MARGIN = 2, threshold = 5)
+iheatmapr::main_heatmap(mymat, show_colorbar = FALSE, layout = list(font = list(size = 2))) %>%
+  iheatmapr::add_col_labels() %>%
+  iheatmapr::add_row_labels() %>%
+  iheatmapr::add_col_annotation(ann5) %>%
+  iheatmapr::add_main_heatmap(ann3, name = "signal3", size = 0.2, show_colorbar = FALSE) %>%
+  iheatmapr::add_main_heatmap(ann2, name = "signal2", size = 0.5, show_colorbar = FALSE) %>%
+  iheatmapr::add_main_heatmap(ann1, name = "signal1", size = 0.5, show_colorbar = FALSE) %>%
+  iheatmapr::add_row_annotation(ann4) %>%
+  iheatmapr::add_row_annotation(ann5, side = "left") %>%
+  iheatmapr::add_row_annotation(ann6, side = "left")
+
+```
+
+By using distinct `name` values in the calls to `main_heatmap`, I can get the two heatmaps together and with different color schemes.
+
+```{r sess_info}
+devtools::session_info()
+```

### 2019-08-11 analysis/paper/Rmd/hotspots-paper-2.Rmd
R coding revisions
ignored accessory files

@@
     code_folding: hide
+    self_contained: no
+    keep_md: true
+  bookdown::pdf_document2:
+    fig_caption: true
+    keep_tex: true
 ---
@@
 
-```{r load_heatmaply}
-library(heatmaply)
-```
-
 ```{r saveRDS}
@@
 probs <- readRDS(here("analysis", "data", "derived_data", "genoprobs.rds"))
-```
-
-```{r, load_purrr}
-library(purrr)
+K <- readRDS(here("analysis", "data", "derived_data", "kinship.rds"))
+covar <- readRDS(here("analysis", "data", "derived_data", "covar.rds"))
+map <- readRDS(here("analysis", "data", "derived_data", "map.rds"))
 ```

### 2019-08-11 analysis/paper/Rmd/hotspot-one-chromosome.Rmd
increased out.width for iheatmapr chunk output

@@
 
-```{r iheatmapr-{{hot_chr}}}
+```{r iheatmapr-{{hot_chr}}, out.width="20in"}
 ann3 <- annots_ord[[3]] %>%

### 2019-08-12 analysis/paper/Rmd/hotspot-one-chromosome.Rmd
fixed column & row labels

--- 2019-08-11 analysis/paper/Rmd/hotspot-one-chromosome.Rmd
@@
   iheatmapr::add_col_labels() %>%
-  iheatmapr::add_row_labels() %>%
   iheatmapr::add_col_annotation(ann5) %>%
   iheatmapr::add_main_heatmap(ann3, name = "signal3", size = 0.2, show_colorbar = FALSE) %>%
+    iheatmapr::add_col_labels() %>%
+
   iheatmapr::add_main_heatmap(ann2, name = "signal2", size = 0.5, show_colorbar = FALSE) %>%
+  iheatmapr::add_col_labels() %>%
   iheatmapr::add_main_heatmap(ann1, name = "signal1", size = 0.5, show_colorbar = FALSE) %>%
-  iheatmapr::add_row_annotation(ann4) %>%
+  iheatmapr::add_row_signal(signal = as.vector(ann4), name = "foo1", show_colorbar = FALSE) %>%
   iheatmapr::add_row_annotation(ann5, side = "left") %>%
-  iheatmapr::add_row_annotation(ann6, side = "left")
+    iheatmapr::add_col_labels() %>%
+
+  iheatmapr::add_row_annotation(ann6, side = "left") %>%
+    iheatmapr::add_row_labels()
 

### 2019-08-12 analysis/paper/Rmd/methods.Rmd
revised first subsection of Methods

@@
 
-We used the publicly available data from @keller2018genetic to demonstrate our strategy. @keller2018genetic
-measured genome-wide gene expression levels with RNA sequencing of pancreatic islets in 378 Diversity Outbred mice.
-From the RNA sequencing counts, they obtained normalized gene expression levels for 21,771 transcripts, which they shared publicly via DataDryad [@dryad].
-@keller2018genetic also included founder allele probabilities at 69,005 markers across the genome, covariates, and leave-one-chromosome-out estimated kinship matrices.
-@keller2018genetic calculated leave-one-chromosome-out kinship matrices for every chromosome [@broman2019rqtl2; @yang2014advantages].
+We used the publicly available data from @keller2018genetic to demonstrate our approach for 
+expression trait hotspot dissection. @keller2018genetic measured genome-wide gene expression 
+levels with RNA sequencing of pancreatic islets in 378 
+Diversity Outbred mice. From the RNA sequencing counts, they obtained normalized gene 
+expression levels for 21,771 transcripts, which they shared publicly via DataDryad [@dryad]. 
+@keller2018genetic also included founder allele probabilities at 69,005 markers across the 
+genome, covariates, and leave-one-chromosome-out estimated kinship matrices [@broman2019rqtl2; @yang2014advantages].
 
-@keller2018genetic performed univariate QTL mapping for all expression traits. 
+@keller2018genetic performed univariate QTL mapping for all expression traits. They
+defined an expression trait hotspot to be a 4-Mb region that affects at least 100 nonlocal 
+expression traits.
 They identified five expression trait hotspots in their
-study of pancreatic islet gene expression in 378 Diversity Outbred mice. They
-defined a hotspot to be a 4-Mb region that affects at least 100 nonlocal expression
-traits.
+study of pancreatic islet gene expression in 378 Diversity Outbred mice .
 
-For each hotspot, we identified a set of local expression traits and a
-set of nonlocal expression traits. We focused on those local 
+In analyzing the data from @keller2018genetic, we first established criteria for identifying local and nonlocal transcripts at each of the five hotspots. We focused on those local 
 traits that strongly associated with QTL within the hotspot. We set the univariate LOD threshold for 
@@
 on other chromosomes, while the local traits reflected expression levels for
-genes on the same chromosome as the hotspot . We ignored transcripts for genes on the hotspot chromosome 
+genes on the same chromosome as the hotspot. We ignored transcripts for genes on the hotspot chromosome 
 but outside the hotspot interval.
+
+For each hotspot, we performed both pairwise pleiotropy tests and linear regression-based 
+mediation analyses. For every pair of traits at a hotspot, we calculated a pleiotropy test 
+statistic[@boehm2019testing]. Due to the computational burden, we didn't determine bootstrap p-values for these 
+tests. For every local trait - nonlocal trait pair, we also performed a linear 
+regression-based mediation analysis [@chick2016defining,@keller2018genetic]. We describe in detail both procedures below.
 

### 2019-08-12 analysis/paper/Rmd/methods.Rmd
added details of annotation info for heatmaps

@@
 They identified five expression trait hotspots in their
-study of pancreatic islet gene expression in 378 Diversity Outbred mice .
+study of pancreatic islet gene expression in 378 Diversity Outbred mice.
 
@@
 We created scatter plots with the `ggplot2` package [@ggplot2] for the R
-statistical computing environment [@r].
-We plotted all local-nonlocal pairs for each hotspot.
-We highlighted in blue those points that correspond to pairs involving the putative mediator that @keller2018genetic identified.
+statistical computing environment [@r]. We plotted pleiotropy test statistics against 
+mediation LOD differences for all local-nonlocal pairs at each hotspot. We highlighted in blue 
+those points that correspond to pairs involving the putative mediator that @keller2018genetic 
+identified. We used the same results in plots of pleiotropy test statistics against mediation LOD difference for analyses involving a single local trait.
 
@@
 To create a heatmap, we first performed two hierarchical clusterings of rows. 
-One hierarchical clustering involved only the rows and columns containing local traits, while the other contained only 
-rows and columns containing nonlocal traits.
+One hierarchical clustering involved only the rows containing local traits, while the other contained only 
+rows corresponding to nonlocal traits.
 We calculated the Euclidean distances between all row pairs (in the respective 
-submatrices) before clustering. Hierarchical clustering was performed using the complete 
-linkage method. Initially, each row is assigned to its own cluster. An iterative 
+submatrices) before clustering. Hierarchical clustering was performed using the iterative complete 
+linkage method. Initially, each row is assigned to its own cluster. The iterative 
 algorithm then combines the two most similar clusters at each step. Distances between 
@@
 To achieve coloring contrasts in the heatmaps, we
-set the color spectrum limits at 0 (dark blue) and 5 (yellow). Pleiotropy test statistics above 5, too, are 
-represented by
-yellow. We annotated rows by a collection of columns on the right-hand side of the heatmap. This binary matrix 
-consists of columns that summarize mediation results. Each column represents a local trait. If a row 
-corresponds to a nonlocal trait that is mediated by the local trait (*i.e.*, with mediation LOD 
-difference greater than 1.5), then the cell is colored dark green; 
-otherwise, the cell is light green. We didn't examine the possibility that a local trait might 
-mediate the relationship between a QTL and another local trait; thus, all local trait rows are 
-colored light green. The last annotation column and the annotation row above the heatmap both 
-designate local traits with dark blue and nonlocal traits with light blue.
+set the color spectrum limits at 0 and 5. 
+We annotated rows with six pieces of information: 1. mediation LOD differences, 2. mediation 
+LOD difference proportions, 3. univariate QTL peak positions, 4. estimated allele effects 
+vector at the univariate peak position, 5. univariate LOD score, and 6. whether the row 
+corresponds to a local trait. For clarity of presentation, we also indicated those columns that  
+correspond to a local traits. We didn't examine the possibility that a local trait might 
+mediate the relationship between a QTL and another local trait.

### 2019-08-12 analysis/paper/Rmd/hotspot-one-chromosome.Rmd
fixed conditional statement for downloading from datadryad

@@
 
-```{r iheatmapr-{{hot_chr}}, out.width="20in"}
+```{r iheatmapr-{{hot_chr}}, out.width="25in", out.height = "10in"}
 ann3 <- annots_ord[[3]] %>%

### 2019-08-12 analysis/paper/Rmd/hotspots-paper.Rmd
added iheatmapr heatmaps to hotspots-paper.Rmd
fixed conditional statement for downloading from datadryad

@@
 
-```{r load_heatmaply}
-library(heatmaply)
-```
-
 ```{r saveRDS}
-if (!file.exists(here("analysis", "data", "derived_data", "expr.rds"))){
+if (!file.exists(here("analysis", "data", "derived_data", "genoprobs.rds"))){
   download.file("https://datadryad.org/bitstream/handle/10255/dryad.166774/Attie_DO378_eQTL_viewer_v1.Rdata?sequence=2",

### 2019-08-13 analysis/paper/Rmd/hotspot-one-chromosome.Rmd
fixed interactive heatmaps

--- 2019-08-11 analysis/paper/Rmd/hotspot-one-chromosome.Rmd
@@
   apply(FUN = trunc2, MARGIN = 2, threshold = 5)
-iheatmapr::main_heatmap(mymat, show_colorbar = FALSE, layout = list(font = list(size = 2))) %>%
+iheatmapr::main_heatmap(mymat, show_colorbar = FALSE, layout = list(font = list(size = 4))) %>%
   iheatmapr::add_col_labels() %>%
-  iheatmapr::add_row_labels() %>%
   iheatmapr::add_col_annotation(ann5) %>%
   iheatmapr::add_main_heatmap(ann3, name = "signal3", size = 0.2, show_colorbar = FALSE) %>%
+    iheatmapr::add_col_labels() %>%
+
   iheatmapr::add_main_heatmap(ann2, name = "signal2", size = 0.5, show_colorbar = FALSE) %>%
+  iheatmapr::add_col_labels() %>%
   iheatmapr::add_main_heatmap(ann1, name = "signal1", size = 0.5, show_colorbar = FALSE) %>%
-  iheatmapr::add_row_annotation(ann4) %>%
   iheatmapr::add_row_annotation(ann5, side = "left") %>%
-  iheatmapr::add_row_annotation(ann6, side = "left")
+    iheatmapr::add_col_labels() %>%
+  iheatmapr::add_row_annotation(ann4, side = "left") %>%
+  iheatmapr::add_col_labels() %>%
+  iheatmapr::add_row_annotation(ann6, side = "left") %>%
+    iheatmapr::add_row_labels()
 

### 2019-08-13 analysis/paper/Rmd/hotspots-paper.Rmd
switched to eval = TRUE for git2r call
removed dependencies on git2r (and on libgit2 library)
fixed inline code

@@
 
-This report was generated on ` r Sys.time()` using the following computational environment and dependencies:
+This report was generated on `r Sys.time()` using the following computational environment and dependencies:
 
@@
 
-```{r git2r, cache = FALSE, echo = TRUE}
+```{r git2r, cache = FALSE, echo = TRUE, eval = FALSE}
 # what commit is this file at?

### 2019-08-13 analysis/paper/Rmd/hotspot-one-chromosome.Rmd
switched to eval = TRUE for git2r code

@@
   iheatmapr::add_row_annotation(ann4, side = "left") %>%
-  iheatmapr::add_col_labels() %>%
   iheatmapr::add_row_annotation(ann6, side = "left") %>%

### 2019-08-15 tests/testthat/test-bin_peak_position.R
fixed test files

@@ -1,4 +1,4 @@
-library(hotspots)
+library(qtl2hotspots)
 library(testthat)
 
 

### 2019-08-15 tests/testthat/test-tibble_to_matrix.R

@@ -1,4 +1,4 @@
-library(hotspots)
+library(qtl2hotspots)
 library(testthat)
 
 foo <- tibble::tibble(v1 = letters[c(1, 1, 2, 2)], v2 = letters[c(3, 4, 3, 4)], lrt = runif(4))

### 2019-08-18 R/listcol_to_cols.R
fixed listcol_to_cols function

@@ -9,8 +9,11 @@
 #' listcol_to_cols(list_column = dplyr::quo(b), another_column = dplyr::quo(a))
 #' @export
 listcol_to_cols <- function(tibble, list_column, another_column){
+  if (!rlang::is_quosure(list_column) | !rlang::is_quosure(another_column)){
+    stop("Both `list_column` and `another_column` arguments must be quosures")
+    }
   tibble %>%
-    tidyr::unnest() %>%
+    tidyr::unnest(!! list_column, .drop = FALSE) %>%
     dplyr::group_by(!! another_column) %>%
     dplyr::mutate(col=seq_along(!! another_column)) %>%
     tidyr::spread(key = col, value = !! list_column)

### 2019-08-18 tests/testthat/test-listcol_to_cols.R
added tests for listcol_to_cols

@@ -1,0 +1,19 @@
+library(qtl2hotspots)
+library(testthat)
+
+myfoo <- tibble::tibble(num = 1:4, numlist = list(1:10, 11:20, 21:30, 31:40))
+myout <- listcol_to_cols(tibble = myfoo,
+                         list_column = dplyr::quo(numlist),
+                         another_column = dplyr::quo(num)
+                         )
+
+
+test_that("correct dimensions & rownames in simple cases", {
+  expect_equal(dim(myout), c(4, 11))
+  expect_equal(colnames(myout)[1], "num")
+  expect_equal(colnames(myout)[2:11], as.character(1:10))
+})
+
+test_that("non-quosures give errors", {
+  expect_error(listcol_to_cols(myfoo, num, numlist))
+})

### 2019-08-18 analysis/paper/preliminary/allele-effects-exploration.Rmd
added file for exploration of allele effects contrasts

---
title: "Founder allele effects for hotspot traits"
author: "Frederick Boehm"
date: "8/14/2019"
output: html_document
---

## Goal

The goals here are to follow up on Karl's suggestion to examine contrasts among allele effects. Specifically, can we see patterns of founder alleles near each trait's univariate peak.

The heatmap of allele effects itself doesn't really reveal patterns that are easily interpreted. Hence, the suggestion to examine contrasts. Specifically, for each diagonal block of traits, is there a single, informative contrast? (or, perhaps a set of contrasts for each diagonal block.)

## Load the needed objects

Here, we reuse code from ../Rmd/hotspots-paper.Rmd.

```{r presetup}
eval_code <- TRUE
knitr::opts_chunk$set(comment=NA, fig.width=6, fig.height=6, eval = eval_code, echo = TRUE)
```

```{r define}
library(qtl2hotspots)

ult <- 15 
# univariate lod threshold
```

```{r load_tidyverse}
library(tidyverse)
options(tibble.print_max = 50, tibble.print_min = 20)
```

```{r readRDS}
lod_peaks <- readRDS(here::here("analysis", "data", "derived_data", "lod_peaks.rds"))
annots <- readRDS(here::here("analysis", "data", "derived_data", "annots.rds"))
expr <- readRDS(here::here("analysis", "data", "derived_data", "expr.rds"))
probs <- readRDS(here::here("analysis", "data", "derived_data", "genoprobs.rds"))
K <- readRDS(here::here("analysis", "data", "derived_data", "kinship.rds"))
covar <- readRDS(here::here("analysis", "data", "derived_data", "covar.rds"))
map <- readRDS(here::here("analysis", "data", "derived_data", "map.rds"))
```

```{r setup_continued}
hot_chr <- c(2, 5, 7, 11, 13)
hot_mid <- c(165.5, 146, 46, 71, 112.5)
keller_mediator <- c("Hnf4a", "Pdx1", "Fam83e", "Sat2", "Il6st")
inputs <- tibble(hot_chr, hot_mid, keller_mediator)
out <- pmap_dfc(inputs, .f = function(hot_chr, hot_mid, keller_mediator) knitr::knit_expand(text = readLines(here::here("analysis", "paper", "Rmd", "hotspot-one-chromosome.Rmd"))))
# note that ... argument must be a list, ie, it must be a single argument, not multiple args
```

`r if (eval_code) knitr::knit(text = out[[2]])`

```{r, define-submap}
figdir <- here::here("analysis", "figures", "Chr5-allele-effects-plots")
names(s1coef_out) -> ids
# find indices on Chr5 for boundaries of hotspot
which(map$`5` > hot_lower)[1] -> start_index
which(map$`5` > hot_upper)[1] -> stop_index
submap <- list()
submap$`5` <- map$`5`[start_index:stop_index]
mylist <- list()
for (i in 1:203){
  mylist[[i]] <- s1out[ , i, drop = FALSE]
}
```

```{r effects-plots}
purrr::pmap(.l = list(s1coef_out, as.list(ids), mylist), 
           .f = function(x, y, z){
             main <- local_nonlocal_annot %>%
               ungroup() %>%
               filter(annot.id == y) %>%
               select(annot.id, symbol) %>%
               paste0()
             qtl2::plot_coefCC(x, map = submap, main = main, scan1_output = z)
           }
)
```

```{r blup-plots}
purrr::pmap(.l = list(s1coef_blup, as.list(ids), mylist), 
           .f = function(x, y, z){
             main <- local_nonlocal_annot %>%
               ungroup() %>%
               filter(annot.id == y) %>%
               select(annot.id, symbol) %>%
               paste0()
             qtl2::plot_coefCC(x, map = submap, main = main, scan1_output = z)
           }
)
```

## Barplots for effects at peak positions

Now, let's focus only on the 8 values at the LOD peak for each trait.

```{r, barplots-scan1coef-effects, fig.height = 20, fig.width = 8}
ord2 <- match(table = local_nonlocal_annot$symbol, x = rn)
foobar <- local_nonlocal_annot %>%
  select(symbol, cA:cH) %>%
  (function(x) x[ord2, ]) %>%
  tidyr::gather(key = "founder_allele", value = "effect", cA: cH)
foobar2 <- arrange(transform(foobar,
             symbol=factor(symbol,levels=rn)),symbol)
foobar2 %>% 
  ggplot() + geom_col(aes(y = effect, x = founder_allele, fill = founder_allele)) +
  ylim(c(-1.5, 1.5)) + 
  facet_wrap(symbol ~ ., ncol = 4) #+ theme(strip.text.x = element_blank())
ggsave(here::here("analysis", "figures", "allele_blupss_at_peaks.png"), limitsize = FALSE)
```

## Bar plots for blup effects

```{r, barplots-blup, fig.height = 20, fig.width = 8}
foobar <- local_nonlocal_annot %>%
  ungroup() %>%
  (function(x) x[ord2, ]) %>%
  select(symbol, 26:33) %>%
  tidyr::gather(paste0("b", LETTERS[1:8]), key = "founder_allele", value = "blup_effect")
foobar2 <- arrange(transform(foobar,
             symbol=factor(symbol,levels=rn)),symbol)
foobar2 %>% 
  ggplot() + geom_col(aes(y = effect, x = founder_allele, fill = founder_allele)) +
  ylim(c(-1.5, 1.5)) + 
  facet_wrap(symbol ~ ., ncol = 4) #+ theme(strip.text.x = element_blank())
ggsave(here::here("analysis", "figures", "allele_effects_at_peaks.png"), limitsize = FALSE)
```

```{r correlations}
local_nonlocal_annot %>%
  ungroup() %>%
# 
  select(18:25) %>% 
  as.matrix() %>%
  apply(FUN = as.numeric, MARGIN = 2) %>%
  (function(x){rownames(x) <- local_nonlocal_annot$symbol; return(x)}) %>%
  (function(x) x[ord2, ]) %>%
  t() %>%
  cor()
which(rn == "Igfbp3")
```

```{r BC-v-FG}
local_nonlocal_annot %>%
  ungroup() %>%
  (function(x) x[ord2, ]) %>%
  select(symbol, 18:25) %>%
    rename("A" = `1`, "B" = `2`, "C" = `3`, "D" = `4`, "E" = `5`, "F" = `6`, G = `7`, H = `8`
         ) %>%
  mutate(contrast_BC_FG = abs((B + C) - (F + G)), row_num = 1:203) %>%
  ggplot() + geom_point(aes(y = contrast_BC_FG, x = row_num))
```

```{r BC-v-G}
p <- local_nonlocal_annot %>%
  ungroup() %>%
  (function(x) x[ord2, ]) %>%
  select(symbol, 18:25) %>%
    rename("A" = `1`, "B" = `2`, "C" = `3`, "D" = `4`, "E" = `5`, "F" = `6`, G = `7`, H = `8`
         ) %>%
  mutate(contrast = abs((B + C) / 2 - G), row_num = 1:203) %>%
  ggplot() + geom_point(aes(y = contrast, x = row_num)) + geom_smooth(aes(y = contrast, x = row_num))
plotly::ggplotly(p)
```

```{r BCGH:ADEF}
p <- local_nonlocal_annot %>%
  ungroup() %>%
  select(symbol, cA:cH) %>%
  (function(x) x[ord2, ]) %>%
  mutate(contrast = abs((cB + cC + cG + cH) - (cA + cD + cE + cF)), row_num = 1:203) %>%
  ggplot() + geom_point(aes(y = contrast, x = row_num)) + geom_smooth(aes(y = contrast, x = row_num))
plotly::ggplotly(p)

```

```{r BCD:F}
p <- local_nonlocal_annot %>%
  ungroup() %>%
  select(symbol, cA:cH, local_tx) %>%
  (function(x) x[ord2, ]) %>%
  mutate(contrast = abs((cB + cC + cD) / 3 - cF), row_num = 1:203) %>%
  ggplot() + geom_point(aes(y = contrast, x = row_num, colour = local_tx, symbol = symbol)) + geom_smooth(aes(y = contrast, x = row_num))
plotly::ggplotly(p)
```

```{r BCGH:ADEF}
med_small <- med_tib %>% 
  filter(local_symbol == "Rpl21")
p <- local_nonlocal_annot %>%
  ungroup() %>%
  select(symbol, cA:cH, local_tx) %>%
  (function(x) x[ord2, ]) %>%
  inner_join(med_small, by = c("symbol" = "nonlocal_symbol")) %>%
  mutate(contrast = abs((cB + cC + cG + cH) - (cA + cD + cE + cF))) %>%
  ggplot() + geom_point(aes(y = contrast, x = lod_diff_proportion)) + geom_smooth(aes(y = contrast, x = lod_diff_proportion))
plotly::ggplotly(p)

```
Cxcl12 shows a strong contrast between B & G.

```{r B:G}
med_small <- med_tib %>% 
  filter(local_symbol == "Rpl21")
p <- local_nonlocal_annot %>%
  ungroup() %>%
  select(symbol, cA:cH, local_tx) %>%
  (function(x) x[ord2, ]) %>%
  inner_join(med_small, by = c("symbol" = "nonlocal_symbol")) %>%
  mutate(contrast = abs(cB - cG)) %>%
  ggplot() + geom_point(aes(y = contrast, x = lod_diff_proportion)) + geom_smooth(aes(y = contrast, x = lod_diff_proportion))
plotly::ggplotly(p)

```

```{r B:G}
p <- local_nonlocal_annot %>%
  ungroup() %>%
  select(symbol, cA:cH, local_tx) %>%
  (function(x) x[ord2, ]) %>%
  mutate(contrast = abs(cB - cG), row_num = 1:203) %>%
  ggplot() + geom_point(aes(y = contrast, x = row_num, colour = local_tx, symbol = symbol)) + geom_smooth(aes(y = contrast, x = row_num))
plotly::ggplotly(p)
```

## Calculate all pairwise contrasts for main effects

```{r}
contrasts <- local_nonlocal_annot %>%
  ungroup() %>%
  select(symbol, cA:cH, local_tx, lod) %>%
  (function(x) x[ord2, ]) %>%
  mutate(row_num = 1:203, 
         AB = abs(cA - cB),
         AC = abs(cA - cC),
         AD = abs(cA - cD),
         AE = abs(cA - cE),
         AF = abs(cA - cF),
         AG = abs(cA - cG),
         AH = abs(cA - cH),
         BC = abs(cB - cC),
         BD = abs(cB - cD),
         BE = abs(cB - cE),
         BF = abs(cB - cF),
         BG = abs(cB - cG),
         BH = abs(cB - cH),
         CD = abs(cC - cD),
         CE = abs(cC - cE),
         CF = abs(cC - cF),
         CG = abs(cC - cG),
         CH = abs(cC - cH),
         DE = abs(cD - cE),
         DF = abs(cD - cF),
         DG = abs(cD - cG),
         DH = abs(cD - cH),
         EF = abs(cE - cF),
         EG = abs(cE - cG),
         EH = abs(cE - cH),
         FG = abs(cF - cG),
         FH = abs(cF - cH),
         GH = abs(cG - cH)
  ) 
ctr <- contrasts %>% 
  tidyr::gather(contrast, value, AB:GH) %>%
  group_by(symbol) %>%
  filter(value == max(value)) %>%
  arrange(row_num) %>%
  ungroup() %>%
  select(contrast)
ctr %>%
  unlist() %>% 
  table()
ctr_runs <- ctr %>%
  unlist() %>% 
  rle()
con2 <- contrasts %>%
  mutate(BC_A = abs((cB + cC) / 2 - cA),
         BC_D = abs((cB + cC) / 2 - cD),
         BC_E = abs((cB + cC) / 2 - cE),
         BC_F = abs((cB + cC) / 2 - cF),
         BC_G = abs((cB + cC) / 2 - cG),
         BC_H = abs((cB + cC) / 2 - cH),
         BC_FG = abs((cB + cC) - (cF + cG))
         )
  con2 %>% 
    ggplot() + geom_point(aes(x = row_num, y = BC_E, colour = local_tx)) + geom_smooth(aes(x = row_num, y = BC_E))
```

```{r}
# which pairwise contrast has the greatest variability?
contrasts %>%
  select(local_tx, 13:40) %>% #select only the contrast column
  filter(!local_tx) %>%
  select(- local_tx) %>%
  purrr::map(.f = function(x)var(x)) %>% unlist()
```

```{r}
con2 %>% 
  filter(!local_tx) %>%
    ggplot() + geom_point(aes(x = row_num, y = BC_FG, colour = local_tx)) + geom_smooth(aes(x = row_num, y = BC_FG))
```

Now, add vertical lines that correspond to partitions between diagonal "blocks" in the pleiotropy statistics heatmap.

The lower left block ends just before transcript "Elmo1". That is, "Elmo1" is the first component of the second block. What is the row_num value of the entry that precedes "Elmo1"?

```{r}
con2 %>%
  filter(symbol == "Elmo1")
# 68 is the row_num value for "Elmo1". Let's draw a vertical line at 67.5
con2 %>%
  filter(symbol == "Cep68") # row_num 129, so draw a vertical line at 129.5

```

```{r}
con2 %>% 
  filter(!local_tx) %>%
    ggplot() + geom_point(aes(x = row_num, y = BC_FG, colour = local_tx)) + geom_smooth(aes(x = row_num, y = BC_FG)) + geom_vline(aes(xintercept = 67.5)) + geom_vline(aes(xintercept = 129.5))
```

```{r}
con2 %>%
  filter(!local_tx) %>%
  mutate(small = row_num < 67.5, big = row_num > 129.5) %>%
  ggplot() + geom_histogram(aes(x = BC_FG)) + facet_wrap(~ big + small)
```

```{r}
p <- contrasts %>%
  select(1, row_num, local_tx, 13:40) %>%
  tidyr::gather(key = "contrast", value = "value", AB:GH) %>%
  filter(!local_tx) %>%
  ggplot() + geom_point(aes(x = row_num, y = value, colour = symbol)) + facet_wrap(contrast ~ .) + theme(legend.position = "none")
plotly::ggplotly(p)
```

```{r}
  con2 %>% 
    ggplot() + geom_point(aes(x = row_num, y = BC_G, colour = local_tx)) + geom_smooth(aes(x = row_num, y = BC_G))

```

```{r}
  con2 %>% 
    ggplot() + geom_point(aes(x = row_num, y = BC_F, colour = local_tx)) + geom_smooth(aes(x = row_num, y = BC_F))
```

```{r}
  con2 %>% 
    ggplot() + geom_point(aes(x = row_num, y = BC_H, colour = local_tx)) + geom_smooth(aes(x = row_num, y = BC_H))
```

```{r}
  con2 %>% 
    ggplot() + geom_point(aes(x = row_num, y = BC_D, colour = local_tx)) + geom_smooth(aes(x = row_num, y = BC_D))
```

```{r}
  con2 %>% 
    ggplot() + geom_point(aes(x = row_num, y = BC_A, colour = local_tx)) + geom_smooth(aes(x = row_num, y = BC_A))
```

### 2019-08-18 analysis/paper/Rmd/hotspot-one-chromosome.Rmd
removed code for working with blups

@@
 ```{r, readRDS-{{hot_chr}}}
-pleio_mixed_list <- readRDS(here("analysis", "data", "derived_data", paste0("Chr", {{hot_chr}}, "_pleio.rds")))
-pleio_local_list <- readRDS(here("analysis", "data", "derived_data", paste0("Chr", {{hot_chr}}, "_pleio_local.rds")))
-pleio_nonlocal_list <- readRDS(here("analysis", "data", "derived_data", paste0("Chr", {{hot_chr}}, "_pleio_nonlocal.rds")))
-med_list <- readRDS(here("analysis", "data", "derived_data", paste0("Chr", {{hot_chr}}, "_mediations.rds")))
+pleio_mixed_list <- readRDS(here::here("analysis", "data", "derived_data", paste0("Chr", {{hot_chr}}, "_pleio.rds")))
+pleio_local_list <- readRDS(here::here("analysis", "data", "derived_data", paste0("Chr", {{hot_chr}}, "_pleio_local.rds")))
+pleio_nonlocal_list <- readRDS(here::here("analysis", "data", "derived_data", paste0("Chr", {{hot_chr}}, "_pleio_nonlocal.rds")))
+med_list <- readRDS(here::here("analysis", "data", "derived_data", paste0("Chr", {{hot_chr}}, "_mediations.rds")))
 ```
@@
   mutate(peak_effects = purrr::map2(.f = qtl2pleio::get_effects, .x = peak_index, .y = s1coef_out, map = map[["{{hot_chr}}"]])) %>%
-  qtl2hotspots::listcol_to_cols(list_column = quo(peak_effects), another_column = quo(symbol))
+  qtl2hotspots::listcol_to_cols(list_column = quo(peak_effects), another_column = quo(symbol)) %>%
+  rename(cA = `1`, cB = `2`, cC = `3`, cD = `4`, cE = `5`, cF = `6`, cG = `7`, cH = `8`
+         ) %>%
+```
+
+```{r, define-contrasts-{{hot_chr}}}
+contrasts_BC_FG <- local_nonlocal_annot %>%
+  ungroup() %>%
+  select(symbol, cA:cH, local_tx, lod) %>%
+  (function(x) x[ord2, ]) %>%
+  mutate(row_num = 1:nrow(local_nonlocal_annot), 
+         AB = abs(cA - cB),
+         AC = abs(cA - cC),
+         AD = abs(cA - cD),
+         AE = abs(cA - cE),
+         AF = abs(cA - cF),
+         AG = abs(cA - cG),
+         AH = abs(cA - cH),
+         BC = abs(cB - cC),
+         BD = abs(cB - cD),
+         BE = abs(cB - cE),
+         BF = abs(cB - cF),
+         BG = abs(cB - cG),
+         BH = abs(cB - cH),
+         CD = abs(cC - cD),
+         CE = abs(cC - cE),
+         CF = abs(cC - cF),
+         CG = abs(cC - cG),
+         CH = abs(cC - cH),
+         DE = abs(cD - cE),
+         DF = abs(cD - cF),
+         DG = abs(cD - cG),
+         DH = abs(cD - cH),
+         EF = abs(cE - cF),
+         EG = abs(cE - cG),
+         EH = abs(cE - cH),
+         FG = abs(cF - cG),
+         FH = abs(cF - cH),
+         GH = abs(cG - cH)
+  ) %>%
+  mutate(BC_A = abs((cB + cC) / 2 - cA),
+         BC_D = abs((cB + cC) / 2 - cD),
+         BC_E = abs((cB + cC) / 2 - cE),
+         BC_F = abs((cB + cC) / 2 - cF),
+         BC_G = abs((cB + cC) / 2 - cG),
+         BC_H = abs((cB + cC) / 2 - cH),
+         BC_FG = abs((cB + cC) - (cF + cG))
+         ) %>%
+  select(symbol, BC_FG)
+
 ```
@@
                    local_tibble,
-                   lod_tibble
+                   lod_tibble, 
+                   contrasts_BC_FG
                     ) %>% 
@@
   (function(x){rownames(x) <- annots_ord[[6]]$symbol; return(x)})
+ann7 <- annots_ord[[7]] %>% # BC_FG contrast
+  select(-symbol) %>%
+  as.matrix() %>%
+  (function(x){rownames(x) <- annots_ord[[7]]$symbol; return(x)})
 ann4 <- annots_ord[[4]] %>%
@@
   iheatmapr::add_row_annotation(ann6, side = "left") %>%
+  iheatmapr::add_row_annotation(ann7, side = "left") %>%
   iheatmapr::add_row_labels()

### 2019-08-18 analysis/paper/Rmd/hotspots-paper.Rmd
switched to eval = TRUE for git2r code
removed code for working with blups

@@
 
-```{r load-broman}
-library(broman)
-library(here)
-```
-
 ```{r child="methods.Rmd", eval = TRUE}
@@
 ```{r setup}
-library(knitr)
 library(qtl2hotspots)
@@
 ```{r saveRDS}
-if (!file.exists(here("analysis", "data", "derived_data", "genoprobs.rds"))){
+if (!file.exists(here::here("analysis", "data", "derived_data", "genoprobs.rds"))){
   download.file("https://datadryad.org/bitstream/handle/10255/dryad.166774/Attie_DO378_eQTL_viewer_v1.Rdata?sequence=2",
-              destfile = here("analysis", "data", "raw_data", "Attie_DO378_eQTL_viewer_v1.Rdata")
+              destfile = here::here("analysis", "data", "raw_data", "Attie_DO378_eQTL_viewer_v1.Rdata")
               )
-  load(here("analysis", "data", "raw_data", "Attie_DO378_eQTL_viewer_v1.Rdata"))
-  saveRDS(dataset.islet.rnaseq$lod.peaks, here("analysis", "data", "derived_data", "lod_peaks.rds"))
-  saveRDS(dataset.islet.rnaseq$annots, here("analysis", "data", "derived_data", "annots.rds"))
-  saveRDS(dataset.islet.rnaseq$expr, here("analysis", "data", "derived_data", "expr.rds"))
-  saveRDS(K, here("analysis", "data", "derived_data", "kinship.rds"))
-  saveRDS(map, here("analysis", "data", "derived_data", "map.rds"))
-  saveRDS(genoprobs, here("analysis", "data", "derived_data", "genoprobs.rds"))
-  saveRDS(dataset.islet.rnaseq$covar, here("analysis", "data", "derived_data", "covar.rds"))
+  load(here::here("analysis", "data", "raw_data", "Attie_DO378_eQTL_viewer_v1.Rdata"))
+  saveRDS(dataset.islet.rnaseq$lod.peaks, here::here("analysis", "data", "derived_data", "lod_peaks.rds"))
+  saveRDS(dataset.islet.rnaseq$annots, here::here("analysis", "data", "derived_data", "annots.rds"))
+  saveRDS(dataset.islet.rnaseq$expr, here::here("analysis", "data", "derived_data", "expr.rds"))
+  saveRDS(K, here::here("analysis", "data", "derived_data", "kinship.rds"))
+  saveRDS(map, here::here("analysis", "data", "derived_data", "map.rds"))
+  saveRDS(genoprobs, here::here("analysis", "data", "derived_data", "genoprobs.rds"))
+  saveRDS(dataset.islet.rnaseq$covar, here::here("analysis", "data", "derived_data", "covar.rds"))
 }
@@
 ```{r readRDS}
-lod_peaks <- readRDS(here("analysis", "data", "derived_data", "lod_peaks.rds"))
-annots <- readRDS(here("analysis", "data", "derived_data", "annots.rds"))
-expr <- readRDS(here("analysis", "data", "derived_data", "expr.rds"))
-probs <- readRDS(here("analysis", "data", "derived_data", "genoprobs.rds"))
-K <- readRDS(here("analysis", "data", "derived_data", "kinship.rds"))
-covar <- readRDS(here("analysis", "data", "derived_data", "covar.rds"))
-map <- readRDS(here("analysis", "data", "derived_data", "map.rds"))
+lod_peaks <- readRDS(here::here("analysis", "data", "derived_data", "lod_peaks.rds"))
+annots <- readRDS(here::here("analysis", "data", "derived_data", "annots.rds"))
+expr <- readRDS(here::here("analysis", "data", "derived_data", "expr.rds"))
+probs <- readRDS(here::here("analysis", "data", "derived_data", "genoprobs.rds"))
+K <- readRDS(here::here("analysis", "data", "derived_data", "kinship.rds"))
+covar <- readRDS(here::here("analysis", "data", "derived_data", "covar.rds"))
+map <- readRDS(here::here("analysis", "data", "derived_data", "map.rds"))
 ```
@@
 
-```{r git2r, cache = FALSE, echo = TRUE, eval = FALSE}
+```{r git2r, cache = FALSE, echo = TRUE, eval = TRUE}
 # what commit is this file at?

### 2019-08-18 analysis/paper/Rmd/hotspot-one-chromosome.Rmd
fixed R code

@@
   rename(cA = `1`, cB = `2`, cC = `3`, cD = `4`, cE = `5`, cF = `6`, cG = `7`, cH = `8`
-         ) %>%
+         )
 ```

### 2019-08-18 analysis/paper/Rmd/hotspot-one-chromosome.Rmd
reordered code chunks & removed call with 'ord2'

@@
 
-```{r, define-contrasts-{{hot_chr}}}
-contrasts_BC_FG <- local_nonlocal_annot %>%
-  ungroup() %>%
-  select(symbol, cA:cH, local_tx, lod) %>%
-  (function(x) x[ord2, ]) %>%
-  mutate(row_num = 1:nrow(local_nonlocal_annot), 
-         AB = abs(cA - cB),
-         AC = abs(cA - cC),
-         AD = abs(cA - cD),
-         AE = abs(cA - cE),
-         AF = abs(cA - cF),
-         AG = abs(cA - cG),
-         AH = abs(cA - cH),
-         BC = abs(cB - cC),
-         BD = abs(cB - cD),
-         BE = abs(cB - cE),
-         BF = abs(cB - cF),
-         BG = abs(cB - cG),
-         BH = abs(cB - cH),
-         CD = abs(cC - cD),
-         CE = abs(cC - cE),
-         CF = abs(cC - cF),
-         CG = abs(cC - cG),
-         CH = abs(cC - cH),
-         DE = abs(cD - cE),
-         DF = abs(cD - cF),
-         DG = abs(cD - cG),
-         DH = abs(cD - cH),
-         EF = abs(cE - cF),
-         EG = abs(cE - cG),
-         EH = abs(cE - cH),
-         FG = abs(cF - cG),
-         FH = abs(cF - cH),
-         GH = abs(cG - cH)
-  ) %>%
-  mutate(BC_A = abs((cB + cC) / 2 - cA),
-         BC_D = abs((cB + cC) / 2 - cD),
-         BC_E = abs((cB + cC) / 2 - cE),
-         BC_F = abs((cB + cC) / 2 - cF),
-         BC_G = abs((cB + cC) / 2 - cG),
-         BC_H = abs((cB + cC) / 2 - cH),
-         BC_FG = abs((cB + cC) - (cF + cG))
-         ) %>%
-  select(symbol, BC_FG)
-
-```
-
 ```{r, scols-peak-position-{{hot_chr}}}
@@
   ungroup()
+```
+
+```{r, define-contrasts-{{hot_chr}}}
+contrasts_BC_FG <- local_nonlocal_annot %>%
+  ungroup() %>%
+  select(symbol, cA:cH, local_tx, lod) %>%
+  mutate(row_num = 1:nrow(local_nonlocal_annot), 
+         AB = abs(cA - cB),
+         AC = abs(cA - cC),
+         AD = abs(cA - cD),
+         AE = abs(cA - cE),
+         AF = abs(cA - cF),
+         AG = abs(cA - cG),
+         AH = abs(cA - cH),
+         BC = abs(cB - cC),
+         BD = abs(cB - cD),
+         BE = abs(cB - cE),
+         BF = abs(cB - cF),
+         BG = abs(cB - cG),
+         BH = abs(cB - cH),
+         CD = abs(cC - cD),
+         CE = abs(cC - cE),
+         CF = abs(cC - cF),
+         CG = abs(cC - cG),
+         CH = abs(cC - cH),
+         DE = abs(cD - cE),
+         DF = abs(cD - cF),
+         DG = abs(cD - cG),
+         DH = abs(cD - cH),
+         EF = abs(cE - cF),
+         EG = abs(cE - cG),
+         EH = abs(cE - cH),
+         FG = abs(cF - cG),
+         FH = abs(cF - cH),
+         GH = abs(cG - cH)
+  ) %>%
+  mutate(BC_A = abs((cB + cC) / 2 - cA),
+         BC_D = abs((cB + cC) / 2 - cD),
+         BC_E = abs((cB + cC) / 2 - cE),
+         BC_F = abs((cB + cC) / 2 - cF),
+         BC_G = abs((cB + cC) / 2 - cG),
+         BC_H = abs((cB + cC) / 2 - cH),
+         BC_FG = abs((cB + cC) - (cF + cG))
+         ) %>%
+  select(symbol, BC_FG)
 ```

### 2019-08-18 analysis/paper/Rmd/hotspot-one-chromosome.Rmd
removed blank lines

@@
   iheatmapr::add_row_labels()
-
 ```

### 2019-08-19 analysis/paper/preliminary/allele-effects-exploration.Rmd
added partitions fctn

@@
 ```
+
+## Calculate all partitions & contrasts
+
+```{r}
+pp <- partitions::setparts(8) %>% as.matrix() # 8 founder lines
+dim(pp)
+# calculate set of contrasts for a single group_assignments vector
+calculate_contrasts <- function(group_assignments, effects_vector){
+  if (max(group_assignments) == 1){
+    stop("group_assignments vector must have at least one entry greater than 1.")
+  }
+  n <- max(group_assignments) # n subsets constitute the partition.
+  ncontr <- choose(n, 2) # we'll always compare pairs of subsets as contrasts
+  # calculate the rowwise means for each of n groups
+  means <- numeric(length = n) # one mean per founder allele
+  for (i in 1:n){
+    means[i] <- mean(effects_vector[which(group_assignments == i)])
+  }
+  combn(n, 2) -> cmat # indices for contrasts
+  out <- apply(X = cmat, MARGIN = 2, FUN = function(x){abs(means[x[1]] - means[x[2]])})
+  mynames <- list()
+  for (i in 1:n){
+    mynames[[i]] <- paste(LETTERS[which(group_assignments == i)], collapse = "")
+  }
+  for (i in 1:ncontr){
+    names(out)[i]<- paste(mynames[cmat[, i]], collapse = "_")
+  }
+  return(out)
+}
+effects_mat <- local_nonlocal_annot %>%
+  ungroup() %>%
+  select(cA:cH) %>%
+  as.matrix()
+```
+
+```{r}
+list(
+  calculate_contrasts(pp[,4001], effects_vector = as.numeric(effects_mat[1, ])),
+  calculate_contrasts(pp[,4000], effects_vector = as.numeric(effects_mat[1, ]))
+)
+```

### 2019-08-19 R/calc_contrasts.R
added code for calculating contrast between 2 groups

@@ -1,0 +1,15 @@
+#' Calculate all contrasts for 8 founder alleles
+#'
+#' @param grouping a numeric vector, length 8, assigning founder alleles to each of 3 groups: 1, 2, or 3.
+#' @param effects founder allele effects vector of length 8
+#'
+#' @export
+calc_contrasts <- function(grouping, effects){
+  means <- numeric(length = 2) # one mean per founder allele
+  for (i in 1:2){
+    means[i] <- mean(effects_vector[which(grouping == i)])
+    names(means)[i] <- LETTERS[which(grouping == i)]
+  }
+  diff <- abs(means[1] - means[2])
+  return(tibble::tibble(contrast = contr, value = diff))
+}

### 2019-08-19 R/calc_contrast.R
fixed filename

@@ -4,7 +4,7 @@
 #' @param effects founder allele effects vector of length 8
 #'
 #' @export
-calc_contrasts <- function(grouping, effects){
+calc_contrast <- function(grouping, effects){
   means <- numeric(length = 2) # one mean per founder allele
   for (i in 1:2){
     means[i] <- mean(effects_vector[which(grouping == i)])

### 2019-08-19 R/calc_contrast.R
fixed calc_contrast fctn

@@ -5,11 +5,18 @@
 #'
 #' @export
 calc_contrast <- function(grouping, effects){
-  means <- numeric(length = 2) # one mean per founder allele
-  for (i in 1:2){
-    means[i] <- mean(effects_vector[which(grouping == i)])
-    names(means)[i] <- LETTERS[which(grouping == i)]
-  }
-  diff <- abs(means[1] - means[2])
-  return(tibble::tibble(contrast = contr, value = diff))
+  # return NA if contrast can't be calculated, say, for a vector of all 1's.
+  if (length(unique(grouping[grouping < 3])) <= 1) {
+    out <- NA
+    } else {
+      means <- numeric(length = 2) # one mean per founder allele
+      for (i in 1:2){
+        means[i] <- mean(effects_vector[which(grouping == i)])
+        names(means)[i] <- paste(LETTERS[which(grouping == i)], collapse = "")
+        }
+      diff <- abs(means[1] - means[2])
+      contr <- paste(names(means), collapse = "_")
+      out <- tibble::tibble(contrast = contr, value = diff)
+    }
+  return(out)
 }

### 2019-08-19 analysis/paper/preliminary/allele-effects-exploration.Rmd
deleted calculate_contrasts function.
defined 'ef' as list of lists
fixed calcs

--- 2019-08-18 analysis/paper/preliminary/allele-effects-exploration.Rmd
@@
 ```
+
+## Calculate all partitions & contrasts
+
+```{r}
+pp <- expand.grid(1:3, 
+                  1:3,
+                  1:3, 
+                  1:3,
+                  1:3, 
+                  1:3,
+                  1:3, 
+                  1:3
+                  )
+effects_mat <- local_nonlocal_annot %>%
+  ungroup() %>%
+  select(cA:cH) %>%
+  as.matrix()
+
+```
+
+```{r}
+pp_tib <- pp %>%
+  as_tibble()
+effects_tib <- effects_mat %>%
+  as_tibble()
+
+ef <- pp_tib %>%
+  apply(MARGIN = 1, 
+        FUN = function(x){apply(X = effects_tib, 
+                                MARGIN = 1, 
+                                FUN = function(a, b = x){calc_contrast(grouping = b, effects = a)}
+                                )})
+
+ef %>%  apply(FUN = function(x) x[!is.na(x)], MARGIN = 1) %>% 
+  lapply(FUN = function(x) bind_rows(x)) %>%
+  purrr::map2(local_nonlocal_annot$symbol, .f = function(x,y){x %>% mutate(symbol = y)}) %>%
+  bind_rows() %>%
+  filter(contrast == "A_B") 
+  summarise(v = var(value)) %>%
+  arrange(desc(v))
+```

### 2019-09-07 analysis/paper/Rmd/methods.Rmd
removed heatmaply dependency; added iheatmapr & pheatmap
updated methods

@@
 expression traits.
-They identified five expression trait hotspots in their
-study of pancreatic islet gene expression in 378 Diversity Outbred mice.
+They identified five expression trait hotspots - on Chromosomes 2, 5, 7, 11, and 13 - in their data.
 
-In analyzing the data from @keller2018genetic, we first established criteria for identifying local and nonlocal transcripts at each of the five hotspots. We focused on those local 
-traits that strongly associated with QTL within the hotspot. We set the univariate LOD threshold for 
-local traits at 15. We limited consideration to those local traits for which the middle of the gene lay 
-within the hotspot interval. 
-All nonlocal traits demonstrated univariate LOD peaks that met
-the genome-wide significance threshold, 7.18 [@keller2018genetic]. The nonlocal expression traits arose 
-from genes
-on other chromosomes, while the local traits reflected expression levels for
-genes on the same chromosome as the hotspot. We ignored transcripts for genes on the hotspot chromosome 
+In analyzing the data from @keller2018genetic, we first established criteria for identifying local and 
+nonlocal transcripts at each of the five hotspots. We defined a "local trait" to be one whose corresponding gene lay within the hotspot. We focused on those local 
+traits that strongly associated with QTL within the hotspot. To limit the scope of our analyses, 
+we only studied those expression traits with univariate LOD greater than 15.
+
+We defined a "nonlocal trait" to be one with univariate evidence of association (LOD at least 7.18, the genome-wide threshold)
+with a QTL in the hotspot. Additionally, we required that the nonlocal trait's gene not lie on the hotspot's chromosome. We ignored transcripts for genes on the hotspot chromosome 
 but outside the hotspot interval.
@@
 tests. For every local trait - nonlocal trait pair, we also performed a linear 
-regression-based mediation analysis [@chick2016defining,@keller2018genetic]. We describe in detail both procedures below.
+regression-based mediation analysis [@chick2016defining,@keller2018genetic]. We describe both procedures below.
 
@@
 
-For each hotspot, we tested all pairs of traits, regardless of whether the traits were local or nonlocal. 
-We performed a
+For each hotspot, we tested all pairs of traits, regardless of whether the traits were local or nonlocal. For each test, we performed a
 two-dimensional, two-QTL scan over a genomic region that included the entirety of the
@@
 
-From the two-dimensional scans, we calculated pleiotropy likelihood ratio test (LOD) statistics for each
+From the two-dimensional scan results, we calculated pleiotropy likelihood ratio test (LOD) statistics for each
 pair of traits [@jiang1995multiple;@boehm2019testing;@boehm2019thesis]. Because of the required
@@
 
-In all of our mediation analyses, the nonlocal transcripts are the targets, while the local
-transcripts serve as candidate mediators of the QTL-target relationship. 
+In all of our mediation analyses, the nonlocal traits are the targets (Y), while the local
+transcripts serve as candidate mediators (M) of the QTL-target relationship. 
 We only considered as potential mediators those local traits with strong univariate LODs.

### 2019-09-07 analysis/paper/preliminary/allele-effects-exploration.Rmd
updated methods

--- 2019-08-18 analysis/paper/preliminary/allele-effects-exploration.Rmd
@@
 
-```{r define}
+```{r define-ult}
 library(qtl2hotspots)
@@
 figdir <- here::here("analysis", "figures", "Chr5-allele-effects-plots")
-names(s1coef_out) -> ids
+names(s1coef_out) -> ids # need to run inline code above to define s1coef_out
 # find indices on Chr5 for boundaries of hotspot
@@
 purrr::pmap(.l = list(s1coef_out, as.list(ids), mylist), 
-           .f = function(x, y, z){
-             main <- local_nonlocal_annot %>%
-               ungroup() %>%
-               filter(annot.id == y) %>%
-               select(annot.id, symbol) %>%
-               paste0()
-             qtl2::plot_coefCC(x, map = submap, main = main, scan1_output = z)
-           }
-)
-```
-
-```{r blup-plots}
-purrr::pmap(.l = list(s1coef_blup, as.list(ids), mylist), 
            .f = function(x, y, z){
@@
 
-## Bar plots for blup effects
-
-```{r, barplots-blup, fig.height = 20, fig.width = 8}
-foobar <- local_nonlocal_annot %>%
-  ungroup() %>%
-  (function(x) x[ord2, ]) %>%
-  select(symbol, 26:33) %>%
-  tidyr::gather(paste0("b", LETTERS[1:8]), key = "founder_allele", value = "blup_effect")
-foobar2 <- arrange(transform(foobar,
-             symbol=factor(symbol,levels=rn)),symbol)
-foobar2 %>% 
-  ggplot() + geom_col(aes(y = effect, x = founder_allele, fill = founder_allele)) +
-  ylim(c(-1.5, 1.5)) + 
-  facet_wrap(symbol ~ ., ncol = 4) #+ theme(strip.text.x = element_blank())
-ggsave(here::here("analysis", "figures", "allele_effects_at_peaks.png"), limitsize = FALSE)
-```
-
 ```{r correlations}
@@
   (function(x) x[ord2, ]) %>%
-  select(symbol, 18:25) %>%
-    rename("A" = `1`, "B" = `2`, "C" = `3`, "D" = `4`, "E" = `5`, "F" = `6`, G = `7`, H = `8`
-         ) %>%
-  mutate(contrast_BC_FG = abs((B + C) - (F + G)), row_num = 1:203) %>%
+  dplyr::select(symbol, 18:25) %>%
+  mutate(contrast_BC_FG = abs((cB + cC) - (cF + cG)), row_num = 1:203) %>%
   ggplot() + geom_point(aes(y = contrast_BC_FG, x = row_num))
@@
   (function(x) x[ord2, ]) %>%
-  select(symbol, 18:25) %>%
-    rename("A" = `1`, "B" = `2`, "C" = `3`, "D" = `4`, "E" = `5`, "F" = `6`, G = `7`, H = `8`
-         ) %>%
-  mutate(contrast = abs((B + C) / 2 - G), row_num = 1:203) %>%
+  dplyr::select(symbol, 18:25) %>%
+  mutate(contrast = abs((cB + cC) / 2 - cG), row_num = 1:203) %>%
   ggplot() + geom_point(aes(y = contrast, x = row_num)) + geom_smooth(aes(y = contrast, x = row_num))
@@
 
-```{r B:G}
+```{r B:G2}
 p <- local_nonlocal_annot %>%
@@
 ```
+
+## Calculate all partitions & contrasts
+
+```{r}
+pp <- expand.grid(1:3, 
+                  1:3,
+                  1:3, 
+                  1:3,
+                  1:3, 
+                  1:3,
+                  1:3, 
+                  1:3
+                  )
+effects_mat <- local_nonlocal_annot %>%
+  ungroup() %>%
+  select(cA:cH) %>%
+  as.matrix()
+
+```
+
+```{r}
+pp_tib <- pp %>%
+  as_tibble()
+effects_tib <- effects_mat %>%
+  as_tibble()
+
+ef <- pp_tib %>%
+  apply(MARGIN = 1, 
+        FUN = function(x){apply(X = effects_tib, 
+                                MARGIN = 1, 
+                                FUN = function(a, 
+                                               b = x){
+                                  qtl2hotspots::calc_contrast(grouping = b, 
+                                                              effects = a)
+                                  }
+                                )}
+        )
+
+ef2 <- ef %>%  apply(FUN = function(x) x[!is.na(x)], MARGIN = 1) %>% 
+  lapply(FUN = function(x) bind_rows(x)) %>%
+  purrr::map2(local_nonlocal_annot$symbol, .f = function(x,y){x %>% mutate(symbol = y)}) %>%
+  bind_rows() 
+ef2 %>%
+  group_by(contrast) %>%
+  summarise(v = var(value)) %>%
+  arrange(desc(v))
+```
+
+```{r}
+rn_tib <- tibble::tibble(symbol = rn, row_number = 1:203)
+ef2 %>%
+  filter(contrast == "B_F") %>%
+  inner_join(rn_tib) %>%
+  inner_join(local_nonlocal_annot, by = "symbol") %>%
+  filter(!local_tx) %>%
+  ggplot() + geom_point(aes(x = row_number, y = value)) + geom_vline(aes(xintercept = 68.5))
+```
+
+## Defining the boundaries from the heatmap
+
+I need to examine the heatmap to see how many diagonal "blocks" there are, and where the endpoints are, from 1:203, for every block.
+
+First, note that 1:182 are all nonlocal traits. Now, Adamts15 through Cntnap2, inclusive, seems to define the first block.
+
+The second block starts right above Cntnap2, ie, with Elmo1 and ends roughtly at Cep68.
+
+Let's look at the first "block".
+
+```{r}
+which(rn_tib$symbol == "Elmo1")
+which(rn_tib$symbol == "Adamts15")
+abs_diff <- function(x){abs(x[1] - x[2])}
+notgroup1 <- rn_tib %>%
+  mutate(group1_member = row_number <= 68) %>%
+  inner_join(ef2) %>%
+  group_by(contrast, group1_member) %>%
+  summarise(variance = var(value), mean = mean(value)) %>%
+  arrange(contrast) %>%
+  ungroup() %>%
+  group_by(contrast) %>%
+  filter(!group1_member) %>%
+  arrange(variance) %>% 
+  rename(variance_notgroup1 = variance, mean_notgroup1 = mean)
+
+rn_tib %>%
+  mutate(group1_member = row_number <= 68) %>%
+  inner_join(ef2) %>%
+  group_by(contrast, group1_member) %>%
+  summarise(variance = var(value), mean = mean(value)) %>%
+  arrange(contrast) %>%
+  ungroup() %>%
+  group_by(contrast) %>%
+  filter(group1_member) %>%
+  arrange(variance) %>% 
+  rename(variance_group1 = variance, mean_group1 = mean) %>%
+  inner_join(notgroup1, by = "contrast") %>%
+  dplyr::select(- group1_member.x, - group1_member.y)
+
+```
+
+```{r}
+ef2 %>%
+  filter(contrast == "B_G") %>%
+  inner_join(rn_tib) %>%
+  inner_join(local_nonlocal_annot, by = "symbol") %>%
+  filter(!local_tx) %>%
+  ggplot() + geom_point(aes(x = row_number, y = value)) + geom_vline(aes(xintercept = 68.5)) + geom_smooth(aes(x = row_number, y = value))
+
+```
+
+```{r}
+rn_tib %>%
+  mutate(group1_member = row_number <= 68) %>%
+  inner_join(ef2) %>%
+  group_by(contrast) %>%
+  summarise(t_stat = t.test(value ~ group1_member)$statistic) %>%
+  arrange(desc(t_stat))
+```
+
+```{r}
+ef2 %>%
+  filter(contrast == "AGH_CEF") %>%
+  inner_join(rn_tib) %>%
+  inner_join(local_nonlocal_annot, by = "symbol") %>%
+  filter(!local_tx) %>%
+  ggplot() + geom_point(aes(x = row_number, y = value)) + geom_vline(aes(xintercept = 68.5))
+
+```
+
+```{r, define-contrasts-5-1}
+contr <- ef2 %>%
+  filter(contrast == "AGH_CEF") %>%
+  inner_join(rn_tib) %>%
+  arrange(row_number)
+```
+
+```{r, order-annotations-5-1}
+# all annots 
+annots_ord <- list(lod_diffs_tibble, 
+                   lod_diff_proportions_tibble, 
+                   allele_effects_tibble,
+                   peak_positions, 
+                   local_tibble,
+                   lod_tibble, 
+                   contr
+                    ) %>% 
+  purrr::map(.f = function(df, rn){
+    out <- df[match(x = rn, table = df$symbol), ]
+    return(out)
+  }, 
+  rn = rn)
+ann7 <- annots_ord[[7]] %>% # contrast
+  dplyr::select(value) %>%
+  as.matrix() %>%
+  (function(x){rownames(x) <- annots_ord[[7]]$symbol; return(x)})
+```
+
+```{r}
+iheatmapr::main_heatmap(mymat, show_colorbar = FALSE, layout = list(font = list(size = 4))) %>%
+  iheatmapr::add_col_labels() %>%
+  iheatmapr::add_col_annotation(ann5, show_colorbar = FALSE) %>%
+  iheatmapr::add_main_heatmap(ann3, name = "signal3", size = 0.2, show_colorbar = FALSE) %>%
+    iheatmapr::add_col_labels() %>%
+
+  iheatmapr::add_main_heatmap(ann2, name = "signal2", size = 0.5, show_colorbar = FALSE) %>%
+  iheatmapr::add_col_labels() %>%
+  iheatmapr::add_main_heatmap(ann1, name = "signal1", size = 0.5, show_colorbar = FALSE) %>%
+  iheatmapr::add_row_annotation(ann5, side = "left", show_colorbar = TRUE) %>%
+  iheatmapr::add_col_labels() %>%
+  iheatmapr::add_row_annotation(ann4, side = "left", show_colorbar = TRUE) %>%
+  iheatmapr::add_row_annotation(ann6, side = "left", show_colorbar = TRUE) %>%
+  iheatmapr::add_row_annotation(ann7, side = "left", show_colorbar = TRUE) %>%
+  iheatmapr::add_row_labels()
+
+```
+
+# Other ideas
+
+What results do we get when we try to "find the contrast" for other hotspots: 2, 7, 11, 13?

### 2019-09-07 R/calc_contrast.R

@@ -11,7 +11,7 @@ calc_contrast <- function(grouping, effects){
     } else {
       means <- numeric(length = 2) # one mean per founder allele
       for (i in 1:2){
-        means[i] <- mean(effects_vector[which(grouping == i)])
+        means[i] <- mean(effects[which(grouping == i)])
         names(means)[i] <- paste(LETTERS[which(grouping == i)], collapse = "")
         }
       diff <- abs(means[1] - means[2])

### 2019-09-17 analysis/paper/Rmd/hotspot-one-chromosome.Rmd
updated README

@@
 
-```{r, scols-peak-position-{{hot_chr}}}
-
-binning <- local_nonlocal_annot %>%
-  dplyr::select(symbol, pos, lod) %>%
-  mutate(bin = unlist(pos) %>%
-           map(.f = function(x) {x %>% 
-               bin_peak_position(breaks = hot_lower + (1:9) * 0.4) %>%
-               binarize_bin_number(total_number_of_bins = 10L)
-             })
-  ) %>%
-  listcol_to_cols(list_column = quo(bin), another_column = quo(symbol))
-```
-
 ```{r, n-med-{{hot_chr}}}
@@
 8 allele effects per trait, 5.
-peak position per trait, encoded as binned position
+peak position per trait, as a one-column heatmap
 # 2,3,4 are numeric; 1 & 5 are binary
@@
   ungroup()
-```
-
-```{r, define-contrasts-{{hot_chr}}}
-contrasts_BC_FG <- local_nonlocal_annot %>%
-  ungroup() %>%
-  select(symbol, cA:cH, local_tx, lod) %>%
-  mutate(row_num = 1:nrow(local_nonlocal_annot), 
-         AB = abs(cA - cB),
-         AC = abs(cA - cC),
-         AD = abs(cA - cD),
-         AE = abs(cA - cE),
-         AF = abs(cA - cF),
-         AG = abs(cA - cG),
-         AH = abs(cA - cH),
-         BC = abs(cB - cC),
-         BD = abs(cB - cD),
-         BE = abs(cB - cE),
-         BF = abs(cB - cF),
-         BG = abs(cB - cG),
-         BH = abs(cB - cH),
-         CD = abs(cC - cD),
-         CE = abs(cC - cE),
-         CF = abs(cC - cF),
-         CG = abs(cC - cG),
-         CH = abs(cC - cH),
-         DE = abs(cD - cE),
-         DF = abs(cD - cF),
-         DG = abs(cD - cG),
-         DH = abs(cD - cH),
-         EF = abs(cE - cF),
-         EG = abs(cE - cG),
-         EH = abs(cE - cH),
-         FG = abs(cF - cG),
-         FH = abs(cF - cH),
-         GH = abs(cG - cH)
-  ) %>%
-  mutate(BC_A = abs((cB + cC) / 2 - cA),
-         BC_D = abs((cB + cC) / 2 - cD),
-         BC_E = abs((cB + cC) / 2 - cE),
-         BC_F = abs((cB + cC) / 2 - cF),
-         BC_G = abs((cB + cC) / 2 - cG),
-         BC_H = abs((cB + cC) / 2 - cH),
-         BC_FG = abs((cB + cC) - (cF + cG))
-         ) %>%
-  select(symbol, BC_FG)
 ```
@@
 ```
+
+### Allele effects contrasts
+
+```{r, define-submap-{{hot_chr}}}
+figdir <- here::here("analysis", "figures", "Chr{{hot_chr}}-allele-effects-plots")
+names(s1coef_out) -> ids # need to run inline code above to define s1coef_out
+# find indices on Chr5 for boundaries of hotspot
+which(map[[hot_chr]] > hot_lower)[1] -> start_index
+which(map[[hot_chr]] > hot_upper)[1] -> stop_index
+submap <- list()
+submap$`5` <- map[[hot_chr]][start_index:stop_index]
+mylist <- list()
+for (i in 1:(nlocal + nnonlocal)) {
+  mylist[[i]] <- s1out[ , i, drop = FALSE]
+}
+```
+
+```{r effects-plots-{{hot_chr}}}
+purrr::pmap(.l = list(s1coef_out, as.list(ids), mylist), 
+           .f = function(x, y, z){
+             main <- local_nonlocal_annot %>%
+               ungroup() %>%
+               filter(annot.id == y) %>%
+               select(annot.id, symbol) %>%
+               paste0()
+             fn <- here::here("analysis", "figures", "Chr{{hot_chr}}-allele-effects", paste0(main, ".pdf"))
+             pdf(file = fn)
+             qtl2::plot_coefCC(x, map = submap, main = main, scan1_output = z)
+             dev.off()
+           }
+)
+
+```
+
+## Barplots for effects at peak positions
+
+Now, let's focus only on the 8 values at the LOD peak for each trait.
+
+```{r, barplots-scan1coef-effects-{{hot_chr}}, fig.height = 20, fig.width = 8}
+ord2 <- match(table = local_nonlocal_annot$symbol, x = rn)
+foobar <- local_nonlocal_annot %>%
+  select(symbol, cA:cH) %>%
+  (function(x) x[ord2, ]) %>%
+  tidyr::gather(key = "founder_allele", value = "effect", cA:cH)
+foobar2 <- arrange(transform(foobar,
+             symbol = factor(symbol, levels = rn)), symbol)
+foobar2 %>% 
+  ggplot() + geom_col(aes(y = effect, x = founder_allele, fill = founder_allele)) +
+  ylim(c(-1.5, 1.5)) + 
+  facet_wrap(symbol ~ ., ncol = 4) #+ theme(strip.text.x = element_blank())
+ggsave(here::here("analysis", "figures", "allele_effects_at_peaks-{{hot_chr}}.png"), limitsize = FALSE)
+```

### 2019-09-17 analysis/paper/preliminary/allele-effects-exploration.Rmd
updated README

@@
 out <- pmap_dfc(inputs, .f = function(hot_chr, hot_mid, keller_mediator) knitr::knit_expand(text = readLines(here::here("analysis", "paper", "Rmd", "hotspot-one-chromosome.Rmd"))))
-# note that ... argument must be a list, ie, it must be a single argument, not multiple args
+# note that ... argument for knit_expand must be a list, ie, it must be a single argument, not multiple args
 ```
@@
 `r if (eval_code) knitr::knit(text = out[[2]])`
-
-```{r, define-submap}
-figdir <- here::here("analysis", "figures", "Chr5-allele-effects-plots")
-names(s1coef_out) -> ids # need to run inline code above to define s1coef_out
-# find indices on Chr5 for boundaries of hotspot
-which(map$`5` > hot_lower)[1] -> start_index
-which(map$`5` > hot_upper)[1] -> stop_index
-submap <- list()
-submap$`5` <- map$`5`[start_index:stop_index]
-mylist <- list()
-for (i in 1:203){
-  mylist[[i]] <- s1out[ , i, drop = FALSE]
-}
-```
-
-```{r effects-plots}
-purrr::pmap(.l = list(s1coef_out, as.list(ids), mylist), 
-           .f = function(x, y, z){
-             main <- local_nonlocal_annot %>%
-               ungroup() %>%
-               filter(annot.id == y) %>%
-               select(annot.id, symbol) %>%
-               paste0()
-             qtl2::plot_coefCC(x, map = submap, main = main, scan1_output = z)
-           }
-)
-```
-
-## Barplots for effects at peak positions
-
-Now, let's focus only on the 8 values at the LOD peak for each trait.
-
-```{r, barplots-scan1coef-effects, fig.height = 20, fig.width = 8}
-ord2 <- match(table = local_nonlocal_annot$symbol, x = rn)
-foobar <- local_nonlocal_annot %>%
-  select(symbol, cA:cH) %>%
-  (function(x) x[ord2, ]) %>%
-  tidyr::gather(key = "founder_allele", value = "effect", cA: cH)
-foobar2 <- arrange(transform(foobar,
-             symbol=factor(symbol,levels=rn)),symbol)
-foobar2 %>% 
-  ggplot() + geom_col(aes(y = effect, x = founder_allele, fill = founder_allele)) +
-  ylim(c(-1.5, 1.5)) + 
-  facet_wrap(symbol ~ ., ncol = 4) #+ theme(strip.text.x = element_blank())
-ggsave(here::here("analysis", "figures", "allele_blupss_at_peaks.png"), limitsize = FALSE)
-```
 
@@
 What results do we get when we try to "find the contrast" for other hotspots: 2, 7, 11, 13?
+
+What if more than one contrast is "responsible" for a given block in a hotspot? Maybe max of two contrasts, for example? Or min of two contrasts?

### 2019-09-22 analysis/paper/Rmd/hotspot-one-chromosome.Rmd
removed, partially, code dealing with allele effects.

--- 2019-08-18 analysis/paper/Rmd/hotspot-one-chromosome.Rmd
@@
 # peak positions tibble
-peak_positions <- binning %>%
+peak_positions <- local_nonlocal_annot %>%
   select(symbol, pos)
@@
 
-```{r, define-contrasts-{{hot_chr}}}
-contrasts_BC_FG <- local_nonlocal_annot %>%
-  ungroup() %>%
-  select(symbol, cA:cH, local_tx, lod) %>%
-  mutate(row_num = 1:nrow(local_nonlocal_annot), 
-         AB = abs(cA - cB),
-         AC = abs(cA - cC),
-         AD = abs(cA - cD),
-         AE = abs(cA - cE),
-         AF = abs(cA - cF),
-         AG = abs(cA - cG),
-         AH = abs(cA - cH),
-         BC = abs(cB - cC),
-         BD = abs(cB - cD),
-         BE = abs(cB - cE),
-         BF = abs(cB - cF),
-         BG = abs(cB - cG),
-         BH = abs(cB - cH),
-         CD = abs(cC - cD),
-         CE = abs(cC - cE),
-         CF = abs(cC - cF),
-         CG = abs(cC - cG),
-         CH = abs(cC - cH),
-         DE = abs(cD - cE),
-         DF = abs(cD - cF),
-         DG = abs(cD - cG),
-         DH = abs(cD - cH),
-         EF = abs(cE - cF),
-         EG = abs(cE - cG),
-         EH = abs(cE - cH),
-         FG = abs(cF - cG),
-         FH = abs(cF - cH),
-         GH = abs(cG - cH)
-  ) %>%
-  mutate(BC_A = abs((cB + cC) / 2 - cA),
-         BC_D = abs((cB + cC) / 2 - cD),
-         BC_E = abs((cB + cC) / 2 - cE),
-         BC_F = abs((cB + cC) / 2 - cF),
-         BC_G = abs((cB + cC) / 2 - cG),
-         BC_H = abs((cB + cC) / 2 - cH),
-         BC_FG = abs((cB + cC) - (cF + cG))
-         ) %>%
-  select(symbol, BC_FG)
-```
-
 ```{r, order-annotations-{{hot_chr}}}
@@
                    local_tibble,
-                   lod_tibble, 
-                   contrasts_BC_FG
+                   lod_tibble 
                     ) %>% 
@@
 
-```{r iheatmapr-{{hot_chr}}, out.width="25in", out.height = "10in"}
+```{r {{hot_chr}}-ann3}
 ann3 <- annots_ord[[3]] %>%
@@
 colnames(ann3) <- LETTERS[1:8]
+```
+
+```{r {{hot_chr}}-ann2}
 ann2 <- annots_ord[[2]] %>%
@@
   (function(x){rownames(x) <- annots_ord[[2]]$symbol; return(x)})
+```
+
+```{r {{hot_chr}}-ann1}
 ann1 <- annots_ord[[1]] %>%
@@
   (function(x){rownames(x) <- annots_ord[[1]]$symbol; return(x)})
+```
+
+```{r {{hot_chr}}-ann5}
 ann5 <- annots_ord[[5]] %>%
@@
   (function(x){rownames(x) <- annots_ord[[5]]$symbol; return(x)})
+```
+
+```{r {{hot_chr}}-ann6}
 ann6 <- annots_ord[[6]] %>%
@@
   (function(x){rownames(x) <- annots_ord[[6]]$symbol; return(x)})
-ann7 <- annots_ord[[7]] %>% # BC_FG contrast
-  select(-symbol) %>%
-  as.matrix() %>%
-  (function(x){rownames(x) <- annots_ord[[7]]$symbol; return(x)})
+```
+
+```{r {{hot_chr}}-ann4}
 ann4 <- annots_ord[[4]] %>%
@@
   (function(x){rownames(x) <- annots_ord[[4]]$symbol; return(x)})
+```
+
+```{r iheatmapr-{{hot_chr}}, out.width="25in", out.height = "10in"}
 mymat <- pm[ord, ord] %>% # order both cols and rows by 'ord'
@@
   iheatmapr::add_row_annotation(ann6, side = "left") %>%
-  iheatmapr::add_row_annotation(ann7, side = "left") %>%
   iheatmapr::add_row_labels()
-
 ```

### 2019-09-29 analysis/paper/Rmd/hotspots-paper.Rmd
fixed url for attie data

@@
 if (!file.exists(here::here("analysis", "data", "derived_data", "genoprobs.rds"))){
-  download.file("https://datadryad.org/bitstream/handle/10255/dryad.166774/Attie_DO378_eQTL_viewer_v1.Rdata?sequence=2",
+  download.file("https://datadryad.org/stash/downloads/file_stream/63306",
               destfile = here::here("analysis", "data", "raw_data", "Attie_DO378_eQTL_viewer_v1.Rdata")

### 2019-09-29 analysis/paper/Rmd/hotspot-one-chromosome.Rmd
fixed number of cores

@@
                        reml = TRUE, 
-                       cores = 7
+                       cores = 1
                        )
@@
                              reml = TRUE, 
-                             cores = 7
+                             cores = 1
                              )

### 2019-10-02 analysis/paper/Rmd/abstract.Rmd
updated abstr & methods

@@
 Diversity Outbred mice.
+@keller2018genetic defined the hotspots in their initial report.
+We now apply our newly developed pleiotropy test and mediation analysis to understand biological relationships among traits that map to the hotspots.
+
+Four of five hotspots show patterns consistent with a single local trait causing the hotspot associations.
+For these hotspots, mediation analyses revealed that the local trait mediates most of the associated nonlocal traits, while pleiotropy test results were consistent with presence of a single QTL at each hotspot.
+However, the fifth hotspot demonstrated distinct patterns in both mediation analyses and pleiotropy testing.

### 2019-10-02 analysis/paper/Rmd/methods.Rmd
updated abstr & methods

@@
 In analyzing the data from @keller2018genetic, we first established criteria for identifying local and 
-nonlocal transcripts at each of the five hotspots. We defined a "local trait" to be one whose corresponding gene lay within the hotspot. We focused on those local 
+nonlocal transcripts at each of the five hotspots. We defined a "local trait" to be a transcript whose corresponding gene lay within the hotspot. We focused on those local 
 traits that strongly associated with QTL within the hotspot. To limit the scope of our analyses, 
-we only studied those expression traits with univariate LOD greater than 15.
+we limited local traits to those with univariate LOD greater than 15.
 
 We defined a "nonlocal trait" to be one with univariate evidence of association (LOD at least 7.18, the genome-wide threshold)
-with a QTL in the hotspot. Additionally, we required that the nonlocal trait's gene not lie on the hotspot's chromosome. We ignored transcripts for genes on the hotspot chromosome 
+with a QTL in the hotspot [@keller2018genetic].
+Additionally, we required that the nonlocal trait's gene not lie on the hotspot's chromosome.
+We ignored transcripts for genes on the hotspot chromosome
 but outside the hotspot interval.
@@
 For each hotspot, we performed both pairwise pleiotropy tests and linear regression-based 
-mediation analyses. For every pair of traits at a hotspot, we calculated a pleiotropy test 
+mediation analyses. For every pair of traits (local or nonlocal) at a hotspot, we calculated a pleiotropy test 
 statistic[@boehm2019testing]. Due to the computational burden, we didn't determine bootstrap p-values for these 
@@
 
-For $n$ mice, $\Sigma$ is a $2n$ by $2n$ covariance matrix, while $V_g$ and $V_e$ are $2$ by $2$ covariance matrices.
+For $n$ mice and two traits, $\Sigma$ is a $2n$ by $2n$ covariance matrix, while $V_g$ and $V_e$ are $2$ by $2$ covariance matrices (\@ref(eq:Sigmahat)).
 Estimates $\hat V_g$ and $\hat V_e$ are obtained via an expectation-maximization algorithm [@zhou2014efficient;@gemma2].
@@
 From the two-dimensional scan results, we calculated pleiotropy likelihood ratio test (LOD) statistics for each
-pair of traits [@jiang1995multiple;@boehm2019testing;@boehm2019thesis]. Because of the required
-computing time for bootstrap analyses, we didn't obtain p-values for the pleiotropy tests. Instead,
+pair of traits [@jiang1995multiple;@boehm2019testing;@boehm2019thesis]. Because of computing limitations, we didn't obtain bootstrap p-values for the pleiotropy tests. Instead,
 we worked directly with the test statistic values.
@@
 effects, $M$ is a vector of values for a single candidate mediator (a local expression trait, in our case), $\beta$ is the effect
-of the mediator on the complex trait (after adjusting for other covariates in the model), 1 is a vector with all entries set to 1, $\alpha$
+of the mediator on the complex trait (after adjusting for other covariates in the model), $1$ is a vector with all entries set to 1, $\alpha$
 is a parameter that corresponds to the mean trait value (after adjusting for other covariates in the model), and $E$ is a vector of normally distributed random errors.
@@
 To create a heatmap, we first performed two hierarchical clusterings of rows. 
-One hierarchical clustering involved only the rows containing local traits, while the other contained only 
+One hierarchical clustering involved only the rows for local traits, while the other contained only 
 rows corresponding to nonlocal traits.

### 2020-03-18 analysis/paper/Rmd/hotspot-one-chromosome-TIMBR.Rmd
added TIMBR R code

```{r setup}
library(qtl2hotspots)
ult <- 15 # univariate lod threshold
```

```{r load_tidyverse}
library(tidyverse)
options(tibble.print_max = 50, tibble.print_min = 20)
```

```{r saveRDS}
if (!file.exists(here::here("analysis", "data", "derived_data", "genoprobs.rds"))){
  download.file("https://datadryad.org/stash/downloads/file_stream/63306",
              destfile = here::here("analysis", "data", "raw_data", "Attie_DO378_eQTL_viewer_v1.Rdata")
              )
  load(here::here("analysis", "data", "raw_data", "Attie_DO378_eQTL_viewer_v1.Rdata"))
  saveRDS(dataset.islet.rnaseq$lod.peaks, here::here("analysis", "data", "derived_data", "lod_peaks.rds"))
  saveRDS(dataset.islet.rnaseq$annots, here::here("analysis", "data", "derived_data", "annots.rds"))
  saveRDS(dataset.islet.rnaseq$expr, here::here("analysis", "data", "derived_data", "expr.rds"))
  saveRDS(K, here::here("analysis", "data", "derived_data", "kinship.rds"))
  saveRDS(map, here::here("analysis", "data", "derived_data", "map.rds"))
  saveRDS(genoprobs, here::here("analysis", "data", "derived_data", "genoprobs.rds"))
  saveRDS(dataset.islet.rnaseq$covar, here::here("analysis", "data", "derived_data", "covar.rds"))
}
```

```{r readRDS}
lod_peaks <- readRDS(here::here("analysis", "data", "derived_data", "lod_peaks.rds"))
annots <- readRDS(here::here("analysis", "data", "derived_data", "annots.rds"))
expr <- readRDS(here::here("analysis", "data", "derived_data", "expr.rds"))
probs <- readRDS(here::here("analysis", "data", "derived_data", "genoprobs.rds"))
K <- readRDS(here::here("analysis", "data", "derived_data", "kinship.rds"))
covar <- readRDS(here::here("analysis", "data", "derived_data", "covar.rds"))
map <- readRDS(here::here("analysis", "data", "derived_data", "map.rds"))
```

```{r}
hot_chr <- 2
hot_mid <- 165.5
keller_mediator <- "Hnf4a"
```

```{r}
hot_lower <- hot_mid - 2
hot_upper <- hot_mid + 2
```

```{r, define_local_annot}
local_annot <- lod_peaks %>%
  filter(chrom == hot_chr, pos <= hot_upper, pos >= hot_lower) %>%
  left_join(y = annots, by = c("annot.id" = "gene_id")) %>%
  filter(!duplicated(annot.id)) %>%
  filter(lod >= ult) %>%
  mutate(local_tx = chr == hot_chr) %>%
  filter(local_tx)
```

```{r, define_nonlocal_annot}
nonlocal_annot <- lod_peaks %>%
  filter(chrom == hot_chr, pos <= hot_upper, pos >= hot_lower) %>%
  inner_join(y = annots, by = c("annot.id" = "gene_id")) %>%
  filter(!duplicated(annot.id)) %>%
  filter(lod >= 7.18) %>%
  mutate(local_tx = chr == hot_chr) %>%
  filter(!local_tx) %>%
  filter(!is.na(hotspot))
```

```{r, local_annot2}
local_annot2 <- local_annot %>%
  filter(middle <= hot_upper, middle >= hot_lower) %>% 
  arrange(desc(lod))
```

```{r hot_expr}
hot_local <- expr[, colnames(expr) %in% local_annot2$annot.id]
hot_nonlocal <- expr[, colnames(expr) %in% nonlocal_annot$annot.id]
nnonlocal <- nrow(nonlocal_annot)
nlocal <- nrow(local_annot2)
```

```{r, hot_annot-hot_chr}
hot_annot <- local_annot2 %>%
  bind_rows(nonlocal_annot) %>%
  dplyr::select(annot.id, symbol)
```

```{r, scan1}
fn <- here::here("analysis", "data", "derived_data", "Chr2_scan1_out.rds")
if (file.exists(fn)) {
  s1out <- readRDS(fn)
} else {
  s1out <- qtl2::scan1(genoprobs = probs, 
                       pheno = cbind(hot_local, hot_nonlocal), 
                       kinship = K[["{{hot_chr}}"]], 
                       addcovar = covar, 
                       reml = TRUE, 
                       cores = 1
                       )
  saveRDS(s1out, fn)
}
```

```{r allele-effects}
fn <- here::here("analysis", "data", "derived_data", "Chr2_scan1coef_out.rds")
if (file.exists(fn)) {
  s1coef_out <- readRDS(fn)
} else {
s1coef_out <- purrr::map(.x = as_tibble(cbind(hot_local, hot_nonlocal)), 
           .f = function(pheno) {
             ph <- as.matrix(pheno)
             rownames(ph) <- rownames(expr)
             qtl2::scan1coef(genoprobs = list(probs[[hot_chr]]), 
                             pheno = ph, 
                             kinship = K[[hot_chr]], 
                             addcovar = covar, 
                             reml = TRUE, 
                             cores = 1
                             )
           }
                )
saveRDS(s1coef_out, fn)
}
```

```{r, get-allele-effects}
# add index values for peak position
local_nonlocal_annot <- dplyr::bind_rows(local_annot2, nonlocal_annot) %>%
  mutate(peak_index = match(table = names(map[[hot_chr]]), x = marker.id)) 
```

### TIMBR analyses

```{r, timbr-setup}
tr <- cbind(hot_local, hot_nonlocal)
prior_M <- list(model.type = "crp",
                prior.alpha.type = "gamma",
                prior.alpha.shape = 2.3009322,
                prior.alpha.rate = 0.7488104
                )
library(TIMBR)
data(mcv.data) # get A matrix
peaks <- qtl2::find_peaks(s1out, map)
```

```{r}
fn <- "attie_DO500_genoprobs_v5_36state_chr2.rds"
gp <- readRDS(here::here("analysis", "data", "derived_data", fn))
```

```{r, timbr}
pheno <- cbind(hot_local, hot_nonlocal)
set.seed(3411192) # to ensure getting the same samples with TIMBR
t_out <- peaks %>%
  dplyr::filter(chr == hot_chr, pos <= hot_upper, pos >= hot_lower) %>%
  dplyr::mutate(marker_index = purrr::map2_int(.x = pos, .y = chr, .f = function(x, y) which(map[[y]] == x))) %>%
  dplyr::mutate(t_out = purrr::pmap(.l = list(lodcolumn, chr, marker_index), .f = function(lodcolumn, chr, marker_index){
    phe_pre <- pheno[ , colnames(pheno) == lodcolumn, drop = FALSE]
    phe <- phe_pre[!is.na(phe_pre), , drop = FALSE]
    pr_pre <- gp[[1]][, , marker_index] # use first (and only!) entry in gp
    pr <- pr_pre[rownames(pr_pre) %in% rownames(phe), ]
    cov_pre <- cbind(1, covar)
    cov_pre2 <- cov_pre[rownames(cov_pre) %in% rownames(phe), ]
    cov <- qtl2:::drop_depcols(cov_pre2)
    TIMBR::TIMBR(y = phe, 
                 prior.D = list(P = pr, 
                                A = mcv.data$prior.D$A, # works for autosomes
                                fixed.diplo = FALSE), 
                 prior.M = prior_M, 
                 samples = 10000, 
                 samples.ml = 10000,
                 Z = cov, 
                 verbose = TRUE
                 )  
}))
saveRDS(t_out, here::here("analysis", "data", "derived_data", "timbr-chr2-10ksamples.rds"))
```

### 2020-03-18 analysis/paper/Rmd/intro.Rmd
added TIMBR R code

@@
 genomic regions that affect complex quantitative traits, such as systolic blood pressure, body 
-weight, and, more recently, biomolecular traits such as gene expression levels and protein 
-concentrations [@sax1923association;@soller1976power;@lander1989mapping;@broman2009guide;@jansen2007quantitative;@chick2016defining]. 
+weight, and, more recently, biomolecular traits such as gene expression levels and protein levels [@sax1923association;@soller1976power;@lander1989mapping;@broman2009guide;@jansen2007quantitative;@chick2016defining;@keller2018genetic].
 These genomic regions that affect variability in one or more complex traits are called "quantitative trait loci" or "QTL".
@@
 challenging and may require costly and time-consuming experiments.
+
 The growing need for greater QTL mapping resolution fueled development over the
@@
 
-Other researchers have identified and examined QTL hotspots in model organism genetics studies. @tian2016dissection 
+Other researchers have identified and examined QTL hotspots in model organism genetics studies [@neto2012quantile]. @tian2016dissection 
 presented quantitative methods for hotspot dissection in F$_2$ intercross mice.
@@
 hotspots in a publicly available data set. At each hotspot, we identify 
-a set of local traits that have QTL within the hotspot. We also identify a set of nonlocal traits with QTL 
+a set of local traits that have QTL within the hotspot region. We also identify a set of nonlocal traits with QTL 
 arising from the hotspot. We perform pleiotropy tests and mediation analyses for all local-nonlocal trait pairs. 
@@
 - who else has studied hotspots? And what did they do?
+- reread intros of Neto 2012 and Tian 2016

### 2020-03-25 analysis/paper/Rmd/hotspot-one-chromosome-TIMBR-Chr2.Rmd
set up for using child processes with TIMBR analyses

--- 2020-03-18 analysis/paper/Rmd/hotspot-one-chromosome-TIMBR.Rmd
@@
+---
+title: "QTL mapping and TIMBR for Chr 2"
+author: "Frederick J. Boehm"
+date: "03/18/2020"
+output: html_document
+params:
+  lastmod: !r lubridate::now()
+---
+
+Last modified: `r params$lastmod`.
+
+```{r setupx, include=FALSE, dev="CairoPNG"}
+.libPaths("~/work/libs")
+knitr::opts_chunk$set(dev="CairoPNG")
+```
+
 ```{r setup}
@@
 local_annot <- lod_peaks %>%
-  filter(chrom == hot_chr, pos <= hot_upper, pos >= hot_lower) %>%
-  left_join(y = annots, by = c("annot.id" = "gene_id")) %>%
-  filter(!duplicated(annot.id)) %>%
-  filter(lod >= ult) %>%
-  mutate(local_tx = chr == hot_chr) %>%
-  filter(local_tx)
+  dplyr::filter(chrom == hot_chr, pos <= hot_upper, pos >= hot_lower) %>%
+  dplyr::left_join(y = annots, by = c("annot.id" = "gene_id")) %>%
+  dplyr::filter(!duplicated(annot.id)) %>%
+  dplyr::filter(lod >= ult) %>%
+  dplyr::mutate(local_tx = chr == hot_chr) %>%
+  dplyr::filter(local_tx)
 ```
@@
 nonlocal_annot <- lod_peaks %>%
-  filter(chrom == hot_chr, pos <= hot_upper, pos >= hot_lower) %>%
-  inner_join(y = annots, by = c("annot.id" = "gene_id")) %>%
-  filter(!duplicated(annot.id)) %>%
-  filter(lod >= 7.18) %>%
-  mutate(local_tx = chr == hot_chr) %>%
-  filter(!local_tx) %>%
-  filter(!is.na(hotspot))
+  dplyr::filter(chrom == hot_chr, pos <= hot_upper, pos >= hot_lower) %>%
+  dplyr::inner_join(y = annots, by = c("annot.id" = "gene_id")) %>%
+  dplyr::filter(!duplicated(annot.id)) %>%
+  dplyr::filter(lod >= 7.18) %>%
+  dplyr::mutate(local_tx = chr == hot_chr) %>%
+  dplyr::filter(!local_tx) %>%
+  dplyr::filter(!is.na(hotspot))
 ```
@@
 local_annot2 <- local_annot %>%
-  filter(middle <= hot_upper, middle >= hot_lower) %>% 
-  arrange(desc(lod))
+  dplyr::filter(middle <= hot_upper, middle >= hot_lower) %>% 
+  dplyr::arrange(desc(lod))
 ```
@@
 hot_annot <- local_annot2 %>%
-  bind_rows(nonlocal_annot) %>%
+  dplyr::bind_rows(nonlocal_annot) %>%
   dplyr::select(annot.id, symbol)
@@
 ```
+
+```{r}
+t_out <- readRDS(here::here("analysis", "data", "derived_data", "timbr-chr2-10ksamples.rds")) %>%
+  dplyr::rename(timbr = t_out)
+```

### 2020-03-25 analysis/paper/Rmd/hotspot-one-chromosome-timbr.Rmd
set up for using child processes with TIMBR analyses

--- 2019-09-29 analysis/paper/Rmd/hotspot-one-chromosome.Rmd
@@
 hot_upper <- {{hot_mid}} + 2
-```
-
-```{r, readRDS-{{hot_chr}}}
-pleio_mixed_list <- readRDS(here::here("analysis", "data", "derived_data", paste0("Chr", {{hot_chr}}, "_pleio.rds")))
-pleio_local_list <- readRDS(here::here("analysis", "data", "derived_data", paste0("Chr", {{hot_chr}}, "_pleio_local.rds")))
-pleio_nonlocal_list <- readRDS(here::here("analysis", "data", "derived_data", paste0("Chr", {{hot_chr}}, "_pleio_nonlocal.rds")))
-med_list <- readRDS(here::here("analysis", "data", "derived_data", paste0("Chr", {{hot_chr}}, "_mediations.rds")))
 ```
@@
 
-```{r, med_tib-{{hot_chr}}, message=FALSE, warning = FALSE}
-med_tib <- med_list %>%
-  purrr::map(bind_rows) %>%
-  bind_rows() %>%
-  as_tibble() %>%
-  rename(lod_without_med = lod_no_med, lod_with_med = lod_med) %>%
-  mutate(lod_diff = lod_without_med - lod_with_med, lod_diff_proportion = lod_diff / lod_without_med) %>%
-  left_join(local_annot2, by = c("local_gene_id" = "annot.id")) %>%
-  left_join(nonlocal_annot, by = c("nonlocal_gene_id" = "annot.id")) %>%
-  dplyr::select(symbol.x, symbol.y, lod_without_med, lod_with_med, lod_diff, lod_diff_proportion, local_gene_id, nonlocal_gene_id) %>%
-  rename(local_symbol = symbol.x, nonlocal_symbol = symbol.y) %>%
-  filter(!is.na(local_symbol))
-```
-
 ```{r, hot_annot-{{hot_chr}}}
@@
   dplyr::select(annot.id, symbol)
-pleio_tib <- pleio_mixed_list %>%
-  purrr::map(.f = function(x) rename(x, gene1_id = local_gene_id, gene2_id = nonlocal_gene_id)) %>%
-  c(pleio_local_list, pleio_nonlocal_list) %>%
-  bind_rows() %>%
-  rename(pleiotropy_lod = lod) %>%
-  inner_join(hot_annot, by = c("gene1_id" = "annot.id")) %>%
-  rename(gene1_symbol = symbol) %>%
-  inner_join(hot_annot, by = c("gene2_id" = "annot.id")) %>%
-  rename(gene2_symbol = symbol) %>%
-  dplyr::select(gene1_symbol, gene2_symbol, pleiotropy_lod)
-```
-
-```{r pm-{{hot_chr}}}
-pm <- pleio_tib %>%
-  arrange(gene1_symbol, gene2_symbol) %>%
-  tibble_to_matrix(symmetric = TRUE)
-```
-
-```{r, scatter-{{hot_chr}}, fig.cap="Pleiotropy test statistic vs. mediation LOD difference for Chr {{hot_chr}} hotspot."}
-p <- pleio_mixed_list %>%
-  bind_rows() %>%
-  rename(pleiotropy_lod = lod) %>%
-  inner_join(med_tib, by = c("local_gene_id", "nonlocal_gene_id")) %>%
-  dplyr::select(local_symbol, nonlocal_symbol, pleiotropy_lod, lod_diff, lod_diff_proportion, lod_without_med, lod_with_med, filename, nonlocal_gene_id, local_gene_id) %>%
-  ggplot() + geom_point(mapping = aes(x = pleiotropy_lod, y = lod_diff, colour = local_symbol == "{{keller_mediator}}", nonlocal = nonlocal_symbol, local = local_symbol), size = 0.05) + theme(legend.position = "none") + xlab("Pleiotropy LOD") + ylab("Mediation LOD difference")
-plotly::ggplotly(p)
-```
-
-```{r, scatter-by-local-{{hot_chr}}, fig.height = ceiling(nlocal / 2), fig.cap="Pleiotropy test statistic vs. mediation LOD difference by local expression trait for Chr {{hot_chr}} hotspot."}
-foo <- pleio_mixed_list %>%
-  bind_rows() %>%
-  rename(pleiotropy_lod = lod) %>%
-  inner_join(med_tib, by = c("local_gene_id", "nonlocal_gene_id")) %>%
-  dplyr::select(local_symbol, nonlocal_symbol, pleiotropy_lod, lod_diff, lod_diff_proportion, lod_without_med, lod_with_med, filename, nonlocal_gene_id, local_gene_id)
-p <- foo %>%
-  ggplot() +
-  geom_point(mapping = aes(x = pleiotropy_lod, y = lod_diff, colour = nonlocal_symbol), size = 0.1) +
-  facet_wrap(. ~ local_symbol, nrow = ceiling(nlocal / 2), ncol = 2) +
-  broman::karl_theme(strip.placement = "inside") +
-  theme(legend.position = "none") +
-  theme(
-    strip.background = element_blank(),
-    strip.text.x = element_blank()
-) +
-  xlab("Pleiotropy LOD") +
-  ylab("Mediation LOD difference") +
-  geom_text(aes(x = max(foo$pleiotropy_lod), y = 1, label = local_symbol), inherit.aes=FALSE)
-plotly::ggplotly(p)
-```
-
-```{r, scols-{{hot_chr}}}
-scols1 <- tibble(local = 1L + (rownames(pm) %in% local_annot2$symbol), 
-                 symbol = rownames(pm), 
-                 mediated_by_keller_mediator = rownames(pm) %in% (med_tib %>% filter(local_symbol == "{{keller_mediator}}", lod_diff > 1.5))$nonlocal_symbol
-                 )
-```
-
-```{r, ordering-{{hot_chr}}}
-# order the nonlocal traits first
-## indicator for being a nonlocal trait
-indic_nonlocal <- rownames(pm) %in% nonlocal_annot$symbol
-# define ord_nonlocal
-ord_nonlocal <- pm[indic_nonlocal, ] %>%
-  dist() %>%
-  hclust() %>%
-  as.dendrogram() %>%
-  dendextend::seriate_dendrogram(dist(pm[indic_nonlocal, ])) %>%
-  order.dendrogram()
-rn_nonlocal <- rownames(pm[indic_nonlocal, ][ord_nonlocal, ])
-# define ord_local
-ord_local<- pm[!indic_nonlocal, ] %>%
-  dist() %>%
-  hclust() %>%
-  as.dendrogram() %>%
-  dendextend::seriate_dendrogram(dist(pm[!indic_nonlocal, ])) %>%
-  order.dendrogram()
-rn_local <- rownames(pm[!indic_nonlocal, ][ord_local, ])
-rn <- c(rn_nonlocal, rn_local) # order for dendrogram
 ```
@@
 
-```{r allele-effects-{{hot_chr}}}
-fn <- here::here("analysis", "data", "derived_data", "Chr{{hot_chr}}_scan1coef_out.rds")
-if (file.exists(fn)) {
-  s1coef_out <- readRDS(fn)
-} else {
-s1coef_out <- purrr::map(.x = as_tibble(cbind(hot_local, hot_nonlocal)), 
-           .f = function(pheno) {
-             ph <- as.matrix(pheno)
-             rownames(ph) <- rownames(expr)
-             qtl2::scan1coef(genoprobs = list(probs[["{{hot_chr}}"]]), 
-                             pheno = ph, 
-                             kinship = K[["{{hot_chr}}"]], 
-                             addcovar = covar, 
-                             reml = TRUE, 
-                             cores = 1
+### TIMBR analyses
+
+```{r, timbr-setup}
+tr <- cbind(hot_local, hot_nonlocal)
+prior_M1 <- list(model.type = "crp", # crp - Chinese Restaurant Process
+  prior.alpha.type = "gamma",
+  prior.alpha.shape = 1,
+  prior.alpha.rate = 2.333415)
+prior_M2 <- list(model.type = "crp",
+                prior.alpha.type = "gamma",
+                prior.alpha.shape = 2.3009322,
+                prior.alpha.rate = 0.7488104
                 )
-           }
-                )
-saveRDS(s1coef_out, fn)
-}
+library(TIMBR)
+data(mcv.data) # get A matrix
+peaks <- qtl2::find_peaks(s1out, map)
 ```
 
-```{r, get-allele-effects-{{hot_chr}}}
-# add index values for peak position
-local_nonlocal_annot <- dplyr::bind_rows(local_annot2, nonlocal_annot) %>%
-  mutate(peak_index = match(table = names(map[["{{hot_chr}}"]]), x = marker.id)) %>%
-  mutate(peak_effects = purrr::map2(.f = qtl2pleio::get_effects, .x = peak_index, .y = s1coef_out, map = map[["{{hot_chr}}"]])) %>%
-  qtl2hotspots::listcol_to_cols(list_column = quo(peak_effects), another_column = quo(symbol)) %>%
-  rename(cA = `1`, cB = `2`, cC = `3`, cD = `4`, cE = `5`, cF = `6`, cG = `7`, cH = `8`
-         )
+```{r}
+fn <- "attie_DO500_genoprobs_v5_36state_chr{{hot_chr}}.rds"
+gp <- readRDS(here::here("analysis", "data", "derived_data", fn))
 ```
 
-```{r, n-med-{{hot_chr}}}
-n_med_tr <- 10 # arbitrary threshold for number of mediated traits per local
-# only those local traits with n_med_tr or more mediated nonlocal traits 
-# appear as annotation columns in the heatmap
-n_med_per_local <- med_tib %>%
-  group_by(local_symbol) %>%
-  filter(lod_diff > 1.5) %>%
-  count() %>%
-  filter(n >= n_med_tr)
-n_med_per_local %>%
-  arrange(desc(n)) %>%
-  knitr::kable(caption = "Number of mediated traits per local trait for Chr {{hot_chr}} hotspot.")
+```{r, timbr}
+pheno <- cbind(hot_local, hot_nonlocal)
+set.seed(3411192) # to ensure getting the same samples with TIMBR
+future::plan(cluster)
+t_out <- peaks %>%
+  dplyr::filter(chr == hot_chr, pos <= hot_upper, pos >= hot_lower) %>%
+  dplyr::mutate(marker_index = purrr::map2_int(.x = pos, .y = chr, .f = function(x, y) which(map[[y]] == x))) %>%
+  dplyr::mutate(t_out = furrr::future_pmap(.l = list(lodcolumn, chr, marker_index), .f = function(lodcolumn, chr, marker_index){
+    phe_pre <- pheno[ , colnames(pheno) == lodcolumn, drop = FALSE]
+    phe <- phe_pre[!is.na(phe_pre), , drop = FALSE]
+    pr_pre <- gp[[1]][, , marker_index] # use first (and only!) entry in gp
+    pr <- pr_pre[rownames(pr_pre) %in% rownames(phe), ]
+    cov_pre <- cbind(1, covar)
+    cov_pre2 <- cov_pre[rownames(cov_pre) %in% rownames(phe), ]
+    cov <- qtl2:::drop_depcols(cov_pre2)
+    TIMBR::TIMBR(y = phe, 
+                 prior.D = list(P = pr, 
+                                A = mcv.data$prior.D$A, # works for autosomes
+                                fixed.diplo = FALSE), 
+                 prior.M = prior_M1, 
+                 samples = 10000, 
+                 samples.ml = 10000,
+                 Z = cov, 
+                 verbose = TRUE
+                 )  
+}))
+saveRDS(t_out, here::here("analysis", "data", "derived_data", "timbr-chr{{hot_chr}}-10ksamples-M1.rds"))
 ```
-
-```{r, prepare-annotations-{{hot_chr}}}
-# need 5 annotation object: 1.
-local v nonlocal, 2.
-mediation LOD difference for all local traits, 3.
-mediation LOD difference proportion for all local traits, 4.
-8 allele effects per trait, 5.
-peak position per trait, as a one-column heatmap
-# 2,3,4 are numeric; 1 & 5 are binary
-## local_nonlocal_annot already has the allele effect and local v nonlocal; binning has the peak position. Need to join these two on symbol
-### med_tib has lod diff & lod diff proportion for all local - nonlocal pairs
-fake_tibble_mediation <- tidyr::crossing(symbol = local_annot2$symbol, 
-                                         symbol2 = local_annot2$symbol) %>%
-  mutate(dat = NA) %>%
-  spread(key = symbol, value = dat) %>%
-  rename(symbol = symbol2)
-# to add to mediation tibbles
-
-# first med tibble
-lod_diffs_tibble <- med_tib %>%
-  select( - c(lod_without_med, 
-              lod_with_med, 
-              lod_diff_proportion, 
-              local_gene_id, 
-              nonlocal_gene_id)
-          ) %>%
-  tidyr::spread(key = local_symbol, 
-                value = lod_diff) %>% 
-  dplyr::rename(symbol = nonlocal_symbol) %>%
-  dplyr::bind_rows(fake_tibble_mediation)
-# second med tibble
-lod_diff_proportions_tibble <- med_tib %>%
-  select( - c(lod_without_med, 
-              lod_with_med, 
-              lod_diff, 
-              local_gene_id, 
-              nonlocal_gene_id)
-          ) %>%
-  tidyr::spread(key = local_symbol, 
-                value = lod_diff_proportion) %>% # 147 rows by 32 columns
-  rename(symbol = nonlocal_symbol) %>%
-  bind_rows(fake_tibble_mediation)
-# effects tibble
-allele_effects_tibble <- local_nonlocal_annot %>%
-  dplyr::select(symbol, 18:25)
-# peak positions tibble
-peak_positions <- local_nonlocal_annot %>%
-  select(symbol, pos)
-# local indicator
-local_tibble <- local_nonlocal_annot %>%
-  select(symbol, local_tx) %>% 
-  ungroup()
-# univariate lods
-lod_tibble <- local_nonlocal_annot %>%
-  select(symbol, lod) %>%
-  ungroup()
-```
-
-```{r, order-annotations-{{hot_chr}}}
-# all annots 
-annots_ord <- list(lod_diffs_tibble, 
-                   lod_diff_proportions_tibble, 
-                   allele_effects_tibble,
-                   peak_positions, 
-                   local_tibble,
-                   lod_tibble 
-                    ) %>% 
-  purrr::map(.f = function(df, rn){
-    out <- df[match(x = rn, table = df$symbol), ]
-    return(out)
-  }, 
-  rn = rn)
-```
-
-```{r, heat-{{hot_chr}}, fig.cap="Heatmap for pleiotropy test statistics for Chr {{hot_chr}}."}
-ord <- match(x = rn, table = rownames(pm))
-
-pm[ord, ord] %>% # order both cols and rows by 'ord'
-  apply(FUN = trunc2, MARGIN = 2, threshold = 5) %>%
-  pheatmap::pheatmap(cluster_rows = FALSE,
-                     cluster_cols = FALSE,
-                     fontsize = 2
-                     )
-```
-
-```{r, index-v-position-{{hot_chr}}, fig.cap="QTL peak position vs. dendrogram index for Chr {{hot_chr}}."}
-bar <- foo %>%
-  filter(local_symbol == "{{keller_mediator}}")
-
-p <- tibble(nonlocal_symbol = rownames(pm[indic_nonlocal, ])[ord_nonlocal], index = 1:length(nonlocal_symbol)) %>%
-  inner_join(bar, by = "nonlocal_symbol") %>% 
-  left_join(nonlocal_annot, by = c("nonlocal_symbol" = "symbol")) %>%
-  ggplot() + geom_point(aes(y = pos, x = index, color = lod_diff > 1.5, nonlocal = nonlocal_symbol, lod_diff = lod_diff))
-
-plotly::ggplotly(p)
-```
-
-```{r {{hot_chr}}-ann3}
-ann3 <- annots_ord[[3]] %>%
-  ungroup() %>%
-  select(- symbol) %>%
-  as.matrix() %>%
-  (function(x){rownames(x) <- annots_ord[[3]]$symbol; return(x)})
-colnames(ann3) <- LETTERS[1:8]
-```
-
-```{r {{hot_chr}}-ann2}
-ann2 <- annots_ord[[2]] %>%
-  select(-symbol) %>%
-  as.matrix() %>%
-  (function(x){rownames(x) <- annots_ord[[2]]$symbol; return(x)})
-```
-
-```{r {{hot_chr}}-ann1}
-ann1 <- annots_ord[[1]] %>%
-  select(-symbol) %>%
-  as.matrix() %>%
-  (function(x){rownames(x) <- annots_ord[[1]]$symbol; return(x)})
-```
-
-```{r {{hot_chr}}-ann5}
-ann5 <- annots_ord[[5]] %>%
-  select(-symbol) %>%
-  as.matrix() %>%
-  (function(x){rownames(x) <- annots_ord[[5]]$symbol; return(x)})
-```
-
-```{r {{hot_chr}}-ann6}
-ann6 <- annots_ord[[6]] %>%
-  select(-symbol) %>%
-  as.matrix() %>%
-  (function(x){rownames(x) <- annots_ord[[6]]$symbol; return(x)})
-```
-
-```{r {{hot_chr}}-ann4}
-ann4 <- annots_ord[[4]] %>%
-  ungroup() %>%
-  select(-symbol) %>%
-  as.matrix() %>%
-  (function(x){rownames(x) <- annots_ord[[4]]$symbol; return(x)})
-```
-
-```{r iheatmapr-{{hot_chr}}, out.width="25in", out.height = "10in"}
-mymat <- pm[ord, ord] %>% # order both cols and rows by 'ord'
-  apply(FUN = trunc2, MARGIN = 2, threshold = 5)
-iheatmapr::main_heatmap(mymat, show_colorbar = FALSE, layout = list(font = list(size = 4))) %>%
-  iheatmapr::add_col_labels() %>%
-  iheatmapr::add_col_annotation(ann5) %>%
-  iheatmapr::add_main_heatmap(ann3, name = "signal3", size = 0.2, show_colorbar = FALSE) %>%
-    iheatmapr::add_col_labels() %>%
-
-  iheatmapr::add_main_heatmap(ann2, name = "signal2", size = 0.5, show_colorbar = FALSE) %>%
-  iheatmapr::add_col_labels() %>%
-  iheatmapr::add_main_heatmap(ann1, name = "signal1", size = 0.5, show_colorbar = FALSE) %>%
-  iheatmapr::add_row_annotation(ann5, side = "left") %>%
-    iheatmapr::add_col_labels() %>%
-  iheatmapr::add_row_annotation(ann4, side = "left") %>%
-  iheatmapr::add_row_annotation(ann6, side = "left") %>%
-  iheatmapr::add_row_labels()
-```

### 2020-03-25 analysis/paper/Rmd/hotspots-paper-timbr.Rmd
set up for using child processes with TIMBR analyses

--- 2019-09-29 analysis/paper/Rmd/hotspots-paper.Rmd
@@
 author: "Frederick J. Boehm, Brian S. Yandell, and Karl W. Broman"
-date: "August 3, 2019"
+date: "March 25, 2020"
 bibliography: research.bib
@@
 
-```{r child = "abstract.Rmd", eval = TRUE}
-```
-
-```{r child="intro.Rmd", eval = TRUE}
+```{r setupx, include=FALSE, dev="CairoPNG"}
+.libPaths("~/work/libs")
+knitr::opts_chunk$set(dev="CairoPNG")
 ```
@@
 knitr::opts_chunk$set(comment=NA, fig.width=6, fig.height=6, eval = eval_code)
-```
-
-```{r child="methods.Rmd", eval = TRUE}
 ```
@@
 inputs <- tibble(hot_chr, hot_mid, keller_mediator)
-out <- pmap_dfc(inputs, .f = function(hot_chr, hot_mid, keller_mediator) knitr::knit_expand(text = readLines(here::here("analysis", "paper", "Rmd", "hotspot-one-chromosome.Rmd"))))
+out <- pmap_dfc(inputs, .f = function(hot_chr, hot_mid, keller_mediator) knitr::knit_expand(text = readLines(here::here("analysis", "paper", "Rmd", "hotspot-one-chromosome-timbr.Rmd"))))
 # note that ... argument must be a list, ie, it must be a single argument, not multiple args
@@
 `r if (eval_code) knitr::knit(text = out[[5]])`
-
-```{r read_results, child = "results.Rmd", eval = TRUE}
-```
-
-```{r read_discussion, child = "discussion.Rmd", eval = TRUE}
-```
 

### 2020-03-25 analysis/paper/Rmd/hotspot-one-chromosome-timbr.Rmd
added quotations to cluster argument

@@
 set.seed(3411192) # to ensure getting the same samples with TIMBR
-future::plan(cluster)
+future::plan("cluster")
 t_out <- peaks %>%

### 2020-03-25 analysis/paper/Rmd/hotspot-one-chromosome-timbr.Rmd
cluster -> multiprocess

--- 2020-03-25 analysis/paper/Rmd/hotspot-one-chromosome-timbr.Rmd
@@
 set.seed(3411192) # to ensure getting the same samples with TIMBR
-future::plan(cluster)
+future::plan("multiprocess")
 t_out <- peaks %>%

### 2020-03-25 analysis/paper/Rmd/hotspot-one-chromosome-timbr.Rmd
updated future.globals.maxSize

--- 2020-03-25 analysis/paper/Rmd/hotspot-one-chromosome-timbr.Rmd
@@
 
-```{r, timbr}
+```{r}
+library(future)
+options(future.globals.maxSize= 891289600)
+```
+
+```{r}
 pheno <- cbind(hot_local, hot_nonlocal)
 set.seed(3411192) # to ensure getting the same samples with TIMBR
-future::plan(cluster)
+future::plan("multiprocess")
 t_out <- peaks %>%

### 2020-03-26 analysis/paper/Rmd/intro.Rmd
updated intro

@@
 arising from the hotspot. We perform pleiotropy tests and mediation analyses for all local-nonlocal trait pairs. 
-Graphical summaries of our results enable us to conclude that at least one of the five hotspots is due to multiple QTL, 
+Graphical summaries of our results enable us to conclude that one of the five hotspots is due to multiple QTL, 
 while the evidence at the four other hotspots is consistent with a single local trait affecting most or all of the 

### 2020-03-26 analysis/paper/Rmd/hotspot-one-chromosome-timbr.Rmd
added dplyr:: prefix; deleted unneeded code

--- 2020-03-25 analysis/paper/Rmd/hotspot-one-chromosome-timbr.Rmd
@@
 local_annot <- lod_peaks %>%
-  filter(chrom == {{hot_chr}}, pos <= hot_upper, pos >= hot_lower) %>%
-  left_join(y = annots, by = c("annot.id" = "gene_id")) %>%
-  filter(!duplicated(annot.id)) %>%
-  filter(lod >= ult) %>%
-  mutate(local_tx = chr == {{hot_chr}}) %>%
-  filter(local_tx)
+  dplyr::filter(chrom == {{hot_chr}}, pos <= hot_upper, pos >= hot_lower) %>%
+  dplyr::left_join(y = annots, by = c("annot.id" = "gene_id")) %>%
+  dplyr::filter(!duplicated(annot.id)) %>%
+  dplyr::filter(lod >= ult) %>%
+  dplyr::mutate(local_tx = chr == {{hot_chr}}) %>%
+  dplyr::filter(local_tx)
 ```
@@
 nonlocal_annot <- lod_peaks %>%
-  filter(chrom == {{hot_chr}}, pos <= hot_upper, pos >= hot_lower) %>%
-  inner_join(y = annots, by = c("annot.id" = "gene_id")) %>%
-  filter(!duplicated(annot.id)) %>%
-  filter(lod >= 7.18) %>%
-  mutate(local_tx = chr == {{hot_chr}}) %>%
-  filter(!local_tx) %>%
-  filter(!is.na(hotspot))
+  dplyr::filter(chrom == {{hot_chr}}, pos <= hot_upper, pos >= hot_lower) %>%
+  dplyr::inner_join(y = annots, by = c("annot.id" = "gene_id")) %>%
+  dplyr::filter(!duplicated(annot.id)) %>%
+  dplyr::filter(lod >= 7.18) %>%
+  dplyr::mutate(local_tx = chr == {{hot_chr}}) %>%
+  dplyr::filter(!local_tx) %>%
+  dplyr::filter(!is.na(hotspot))
 ```
@@
 hot_annot <- local_annot2 %>%
-  bind_rows(nonlocal_annot) %>%
-  dplyr::select(annot.id, symbol)
-```
-
-```{r, scan1-{{hot_chr}}}
-fn <- here::here("analysis", "data", "derived_data", "Chr{{hot_chr}}_scan1_out.rds")
-if (file.exists(fn)) {
-  s1out <- readRDS(fn)
-} else {
-  s1out <- qtl2::scan1(genoprobs = probs, 
-                       pheno = cbind(hot_local, hot_nonlocal), 
-                       kinship = K[["{{hot_chr}}"]], 
-                       addcovar = covar, 
-                       reml = TRUE, 
-                       cores = 1
-                       )
-  saveRDS(s1out, fn)
-}
+  bind_rows(nonlocal_annot) 
 ```
@@
 ```{r, timbr-setup}
-tr <- cbind(hot_local, hot_nonlocal)
 prior_M1 <- list(model.type = "crp", # crp - Chinese Restaurant Process
@@
 data(mcv.data) # get A matrix
-peaks <- qtl2::find_peaks(s1out, map)
 ```
@@
 
-```{r, timbr}
+```{r}
+library(future)
+options(future.globals.maxSize= 10 * 1024 ^ 3)
+```
+
+```{r}
 pheno <- cbind(hot_local, hot_nonlocal)
 set.seed(3411192) # to ensure getting the same samples with TIMBR
-future::plan("cluster")
-t_out <- peaks %>%
-  dplyr::filter(chr == hot_chr, pos <= hot_upper, pos >= hot_lower) %>%
-  dplyr::mutate(marker_index = purrr::map2_int(.x = pos, .y = chr, .f = function(x, y) which(map[[y]] == x))) %>%
-  dplyr::mutate(t_out = furrr::future_pmap(.l = list(lodcolumn, chr, marker_index), .f = function(lodcolumn, chr, marker_index){
+future::plan("multiprocess")
+t_out <- hot_annot %>%
+  dplyr::filter(pos <= hot_upper, pos >= hot_lower) %>%
+  dplyr::mutate(marker_index = purrr::map_int(.x = pos, .f = function(x) which(map[[{{hot_chr}}]] == x))) %>%
+  dplyr::mutate(t_out = furrr::future_pmap(.l = list(lodcolumn, marker_index), .f = function(lodcolumn, marker_index){
     phe_pre <- pheno[ , colnames(pheno) == lodcolumn, drop = FALSE]
@@
 }))
+dim(t_out)
 saveRDS(t_out, here::here("analysis", "data", "derived_data", "timbr-chr{{hot_chr}}-10ksamples-M1.rds"))

### 2020-03-26 analysis/paper/Rmd/hotspots-paper-timbr.Rmd
removed whitespace
added dplyr:: prefix; deleted unneeded code

@@
 
-```{r saveRDS}
-if (!file.exists(here::here("analysis", "data", "derived_data", "genoprobs.rds"))){
-  download.file("https://datadryad.org/stash/downloads/file_stream/63306",
-              destfile = here::here("analysis", "data", "raw_data", "Attie_DO378_eQTL_viewer_v1.Rdata")
-              )
-  load(here::here("analysis", "data", "raw_data", "Attie_DO378_eQTL_viewer_v1.Rdata"))
-  saveRDS(dataset.islet.rnaseq$lod.peaks, here::here("analysis", "data", "derived_data", "lod_peaks.rds"))
-  saveRDS(dataset.islet.rnaseq$annots, here::here("analysis", "data", "derived_data", "annots.rds"))
-  saveRDS(dataset.islet.rnaseq$expr, here::here("analysis", "data", "derived_data", "expr.rds"))
-  saveRDS(K, here::here("analysis", "data", "derived_data", "kinship.rds"))
-  saveRDS(map, here::here("analysis", "data", "derived_data", "map.rds"))
-  saveRDS(genoprobs, here::here("analysis", "data", "derived_data", "genoprobs.rds"))
-  saveRDS(dataset.islet.rnaseq$covar, here::here("analysis", "data", "derived_data", "covar.rds"))
-}
-```
-
 ```{r readRDS}
@@
 
-`r if (eval_code) knitr::knit(text = out[[2]])`
-
-`r if (eval_code) knitr::knit(text = out[[3]])`
-
-`r if (eval_code) knitr::knit(text = out[[4]])`
-
-`r if (eval_code) knitr::knit(text = out[[5]])`
-
 # Colophon

### 2020-03-26 analysis/paper/Rmd/hotspot-one-chromosome-timbr.Rmd
fixed inputted col name

@@
   dplyr::mutate(marker_index = purrr::map_int(.x = pos, .f = function(x) which(map[[{{hot_chr}}]] == x))) %>%
-  dplyr::mutate(t_out = furrr::future_pmap(.l = list(lodcolumn, marker_index), .f = function(lodcolumn, marker_index){
+  dplyr::mutate(t_out = furrr::future_pmap(.l = list(annot.id, marker_index), .f = function(lodcolumn, marker_index){
     phe_pre <- pheno[ , colnames(pheno) == lodcolumn, drop = FALSE]

### 2020-03-27 analysis/paper/Rmd/hotspot-one-chromosome-timbr.Rmd
fixed code for timbr analysis and setup

@@
 
-```{r, define_local_annot-{{hot_chr}}}
-local_annot <- lod_peaks %>%
+```{r, define_hot_annot-{{hot_chr}}}
+hot_annot <- lod_peaks %>%
   dplyr::filter(chrom == {{hot_chr}}, pos <= hot_upper, pos >= hot_lower) %>%
@@
   dplyr::filter(!duplicated(annot.id)) %>%
-  dplyr::filter(lod >= ult) %>%
-  dplyr::mutate(local_tx = chr == {{hot_chr}}) %>%
-  dplyr::filter(local_tx)
-```
-
-```{r, define_nonlocal_annot-{{hot_chr}}}
-nonlocal_annot <- lod_peaks %>%
-  dplyr::filter(chrom == {{hot_chr}}, pos <= hot_upper, pos >= hot_lower) %>%
-  dplyr::inner_join(y = annots, by = c("annot.id" = "gene_id")) %>%
-  dplyr::filter(!duplicated(annot.id)) %>%
   dplyr::filter(lod >= 7.18) %>%
   dplyr::mutate(local_tx = chr == {{hot_chr}}) %>%
-  dplyr::filter(!local_tx) %>%
-  dplyr::filter(!is.na(hotspot))
-```
-
-```{r, local_annot2-{{hot_chr}}}
-local_annot2 <- local_annot %>%
-  dplyr::filter(middle <= hot_upper, middle >= hot_lower) %>% 
-  dplyr::arrange(desc(lod))
+  dplyr::filter((!local_tx) | (lod >= ult & middle <= hot_upper & middle >= hot_lower))
 ```
@@
 ```{r hot_expr-{{hot_chr}}}
-hot_local <- expr[, colnames(expr) %in% local_annot2$annot.id]
-hot_nonlocal <- expr[, colnames(expr) %in% nonlocal_annot$annot.id]
-nnonlocal <- nrow(nonlocal_annot)
-nlocal <- nrow(local_annot2)
-```
-
-```{r, hot_annot-{{hot_chr}}}
-hot_annot <- local_annot2 %>%
-  bind_rows(nonlocal_annot) 
+hot_expr <- expr[, colnames(expr) %in% hot_annot$annot.id]
 ```
@@
 ### TIMBR analyses
-
-```{r, timbr-setup}
-prior_M1 <- list(model.type = "crp", # crp - Chinese Restaurant Process
-  prior.alpha.type = "gamma",
-  prior.alpha.shape = 1,
-  prior.alpha.rate = 2.333415)
-prior_M2 <- list(model.type = "crp",
-                prior.alpha.type = "gamma",
-                prior.alpha.shape = 2.3009322,
-                prior.alpha.rate = 0.7488104
-                )
-library(TIMBR)
-data(mcv.data) # get A matrix
-```
 
@@
 
-```{r}
-pheno <- cbind(hot_local, hot_nonlocal)
+```{r, timbr-{{hot_chr}}}
 set.seed(3411192) # to ensure getting the same samples with TIMBR
@@
 t_out <- hot_annot %>%
-  dplyr::filter(pos <= hot_upper, pos >= hot_lower) %>%
-  dplyr::mutate(marker_index = purrr::map_int(.x = pos, .f = function(x) which(map[[{{hot_chr}}]] == x))) %>%
-  dplyr::mutate(t_out = furrr::future_pmap(.l = list(annot.id, marker_index), .f = function(lodcolumn, marker_index){
-    phe_pre <- pheno[ , colnames(pheno) == lodcolumn, drop = FALSE]
+  dplyr::mutate(timbr = furrr::future_pmap(.l = list(annot.id, marker.id), .f = function(annot.id, marker.id){
+    phe_pre <- hot_expr[ , colnames(hot_expr) == annot.id, drop = FALSE]
     phe <- phe_pre[!is.na(phe_pre), , drop = FALSE]
-    pr_pre <- gp[[1]][, , marker_index] # use first (and only!) entry in gp
+    pr_pre <- gp[[1]][, , marker.id] # use first (and only!) entry in gp
     pr <- pr_pre[rownames(pr_pre) %in% rownames(phe), ]

### 2020-03-27 analysis/paper/Rmd/hotspots-paper-timbr.Rmd
added back the calls to knitr::knit for all 5 hotsptos
fixed code for timbr analysis and setup

@@
 
+```{r, timbr-setup}
+prior_M1 <- list(model.type = "crp", # crp - Chinese Restaurant Process
+  prior.alpha.type = "gamma",
+  prior.alpha.shape = 1,
+  prior.alpha.rate = 2.333415)
+prior_M2 <- list(model.type = "crp",
+                prior.alpha.type = "gamma",
+                prior.alpha.shape = 2.3009322,
+                prior.alpha.rate = 0.7488104
+                )
+library(TIMBR)
+data(mcv.data) # get A matrix
+```
+
 `r if (eval_code) knitr::knit(text = out[[1]])`
+
+`r if (eval_code) knitr::knit(text = out[[2]])`
+
+`r if (eval_code) knitr::knit(text = out[[3]])`
+
+`r if (eval_code) knitr::knit(text = out[[4]])`
+
+`r if (eval_code) knitr::knit(text = out[[5]])`
 

### 2020-05-19 analysis/paper/Rmd/intro.Rmd
updated intro

@@
 A genomic region with strong evidence of association with a complex quantitative trait, then, defines a QTL.
+
 Because nearby markers have correlated
@@
 last two decades of model organism multiparental populations for high-resolution QTL mapping [@de2017back;@churchill2004collaborative;@svenson2012high;@huang2012multiparent;@shivakumar2018soybean;@huang2011analysis;@kover2009multiparent;@tisne2017identification;@stanley2017genetic].
+In contrast to traditional, two-parent crosses, a multiparental population arises from more than two founders.
+The widely studied Collaborative Cross and Diversity Outbred mice share the same eight inbred founders.
 
@@
 
-Other researchers have identified and examined QTL hotspots in model organism genetics studies [@neto2012quantile]. @tian2016dissection 
-presented quantitative methods for hotspot dissection in F$_2$ intercross mice.
+Other researchers have identified and examined QTL hotspots in model organism genetics studies [@neto2012quantile]. 
+@tian2016dissection presented quantitative methods for hotspot dissection in F$_2$ intercross mice.
 
-@boehm2019thesis and @boehm2019testing argued for complementary roles for mediation analysis and pleiotropy testing in hotspot 
+@boehm2019thesis and @boehm2019testing argued for complementary roles for mediation analysis 
+and pleiotropy testing in hotspot 
 dissection. Here, we clarify this argument by presenting a study of five expression trait 
 hotspots in a publicly available data set. At each hotspot, we identify 
-a set of local traits that have QTL within the hotspot region. We also identify a set of nonlocal traits with QTL 
-arising from the hotspot. We perform pleiotropy tests and mediation analyses for all local-nonlocal trait pairs. 
-Graphical summaries of our results enable us to conclude that one of the five hotspots is due to multiple QTL, 
-while the evidence at the four other hotspots is consistent with a single local trait affecting most or all of the 
-nonlocal associated traits. The Chr 5 hotspot
-
-I
+a set of local traits that have QTL within the hotspot region. We also identify a set of 
+nonlocal traits affected by the hotspot. We perform pleiotropy tests and mediation analyses 
+for all local-nonlocal trait pairs. 
+Graphical summaries of our results enable us to conclude that one of the five hotspots is 
+due to multiple QTL, 
+while the evidence at the four other hotspots is consistent with a single local trait 
+affecting most or all of the 
+nonlocal associated traits. The Chr 5 hotspot, initially attributed to *Pdx1*, likely 
+arises from multiple causal loci. Evidence for this includes the pleiotropy test statistics 
+(Figure ), which support presence of more than one QTL and the allelic series inferences, 
+which demonstrate multiple allelic series at the Chr 5 hotspot.
 
